---
title: 'OHIBC: data prep for wild-capture fisheries'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

here is a setup code chunk with many of the common parameters:
``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = FALSE, message = FALSE, warning = FALSE)

library(sp)
library(rgdal)
library(raster)
library(DT)

dir_git <- '~/github/ohibc'
source(file.path(dir_git, 'src/R/common.R'))  ### an OHIBC specific version of common.R
dir_rgn <- file.path(dir_git, 'prep/regions')  ### github: general buffer region shapefiles
dir_anx <- file.path(dir_M, 'git-annex/bcprep')


### goal specific folders and info
goal      <- 'fis'
scenario  <- 'v2016'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario)
dir_goal_anx <- file.path(dir_anx, goal, scenario)

### provenance tracking
source(file.path(dir_git, 'src/R/prov.R'))   
  ### Provenance tracking functions: must source at start to initialize prov_track
prov_run_tag <- 'standard run'

### support scripts
source(file.path(dir_git, 'src/R/poly_plot_scores.R')) 
  ### score plotting scripts
source(file.path(dir_git, 'src/R/rast_tools.R')) 
  ### raster plotting and analyzing scripts
source(file.path(dir_goal, 'fis_fxns.R'))
  ### goal- or script-specific support functions

### set up proj4string options: BC Albers and WGS84
p4s_wgs84 <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
p4s_bcalb <- '+proj=aea +lat_1=50 +lat_2=58.5 +lat_0=45 +lon_0=-126 +x_0=1000000 +y_0=0 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0'

```

fisheries reporting:
* fisheries reporting is in zones (denoted by polygons)
* some fisheries are reported by cells (4 km cells?)
  * dungeness crabs
  
  
We will work at 1 km resolution (?)
* create a raster of rgn_id at 1 km resolution
  * does this already exist?
    * yes: prep/spatial/ohibc_rgn_raster_1000m.tif
      * NOTE: the raster has some problems; e.g. Haida Gwaii is filled in
    * also 500 m raster if desired
* create a raster of fishery zones (for csv reporting)
  * at 1 km resolution
* rasterize layers for each year:
  * use nearest neighbor if reclassifying
  * use cell center (default) otherwise; this allows for gdal_rasterize
  * stack together all years of a single fishery for processing

Explore directories and files from DFO
``` {r explore_dfo_directories, eval = FALSE}
dir_dfo <- file.path(dir_anx, '_raw_data/dfo_khunter')
dir_list <- list.files(dir_dfo)
dir_list <- dir_list[file.info(file.path(dir_dfo, dir_list))$isdir] %>%
  file.path(dir_dfo, .)
#  [1] "aquaculture"               "dungeness_crab"            "ferry_terminals"          
#  [4] "geoduck"                   "groundfish_longline"       "groundfish_mgmt_areas"    
#  [7] "groundfish_trawl"          "marina_pnt"                "pac_fishery_mgmt_areas"   
# [10] "pac_fishery_mgmt_subareas" "pacific_sardine"           "salmon_enhancement"       
# [13] "spot_prawn"  

```

``` {r setup_fishery_dirs_and_fields}

fish_dirs <- c("dungeness_crab", "geoduck", "groundfish_longline", "groundfish_trawl", "pacific_sardine", "spot_prawn")

get_hdrs <- function(fish_shps, dir, dir_dfo) {
  hdrs <- vector(mode = 'list', length = length(fish_shps))
  for (i in 1:length(fish_shps)) { # i = 1
    shp <- fish_shps[i]
    dbf_fn <- file.path(dir_dfo, dir, paste0(shp, '.dbf'))
    tmp <- foreign::read.dbf(dbf_fn) %>% names()
    hdrs[[i]] <- tmp
    names(hdrs)[i] <- shp
  }
  if(length(unique(hdrs)) > 1) {
    message('Headers are not identical for all shapefiles in ', dir)
    return(FALSE)    
  }
  message('all headers are identical in ', dir)
  return(hdrs[[1]])
}

for (dir in fish_dirs) { # dir = 'groundfish_longline' # dir = 'spot_prawn'
  fish_fns <- list.files(file.path(dir_dfo, dir))
  fish_shps <- fish_fns[str_detect(fish_fns, '.shp$')] %>% ### $ indicates end of line!
    str_replace('.shp', '')
  hdrs <- get_hdrs(fish_shps, dir, dir_dfo)
  cat(sprintf('%s: %s', dir, paste(hdrs, collapse = ', ')))
}

# dungeness_crab: ID, ET_ID, GRIDCELL, PNT_COUNT, Catch_kg, Effort_hr, SEASON, Shape_Leng, Shape_Area
#   2001 - 2011; 16 digit time stamp (e.g. CrabTrap_Dungeness2001_20110222_115633)
# geoduck:        ID, ET_ID, GRIDCELL, PNT_COUNT, Catch_kg, Effort_hr, SEASON, Shape_Leng, Shape_Area
#   2003 - 2011; 16-digit time stamp (e.g. Dive_Geoduck2004_20110222_132708)
# groundfish_longline: BlockId, Sector, NumVessels, EffortHrs, SableWt, HalibWt, POPWt, HakeWt, OtherWt, Shape_Leng, Shape_Area
#   2006 - 2012; 4-digit extension (e.g. Longline_2006_BCA)
# groundfish_trawl:    BlockId, Sector, NumVessels, EffortHrs, SableWt, HalibWt, POPWt, HakeWt, OtherWt, Shape_Leng, Shape_Area, SableCPUE, HalibCPUE, POPCPUE, HakeCPUE, OtherCPUE
#   1996 - 2012; 4-digit extension (e.g. Trawl_1996_BCA)
# 
# Headers are not identical for all shapefiles in spot_prawn
# spot_prawn:
### in 2001-2009, headers: ID, ET_ID, GRIDCELL, PNT_COUNT, Catch_kg, Effort_trp, SEASON    
### in 2010-2011, headers: ID, ET_ID, GRIDCELL, PNT_COUNT, Catch_kg, Effort_hr, SEASON
#  2001 - 2011; 16-digit time stamp (e.g. PrawnTrap_Prawns2001_20110330_114657)
#
# pacific_sardine: OBJECTID, OID_1, Count_, Sum_EstTon, Shape_Leng, Shape_Area
#   OLD DATA - NEEDS UPDATING

```

``` {r setup_harvest_rasters, eval = FALSE}

if(!exists('rast_rgn')) 
  rast_rgn <- raster(file.path(dir_git, 'prep/spatial/ohibc_rgn_raster_1000m.tif'))

### dungeness_crab: 2001 through 2011
###   ID ET_ID GRIDCELL PNT_COUNT Catch_kg Effort_hr SEASON Shape_Leng Shape_Area
###    0 21178  60--298         1        0         0   2001      16000    1.6e+07
  
dir_shp <- file.path(dir_dfo, 'dungeness_crab')
crab_polys <- create_poly_list(dir_shp)
rasterize_list(crab_polys, fld_name = 'catch_kg')
rasterize_list(crab_polys, fld_name = 'effort_hr')


### geoduck: 2003 through 2011
#   ID ET_ID GRIDCELL PNT_COUNT Catch_kg Effort_hr SEASON Shape_Leng Shape_Area
# 0  0 26405  75--305         3      776        12   2003      16000    1.6e+07

dir_shp <- file.path(dir_dfo, 'geoduck')
geoduck_polys <- create_poly_list(dir_shp)
rasterize_list(geoduck_polys, fld_name = 'catch_kg')
rasterize_list(geoduck_polys, fld_name = 'effort_hr')
```

``` {r setup prawn rasters, eval = FALSE}
### spot_prawn: 2001 through 2011
#   ID ET_ID GRIDCELL PNT_COUNT Catch_kg Effort_trp SEASON Shape_Leng Shape_Area
# 0  0 23574  67--258         1        0          0   2001      16000    1.6e+07

dir_shp <- file.path(dir_dfo, 'spot_prawn')
prawn_polys <- create_poly_list(dir_shp)
rasterize_list(prawn_polys, fld_name = 'catch_kg', shp_area = 1.6e7) ### no shape_area field; assign manually assuming 4 km grid cells
rasterize_list(prawn_polys[1:9], fld_name = 'effort_trp', shp_area = 1.6e7) # 2001-2009
rasterize_list(prawn_polys[10:11], fld_name = 'effort_hr', shp_area = 1.6e7)  # 2010-2011
```

``` {r setup groundfish rasters, eval = FALSE}
# groundfish_longline: BlockId, Sector, NumVessels, EffortHrs, SableWt, HalibWt, POPWt, HakeWt, OtherWt, Shape_Leng, Shape_Area
#   2006 - 2012; 4-digit extension (e.g. Longline_2006_BCA)
# groundfish_trawl:    BlockId, Sector, NumVessels, EffortHrs, SableWt, HalibWt, POPWt, HakeWt, OtherWt, Shape_Leng, Shape_Area, SableCPUE, HalibCPUE, POPCPUE, HakeCPUE, OtherCPUE
#   1996 - 2012; 4-digit extension (e.g. Trawl_1996_BCA)

dir_shp <- file.path(dir_dfo, 'groundfish_longline')
gf_long_polys <- create_poly_list(dir_shp, trim_end = 4)
rasterize_list(gf_long_polys, fld_name = 'sablewt')
rasterize_list(gf_long_polys, fld_name = 'halibwt')
rasterize_list(gf_long_polys, fld_name = 'popwt')
rasterize_list(gf_long_polys, fld_name = 'hakewt')
rasterize_list(gf_long_polys, fld_name = 'otherwt')

dir_shp <- file.path(dir_dfo, 'groundfish_trawl')
gf_trawl_polys <- create_poly_list(dir_shp, trim_end = 4)
rasterize_list(gf_trawl_polys, fld_name = 'sablewt')
rasterize_list(gf_trawl_polys, fld_name = 'halibwt')
rasterize_list(gf_trawl_polys, fld_name = 'popwt')
rasterize_list(gf_trawl_polys, fld_name = 'hakewt')
rasterize_list(gf_trawl_polys, fld_name = 'otherwt')

```

``` {r setup sardine rasters, eval = FALSE}
### pacific_sardine: 2002, 2003, 2004 (?)
#   OBJECTID OID_1 Count_ Sum_EstTon Shape_Leng Shape_Area
# 0     2893  2893      3         50      16000    1.6e+07

dir_shp <- file.path(dir_anx, '_raw_data/dfo_khunter',
                      'pacific_sardine')
sardine_polys <- create_poly_list(dir_shp, trim_end = -1)
names(sardine_polys) <- paste0('sardines', c(2002:2004))
rasterize_list(sardine_polys, fld_name = 'sum_estton')

```

``` {r aggregate_to_rgn, eval = FALSE}
if(!exists('rast_rgn'))
  rast_rgn <- raster(file.path(dir_git, 'prep/spatial/ohibc_rgn_raster_1000m.tif'))

dir_rast <- file.path(dir_anx, goal, scenario, 'spatial')

zonestats <- function(stock, value, yrs, fun = 'sum') {
  fis_df <- data.frame()
  for (yr in yrs) { # yr <- 2001
    tmp_rast <- raster(file.path(dir_rast, sprintf('%s%s_%s.tif', stock, yr, value)))
    tmp_zonal <- zonal(tmp_rast, rast_rgn, fun = fun, digits=0, na.rm=TRUE) %>%
      as.data.frame()
    fis_df <- bind_rows(tmp_zonal, fis_df)
  }
names(fis_df) <- c('rgn_id', value, 'year')
return(fis_df)
}

### dungeness crabs!
fis_stock <- 'crabtrap_dungeness'
fis_value <- 'catch_kg'
fis_fun   <- 'sum'
yrs <- c(2001:2011)

catch_df  <- zonestats(fis_stock, 'catch_kg',  yrs, fis_fun)
effort_df <- zonestats(fis_stock, 'effort_hr', yrs, fis_fun)
write_csv(catch_df, file.path(dir_goal, sprintf('int/%s_catch_kg.csv', fis_stock)))
write_csv(effort_df, file.path(dir_goal, sprintf('int/%s_effort_hr.csv', fis_stock)))

### Geoduck!
fis_stock <- 'dive_geoduck'
fis_fun   <- 'sum'
yrs <- c(2003:2011)

catch_df  <- zonestats(fis_stock, 'catch_kg',  yrs, fis_fun)
effort_df <- zonestats(fis_stock, 'effort_hr', yrs, fis_fun)
write_csv(catch_df, file.path(dir_goal, sprintf('int/%s_catch_kg.csv', fis_stock)))
write_csv(effort_df, file.path(dir_goal, sprintf('int/%s_effort_hr.csv', fis_stock)))

### Spot prawns!
fis_stock <- 'prawntrap_prawns'
fis_fun   <- 'sum'
yrs <- c(2001:2002)

catch_df  <- zonestats(fis_stock, 'catch_kg',  yrs, fis_fun)
effort_df <- zonestats(fis_stock, 'effort_trp', yrs, fis_fun)
write_csv(catch_df, file.path(dir_goal, sprintf('int/%s_catch_kg.csv', fis_stock)))
write_csv(effort_df, file.path(dir_goal, sprintf('int/%s_effort_trp.csv', fis_stock)))

### Sardines!
fis_stock <- 'sardines'
fis_fun   <- 'sum'
yrs <- c(2002:2004)
  
catch_df  <- zonestats(fis_stock, 'sum_estton',  yrs, fis_fun)
write_csv(catch_df, file.path(dir_goal, sprintf('int/%s_catch_estton.csv', fis_stock)))

```

``` {r fish_mgmt_zones_by_rgn}
if(!exists('rast_rgn'))
  rast_rgn <- raster(file.path(dir_git, 'prep/spatial/ohibc_rgn_raster_1000m.tif'))

### count cells of each fishery management zone that fall within each OHIBC region.
### e.g. zone 100 (300 cells) falls between region 1 (100 cells) and region 2 (200 cells)
### set up df with columns: rgn_id, pfma_id, pfmsa_id, gfma_id using getValues.
### If all rasters are same extents, then the return from getValues() should be a vector
### of cell values that correspond to the same order of cells for each vector.

rast_pfma  <- raster(file.path(dir_anx, 'fis', scenario, 'spatial', 
                              'DFO_BC_PFMA_AREAS_50K_V3_1.tif'))
rast_pfmsa <- raster(file.path(dir_anx, 'fis', scenario, 'spatial', 
                              'DFO_BC_PFMA_SUBAREAS_50K_V3_1.tif'))
rast_gfma  <- raster(file.path(dir_anx, 'fis', scenario, 'spatial', 
                              'MajorMinor_GroundfishManagementAreas.tif'))

if(!(extent(rast_pfma)  == extent(rast_rgn))) {
  message('rgn and pfma extents not equal... fix them!')
} else {
  rgn_pfma_df <- data.frame(
      rgn_id   = round(getValues(rast_rgn)),
      pfma_id  = round(getValues(rast_pfma))) %>%
    group_by(rgn_id) %>%
    mutate(n_cells_rgn = n()) %>%
    group_by(pfma_id) %>%
    mutate(n_pfma_tot = n()) %>%
    group_by(rgn_id, pfma_id) %>%
    mutate(n_pfma_rgn = n()) %>%
    filter(!is.na(rgn_id)) %>%
    distinct()
}

if(!(extent(rast_pfmsa)  == extent(rast_rgn))) {
  message('rgn and pfmsa extents not equal... fix them!')
} else {
  rgn_pfmsa_df <- data.frame(
      rgn_id   = round(getValues(rast_rgn)),
      pfma_id  = round(getValues(rast_pfma)),
      pfmsa_id = round(getValues(rast_pfmsa))) %>%
    group_by(rgn_id) %>%
    mutate(n_cells_rgn = n(),
           pfmsa_label = paste0(pfma_id, '-', pfmsa_id)) %>%
    group_by(pfmsa_label) %>%
    mutate(n_pfmsa_tot = n()) %>%
    group_by(rgn_id, pfmsa_label) %>%
    mutate(n_pfmsa_rgn = n()) %>%
    filter(!is.na(rgn_id)) %>%
    distinct()
}

if(!(extent(rast_gfma)  == extent(rast_rgn))) {
  message('rgn and gfma extents not equal... fix them!')
} else {
  rgn_gfma_df <- data.frame(
      rgn_id   = round(getValues(rast_rgn)),
      gfma_id  = round(getValues(rast_gfma))) %>%
    group_by(rgn_id) %>%
    mutate(n_cells_rgn = n()) %>%
    group_by(gfma_id) %>%
    mutate(n_gfma_tot = n()) %>%
    group_by(rgn_id, gfma_id) %>%
    mutate(n_gfma_rgn = n()) %>%
    filter(!is.na(rgn_id)) %>%
    distinct()
}

x <- xx <- rast_pfma
y <- yy <- rast_pfmsa
values(xx)[is.na(values(x)) & !is.na(values(y))] <- 10000
values(xx)[is.na(values(y)) & !is.na(values(x))]  <- 3000
plot(xx)

### WHY ARE THERE MISSING ZONES?

```

``` {r child = file.path(dir_git, 'src/templates/ohibc_prov_ftr.Rmd')}
```

