---
title: "R Markdown test rmd"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
  pdf_document:
    toc: true
---

-----

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
dir_git <- '~/github/ohibc'
library(dplyr)
library(readr)
library(stringr)

dir_test <- file.path(dir_git, 'rmd_testing')

### set up provenance tracking for this script:
source(file.path(dir_git, 'src/R/prov.R'))          
  ### initializes prov_track and loads functions

```

# data prep spatial layers

read in some data A and data B.  Modify some things, and save as layer 1.

``` {r read two files and save a layer, echo = FALSE, message = FALSE, warning = FALSE}
tableA_file <- file.path(dir_test, 'tableA.csv')
tableA      <- read.csv(tableA_file, stringsAsFactors = FALSE)

tableA1_file <- file.path('~', 'tableA2.csv')
tableA1      <- read.csv(tableA1_file, stringsAsFactors = FALSE)

tableB_file <- file.path(dir_test, 'tableB.csv')
tableB      <- read_csv(tableB_file)

table1_file <- file.path(dir_test, 'table1.csv')
write.csv(tableB, table1_file)

```

# data prep other layers

This sources a script that modifies another new layer (tableC) and saves it as table2 and table3, external to the direct Rmd path.  It also sources a separate script that reads a previously-saved layer (table 1) and gapfills, saving as table1a.

``` {r source a script that processes other layers, echo = FALSE, message = FALSE, warning = FALSE}

source(file.path(dir_test, '1_data_prep_sample.R'))

```


# model

sourced script reads in layers 1a, 2, and 3, and does some stuff, and saves the results as layers D and E.

``` {r run a model, echo = FALSE, message = FALSE, warning = FALSE}
source(file.path(dir_test, '2_model_sample.R'))

# tableA      <- read_csv(tableA_file)
# 
# write_csv(tableA, tableB_file)

```



``` {r display table, echo = FALSE, message = FALSE, warning = FALSE}
DT::datatable(table1,   ### does not display system and session info
              caption  = 'this is a sample table:',              
              rownames = FALSE,
              class    = 'stripe hover compact',
              options  = list(dom = 'tp'))
```

-----

# The next section (below the divider) is the footer, knitted as a child

The code for the R code chunk header looks like this:

`{r child = file.path(dir_git, 'src/templates/ohibc_prov_ftr.Rmd')}`

Note that it is executed as if it were just part of this main document; it numbers the section ("Provenance") as a continuation of the numbering, and the variables are evaluated in the environment of the main document.

``` {r child = file.path(dir_git, 'src/templates/ohibc_prov_ftr.Rmd')}
```
