---
title: "R Markdown test rmd"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
  pdf_document:
    toc: true
---

-----

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
dir_git <- '~/github/ohibc'
source(file.path(dir_git, 'src/R/common.R'))

dir_test <- file.path(dir_git, 'rmd_testing')

### set up provenance tracking for this script:
source(file.path(dir_git, 'src/R/prov.R'))          ### initializes prov_track and loads functions
dir_prov <- file.path(dir_test, 'prov')             ### sets the directory for final provenance .csv (for script_prov())
this_script_file <- file.path(dir_test, 'test.Rmd') ### can't recognize the current script name on its own :(

```

# read a table; capture prov info

``` {r read global species list, echo = FALSE, message = FALSE, warning = FALSE}
table1_file <- file.path(dir_test, 'table1.csv')
table1      <- read.csv(table1_file, stringsAsFactors = FALSE)
git_prov(table1_file)

table2_file <- file.path(dir_test, 'table2.csv')
write.csv(table1, table2_file)
git_prov(table2_file, type = 'Output')

table3_file <- file.path(dir_test, 'table3.csv')
write.csv(table1, table3_file)
git_prov(table3_file, type = 'Output')

table4_file <- file.path(dir_test, 'table4.csv')
git_prov(table4_file, type = 'output')

table3_file <- file.path(dir_neptune_data, 'git-annex/bcprep', 'table3.csv')
write.csv(table1, table2_file)
git_prov(table3_file, type = 'output')

```

``` {r display table, echo = FALSE}
DT::datatable(table1,   ### does not display system and session info
              caption  = 'this is a sample table:',              
              rownames = FALSE,
              class    = 'stripe hover compact',
              options  = list(dom = 'tp'))
```

-----

# The next section (below the divider) is the footer, knitted as a child

The code for the R code chunk header looks like this:

`{r child = file.path(dir_git, 'src/templates/ohibc_prov_ftr.Rmd')}`

Note that it is executed as if it were just part of this main document; it numbers the section ("Provenance") as a continuation of the numbering, and the variables are evaluated in the environment of the main document.

``` {r child = file.path(dir_git, 'src/templates/ohibc_prov_ftr.Rmd')}
```
