---
title: 'OHIBC: Layer check helper'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

source('~/github/ohibc/src/R/common.R')  ### an OHIBC specific version of common.R;
  ### includes library(tidyverse); library(stringr)

dir_calc <- '~/github/ohibc/calc_ohibc'

```


# Summary

Check layers used in toolbox (layers_files, layers_meta, layers_source, layers_targets) to make sure all layers are present and accounted for.

-----

``` {r}

source(file.path(dir_calc, 'explore/create_layer_to_target_lookup.R'))
       
source(file.path(dir_calc, 'explore/create_layer_source_lookup.R'))
       
```

``` {r}

layers_files   <- read_csv(file.path(dir_calc, 'master/layers_files_master.csv')) %>%
  rename(filename = fn_data)
layers_meta    <- read_csv(file.path(dir_calc, 'master/layers_meta_master.csv')) %>%
  select(-targets)
layers_source  <- read_csv(file.path(dir_calc, 'explore/int/layers_source_master.csv'))
layers_targets <- read_csv(file.path(dir_calc, 'explore/int/layers_targets_master.csv'))

layers_full <- layers_files %>%
  full_join(layers_meta,    by = 'layer') %>%
  full_join(layers_source,  by = c('layer', 'filename', 'dir_prep')) %>%
  full_join(layers_targets, by = 'layer')

### Manually fix "elements" layers
layers_elements <- layers_full %>%
  filter(str_detect(layer, '^element')) %>%
  mutate(date = date(),
         prep_script = 'element weighting layers are generated within functions.R',
         target = str_replace(layer, 'element_wts_', ''),
         target = str_replace(toupper(target), '_.+', ''))
layers_elements$dimension <- list(c('pressure', 'resilience'))
layers_elements <- layers_elements %>%
  unnest(dimension)

### Manually fix "FP" wildcaught weighting layer
layers_fp_wt <- layers_full %>%
  filter(str_detect(layer, 'fp_wildcaught')) %>%
  mutate(date = date(),
         prep_script = 'wildcaught weighting layer is generated within functions.R',
         target = 'FP',
         dimension = 'status')

layers_full <- layers_full %>%
  filter(!str_detect(layer, '^element|fp_wildcaught')) %>%
  bind_rows(layers_elements, layers_fp_wt) %>%
  arrange(layer)

write_csv(layers_full, file.path(dir_calc, 'explore/layers_full_list.csv'))

```

## Identify missing layers

Layers with no prep script may indicate:

* dummy layers (placeholders borrowed from global)
* mismatch in layer name or file name among matrices and layer lists
* problem with provenance tracking (e.g. script fails to check provenance for an already-prepared and not-recalculated output file)
* missing layer

``` {r identify_missing_layers}

missing <- layers_full %>%
  filter(is.na(prep_script) & !is.na(layer)) %>%
  group_by(layer, dir_prep, filename, name, 
           description, dimension) %>%
  summarize(target = paste(unique(target), collapse = ', ')) %>%
  ungroup() %>%
  distinct()

write_csv(missing, file.path(dir_calc, 'explore/xxx_missing_layers.csv'))

```

<font size = 1>

`r DT::datatable(missing)`

</font>

## Identify "extra" layers

Layers in 'output' or 'dummy' folders that are not included in the layers master list might be orphans.

``` {r identify_extra_layers}

### ID all .csv files in prep directories
prep_csvs <- list.files('~/github/ohibc/prep', 
                    pattern = '.csv$',
                    recursive = TRUE, 
                    full.names = TRUE)
### narrow down to just those in output folders
output_csvs <- prep_csvs[str_detect(prep_csvs, '/output/') & !str_detect(prep_csvs, '/vHS/')]

### ID all .csv files in prep directories
dummy_csvs <- list.files('~/github/ohibc/calc_ohibc/master/dummy_layers', 
                    pattern = '.csv$',
                    recursive = TRUE, 
                    full.names = TRUE)


  
extras <- c(output_csvs[!basename(output_csvs) %in% layers_full$filename],
            dummy_csvs[!basename(dummy_csvs) %in% layers_full$filename])

```

#### Extra 'output' csvs not used as layers:

`r paste0('\n* ', extras, collapse = '')`

These may be orphaned layers from prior calculations, or they might be legit (e.g. gapfilling).

