---
title: 'OHIBC Howe Sound: Habitat goal (CP, CS, HAB) layers prep'
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(raster)
library(rgdal)
library(rgeos)

source('~/github/ohibc/src/R/common.R')  ### an OHIBC specific version of common.R

dir_git     <- path.expand('~/github/ohibc') %>% path.expand()
dir_spatial <- file.path(dir_git, 'prep/spatial')  ### github: general buffer region shapefiles
dir_anx     <- file.path(dir_M, 'git-annex/bcprep')

### goal specific folders and info
goal      <- 'hab_cs_cp'
scenario  <- 'vHS'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario)
dir_v2016_anx <- file.path(dir_anx, goal, 'v2016')
dir_goal_anx <- file.path(dir_anx, goal, scenario)

### provenance tracking
library(provRmd); prov_setup()

### support scripts
source(file.path(dir_git, 'src/R/rast_tools.R')) 
  ### gdal_rast2, plotting

### set up the default BC projection to be BC Albers
p4s_bcalb <- c('bcalb' = '+init=epsg:3005')


```

# Summary

Clip existing habitat layers (salt marsh/forest raster, salt marsh and seagrass polygons) for HAB, CP, CS goals for Howe Sound assessment.  For CP, calculate values based on exposure, elevation, and protective potential weighting as required for different protective habitats.  For CS, calculate values based on carbon storage potential.

***questions to be answered***: 

* CP coastal forests: buffer, elevation, etc? 
* CS coastal forests: full watershed or buffer?

***

# Data Sources

* **Seagrass and salt marsh polygons**
    * **Reference**: CEC. 2016. North America’s Blue Carbon: Assessing Seagrass, Salt Marsh and Mangrove Distribution and Carbon Sinks. Montreal, Canada: Commission for Environmental Cooperation. 54 pp.
    * **Downloaded**: http://www.cec.org/sites/default/files/Atlas/Files/blue_carbon/BlueCarbon_Shapefile.zip (downloaded 8/25/16)
    * **Description**:  shapefiles of marine seagrass, mangrove, and salt marsh habitat in North America
    * **Native data resolution**: NA   
    * **Time range**: NA (data were collected over many years but no time series available)
    * **Format**:  vector shapefiles
* **Land cover raster 1990-2000-2010**
    * **Reference**: Land Use 1990, 2000 and 2010. 2015. Agriculture and Agri-Foods Canada.
    * **Downloaded**: http://open.canada.ca/data/en/dataset/18e3ef1a-497c-40c6-8326-aac1a34a0dec (downloaded 10/7/16)
    * **Description**:  Land cover raster covering all of Canada south of 60° N
    * **Native data resolution**: 30 meters   
    * **Time range**: 1990, 2000, 2010
    * **Format**:  GeoTIFF
* **Exposure polylines and coastal class polylines**
    * **Reference**: 
        * British Columbia Marine Conservation Analysis Project Team. 2011. Shorezone exposure. Marine Atlas of Pacific Canada: A Product of the British Columbia Marine Conservation Analysis. Available from www.bcmca.ca.
        * British Columbia Marine Conservation Analysis Project Team. 2011. Coastal classes. Marine Atlas of Pacific Canada: A Product of the British Columbia Marine Conservation Analysis. Available from www.bcmca.ca.
    * **Downloaded**: http://bcmca.ca/datafiles/sets/bcmca_eco_set_physical_atlas.zip (downloaded 11/5/15)
    * **Description**:  Marxan-prepared polylines representing exposure and coastal class along BC coastline
    * **Native data resolution**: NA; shorezone reference units are 400 m
    * **Time range**: NA
    * **Format**:  Polyline shapefiles
* **Elevation raster**
    * **Reference**: National Geophysical Data Center. 2006. 2-minute Gridded Global Relief Data (ETOPO2) v2. National Geophysical Data Center, NOAA. doi:10.7289/V5J1012Q [7/26/16].
    * **Downloaded**: https://www.ngdc.noaa.gov/mgg/global/etopo2.html (downloaded 7/26/16)
    * **Description**:  Global elevation and bathymetry
    * **Native data resolution**: 2 arc-minutes   
    * **Time range**: NA
    * **Format**:  NetCDF, GeoTIFF
* **Population density raster**
    * **NOTE**: Population weighting was considered based on InVEST coastal vulnerability model; however, after discussion with Andrew Day and Karin Bodtker of Vancouver Aquarium, population weighting will not be used. 
    * **Reference**: Center for International Earth Science Information Network - CIESIN - Columbia University. 2016. Documentation for the Gridded Population of the World, Version 4 (GPWv4). Palisades NY: NASA Socioeconomic Data and Applications Center (SEDAC). http://dx.doi.org/10.7927/H4D50JX4 Accessed 25 Oct 2016.
    * **Downloaded**: http://sedac.ciesin.columbia.edu/data/collection/gpw-v4/sets/browse (downloaded 10/25/16)
    * **Description**:  Population density raster, adjusted to UN 2015 values; pop density in persons per km^2^
    * **Native data resolution**: ???? 2.5 arc-minutes   
    * **Time range**: 2000, 2005, 2010, 2015, 2020 (only 2015 used for this analysis)
    * **Format**:  GeoTIFF
* **Squamish River watershed**
    * **Reference**:
    * **Downloaded**:  (downloaded 11/30/16)
    * **Description**:  Watershed outlines for British Columbia
    * **Native data resolution**: NA
    * **Time range**: NA
    * **Format**:  ESRI Shapefile modified from geodatabase
    
***
  
# Methods

Define extents for Howe Sound in BC Albers projection.

``` {r set up the extents for Howe Sound, based on Howe Sound region outline and watershed}
# hs_ws <- readOGR(dsn = file.path(dir_goal, 'watershed'), layer = 'howe_sound_watersheds')
# hs_rgn <- readOGR(dsn = file.path(dir_spatial), layer = 'howe_sound_rgn')
# bbox(hs_ws)
# #         min       max
# # x 1141118.3 1238724.7
# # y  471385.5  616826.2
# bbox(hs_rgn)
# #         min       max
# # x 1169571.2 1227753.5
# # y  473548.3  534590.4
ext_howe <- extent(c(1140000, 1240000, 470000, 620000))
```

Crop region rasters (at 500 m resolution) to Howe Sound extents, then use these to process layers:

* crop exposure polylines and coastal class polylines
    * identify appropriate exposure and coastal classes and rank from low to high (0 to 1) in terms of vulnerability
    * buffer out to (2 km? cap ends?)
    * raster::rasterize using the mean function
* reproject elevation raster and population raster (most recent year, just for weighting) to region raster parameters
* crop land cover raster to region raster extents

## Prepare support rasters

### create inland and offshore buffer area rasters

Create rasters at 500 m resolution for inland (2 km buffer) and offshore (3 nm buffer, which effectively captures all of Howe Sound).  These rasters will be used to mask analysis areas.

``` {r prep_region_rasters}

### Use one of the land-use rasters as the base, crop it down to the Howe Sound extents.

rast_base <- raster(file.path(dir_v2016_anx, 'raster/land_cover_1990_30m.tif')) %>%
  crop(ext_howe)
res(rast_base) <- 30

rast_rgn      <- gdal_rast2(src = file.path(dir_spatial, 'howe_sound_rgn'),
                                 rast_base = rast_base,
                                 dst = file.path(dir_goal_anx, 'raster/howe_rgn_30m.tif'),
                                 value = 'id',
                                 override_p4s = TRUE)
rast_inland   <- gdal_rast2(src = file.path(dir_spatial, 'howe_sound_inland_2km'),
                                 rast_base = rast_base,
                                 dst = file.path(dir_goal_anx, 'raster/howe_inland_2km_30m.tif'),
                                 value = 'rgn_id',
                                 override_p4s = TRUE)
rast_watershed <- gdal_rast2(src = file.path(dir_spatial, 'watershed/howe_sound_watersheds'),
                                 rast_base = rast_base,
                                 dst = file.path(dir_goal_anx, 'raster/howe_sound_watershed_30m.tif'),
                                 # value = 'WATERSHED_',
                                 override_p4s = TRUE)
rast_offshore <- gdal_rast2(src = file.path(dir_spatial, layer = 'howe_sound_offshore_3nm'),
                                 rast_base = rast_base,
                                 dst = file.path(dir_goal_anx, 'raster/howe_offshore_3nm_30m.tif'),
                                 value = 'rgn_id',
                                 override_p4s = TRUE)

```

### create population density and elevation rasters

**deprecated**: population weighting not used in OHIBC Howe Sound coastal protection model.

Prepare population density and elevation rasters by reprojecting global rasters to the extents, projection, and resolution of Howe Sound base region raster.  These rasters are used as weighting factors in the Coastal Protection goal.

``` {r prep pop and elev rasters}

rast_rgn <- raster(file.path(dir_goal_anx, 'raster/howe_rgn_30m.tif'))
  
topo_rast_file  <- file.path(dir_goal_anx, 'raster/howe_topo_30m.tif')
# pdens_rast_file <- file.path(dir_goal_anx, 'raster/howe_pdens.tif')

if(!file.exists(topo_rast_file)) {
  message('Creating file: \n  ', topo_rast_file)

  ### Gridded Pop of the World raster at Mazu: 
  ###   git-annex/globalprep/_raw_data/CIESEandCIAT_population/d2015
  
  # pop_gl <- raster(file.path(dir_anx, '../globalprep/_raw_data/CIESEandCIAT_population/d2015/pop_count_data', 'glp15ag.asc'))
  # crs(pop_gl) <- '+init=epsg:4326'
  # rast_pop   <- projectRaster(pop_gl, rast_rgn, method = 'ngb')
  
  # pop_dens_gl <- raster(file.path(dir_anx, 
  #                                 '../globalprep/_raw_data/CIESEandCIAT_population', 
  #                                 'd2015/pop_density_data/glfedens15/glds15ag',
  #                                 'dblbnd.adf'))
  # crs(pop_dens_gl) <- '+init=epsg:4326'
  # rast_pdens <- projectRaster(pop_dens_gl, rast_rgn) ### NOTE: uses bilinear
  # writeRaster(rast_pdens, pdens_rast_file, overwrite = TRUE)  
  
  ### ETOPO v2 in Mazu:
  ### git-annex/bcprep/_raw_data/bathymetry/ETOPO2v2_bathymetry
  etopo <- raster(file.path(dir_anx, '_raw_data/bathymetry/ETOPO2v2_bathymetry',
                            'ETOPO2v2c.tif'))
  ### add CRS and extent info:
  crs(etopo) <- '+init=epsg:4326'
  extent(etopo) <- c(-180, 180, -90, 90)
  
  rast_topo <- projectRaster(etopo, rast_rgn)
  
  writeRaster(rast_topo, topo_rast_file, overwrite = TRUE) 
  
} else {
  
  # message('Files exist: \n  ', paste(c(topo_rast_file, pdens_rast_file), collapse = '\n  '))
  # git_prov(c(topo_rast_file, pdens_rast_file), filetype = 'output')

  message('File exists: \n  ', topo_rast_file)
  git_prov(topo_rast_file, filetype = 'output')
  
}


```

### Prepare exposure raster

Exposure raster is created by extending a 2 km buffer from shoreline polylines to create polygons with values indicating levels of wave exposure, from 1 (very protected) to 6 (very exposed), with only values 1-4 appearing in Howe Sound.  The buffer polygons are rasterized, and where two buffers overlap (at boundaries between two exposure classes), the results are averaged to approximate an intermediate exposure class.

Because the exposure raster is based off a 2 km buffer from shoreline polylines, it also acts as a masking layer to avoid including inland habitats.

``` {r prep_exposure_raster}

rast_rgn <- raster(file.path(dir_goal_anx, 'raster/howe_rgn_30m.tif'))

expos_rast_file  <- file.path(dir_goal_anx, 'raster/howe_expos_30m.tif')

if(!file.exists(expos_rast_file)) {
  message('Creating file: \n  ', expos_rast_file)
  
  dir_bcmca <- file.path(dir_anx, '_raw_data/bcmca/bcmca_eco_set_physical_complete/MarxanData_Physical')
  
  # poly_coastal_class <- readOGR(dir_bcmca, 'BCMCA_ECO_Physical_CoastalClasses_MARXAN') %>%
  #   crop(rast_rgn)
  poly_exposure <- readOGR(dir_bcmca, 'BCMCA_ECO_Physical_Exposure_MARXAN') %>%
    crop(rast_rgn)
  
  poly_exp_2km <- poly_exposure %>%
    gBuffer(width = 2000, byid = TRUE, capStyle = 'FLAT')
  
  ### hunch: third digit is exposure: 
  # 6: Very Exposed
  # 5: Exposed
  # 4: Semi-exposed
  # 3: Semi-protected
  # 2: Protected
  # 1: Very Protected
  
  poly_exp_2km@data <- poly_exp_2km@data %>%
    mutate(exposure = floor((ID - 2400)/10))
  
  rast_expos <- rasterize(poly_exp_2km, rast_rgn, field = 'exposure', fun = 'mean')
  expos_max <- maxValue(rast_expos); expos_min <- 0; # minValue(rast_expos)
  rast_expos <- (rast_expos - expos_min) / (expos_max - expos_min)
  
  rast_expos <- mask(rast_expos, rast_rgn)
  
  writeRaster(rast_expos, expos_rast_file, overwrite = TRUE)
  
} else {
  
  message('File exists: \n  ', expos_rast_file)
  git_prov(expos_rast_file, filetype = 'output')
  
}
```

## Prep Howe Sound land use rasters

Because the land use rasters are not capturing the saltmarsh/wetland area of the Squamish estuary (instead, misclassifying these areas as forest or tree coverage), we will repair those for this analysis.  

* We will capture Howe Sound-specific rasters from the larger BC-wide rasters generated for OHIBC; the extents will capture Howe Sound and the Squamish watershed.
* Using a Google Earth image from 7/8/2010, I manually identified all areas that are wetland as opposed to forest, and created a shapefile of this region.
* Using this shapefile, we will reclassify any "forest" cells (values 41, 42, 45, 46) as undifferentiated wetland (71).
* The raster also misclassifies some of these areas as "settlement" (value 21); these areas will be reclassified as undifferentiated wetland (71).  This is especially a problem in the area north of the log sort.
* For the 2010 raster, the dredge spoils is shown as "other land", though the restoration efforts and the aerial photo seem to indicate that it is now grassland.  This will be fixed here as well.

``` {r fix_landuse_rasts}

howesound_landuse_files <- c('1990' = file.path(dir_goal_anx, 'raster/land_cover_1990_30m.tif'),
                   '2000' = file.path(dir_goal_anx, 'raster/land_cover_2000_30m.tif'),
                   '2010' = file.path(dir_goal_anx, 'raster/land_cover_2010_30m.tif'))

if(any(!file.exists(howesound_landuse_files))) {
  rast_rgn_30m <- raster(file.path(dir_goal_anx, 'raster/howe_rgn_30m.tif'))
  
  ohibc_landuse_files <- c('1990' = file.path(dir_v2016_anx, 'raster/land_cover_1990_30m.tif'),
                     '2000' = file.path(dir_v2016_anx, 'raster/land_cover_2000_30m.tif'),
                     '2010' = file.path(dir_v2016_anx, 'raster/land_cover_2010_30m.tif'))
  
  stack_landuse <- stack(ohibc_landuse_files) %>%
    crop(rast_rgn_30m)
  
  rast_estuary <- gdal_rast2(src = file.path(dir_goal, 'spatial', 'squamish_estuary.shp'),
                             rast_base = rast_rgn_30m,
                             dst = file.path(dir_goal, 'raster', 'squamish_estuary.tif'),
                             override_p4s = TRUE)
  
  values(stack_landuse)[values(stack_landuse) %in% c(21, 41, 42, 45, 46) & !is.na(values(rast_estuary))] <- 71
  
  values(stack_landuse[[3]])[values(stack_landuse[[3]]) == 91 & !is.na(values(rast_estuary))] <- 71
  
  writeRaster(stack_landuse, filename = howesound_landuse_files, bylayer = TRUE, overwrite = TRUE)
} else {
  message('Howe Sound landuse rasters already exist: \n  ', paste(howesound_landuse_files, collapse = '\n  '))
  git_prov(howesound_landuse_files)
}


```

## Analysis of coastal protection

Coastal protection for seagrass and salt marsh will be mitigated by exposure; habitat in areas with higher exposure will be valued as providing more protection.

Coastal protection for trees/forest will be mitigated by exposure and elevation (0 m = 1.0, 5 m or higher = 0).  Maximum inland reach is 2 km.  Carbon storage values will include all trees within the entire Squamish River watershed (including islands and western side of Howe Sound).

**deprecated**: All three will be scaled relative to population density, as log(pop density + 1).

### Seagrass status and trend

Seagrass coastal protection value will currently be left out of the calculation.  Nutrient pressure is insigniicant within Howe Sound; pressures for log booms, development, and sedimentation are likely more significant, but will require serious rethinking of the seagrass model.

### Salt Marsh status and trend

For salt marshes, we will determine rate of salt marsh change within 2 km inland and offshore.  Salt marsh presence will be established using categories 71, 73, and 74 from the Government of Canada 30 m land use rasters for 1990, 2000, and 2010.  

Salt marsh health status for coastal protection is determined by summing the weighted vulnerability value of salt marshes over the whole Howe Sound area.  Vulnerability value is weighted by exposure (0.25 = very protected, 1.0 = semi exposed, categories from BCMCA) and population density (log(ppl/km^2 + 1), then rescaled to highest value = 1).  As saltmarsh is lost, the sum drops; saltmarsh lost in exposed and/or highly populated areas counts as a higher penalty.  The reference point for a score of 1 will be the 1990 salt marsh sum(weighted vulnerability value).

Salt marsh trend will be based on percentage of change from 2000 to 2010 divided across the ten year interval.

Simple area of salt marsh extent will be included as well, for calculation of Carbon Storage and Habitat goals.


``` {r process_saltmarsh_rast}

rast_rgn_30m <- raster(file.path(dir_goal_anx, 'raster/howe_rgn_30m.tif'))
  
landuse_files <- c('1990' = file.path(dir_goal_anx, 'raster/land_cover_1990_30m.tif'),
                   '2000' = file.path(dir_goal_anx, 'raster/land_cover_2000_30m.tif'),
                   '2010' = file.path(dir_goal_anx, 'raster/land_cover_2010_30m.tif'))

stack_landuse <- stack(landuse_files) %>%
  crop(rast_rgn_30m)

stack_sm <- filter_rast(stack_landuse, vals = c(71, 73, 74), revalue = 1)

### multiply saltmarsh values by exposure to de-emphasize protective value
### near already protected areas
rast_expos_30m <- raster(file.path(dir_goal_anx, 'raster/howe_expos_30m.tif'))

stack_sm_expos <- stack_sm * rast_expos_30m

### write final rasters
sm_rast_files <- file.path(dir_goal_anx, sprintf('raster/hs_sm_expos_%s_30m.tif', c(1990, 2000, 2010)))
writeRaster(stack_sm_expos, filename = sm_rast_files, bylayer = TRUE, overwrite = TRUE)

### create dataframe to sum cell values

saltmarsh_df <- data.frame(year = c(1990, 2000, 2010))

for (i in 1:3) {
  saltmarsh_df$sum_sm[i] <- values(stack_sm_expos[[i]]) %>%
    sum(na.rm = TRUE)
  saltmarsh_df$area_sm_km2[i] <- sum(!is.na(values(stack_sm_expos[[i]]))) * (.03 * .03)
}

### create historical status based on 30% historical loss
saltmarsh_hist <- saltmarsh_df %>%
  filter(year == 1990) %>%
  mutate(sum_sm      = sum_sm/(1 - .3),
         area_sm_km2 = area_sm_km2/(1 - .3),
         year = 1900)

### create 2013 status based on this from Karin:
### The map shows a log sort –this is the brownfield being restored. I 
### estimated size based on that map and info in the plan. E.g., Site A is 
### about 30 hectares. Edith reports that from 1999 to 2013 the area labeled 
### dredge spoils was restored and is “over 15 hectares”
saltmarsh_2013 <- saltmarsh_df %>%
  filter(year == 2000) %>%
  mutate(sum_sm      = sum_sm/(1 - .3),
         area_sm_km2 = area_sm_km2/(1 - .3),
         year = 1900)

saltmarsh_df <- bind_rows(saltmarsh_hist, saltmarsh_df)

saltmarsh_df <- saltmarsh_df %>%
  mutate(cp_score_sm = sum_sm / first(sum_sm),
         cp_loss_sm  = (cp_score_sm - lag(cp_score_sm, 1)) / lag(cp_score_sm, 1),
         cp_trend_sm = cp_loss_sm / (year - lag(year)))

write_csv(saltmarsh_df, file.path(dir_goal, 'int/howe_saltmarsh_status_trend.csv'))


```

### Coastal forest status and trend

Coastal protection values for coastal forests are calculated similar to salt marsh; vulnerability factors include exposure and population, and in addition, elevation is considered.  Anything at sea level is vulnerable (1) up to an elevation of 5 m (vulnerability = 0).  This 5 m value is taken from InVEST, due to the inundation/storm surge from an extreme event.

Simple area of extent for coastal forests will be included as well, for use in calculating Carbon Storage and Habitat goals.

``` {r process_coastal_forest_rast}

rast_ws_30m <- raster(file.path(dir_goal_anx, 'raster/howe_sound_watershed_30m.tif'))
  
landuse_files <- c('1990' = file.path(dir_goal_anx, 'raster/land_cover_1990_30m.tif'),
                   '2000' = file.path(dir_goal_anx, 'raster/land_cover_2000_30m.tif'),
                   '2010' = file.path(dir_goal_anx, 'raster/land_cover_2010_30m.tif'))

stack_landuse <- stack(landuse_files) %>%
  crop(rast_ws_30m)

stack_cf <- filter_rast(stack_landuse, vals = c(41, 42, 45, 46), revalue = 1)

### multiply coastal forest values by exposure to de-emphasize protective value
### near already protected areas
rast_expos_30m <- raster(file.path(dir_goal_anx, 'raster/howe_expos_30m.tif'))

stack_cf_expos <- stack_cf * rast_expos_30m

### multiply by elevation vulnerability; areas above 5 m
### in elevation are not vulnerable to typical storm surge (from InVEST)
elev_lim <- 5
rast_elev_30m <- raster(file.path(dir_goal_anx, 'raster/howe_topo_30m.tif')) %>%
  calc(fun = function(x) {ifelse(x <= 0, NA,
                            ifelse(x > elev_lim, 0,
                              (1 - x/elev_lim))
                            )
                          })

stack_cf_expos_elev <- stack_cf_expos * rast_elev_30m

### write final elevation-adjusted exposure rasters
cf_elev_rast_files <- file.path(dir_goal_anx, sprintf('raster/hs_cf_expos_elev_%s_30m.tif', c(1990, 2000, 2010)))
writeRaster(stack_cf_expos_elev, filename = cf_elev_rast_files, bylayer = TRUE, overwrite = TRUE)

### Determine raw area within watershed for CS (no elevation, no exposure)
rast_ws_presence <- rast_ws_30m/rast_ws_30m

stack_cf_ws <- stack_cf * rast_ws_presence
cf_ws_rast_files <- file.path(dir_goal_anx, sprintf('raster/hs_cf_ws_%s_30m.tif', c(1990, 2000, 2010)))
writeRaster(stack_cf_ws, filename = cf_ws_rast_files, bylayer = TRUE, overwrite = TRUE)


### create dataframe to sum cell values

coastalforest_df <- data.frame(year = c(1990, 2000, 2010))

for (i in 1:3) {
  coastalforest_df$sum_cf[i] <- values(stack_cf_expos_elev[[i]]) %>%
    sum(na.rm = TRUE)
  coastalforest_df$area_cf_km2[i] <- sum(!is.na(values(stack_cf_expos_elev[[i]]))) * (.03 * .03)
  coastalforest_df$area_nonzero_cf_km2[i] <- sum(values(stack_cf_expos_elev[[i]]) != 0, na.rm = TRUE) * (.03 * .03)
  coastalforest_df$area_watershed_km2[i]  <- sum(values(stack_cf_ws[[i]]) != 0, na.rm = TRUE) * (.03 * .03)
}

coastalforest_df <- coastalforest_df %>%
  mutate(cp_score_cf = sum_cf / first(sum_cf),
         cp_loss_cf  = (cp_score_cf - lag(cp_score_cf, 1)) / lag(cp_score_cf, 1),
         cp_trend_cf = cp_loss_cf / (year - lag(year)),
         cs_score_cf_ws = area_watershed_km2 / first(area_watershed_km2),
         cs_loss_cf_ws  = (cs_score_cf_ws - lag(cs_score_cf_ws, 1)) / lag(cs_score_cf_ws, 1),
         cs_trend_cf_ws = cs_loss_cf_ws / (year - lag(year)))

write_csv(coastalforest_df, file.path(dir_goal, 'int/howe_forest_status_trend.csv'))

```

## results

### saltmarsh

`r knitr::kable(saltmarsh_df)`

### coastal forests

`r knitr::kable(coastalforest_df)`

***

``` {r prov_footer, results = 'asis'}
prov_wrapup()
```

