---
title: "IUCN vs NatureServe conservation status ranks"
  output: 
    html_document:
    toc: true
  theme: spacelab
    highlight: haddock
---
``` {r setup, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
library(tidyr)
library(dplyr)
library(stringr)
library(ggplot2)
```

# What is R Markdown?

R + Markdown

# How can you use R Markdown?

### Document your code

Use R Markdown to make your code pretty and share it with others.  R code in the .Rmd function exactly the same as in a regular .R script.  A reader can see your code, and what it does.

- Explain your code or teach it to others
- Show all code with outputs
- Basically: human-readable coding

Set variables for solar intensity at Earth's orbital radius, in W * m-2

```{r document code 1, echo = TRUE, eval = TRUE}
i_sol <- 1360 
```

Read planet orbital radii from .csv, and drop Pluto, because really.

```{r document code 2, echo = TRUE, eval = FALSE}
planet_df <- read.csv('data/planets.csv')

planet_df <- planet_df %>%
  filter(planet != 'Pluto')

glimpse(planet_df)
```

```{r document code 3, echo = TRUE, eval = FALSE}


```

### Document your workflow

Use R Markdown to highlight the workflow of your code, to share your process.  The code itself is not so important; instead, you can focus on the steps you used to get from input A to output B.

- Reproducibility and provenance: if they want to see the code, they can dig into the document.
- Focus on the process: inputs, outputs, and the steps in between
- Specific code is not emphasized except where critical

### Document your results

Use R Markdown to document your results.  The code itself is not important; even the intermediate steps are not so critical.  Instead, you can focus on your research question and present the results in nicely-formatted tables and graphs.

- Hide the work: if they want to see the process, they can dig into the document.
- Focus on the big picture: formatted writeup and fancy graphs
- Probably doesn't even look like code


In the actual code:
  
  Styling:
  - bold, italics
- hyperlink
- bullets
- html headers (#, ##, ###, etc.)
  - line breaks and spaces
  - inline r text
  - inline executable r code
  - block quotes
  - show the markdown quick reference in RStudio
  - images
  - headers at the top
  
  Chunks:
    - headers in chunks
  - echo
  - message
  - warning
  - others...
  
  Common Issues:
    - never setwd()
  - preview/rawgit (awesome but only for public repos)
  - ggplot2 needs to use print()
  - spaces/line breaks
  
  Advanced?:
  - tables
  - Themes
  - Set Knit Options
  - Knitchild??? (knitting multiple Rmd together)
  
  
  For the actual script I think we need at least:
    
    - 1 image
  - a couple plots
  - 4-6 code chunks?
  