---
title: 'OHIBC: goal_prep_howesound.Rmd'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = FALSE, message = FALSE, warning = FALSE)

library(raster)
library(rgdal)
library(rgeos)

source('~/github/ohibc/src/R/common.R')  ### an OHIBC specific version of common.R

dir_git     <- path.expand('~/github/ohibc')
dir_spatial <- file.path(dir_git, 'prep/spatial')  ### github: general buffer region shapefiles
dir_anx     <- file.path(dir_M, 'git-annex/bcprep')

### goal specific folders and info
goal      <- 'hab'
scenario  <- 'v2016'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario)
dir_goal_anx <- file.path(dir_anx, goal, scenario)

### provenance tracking
library(provRmd); prov_setup()

### support scripts
source(file.path(dir_git, 'src/R/rast_tools.R')) 
  ### gdal_rast2, plotting

### set up the default BC projection to be BC Albers
p4s_bcalb <- c('bcalb' = '+init=epsg:3005')

```

# Summary

Pull data layers to calculate Howe Sound goals

# Coastal Protection

## Salt marsh

In this model, salt marsh habitat is considered to be more important to coastal protection in areas that are densely populated or highly exposed to wave action.  Loss of habitat in such areas thus is weighed more heavily in the calculations.

$$\frac{P_{sm}}{P_{sm_{ref}}} = \displaystyle\sum_{i=1}^{n_c}(A_i * E_i * \rho_i) \bigg/ \displaystyle\sum_{i=1}^{n_{ref}}(A_i * E_i * \rho_i)$$

where $n_c$ is the set of current presence cells, $n_{ref}$ is the set of reference presence cells, $A$, $E$ and $\rho$ represent cell area, exposure, and population density.

## Seagrass

In this model, seagrass habitat is considered to be more important to coastal protection in areas that are densely populated or highly exposed to wave action.  As a pressure-based model applied to a fixed area, rather than an area-loss-based model, the pressure itself reduces the effective protective value in a given area.

$$\frac{P_{sg}}{P_{sg_{ref}}} = \displaystyle\sum_{j=1}^{n_c}((1 - p_j) * A_j * E_j * \rho_j) / A_{total}$$

where $p$ represents the nutrient pressure (relative to the maximum nutrient pressure found in British Columbia's coastal waters).

## Coastal forests

In this model, coastal forest habitat is considered to be more important to coastal protection in areas that are densely populated or highly exposed to wave action; in addition, coastal forest protective value is diminished as elevation increases.  Coastal forest protection is not considered relevant above a reference elevation of 5 m, beyond which storm surge and inundation is highly unlikely.

$$\frac{P_{cf}}{P_{cf_{ref}}} = \displaystyle\sum_{k=1}^{n_c}(A_k * E_k * \rho_k * h_k) \bigg/ \displaystyle\sum_{k=1}^{n_{ref}}(A_k * E_k * \rho_k * h_k)$$

where $h$ represents a height adjustment factor: $h = \frac{elev_{ref} - elev}{elev_{ref}}$ where $elev_{ref}$ is 5 m. $h = 0$ above 5 m, and $h = 1$ at sea level.

## Goal model

The coastal protection goal model is the weighted mean of the three protective habitats, weighted by reference area (as a proportion of total reference area $A_{ref_T}$ and protection weight factor $w/w_{ref}$.  

$$x_{CP} = \displaystyle\sum_{i=1}^{n}\left(\frac{P_i}{P_{i_{ref}}} * w_i * A_i\right) \bigg/ \displaystyle\sum_{i=1}^{n}\left(w_i * A_i\right)$$

Protection weights are assigned based on vulnerability values from [InVEST Coastal Vulnerability Model](http://data.naturalcapitalproject.org/nightly-build/invest-users-guide/html/coastal_vulnerability.html):  

| Vulnerability     | Very Low   | Low   | Moderate   | High   | Very High   |
| ----              | :--------: | :---: | :--------: | :----: | :---------: |
| Score             | 1          | 2     | 3          | 4      | 5           |
| Natural Habitats  | Coral reef; mangrove; coastal forest | High dune; marsh | Low dune | Seagrass; kelp | No habitat |
| Protection weight | 4          | 3     | 2          | 1      | 0           |

``` {r load_cp_data}

cp_cf <- read_csv(file.path(dir_goal, 'int/howe_forest_status_trend.csv')) %>%
  select(status_yr = year, area = area_nonzero_cf_km2, score = cf_score) %>%
  filter(status_yr > 1990) %>%
  mutate(hab = 'cf')
cp_cf$trend <- lm(score ~ status_yr, data = cp_cf)[['coefficients']]['status_yr']

cp_sm <- read_csv(file.path(dir_goal, 'int/howe_saltmarsh_status_trend.csv')) %>%
  select(status_yr = year, area = area_sm_km2, score = sm_score) %>%
  mutate(hab = 'sm')
cp_sm$trend <- lm(score ~ status_yr, data = cp_sm)[['coefficients']]['status_yr']

cp_sg <- read_csv(file.path(dir_goal, 'int/howe_seagrass_status_trend.csv')) %>%
  select(status_yr = year, area = sg_area_km2, score = score_sg) %>%
  mutate(hab = 'sg')
cp_sg$trend <- lm(score ~ status_yr, data = cp_sg %>% 
                    filter(status_yr %in% c(2003:2013)))[['coefficients']]['status_yr']
cp_weights <- c('sg' = 1, 'sm' = 3, 'cf' = 4)

cp_scores <- cp_cf %>%
  bind_rows(cp_sm) %>%
  bind_rows(cp_sg) %>%
  mutate(hab_weight = cp_weights[hab])
  
trend_yrs <- data.frame('hab' = c('cf', 'sm', 'sg'),
                        'trend_yrs' = c('2000-2010', '2000-2010', '2003:2013'))

cp_scores_summary <- cp_scores %>%
  group_by(hab) %>%
  mutate(a_ref = first(area)) %>%
  filter(status_yr == max(status_yr)) %>%
  ungroup() %>%
  mutate(cp_st_tot = sum(score * hab_weight * area) / sum(hab_weight * area),
         cp_tr_tot = sum(trend * hab_weight * area) / sum(hab_weight * area)) %>%
  left_join(trend_yrs, by = 'hab')
  
write_csv(cp_scores_summary, file.path(dir_goal, 'output', 'cp_scores_summary.csv'))

knitr::kable(cp_scores_summary, caption = 'Coastal Protection')

```               
                  
***

# Carbon Storage

Carbon storage potential depends on the extent and health of a habitat, and the amount of carbon effectively sequestered in a given extent of habitat.

## Salt marsh and Coastal Forests

Condition of each of these habitats will be scored based on its current extent relative to the extent in 1990.

$$\frac{C_{sm}}{C_{sm_{ref}}} = \frac{A_c}{A_{ref}}$$

where $A_c$ is the set of current extent of habitat, $A_{ref}$ is the reference extent of habitat.  Using extent rasters at equal area projection, the condition can be calculated simply using the ratio of cell counts for current and reference time periods.

## Seagrass

Since seagrass condition is based on pressure rather than area, the model is different from saltmarsh and forests.

$$\frac{C_{sg}}{C_{sg_{ref}}} = \displaystyle\sum_{j=1}^{n_c}((1 - p_j) * A_j) / A_{total}$$

where $p$ represents the nutrient pressure.  Using an equal area projection raster and a fixed habitat extent, we can calculate this more simply by $$1 - \bar{p}$$.

## Goal model

The carbon storage goal model is the weighted mean of the three carbon-storing habitats, weighted by reference area (as a proportion of total reference area $A_{ref_T}$ and carbon storage factor $w/w_{ref}$.  

$$x_{CS} = \displaystyle\sum_{i=1}^{n}\left(\frac{C_i}{C_{i_{ref}}} * w_i * A_i\right) \bigg/ \displaystyle\sum_{i=1}^{n}\left(w_i * A_i\right)$$

Carbon burial rates (gC m^-2^ yr^-1^)

* Salt marsh: 218 +/- 24 gC m^-2^ yr^-1^ (mean +/- SE)
* Seagrasses: 138 +/- 38 gC m^-2^ yr^-1^ 
* Coastal forests
    * temperate: 5.1 +/- 1.0 gC m^-2^ yr^-1^ 
    * boreal: 4.6 +/- 2.1 gC m^-2^ yr^-1^ 

Source: Mcleod et al. 2011. A blueprint for blue carbon: toward an improved understanding of the role of vegetated coastal habitats in sequestering C02. Frontiers in Ecology 9(10): 552-560, DOI

``` {r load_cs_data}

cs_cf <- read_csv(file.path(dir_goal, 'int/howe_forest_status_trend.csv')) %>%
  select(status_yr = year, area = area_nonzero_cf_km2, score = cf_score) %>%
  filter(status_yr > 1990) %>%
  mutate(hab = 'cf')
cs_cf$trend <- lm(score ~ status_yr, data = cs_cf)[['coefficients']]['status_yr']

cs_sm <- read_csv(file.path(dir_goal, 'int/howe_saltmarsh_status_trend.csv')) %>%
  select(status_yr = year, area = area_sm_km2, score = sm_score) %>%
  mutate(hab = 'sm')
cs_sm$trend <- lm(score ~ status_yr, data = cs_sm)[['coefficients']]['status_yr']

cs_sg <- read_csv(file.path(dir_goal, 'int/howe_seagrass_status_trend.csv')) %>%
  select(status_yr = year, area = sg_area_km2, score = score_sg) %>%
  mutate(hab = 'sg')
cs_sg$trend <- lm(score ~ status_yr, data = cs_sg %>% 
                    filter(status_yr %in% c(2003:2013)))[['coefficients']]['status_yr']

cs_weights <- c('sg' = 138, 'sm' = 218, 'cf' = 4.6)

cs_scores <- cs_cf %>%
  bind_rows(cs_sm) %>%
  bind_rows(cs_sg) %>%
  mutate(hab_weight = cs_weights[hab])
  
cs_scores_summary <- cs_scores %>%
  group_by(hab) %>%
  mutate(a_ref = first(area)) %>%
  filter(status_yr == max(status_yr)) %>%
  ungroup() %>% 
  mutate(cs_st_tot = sum(score * hab_weight * area) / sum(hab_weight * area),
         cs_tr_tot = sum(trend * hab_weight * area) / sum(hab_weight * area))
  
write_csv(cs_scores_summary, file.path(dir_goal, 'output', 'cs_scores_summary.csv'))


knitr::kable(cs_scores_summary, caption = 'Carbon Storage')

```               

***

# Habitats

Biodiverse habitats are scored based on the relative extent and health.  They are not weighted when combined - each habitat is assumed to have equal importance to the others.  Habitats not present in a region will not be counted against the score.

## Coastal Forests

Since they do not directly support marine biodiversity, coastal forests are not included in this goal.

## Salt marsh

As for carbon storage, condition of salt marsh habitat will be scored based on its current extent relative to the extent in 1990.

$$\frac{C_{sm}}{C_{sm_{ref}}} = \frac{A_c}{A_{ref}}$$

where $A_c$ is the set of current extent of habitat, $A_{ref}$ is the reference extent of habitat.  Using extent rasters at equal area projection, the condition can be calculated simply using the ratio of cell counts for current and reference time periods.

## Seagrasses

As in carbon storage: since seagrass condition is based on pressure rather than area, the model is different from saltmarsh and forests.

$$\frac{C_{sg}}{C_{sg_{ref}}} = \displaystyle\sum_{j=1}^{n_c}((1 - p_j) * A_j) / A_{total}$$

where $p$ represents the nutrient pressure (compared to a reference of zero human-introduced nutrients).  Using an equal area projection raster and a fixed habitat extent, we can calculate this more simply by $$1 - \bar{p}$$.

## Goal model

The habitat goal model is the unweighted mean of the two biodiversity-supporting marine habitats.  

$$x_{HAB} = \displaystyle\sum_{i=1}^{n}\left(\frac{C_i}{C_{i_{ref}}} * \frac{A_i}{A_T}\right)$$

``` {r load_hab_data}

hab_cf <- read_csv(file.path(dir_goal, 'int/howe_forest_status_trend.csv')) %>%
  select(status_yr = year, area = area_nonzero_cf_km2, score = cf_score) %>%
  filter(status_yr > 1990) %>%
  mutate(hab = 'cf')
hab_cf$trend <- lm(score ~ status_yr, data = hab_cf)[['coefficients']]['status_yr']

hab_sm <- read_csv(file.path(dir_goal, 'int/howe_saltmarsh_status_trend.csv')) %>%
  select(status_yr = year, area = area_sm_km2, score = sm_score) %>%
  mutate(hab = 'sm')
hab_sm$trend <- lm(score ~ status_yr, data = hab_sm)[['coefficients']]['status_yr']

hab_sg <- read_csv(file.path(dir_goal, 'int/howe_seagrass_status_trend.csv')) %>%
  select(status_yr = year, area = sg_area_km2, score = score_sg) %>%
  mutate(hab = 'sg')
hab_sg$trend <- lm(score ~ status_yr, data = hab_sg %>% 
                    filter(status_yr %in% c(2003:2013)))[['coefficients']]['status_yr']

hab_scores <- hab_cf %>%
  bind_rows(hab_sm) %>%
  bind_rows(hab_sg)
  
hab_scores_summary <- hab_scores %>%
  group_by(hab) %>%
  filter(hab != 'cf') %>%
  filter(status_yr == max(status_yr)) %>%
  ungroup() %>% 
  mutate(hab_st_tot = sum(score) / n(),
         hab_tr_tot = sum(trend) / n())
  
write_csv(hab_scores_summary, file.path(dir_goal, 'output', 'hab_scores_summary.csv'))

knitr::kable(hab_scores_summary, caption = 'Habitats')

```     

***


``` {r prov_footer, results = 'asis'}
prov_wrapup()
```

