---
title: "CMSY for BC stocks"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---


This script reruns catch-MSY on all SAUP data for stocks in FAO area 67 (British Columbia).

``` {r setup, echo = F, message = FALSE, warning = FALSE, results= "hide"}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/', message = FALSE, warning = FALSE)

dir_git     <- path.expand('~/github/ohibc')
source(file.path(dir_git, 'src/R/common.R'))  
dir_spatial <- file.path(dir_git, 'prep/_spatial')  
dir_anx     <- file.path(dir_M, 'git-annex/bcprep')
dir_goal <- file.path(dir_git, "prep/fis/v2017")

library(datalimited)
library(doParallel)
registerDoParallel(cores = 8)

options(scipen=999)
```

# Get SAUP data for BC prepped for CMSY

Read in the stock catch by region calculated for global OHI 2017 from SAUP data. Filter just for species records (taxon key > 599999). We also need to add resilience measures for the cmsy model. We can use the resilience lookup table from global
```{r}

res <- read_csv("~/github/ohiprep/globalprep/fis/v2017/data/taxon_resilience_lookup.csv")

catch <- read_csv(file.path(dir_M,'git-annex/globalprep/fis/v2017/int/stock_catch_by_rgn.csv')) %>%
  rename(common = taxon_common_name, fao_id = fao_rgn, species=taxon_scientific_name) %>%
  filter(fao_id == 67,
         taxon_key > 599999) %>%
  group_by(year, species, common, stock_id) %>%
  mutate(annual_catch = sum(tons)) %>%
  select(year, sciname = species, common, annual_catch) %>%
  distinct() %>%
  left_join(res, by = c("sciname", "common")) %>%
  select(-X1, -SpecCode)

#how many stocks?

length(unique(catch$sciname)) #92

# how many of these stocks are found in Canadian waters in 67?

ca_stocks <- read_csv(file.path(dir_M,'git-annex/globalprep/fis/v2017/int/stock_catch_by_rgn.csv')) %>%
  filter(fao_rgn == 67,
         rgn_id==218,
         taxon_key > 599999) %>%
  select(taxon_scientific_name, taxon_common_name, stock_id) %>%
  distinct()

nrow(ca_stocks) 

DT::datatable(ca_stocks)
```

There are **81 stocks** in Canadian BC waters. Let's filter out the catch to just incude those found in Canadian waters. Also need to filter out stocks that dont have at least a 20 yr time series and an annual avg catch of 1000 tons


```{r}

stocks_for_cmsy <- catch %>%
  filter(sciname %in% ca_stocks$taxon_scientific_name,
         str_detect(common, "almon")==FALSE) %>% #remove salmon
  group_by(stock_id)%>%
  mutate(nyrs = n(),                       #get the total number of years the stock has records for   
         avg_ann_catch = mean(annual_catch)) %>%    #calculate the mean catch over all catch years
  mutate(run_cmsy = ifelse(nyrs > 19 & avg_ann_catch >= 1000, "yes", "no")) %>%
  select(sciname, common, nyrs, avg_ann_catch, run_cmsy) %>%
  distinct()

DT::datatable(stocks_for_cmsy)

```

There are only **27 stocks** to run through CMSY.


```{r}

#prepping the catch data for CMSY model
catch_for_cmsy <- catch %>%
  filter(sciname %in% ca_stocks$taxon_scientific_name,
         str_detect(common, "almon")==FALSE) %>% #remove salmon
  group_by(stock_id) %>%
  mutate(nyrs = n(),                       #get the total number of years the stock has records for   
         avg_ann_catch = mean(annual_catch)) %>%    #calculate the mean catch over all catch years
  filter(nyrs > 19,
         avg_ann_catch >= 1000)

length(unique(catch_for_cmsy$stock_id))
```

# Run CMSY

```{r, eval=F}

if (!file.exists(file.path(dir_M,"git-annex/bcprep/fis/v2017/cmsy_fits/cmsy-fits.rds"))) {
  
cmsy_fits <- plyr::dlply(catch_for_cmsy, c("stock_id", "common"), function(x) {
  
    #make sure the data is ordered from 1950 to 2010
    x <- arrange(x,year)
    out <- cmsy(ct = x$annual_catch, yr = x$year,  start_r = resilience(x$Resilience[1]), 
      reps = 2e6)
    out$year <- x$year
    out
  }, .parallel = TRUE)
saveRDS(cmsy_fits, file = file.path(dir_M,"git-annex/bcprep/fis/v2017/cmsy_fits/cmsy-fits.rds"))
} else {
  cmsy_fits <- readRDS(file.path(dir_M,"git-annex/bcprep/fis/v2017/cmsy_fits/cmsy-fits.rds"))
}

fake_data <- data.frame(bbmsy_q2.5 = NA, bbmsy_q25 = NA, bbmsy_q50 = NA, 
  bbmsy_q75 = NA, bbmsy_q97.5 = NA)

cmsy_bbmsy <- plyr::ldply(cmsy_fits, function(x) {
  bbmsy_cmsy <- x$biomass[, -1] / x$bmsy
  bbmsy_out <- tryCatch({
    bbmsy_out <- summarize_bbmsy(bbmsy_cmsy)
    bbmsy_out$year <- x$year
    bbmsy_out}, error = function(e) fake_data)
})
cmsy_bbmsy$model <- "CMSY"

write.csv(cmsy_bbmsy,file='v2017/int/cmsy_bbmsy.csv')

```

# Plot bbmsy for all stocks

Remove the stocks for which we have assessments from RAM. This includes:
Pacific Halibut, Lingcod, Pacific ocean perch, rock sole and sablefish

```{r}

cmsy_bbmsy <- read_csv("v2017/int/cmsy_bbmsy.csv") %>%
  filter(!common %in% c("Pacific halibut", "Pacific cod", "Lingcod", "Pacific ocean perch", "Rock sole", "Sablefish"))
  
ggplot(cmsy_bbmsy, aes(x = year, y = bbmsy_mean)) +
  geom_hline(yintercept = 1, col = "red") +
  geom_line() +
  geom_ribbon(aes(ymin=bbmsy_mean-bbmsy_sd,ymax=bbmsy_mean+bbmsy_sd),alpha=0.3) +
  facet_wrap(~common)
```

Compare the stocks we have in RAM

```{r}

cmsy_bbmsy <- read_csv("v2017/int/cmsy_bbmsy.csv") %>%
  filter(common %in% c("Pacific halibut", "Pacific cod", "Lingcod", "Pacific ocean perch", "Rock sole", "Sablefish")) 

#plot RAM species
stocks_ts_df <- bind_rows(read_csv(file.path(dir_goal, 'output/ram_b_bmsy.csv')),
                     read_csv(file.path(dir_goal, 'output/ram_f_fmsy.csv'))) %>%
  left_join(read_csv(file.path(dir_goal, 'raw', '1_ram_ids_to_names.csv'))) %>%
  ungroup() %>%
  mutate(param = str_replace_all(tolower(param), '[^a-z]', '_') %>%
        str_replace_all('u', 'f')) %>%
  filter(!stock_name_mod %in% c("Albacore tuna N Pac", "Bocaccio BC All")) %>%
  mutate(common = case_when(
    stock_name_mod == "Lingcod SoG" ~ "Lingcod",
    stock_name_mod == "Pacific cod Hecate Strait" ~ "Pacific cod",
    stock_name_mod == "Pacific cod QCS" ~ "Pacific cod", 
    stock_name_mod == "Pacific halibut N Pac" ~ "Pacific halibut",
    stock_name_mod == "Pacific Ocean perch QCI" ~ "Pacific ocean perch",
    stock_name_mod == "Pacific Ocean perch WCVanI" ~ "Pacific ocean perch",
    stock_name_mod == "Rock sole Hecate Strait" ~ "Rock sole",
    stock_name_mod == "Rock sole QCS" ~ "Rock sole",
    stock_name_mod == "Sablefish BC All" ~ "Sablefish"
  )) %>%
  filter(param == "b_bmsy") %>%
  left_join(cmsy_bbmsy, by = c("common", "year")) %>%
  filter(year>1990)

  
ggplot(stocks_ts_df, aes(x = year)) +
  geom_hline(yintercept = 1, col = "red") +
  geom_line(aes(y = bbmsy_mean), color = "blue") +
  geom_line(aes(y = value), color = "black") +
  facet_wrap(~stock_name_mod) +
  ylab("B/Bmsy")
```



