---
title: "Exploring fisheries data"
author: "Jamie Afflerbach"
date: "10/6/2017"
output: html_document
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = FALSE, message = FALSE, warning = FALSE)

dir_git     <- path.expand('~/github/ohibc')
source(file.path(dir_git, 'src/R/common.R'))  
dir_spatial <- file.path(dir_git, 'prep/_spatial')  
dir_anx     <- file.path(dir_M, 'git-annex/bcprep')


### goal specific folders and info
goal      <- 'fis'
scenario  <- 'v2017'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario)
dir_goal_anx <- file.path(dir_anx, goal, scenario)

### provenance tracking
library(provRmd); prov_setup()

### Kobe plot functions
source(file.path(dir_goal, 'kobe_fxns.R'))
library(plotly)
```

```{r}

stocks_ts_df <- bind_rows(read_csv(file.path(dir_goal, 'output/ram_b_bmsy.csv')),
                     read_csv(file.path(dir_goal, 'output/ram_f_fmsy.csv'))) %>%
  left_join(read_csv(file.path(dir_goal, 'raw', '1_ram_ids_to_names.csv'))) %>%
  ungroup() %>%
  mutate(param = str_replace_all(tolower(param), '[^a-z]', '_') %>%
           str_replace_all('u', 'f')) %>%
  filter(year > 1994)

```

There are `r length(unique(stocks_ts_df$stock_id))` RAM stocks used for the FIS sub-goal.

# RAM metrics for each stock

```{r}

b_f_plot <- ggplot(stocks_ts_df, aes(x = year, y = value, color = param)) +
  geom_line() +
  facet_wrap(~stock_name_mod)

b_f_plot

```

#Catch over time for each stock
The catch values here come from the RAM database as well.

```{r}

ram_catch <- read_csv(file.path(dir_goal, 'output/ram_catch.csv')) %>%
            filter(stock_name %in% stocks_ts_df$stock_name,
                   year > 1994)

catch_plot <- ggplot(ram_catch, aes(x = year, y = value)) +
  geom_line() +
  facet_wrap(~stock_name)

catch_plot

```

Remove tuna

```{r}

ram_catch_no_tuna <- ram_catch %>%
  filter(stock_id != "ALBANPAC")

catch_plot <- ggplot(ram_catch_no_tuna, aes(x = year, y = value)) +
  geom_line() +
  facet_wrap(~stock_name)

catch_plot

```

```{r}

## plotting fishery catch weighting by region
stock_plot_df <- read_csv(file.path(dir_goal, 'summary/fis_from_functions.csv')) %>%
  filter(!is.na(score)) %>%
  group_by(region_id, year) %>%
  mutate(total_catch = sum(rgn_catch),
         rgn_catch_pct = rgn_catch / total_catch,
         total_score = sum(score * rgn_catch) / total_catch) %>%
  ungroup() %>%
  left_join(get_rgn_names(), by = c('region_id' = 'rgn_id')) %>%
  filter(rgn_catch > 0,
         year > 1994)

```

## Contribution of regional catch by stock

Breaking this up by region and stock

```{r}

stock_plot_df$region_id = as.character(stock_plot_df$region_id)

ggplot(stock_plot_df, aes(x = year, y = rgn_catch_pct, color = stock_id)) + geom_line() + facet_wrap(~ rgn_name)

```

## Tuna

Looking just at albacore tuna, the offshore region is entirely dependent on Albacore tuna while this stock makes up a very small portion of the catch elsewhere.

```{r}

t <- filter(stock_plot_df, stock_id == "ALBANPAC" & basis == "F_Fmsy, B_Bmsy")
t$region_id = as.character(t$region_id)
ggplot(t, aes(x = year, y = rgn_catch_pct, color = rgn_name)) + geom_line()

```


# B/Bmsy estimates from the global CMSY data

Pulling in the B/Bmsy estimates for stocks in FAO area 67 from this year (2017) for consideration in the model.

```{r}

cmsy <- read_csv("~/github/ohiprep/globalprep/fis/v2017/data/cmsy_bbmsy.csv") %>%
        filter(year > 1994)

mean_catch <- read.csv("~/github/ohiprep/globalprep/fis/v2017/data/mean_catch.csv") %>%
   mutate(stock_id_taxonkey = as.character(stock_id_taxonkey)) %>%
   mutate(taxon_key = str_sub(stock_id_taxonkey, -6, -1)) %>%
   mutate(stock_id = substr(stock_id_taxonkey, 1, nchar(stock_id_taxonkey)-7)) 

bc_data <- mean_catch %>%
  left_join(cmsy, by = c("stock_id", "year")) %>%
        filter(!is.na(bbmsy_mean)) %>%
        filter(rgn_id == 218) %>%
        separate(stock_id, int = c("stock", "fao"), sep = "-") %>%
        filter(fao == 67)

ggplot(bc_data, aes(x = year, y = bbmsy_mean, color = common)) +
  geom_line()
        
## why isn't herring in the plot (I assume its because there is too little catch to run the data limited method)

catch <- read_csv(file.path(dir_M,'git-annex/globalprep/fis/v2017/int/stock_catch_by_rgn.csv')) %>%
  rename(common = taxon_common_name, fao_id = fao_rgn, species=taxon_scientific_name) %>%
  filter(species == "Clupea pallasii pallasii",
         fao_id == 67,
         taxon_key > 599999) %>%
  group_by(fao_id, year) %>%
  mutate(annual_catch = sum(tons)) %>%
  select(year, species,common,annual_catch) %>%
  distinct()

ggplot(catch, aes(x = year, y = annual_catch))+
  geom_area()

## need an annual average catch of 1000 tons
mean(catch$annual_catch) #151081.3  this checks out.... and there is a 20 year time history so unsure why this doesn't have a b/bmsy

library(datalimited)

herring_cmsy <- cmsy(yr = catch$year, ct = catch$annual_catch, reps = 2e6) #able to finally get convergence with more reps (2e6)

#lets look at results

h_bbmsy <- herring_cmsy$bbmsy

bmsy <- ggplot(h_bbmsy, aes(x = year)) +
  geom_line(aes(y = bbmsy_mean, colour = "BBmsy")) +
  theme(legend.position = "top") +
  scale_x_continuous(limits = range(h_bbmsy$year)) +
  xlab("")+
  ylab(expression(hat(B/B[MSY])))

h_catch <- ggplot(h_bbmsy, aes(x=year, y=catch)) +
  geom_line(colour="darkblue")+
  scale_x_continuous(limits = range(h_bbmsy$year))+
  ylab("Catch")+
  xlab("Year")

combined<-rbind(ggplotGrob(bmsy), ggplotGrob(h_catch), size="last")
  plot(combined)

```

How much catch is reported for BC stocks in RAM? And of that catch, how much do these stocks contribute to the total?

```{r}

assessment_data_file <- file.path(dir_anx, '_raw_data/ram_fisheries/d2017/RAM_v3.80/DB_Files_With_Assessment_Data/DBdata.RData')
load(assessment_data_file)

#all bc stocks in RAM
bc_stocks <- read_csv(file.path(dir_goal, 'ram/1_bc_stocks_all_ram.csv'))

### get time series from .Rdata
ts <- as.data.frame(timeseries, stringsAsFactors = FALSE) %>%
  mutate(tsyear  = as.integer(tsyear),
         tsvalue = as.numeric(tsvalue))

#remove stocks that we have bbmsy or ffmsy for

bc_stocks_ts <- ts %>%
  inner_join(bc_stocks) %>%
  filter(!is.na(tsvalue)) %>%
  filter(!str_detect(stocklong, "Salmon")) #remove salmon

#over the years, get total catch per stock

bc_stocks_ts_catch <- bc_stocks_ts %>%
  filter(tsid == "Ctouse-MT") %>%
  group_by(tsyear) %>%
  mutate(annual_catch = sum(tsvalue),
         n_stocks = length(unique(stockid))) %>%
  ungroup() %>%
  group_by(stockid, tsyear) %>%
  mutate(catch = sum(tsvalue),
         catch_prop = catch/annual_catch) %>%
  filter(tsyear > 1994) %>%
  select(stockid, tsyear, catch, catch_prop, n_stocks) 
  
a <- ggplot(bc_stocks_ts_catch, aes(x = tsyear, y = catch, fill = stockid)) +
  geom_bar(stat = "identity") + theme(legend.text=element_text(size=6)) +
  labs(x = "Year",
         y = "Catch (tons)")
ggplotly(a)

#this plot adds the total number of stocks in each year. I need to find a way to have it print just once...
ggplot(bc_stocks_ts_catch, aes(x = tsyear, y = catch, fill = stockid, label = n_stocks)) +
 geom_bar(stat = "identity") + theme(legend.text=element_text(size=6))  +
 geom_text(position = "stack", fill = NULL)
```

Plot total catch assessed vs unassessed in a stacked bar chart

```{r}

ass <- bc_stocks_ts_catch %>%
  mutate(assessed = ifelse(stockid %in% stock_plot_df$stock_id, 1, 0)) %>%
  group_by(assessed, tsyear) %>%
  summarize(catch = sum(catch)) %>%
  mutate(assessed_char = ifelse(assessed == 1, "assessed", "unassessed"))

a <- ggplot(ass, aes(x = tsyear, y = catch, fill = assessed_char)) +
  geom_bar(stat = "identity") +
  labs(x = "Year",
         y = "Catch (tons)")
ggplotly(a)
```

***

Let's look at how catch changes by species over time using the SAUP data.

```{r}

library(seaaroundus)

#saup bc data
saup_raw <- read_csv(file.path(dir_goal_anx, 'saup/2_saup_bc_raw.csv'))

#get bc cells from the saup_raw dataset
bc_cells <- unique(saup_raw$cell_id)

bc_saup_catch <- data.frame()

for(i in 1990:2014){
#get catch for those cells with getcelldata from the seaaroundus package
saup_catch <- getcelldata(i, bc_cells) %>%
  filter(taxon_key > 599999) %>%
  select(-functional_group_id, taxon_key, sector_type_id, fishing_entity_name, reporting_status_name, sector_type_name) %>%
  group_by(year, taxon_scientific_name, taxon_common_name) %>%
  summarize(catch = sum(catch_sum)) %>%
  as.data.frame()
  
bc_saup_catch <- rbind(bc_saup_catch, saup_catch)
}

#remove salmon
bc_saup_catch <- bc_saup_catch %>%
  filter(!str_detect(taxon_common_name, "almon")) #removing the s because there are upper and lower case ones, but this should get tme

g <- ggplot(bc_saup_catch, aes(x = year, y = catch, fill = taxon_common_name)) +
  geom_bar(stat = "identity") + theme(legend.text=element_text(size=6)) +
  labs(x = "Year",
         y = "Catch (tons)")
ggplotly(g)

```

Match SAUP data to RAM data to get a better picture of how much catch is assessed vs not assessed

```{r}

#getting all taxa reported for the BC cells from SAUP
saup_bc_taxa <- data.frame()

for(i in 1990:2014){

saup_catch <- getcelldata(i, bc_cells) %>%
  select(taxon_scientific_name, taxon_common_name)
  
saup_bc_taxa <- rbind(saup_bc_taxa, saup_catch)
}


ram_taxa <- stocks_ts_df %>%
  mutate(genus = str_extract(sciname, '.*(?= )')) %>%
  select(stock_id, stock_name, year, sciname, genus)

saup_to_ram_spp <- saup_bc_taxa %>%
  left_join(ram_taxa, by = c('taxon_scientific_name' = 'sciname')) %>%
  select(-genus) %>%
  rename(ram_stock = stock_id)
saup_to_ram_gen <- saup_bc_taxa %>%
   left_join(ram_taxa, by = c('taxon_scientific_name' = 'genus')) %>%
  select(-sciname) %>%
  rename(ram_stock = stock_id)

#this dataset lists all taxa in the SAUP data that are in BC and the assessed column identifies if that species has a ram assessment
saup_to_ram <- rbind(saup_to_ram_spp, saup_to_ram_gen) %>%
  mutate(assessed = ifelse(is.na(ram_stock), 0, 1)) %>%
  filter(assessed == 1) %>%
  select(taxon_scientific_name, taxon_common_name, assessed, year) %>%
  distinct()

#add this to the bc_saup_catch to then plot assessed vs unassessed

saup_assessed <- bc_saup_catch %>%
  filter(year > 1994) %>% #we only use ram data from 1995 on
  left_join(saup_to_ram) %>%
  mutate(assessed = ifelse(is.na(assessed), 0, 1)) %>%
  group_by(year, assessed) %>%
  summarize(catch = sum(catch)) %>%
  mutate(assessed_char = ifelse(assessed == 1, "assessed", "unassessed"))

saup_ass_plot <- ggplot(saup_assessed, aes(x = year, y = catch, fill = assessed_char)) +
  geom_bar(stat = "identity") +
  labs(x = "Year",
         y = "Catch (tons)")
ggplotly(saup_ass_plot)

```

Change to catch prop
```{r}

saup_assessed_prop <- saup_assessed %>%
  group_by(year) %>%
  mutate(annual_catch = sum(catch)) %>%
  ungroup() %>%
  mutate(catch_prop = catch/annual_catch)

saup_ass_prop_plot <- ggplot(saup_assessed_prop, aes(x = year, y = catch_prop, fill = assessed_char)) +
  geom_bar(stat = "identity") +
  labs(x = "Year",
         y = "Proportion of catch")
ggplotly(saup_ass_prop_plot)
```

Can we parse out the SAUP catch to the sub-regional level?

```{r}


```



