---
title: 'OHIBC: herring spawn index prep'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

source('~/github/ohibc/src/R/common.R')  ### an OHIBC specific version of common.R

dir_git     <- '~/github/ohibc'
dir_spatial <- file.path(dir_git, 'prep/spatial')  ### github: general buffer region shapefiles
dir_anx     <- file.path(dir_M, 'git-annex/bcprep')

### goal specific folders and info
goal     <- 'ao'
scenario <- 'v2017'
dir_goal <- file.path(dir_git, 'prep', goal, scenario)
dir_goal_anx <- file.path(dir_anx, goal, scenario)

### provenance tracking
library(provRmd); prov_setup()

### set up the default BC projection to be BC Albers
p4s_bcalb <- c('bcalb' = '+init=epsg:3005')

```

#Summary

This data is incorporated into the OHI British Columbia First Nations Resource Access and Opportunities (AO) goal.

***

#Data Source

**Reference**: 

**Downloaded**: 

**Native Data Resolution**:   

**Values**: 

**Time Range**: N/A

**Format**: CSV

***

# Methods

## Read in Herring Spawn data

### Herring Spawn Index

Herring spawn data is available as Spawn Habitat Index values calculated for sub-regions across BC from 1920s, 30s, and 40s to 2016.

* Data in text tables [here](http://www.pac.dfo-mpo.gc.ca/science/species-especes/pelagic-pelagique/herring-hareng/herspawn/tabsbkm2-eng.html)
* Documentation and background [here](http://www.pac.dfo-mpo.gc.ca/science/species-especes/pelagic-pelagique/herring-hareng/hertags/pdf/project-eng.pdf)

The documentation provides great detail, and some important notes are excerpted below.

> Based on detailed examination of the original records, it is clear that the overwhelming majority
of the fishery officers and divers who report on herring spawns, have done it in a dedicated,
professional and competent manner, often with detailed comments and charts accompanying the
records. There are undoubtedly some errors in individual records, but usually the impact of such
errors is diminished because of the relatively large number of records available for most
locations. Furthermore, these records were gathered by hundreds of different individuals over a
75 year period. Within any single location, there were at least several different data recorders,
and often many more over the years. So it is unlikely to expect a continuous run of poor data
records.

> There are, nevertheless, two kinds of systematic errors in these records which should be
mentioned. The first appears to be a time-dependent trend for fishery officers to under-estimate
the width of egg depositions. This occurred mainly in the earlier years, when it was generally
believed that most of the spawn was confined to the intertidal zone. It has not been until
relatively recent times that the significance of the subtidal fraction has been fully appreciated
(Haegele et al. 1983). Also, there has been systematic changes (see Hay and Kronlund 1987) in
the measurement of spawn density over the years, starting with a 1 to 5 "intensity" or categorical
scale from 1928 to 1950, to a 1 to 9 "intensity" scale from 1951 to 1982 and then eventually to
an egg layer measurement from 1977 to the present date. Each of these intensity measurements
have been converted to average egg layers in the database. Previously, a spawn coefficient (Hay 
and Kronlund 1987) was applied to correct for these time-dependent errors. An area-specific
coefficient was calculated by the mean of the product of spawn width and intensity, pooled at the
herring section geographic scale. Today, a similar correction is applied whereby the median of
the product of spawn width and egg layer measurements are pooled by a range of discrete
kilometre segments along the coastline. This coefficient is incorporated into a cumulative spawn
index described below.

> **Spawn Habitat Index:** We have computed a cumulative spawn habitat index (SHI) to represent the combined, long-term
frequency and magnitude of spawns along each kilometre of coastline over time. The index is
simply a measure of shoreline utilization by spawning herring. This index should not be
confused with the "spawn index" as applied to escapement, age-structure or more recently, catchage
assessment models (Cleary et al. 2010-2012, Schweigert et al. 1993-2009 or Haist et al.
1984-1992). The cumulative Spawn Habitat Index is calculated by the sum of the product of each
measured spawn length (m) and the median of the product of spawn width (m) and egg layers
adjusted by percent cover and pooled geographically. Spawn width and egg layer measurements
were pooled by a range of discrete, coastal shoreline kilometre segments, over all years (1928 to
the current year) and reflect the bathymetric (depth and slope) and vegetative features (seaweed
species compositions) of the spawning sites (area-specific spawn coefficient). The measured
length (m) of shoreline that is utilized by spawners is considered the most reliable of the three,
primary spawn measurements in terms of database completeness and consistency and has the
most influence on the Spawn Habitat Index (SHI = Length x Spawn Coefficient) at a spawning
area.

> ... The only other kind of error in the records which can severely affect conclusions presented in
these analyses, is that of incompleteness. It is clear that for some locations, records were not
collected systematically each year. Sometimes, problems with field communications, equipment
and weather resulted in data deficiencies. It is also known that many small, but observable
spawns, do not reach the doorstep of the database while others do. These spawns may be
important from a habitat perspective but are not high on the list from a stock assessment
viewpoint, where survey priorities are set.

#### extracting data from web

This link is the gateway to the various spawn habitat index pages: 
* http://www.pac.dfo-mpo.gc.ca/science/species-especes/pelagic-pelagique/herring-hareng/herspawn/tabsbkm2-eng.html
* [source for this page](view-source:http://www.pac.dfo-mpo.gc.ca/science/species-especes/pelagic-pelagique/herring-hareng/herspawn/tabsbkm2-eng.html) shows that the links are within [this page in the bottom frame](http://www.pac.dfo-mpo.gc.ca/science/species-especes/pelagic-pelagique/herring-hareng/herspawn/tabindex-eng.html#2).
* scan the link page via XML to extract links for the various sub-area pages.
    * update - not XML, just HTML... extract from the anchor tags

``` {r parse_links_from_web}
library(xml2); library(XML)

links_url <- 'http://www.pac.dfo-mpo.gc.ca/science/species-especes/pelagic-pelagique/herring-hareng/herspawn/tabindex-eng.html#2'
shi_dir <- dirname(links_url)

links_html <- read_html(links_url) %>%
  htmlParse()

links_df_raw <- capture.output(links_html, file=NULL) %>%
  as.data.frame() %>%
  setNames('html_raw') 

links_df <- links_df_raw %>%
  filter(str_detect(tolower(html_raw), 'reg[0-9]roll-eng|statist|section')) %>%
  mutate(html = str_replace_all(html_raw, '<[/]*li>|\\\r', ''),   ### get rid of list tags and \r tags
         html = str_replace_all(html, ' *---+.+', ''),
         link = str_extract(html, '(?<=href\\=\\").+(?=\\")'),
         link = ifelse(is.na(link), NA, file.path(shi_dir, link)),
         link_info = str_extract(html, '(?<=\\">).+')) %>%
  separate(link_info, c('link_area', 'area_name'), '</a>[-, ]*') %>%
  mutate(link_area = ifelse(is.na(link_area), html, link_area)) %>%
  select(-html_raw)

write_csv(links_df, file.path(dir_goal, 'int/shi_links.csv'))

```

``` {r read_linked_pages_from_web}

shi_data_file <- file.path(dir_goal, 'int/shi_data.csv')

if(!file.exists(shi_data_file)) {
  
  links_active <- read_csv(file.path(dir_goal, 'int/shi_links.csv')) %>%
    filter(!is.na(link) & str_detect(tolower(link_area), 'region|statist|section'))
  
  links_vec <- links_active$link %>%
    setNames(paste(links_active$link_area, links_active$area_name, sep = '~'))
  
  table_cols <- c('year', 'tot_records', 'spawn_hab_index', 
                  'l_tot', 'w_mean', 'layers_mean', 'sst_wgt', 
                  'date_mean', 'd_wgt', 'd_min', 'd_max', 
                  'diver_survey_pct')
  
  shi_df <- lapply(links_vec, FUN = function(x) {
    ### x <- links_vec[1]
    y <- read.table(x, skip = 14, fill = TRUE, na.strings = '*') %>%
      setNames(table_cols) %>%
      mutate(year = as.integer(year),
             tot_records = as.integer(tot_records)) %>%
      filter(!is.na(year))
  }) %>%
    setNames(names(links_vec)) %>%
    bind_rows(.id = 'id') %>%
    separate(id, c('area', 'area_name'), '~')
  
  write_csv(shi_df, shi_data_file)
  
} else {
  
  git_prov(shi_data_file, 'output')
  
}

```

#### Herring spawn index by rgn by year

Possible reference point: mean herring spawn index from 1940-1960?  For each DFO larger region, let's plot this.  These are not OHIBC analysis regions, but match somewhat.  These figures are simply for reference.

``` {r plot_herring_spawn_by_dfo_rgn}

shi_df <- read_csv(file.path(dir_goal, 'int/shi_data.csv'))

shi_rgn <- shi_df %>%
  filter(str_detect(area, 'Region')) %>%
  group_by(area) %>%
  mutate(area_name = str_replace_all(area_name, '((?<=\\))).+', ''),
         tmp = ifelse(year <= 1960, 1, NA),
         ref_shi = mean(spawn_hab_index * tmp, na.rm = TRUE)) %>%
  ungroup() %>%
  select(-tmp)

ggplot(shi_rgn, aes(x = year, y = spawn_hab_index, group = area)) +
  ggtheme_plot() +
  geom_line() +
  geom_hline(aes(yintercept = ref_shi), color = 'grey50') +
  facet_wrap( ~ area_name)

ggplot(shi_rgn, aes(x = year, y = date_mean, group = area)) +
  ggtheme_plot() +
  geom_ribbon(aes(ymin = d_min, ymax = d_max), fill = 'red3', color = NA, alpha = .4) +
  geom_line(color = 'black', alpha = .6) +
  geom_line(aes(y = d_wgt), color = 'red4', alpha = .8) +
  facet_wrap( ~ area_name)

```

### Spatialize Spawn Habitat Index data

Most herring sections fall squarely within one of the existing OHIBC regions.  However, a few sections cross between two OHIBC regions (Sections 67, 135, and 136); to spatialize SHI data from these, we explored two methods.

#### Spatialize by SHI totals

Spawn sections do not correspond with PFMAs or PFMSAs.  Instead, we will classify sections into OHIBC regions by inspecting the maps e.g. http://www.pac.dfo-mpo.gc.ca/science/species-especes/pelagic-pelagique/herring-hareng/herspawn/ccst_map-eng.html.  For sections that appear to intersect multiple OHIBC regions, closer inspection (i.e. clicking on the section number on the map) reveals the distribution and sites of spawn within the region; these in turn are linked to time series of SHI values for each site.

Cross-region herring sections: 

* [Section 67](http://www.pac.dfo-mpo.gc.ca/science/species-especes/pelagic-pelagique/herring-hareng/herspawn/067fig-eng.html)
    * Central Coast pools 1, 2, 3, 4; Aristazabal Island pools 5, 6
    * Data starts at [Spnrec06-eng.html#671](http://www.pac.dfo-mpo.gc.ca/science/species-especes/pelagic-pelagique/herring-hareng/herspawn/Spnrec06-eng.html#671)
* [Section 135](http://www.pac.dfo-mpo.gc.ca/science/species-especes/pelagic-pelagique/herring-hareng/herspawn/135fig-eng.html)
    * Strait of Georgia pool 1; North Vancouver Island pool 2
    * data starts at [Spnrec13-eng.html#1351](http://www.pac.dfo-mpo.gc.ca/science/species-especes/pelagic-pelagique/herring-hareng/herspawn/Spnrec13-eng.html#1351)
* [Section 136](http://www.pac.dfo-mpo.gc.ca/science/species-especes/pelagic-pelagique/herring-hareng/herspawn/136fig-eng.html)
    * Strait of Georgia pools 0 and 1; NVI pool 2
    * data starts at [Spnrec13-eng.html#1361](http://www.pac.dfo-mpo.gc.ca/science/species-especes/pelagic-pelagique/herring-hareng/herspawn/Spnrec13-eng.html#1361)
* Sum the SHI values in each pool over all years
* Determine all-time proportion of SHI for each pool for each section.
* Add this to `raw/sect_to_rgn_shi_wt.csv` for weighting calculations.

``` {r determine_spatial_divs, eval = FALSE, include = FALSE}

shi_df <- read_csv(file.path(dir_goal, 'int/shi_data.csv'))

### associate sections with OHIBC regions manually from maps
# shi_df$area %>% unique()
# 
# shi_sec <- shi_df %>%
#   filter(str_detect(tolower(area), 'section')) %>%
#   select(area, area_name) %>%
#   distinct()
# write_csv(shi_sec, file.path(dir_goal, 'herring_maps/shi_sections_raw.csv'))
### ADD rgn_id AND rgn_wgt FIELDS MANUALLY and rename without _raw

shi_fix <- read_csv(file.path(dir_goal, 'herring_maps/shi_secs_67_135_136.csv'),
                              col_types = 'iicnnnnccc') %>%
  group_by(section, pool, location) %>%
  mutate(layers = ifelse(is.na(layers), median(layers, na.rm = TRUE), layers),
         length = ifelse(is.na(length), median(length, na.rm = TRUE), length),
         width  = ifelse(is.na(width),  median(width,  na.rm = TRUE), width),
         shi    = length * width * layers) %>%
  filter(!is.na(shi)) %>%
  group_by(section, pool) %>%
  summarize(shi_sum = sum(shi), n_obs = n()) %>%
  group_by(section) %>%
  mutate(shi_pct = shi_sum / sum(shi_sum))

sect_to_rgn_id <- data.frame(section =              c(67, 67, 67, 67, 67, 67, 135, 135, 136, 136, 136),
                             pool = paste0('pool ', c( 1,  2,  3,  4,  5,  6,   1,   2,   0,   1,   2)),
                             rgn_id  =              c( 3,  3,  3,  3,  8,  8,   6,   4,   6,   6,   4))
shi_fix <- shi_fix %>%
  left_join(sect_to_rgn_id, by = c('section', 'pool')) %>%
  left_join(get_rgn_names(), by = 'rgn_id')

DT::datatable(shi_fix)

### from this, divisions:
###  67 pool 1-4 = CC = .946; 5-6 = AZ = .054
### 135 pool 1 = SG = .321, 2 = NVI = .679
### 136 pool 2 = NVI = .105; pools 0 (?) and 1 = .895
### USE THESE TO INFORM WEIGHTING OF SHI VALUES

```

#### Spatialize by area from herring sections .gdb

Note that this method is more "repeatable" than the previous, using area-weighting of a section across region boundaries to assign harvest values.  For the most part, the results appear nearly indistinguishable from the previous method, except for Aristazabal Island, which comes out looking better than the other version.  However, area weighting seems to be less appropriate than weighting using SHI data as outlined above.

``` {r extract_herring_section_by_rgn}

hsect_to_rgn_file <- file.path(dir_goal, 'herring_maps/sect_to_rgn_area_wt.csv')

if(!file.exists(hsect_to_rgn_file)) {
  
  # read the OGR; check the layers; get the herring layer
  
  boundaries_gdb <- file.path(dir_anx, '_raw_data/dfo_khunter/d2017/boundaries.gdb')
  herring_lyr <- rgdal::ogrListLayers(boundaries_gdb) %>%
    .[str_detect(tolower(.), 'herring')]
  
  hsects <- rgdal::readOGR(boundaries_gdb, herring_lyr) 
  ### is already in BC Albers
  
  ohibc_rast <- raster(file.path(dir_spatial, 'raster/ohibc_rgn_raster_500m.tif'))
  ### in BC Albers
  
  hsects_rgn <- raster::extract(ohibc_rast, hsects, progress = 'text')
  
  hsects_rgn_df <- hsects_rgn %>%
    setNames(hsects@data$Section) %>%
    lapply(FUN = function(x) {
      data.frame(rgn_id = x)
    }) %>%
    bind_rows(.id = 'section') %>%
    mutate(section = paste0('Section ', section)) %>%
    group_by(section, rgn_id) %>%
    summarize(n_cells = n(),
              area_km2 = n_cells/4) %>%
    ungroup() %>%
    filter(!is.na(rgn_id) & !is.na(section))

  write_csv(hsects_rgn_df, hsect_to_rgn_file)
  
} else {
  
  git_prov(hsect_to_rgn_file, filetype = 'output')
  
}
```

### Calculate SHI by region and year

Using the SHI-weighted method described above, we calculate the total SHI per region per year.

``` {r calc_spawn_hab_index_by_rgn}

shi_df <- read_csv(file.path(dir_goal, 'int/shi_data.csv'))

### This part uses the area-weighting based on .gdb spatialization
# shi_sec <- read_csv(file.path(dir_goal, 'herring_maps/sect_to_rgn_area_wt.csv')) %>%
#   group_by(section) %>%
#   mutate(rgn_wgt = area_km2 / sum(area_km2))

### This part reads the manually adjusted lookup table 
shi_sec <- read_csv(file.path(dir_goal, 'herring_maps/sect_to_rgn_shi_wt.csv')) %>%
  select(-area_name) %>%
  distinct()

# x <- shi_df %>% select(section = area) %>% filter(!section %in% shi_sec$section) %>% distinct()
### All sections in shi_df accounted for...
# y <- shi_sec %>% select(section) %>% filter(!section %in% shi_df$area) %>% distinct()
### Sections 031, 071, 081, 101, 171, 201, 211 not in shi_df... OK

shi_ohibc <- shi_df %>%
  filter(str_detect(tolower(area), 'section')) %>%
  select(section = area, area_name, year, tot_records, spawn_hab_index) %>%
  distinct() %>%
  left_join(shi_sec, by = c('section')) %>%
  group_by(rgn_id, year) %>%
  summarize(shi_tot = sum(spawn_hab_index * rgn_wgt, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(get_rgn_names(), by = 'rgn_id')

write_csv(shi_ohibc, file.path(dir_goal, 'output/ao_spawn_hab_index.csv'))

```

``` {r plot_shi_vs_ref}

shi_by_rgn <- read_csv(file.path(dir_goal, 'output/ao_spawn_hab_index.csv'))

ref_yrs <- c(1900:1960)

shi_ohibc <- shi_by_rgn %>%
  group_by(rgn_id) %>%
  mutate(tmp = ifelse(year %in% ref_yrs, 1, NA),
         ref_shi = mean(shi_tot * tmp, na.rm = TRUE)) %>%
  ungroup() %>%
  select(-tmp)
  
ggplot(shi_ohibc, aes(x = year, y = shi_tot, group = rgn_id)) +
  ggtheme_plot() +
  theme(axis.title.x = element_blank()) +
  annotate('rect', xmin=max(min(ref_yrs), min(shi_ohibc$year)), xmax=max(ref_yrs), ymin=0, ymax=Inf, fill = 'red', alpha = .1) +
  geom_line() +
  geom_hline(aes(yintercept = ref_shi), color = 'darkgreen', alpha = .5) +
  labs(y = 'Spawn Habitat Index Total') +
  facet_wrap( ~ rgn_name, scales = 'free')


```


### Herring spawn SSB by year

Additionally, Karen Hunter of the DFO provided spawn data from 2005-2015 divided into PFMA subareas that may be useful to compare to the Spawn Habitat Index, though on its own, its utility is limited by the relatively short time series.  In this analysis, we will not use these data, but here we compare the SSB data to the time-series SHI values, normalizing each dataset to its max value in the overlapping time period (2005-2015).  This strengthens our assumption that SHI is a reasonable proxy for SSB.

In the second plot, the grey line is the regression of SSB/SSB~max~ against SHI/SHI~max~ for the entire OHIBC assessment area, not for each region.

``` {r read_xlsx_from_dfo}

data_file <- file.path(dir_anx, '_raw_data/dfo_khunter/d2016/fisheries', 
                         'Herring Spawn Index_SSBbysection_coastwide_2005to2015.xlsx')

sheets <- readxl::excel_sheets(data_file)

hspawn_dfo_raw <- lapply(sheets, FUN = function(x) {
  # x <- sheets[1]
    capture.output( ### the read_excel function annoyingly prints crap to screen
      {
        y <- readxl::read_excel(data_file, sheet = x)
      },
      file = file.path(dir_goal, 'delete_this.txt')
    )
    unlink(file.path(dir_goal, 'delete_this.txt'))
    
    z <- y %>%
      setNames(tolower(names(.))) %>%
      rename(total_ssb = `total ssb (tonnes)`) %>%
      mutate(year = as.integer(year),
             total_ssb = as.integer(total_ssb)) %>%
      filter(!is.na(year)) %>%
      gather(section, ssb, everything(), -year, -contains('total')) %>%
      mutate(dfo_rgn = x)
    return(z)
  }) %>%
  bind_rows() %>%
  mutate(section = str_replace(section, '-', '') %>% as.integer())

shi_sec <- read_csv(file.path(dir_goal, 'herring_maps/sect_to_rgn_shi_wt.csv')) %>%
  select(-area_name) %>%
  mutate(section = str_replace(section, 'Section ', '') %>% as.integer())

hspawn_dfo_clean <- hspawn_dfo_raw %>%
  left_join(shi_sec, by = 'section')

write_csv(hspawn_dfo_clean, file.path(dir_goal_anx, 'int/hspawn_dfo_clean.csv'))

hspawn_dfo_summary <- hspawn_dfo_clean %>%
  group_by(rgn_id, year) %>%
  summarize(ssb_tot = sum(ssb * rgn_wgt, na.rm = TRUE)) %>%
  ungroup()

  
shi_ohibc_05_15 <- read_csv(file.path(dir_goal, 'output/ao_spawn_hab_index.csv')) %>%
  filter(year >= 2005 & year <= 2015) %>%
  full_join(hspawn_dfo_summary, by = c('year', 'rgn_id')) %>%
  group_by(rgn_id) %>%
  mutate(shi_rel = shi_tot / max(shi_tot),
         ssb_rel = ssb_tot / max(ssb_tot)) %>%
  ungroup()

ggplot(shi_ohibc_05_15, aes(x = year, y = shi_rel, group = rgn_id)) +
  ggtheme_plot() +
  theme(axis.title.x = element_blank()) +
  geom_line() +
  geom_line(aes(x = year, y = ssb_rel, group = rgn_id), color = 'red3') +
  scale_x_continuous(breaks = seq(2005, 2015, 2)) +
  labs(y = 'Relative spawn metric (SSB = black, SHI = red)') +
  facet_wrap( ~ rgn_name, scales = 'free')

ssb_shi_lm <- lm(shi_rel ~ ssb_rel, shi_ohibc_05_15)  

ggplot(shi_ohibc_05_15, aes(x = ssb_rel, y = shi_rel)) +
  ggtheme_plot() +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = 'red3', alpha = .5) +
  geom_abline(slope = ssb_shi_lm$coefficients[2], intercept = ssb_shi_lm$coefficients[1], 
              color = 'grey30', size = 1.5, alpha = .5) +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = 'SSB/SSB max',
       y = 'SHI/SHI max') +
  facet_wrap( ~ rgn_name)

```


***

``` {r prov_footer, results = 'asis'}
prov_wrapup(commit_outputs = FALSE)
```

