---
title: "Comparing Watson and SAUP data"
author: "Jamie Afflerbach"
date: "9/29/2017"
output: html_document
---

While creating the fishing pressure layers, we did this comparison of two different but very similar datasets to try and understand the differences and identify which to use. In the end, we are using the data from Watson (2017) to create the fishing pressure layers.

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = FALSE, message = FALSE, warning = FALSE)

library(rgdal)
library(tidyverse)
library(ggplot2)
library(raster)
library(doParallel)
library(foreach)
library(seaaroundus)
library(sf)

dir_git <- '~/github/ohibc'
source(file.path(dir_git, 'src/R/common.R'))

dir_spatial <- path.expand(file.path(dir_git, 'prep/_spatial'))

### goal specific folders and info
goal      <- '_pressures'
scenario  <- 'v2017'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario)
dir_goal_anx <- file.path(dir_M, 'git-annex/bcprep', goal, scenario)

#setting up provenance
# devtools::install_github('oharac/provRmd')
# library(provRmd)
# prov_setup()

#library(provRmd); prov_setup()

# BC region polygon

bc_poly <- st_read("~/github/ohibc/prep/_spatial", layer = "ohibc_rgn_wgs84") %>%
           as(.,"Spatial")
```


First get the template raster with a resolution of 0.5 degree cells. The `getcells()` function comes from the [seaaroundus R package](https://github.com/ropensci/seaaroundus).

The values of these cells are the Cell ID numbers. In the fisheries catch dataset, these Cell ID numbers match up with the "Seq" numbers.

```{r}

  saup_cells <- getcells("POLYGON ((-180 90, 180 90, 180 -90, -180 -90, -180 90))")

   saup_rast <- raster::raster(ncol = 720, nrow = 360)
   saup_rast[] <- saup_cells

   #crop global cells to just those in the BC area
   bc_rast <- crop(saup_rast, bc_poly) 
   
   bc_cells <- getValues(bc_rast) #get all cells in the area to help filter out spatial catch data
   
   plot(bc_rast)
   plot(bc_poly, add=T)
   
# base raster
rast_base <- raster(file.path(dir_spatial, 'raster/ohibc_rgn_raster_1000m.tif'))
```


# Sea Around Us Project Data

## Download SAUP data using rseaaroundus

```{r}

yrs <- c(2003:2014)

#get all SAUP data for cells in BC for years 2003 - 2014
bc_data <- map_df(yrs, getcelldata, bc_cells)

```

## Aggregate annual catch for landings and discards

```{r}

#register cores for parallel processing
registerDoParallel(cores = 8)

years <- c(2003:2014)

foreach(yr = years) %dopar%{
  
  #landings
  landings_cell_catch <- bc_data %>%
          filter(year == yr,
                 catch_status_name == "Landings") %>%
          rowwise() %>%
          rename(catch = catch_sum)%>% 
          group_by(cell_id)%>%
          summarise(cell_catch = sum(catch))
          
    #rasterize catch by swapping cell ids with 
  raster::subs(bc_rast, landings_cell_catch, by = 1, which = 2, subsWithNA = TRUE, 
               filename = paste0('fish_rasts/saup/landings/landings_', yr, '.tif'), 
               overwrite = TRUE) 
  
  #discards
  discards_cell_catch <- bc_data%>%
          filter(year == yr,
                 catch_status_name == "Discards")%>%
          rowwise() %>%
          rename(catch = catch_sum)%>% 
          group_by(cell_id)%>%
          summarise(cell_catch = sum(catch))

  raster::subs(bc_rast, discards_cell_catch, by = 1, which = 2, subsWithNA = TRUE, 
               filename = paste0('fish_rasts/saup/discards/discards_', yr, '.tif'),
               overwrite = TRUE) 
}

```


# Reg Watson data

For each year read in the raw data (data matched with bycatch category), filter for appropriate type (high, low) and then summarize total catch per cell (in tons/km^2^). The data is large and therefore held on the NCEAS server.

```{r}

#register cores for parallel processing
registerDoParallel(cores = 8)

years <- c(2003:2014)

foreach(yr = years) %dopar%{
  
  #read in raw data for the year
  raw <- readRDS(paste0(file.path(dir_M, 'marine_threats/impact_acceleration', 
                                  'stressors/comm_fish/int/catch_data_'), yr, '.rds')) %>%
          filter(Seq %in% bc_cells) #select only those cells in the BC region
  
  #landings
  landings <- raw %>%
    rowwise() %>%
    mutate(catch = sum(SSF_CR, LSF_CR, IUU_CR)*OceanArea)%>%  
      #multiplying by OceanArea to get tons per cell, making this comparable to SAUP. If we use this data, we will not need to do this
    group_by(Seq)%>%
    summarise(cell_catch = sum(catch))
          
    #rasterize catch by swapping cell ids with 
  raster::subs(bc_rast, landings, by = 1, which = 2, subsWithNA=TRUE, 
               filename = paste0('fish_rasts/watson/landings/landings_', yr, '.tif'),
               overwrite = T) 
  
  #discards
  discards <- raw%>%
          mutate(catch = Discards_CR * OceanArea)%>%
          group_by(Seq)%>%
          summarise(cell_catch = sum(catch))

  raster::subs(bc_rast, discards, by = 1, which = 2, subsWithNA=TRUE, 
               filename = paste0('fish_rasts/watson/discards/discards_', yr, '.tif'),
               overwrite = T) 
}

```

# Comparing SAUP and Watson data

First let's map them using just the 2014 data
```{r}
land_s <- raster('fish_rasts/saup/landings/landings_2014.tif')
land_w <- raster('fish_rasts/watson/landings/landings_2014.tif')
disc_s <- raster('fish_rasts/saup/discards/discards_2014.tif')
disc_w <- raster('fish_rasts/watson/discards/discards_2014.tif')

par(mfrow = c(2,2),
    mar = c(2,1,2,0.5))
plot(land_s, main = "SAUP Landings 2014", axes = F, zlim = c(0,13100))
plot(land_w, main = "Watson Landings 2014", axes = F, zlim = c(0,13100))
plot(disc_s, main = "SAUP Discards 2014", axes = F, zlim = c(0,320))
plot(disc_w, main = "Watson Discards 2014", axes = F, zlim = c(0,320))

```
