---
title: 'OHIBC: Alien Species Pressure layers prep'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = FALSE, message = FALSE, warning = FALSE)

library(sf)

dir_git <- '~/github/ohibc'
source(file.path(dir_git, 'src/R/common.R'))

dir_spatial <- path.expand(file.path(dir_git, 'prep/_spatial'))

### goal specific folders and info
goal      <- '_pressures'
scenario  <- 'v2017'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario)
dir_goal_anx <- file.path(dir_M, 'git-annex/bcprep', goal, scenario)

library(provRmd); prov_setup()

### goal-specific source scripts
source(file.path(dir_goal, 'pressures_lyr_fxns.R'))

### other support functions
source(file.path(dir_git, 'src/R/rast_tools.R'))

reload <- FALSE

```

# Summary

This pressure layer evaluates invasive species pressure on BC regions using the same method as OHI Global and California Current.  From Molnar (2008) data, we have data on the number of invasive species (and number of harmful invasives) found within each Marine Ecoregion of the World (MEOW) region.  For the global assessment and California Current, score is determined based on the proportion of invasives relative to the global maximum.  For OHIBC, we will instead use a species-by-species pressure based on three criteria: _invasive potential_, _ecological impact_, and _management difficulty_.  In Molnar 2008, each of these (as well as _scope_) is ranked on a 1 to 4 scale, with 1 being minimal and 4 being intense.  Invasive pressure of a species is a geometric mean (rescaled to max = 1) of these threat criteria:

$$pressure_{spp} = P_{spp} = \frac{\prod_{T=1}^N(threat_T)^{1/N}}{4}$$
where T is a threat criterion, N is number of threat criteria scores for that species (excluding NAs).

Scores for each OHIBC region are the area-weighted total species pressure from each ecoregion present.  This assumes that invasive species are uniformly distributed throughout the ecoregion.

$$X_{invasive} = \frac{\sum_{ecorgn=1}^{N}((\sum_{spp=1}^M P_{spp})  * A_{ecorgn})}{\sum_{ecorgn=1}^N A_{ecorgn}}$$
where $N$ is the number of ecoregions falling within an OHIBC region; $M$ is the number of species found within that ecoregion; $P_{spp}$ is the pressure score for each species in the ecoregion; and $A_{rgn}$ is the area of the OHIBC region allocated to that ecoregion (so $\sum_{rgn=1}^NA_{ecorgn}$ becomes the total OHIBC region area).

The data has no time scale.

# Data

This layer relies on the same invasive spp database as the global assessment, i.e. Molnar 2008.

- **Reference**: [Molnar et al. (2008)](http://onlinelibrary.wiley.com/doi/10.1890/070064/abstract)
- **Downloaded**: July 31, 2017 from https://www.conservationgateway.org/ConservationPractices/Marine/Pages/marineinvasives.aspx
- **Description**:  Number of alien spp per ecoregion, and threat scores for each species based on several invasive threat criteria.  Many other dataframes are included in the database; however most are not directly relevant for this assessment.
- **Native data resolution**: Ecoregion
- **Format**:  .mdb

# Methods

## Determine invasive-potential-weighted species counts in each ecoregion

Cracking open the .mdb MS Access file reveals 32 dataframes in the relational database.  Many of these relate to pathways which are not relevant to the scope of this pressure layer.  We will use layers that indicate species presence in ecoregions, combined with invasive threat level of each species.

``` {r read_mdb}

# library(RODBC)
# 
# library(Hmisc)
molnar08 <- Hmisc::mdb.get(file.path(dir_M, 'git-annex/bcprep/_raw_data',
                            'invasives/TNC_MarineInvasionsDb_18feb08.mdb'))
### returns a list of dataframes; here are the names of the dataframes:
# sapply(x, nrow)
#           _A_SPECIES_MARINE        A_Geo_Habitat-Marine        A_Occurrences-Marine   A_Species_Pathways-Marine             B_Pathways_List 
#                         338                         338                        2187                         728                          59 
# B_References_Marine (Jan08)         list_db_terminology         list_distr_in_range      list_document pathways          list_documentation 
#                         354                          11                           5                           3                           5 
#               list_FW TR MR           list_introduction               list_pathway2               list_pathways             list_references 
#                           8                           5                          57                          58                         355 
#                 list_scores              list_scores_id                  list_taxon                    list_Y-N                 list_Y-N-UK 
#                           5                          25                          27                           2                           3 
#         Tbl_mr meow list-er       Tbl_mr meow list-prov      Tbl_mr meow list-realm      T-Q_path-er1-summ_Swbd      T-Q_path-er2-summ_Swbd 
#                         196                          57                          12                         452                         796 
#      T-Q_path-er3-summ_Swbd      T-Q_path-er4-summ_Swbd      T-Q_path-er5-summ_Swbd      A_Threat_scores-Marine      B_MEOWv2_draft_08mar07 
#                         481                         279                         337                         338                         232 
#                list_habitat           list_scores_id-EI 
#                          42                           6 

### 
### 

write_csv(molnar08[['A_Occurrences-Marine']], file.path(dir_goal, 'int/invspp_occurrences_raw.csv'))
write_csv(molnar08[['A_Threat_scores-Marine']], file.path(dir_goal, 'int/invspp_threat_raw.csv'))
write_csv(molnar08[['B_MEOWv2_draft_08mar07']], file.path(dir_goal, 'int/invspp_meow_ids.csv'))
write_csv(molnar08[['_A_SPECIES_MARINE']], file.path(dir_goal, 'int/invspp_spp_names.csv'))

```

* `A_Occurrences-Marine` represents the occurrences of each species in each ecoregion.
* `A_Threat_scores-Marine` represents the threat scores of each species (regardless of ecoregion).  This includes a text description justifying each of the four categories of threat scores for that species.

### Threat score descriptions

`r knitr::kable(molnar08[['list_scores_id']][2:25, ])`

### Calculate mean threat score

Using `A_Threat_scores-Marine` (saved as `_pressures/v2017/int/invspp_threat_raw.csv`), we calculate a weighted threat score for each species using a geometric mean (dropping NAs).  This weighted threat score divided by 4 to rescale the highest value to 1.0 - note this means low-threat species still get a minimum score of 0.25.

``` {r calc_threat}

spp_threat_all <- read_csv(file.path(dir_goal, 'int/invspp_threat_raw.csv')) %>%
  setNames(tolower(names(.)) %>% str_replace_all('[^a-z]', '_')) 

spp_threat <- spp_threat_all %>%
  select(species_id, contains('score')) %>%
  gather(threat, score, -species_id) %>%
  filter(threat != 'scope_score') %>%
  mutate(score = as.integer(score),
         threat = str_replace(threat, '_score', ''),
         threat_txt = sprintf('%s = %s', threat, score)) %>%
  group_by(species_id) %>%
  summarize(threat_txt = paste(threat_txt, collapse = '; '),
            score_prod = prod(score, na.rm = TRUE),
            pressure   = score_prod ^ (1 / sum(!is.na(score))) / 4) %>%
  ungroup() %>%
  select(species_id, threat_txt, pressure)

write_csv(spp_threat, file.path(dir_goal, 'int/invspp_threat_scored.csv'))

spp_names <- read_csv(file.path(dir_goal, 'int/invspp_spp_names.csv')) %>%
  setNames(tolower(names(.)) %>% str_replace('[^a-z]', '_')) %>%
  select(species_id, species_name, higher_taxa, common_name)

DT::datatable(spp_names %>% 
                left_join(spp_threat, by = 'species_id') %>%
                mutate(pressure = round(pressure, 3)))
```

### Assign species threat levels to ecoregions

Using `A_Occurrences-Marine`, saved as `int/invspp_occurrences_raw.csv`, join the species threat levels to species occurrences in each ecoregion.  Find total threat-weighted sum of species in each ecoregion.

``` {r assign_spp_to_ecoregions}

spp_threat <- read_csv(file.path(dir_goal, 'int/invspp_threat_scored.csv'))
spp_ecorgn_raw <- read_csv(file.path(dir_goal, 'int/invspp_occurrences_raw.csv')) %>%
  clean_df_names() 

meow_list <- read_csv(file.path(dir_goal, 'int/invspp_meow_ids.csv')) %>%
  clean_df_names() %>%
  select(ecoregion_code = eco_code,
         # eco_code = eco_code_x,
         ecoregion)

spp_ecorgn <- spp_ecorgn_raw %>%
  left_join(spp_threat, by = 'species_id') %>%
  select(ecoregion_code, species_id, pressure) %>%
  group_by(ecoregion_code) %>%
  summarize(n_spp = n(),
            pressure_sum = sum(pressure)) %>%
  ungroup() %>%
  left_join(meow_list, by = 'ecoregion_code')

mdl_fit <- lm(pressure_sum ~ n_spp, data = spp_ecorgn)

r_sq <- mdl_fit %>% summary() %>% .$r.squared %>% round(4)

ggplot(spp_ecorgn, aes(x = n_spp, y = pressure_sum)) +
  geom_point() +
  geom_smooth(method = 'lm')

write_csv(spp_ecorgn, file.path(dir_goal, 'int/invspp_meow_pressures.csv'))

```

Comparing the pressure-weighted sums to the total species count in each region (which was the metric used in OHI Global 2012-2017 and CC), the linear trend indicates that scores shouldn't be too greatly affected using this new method.  

Linear model: intercept = `r mdl_fit$coefficients[1]`, slope = `r mdl_fit$coefficients[2]`, R^2 = `r r_sq`.

## Spatializing invasives to BC regions

Using spatial data for MEOW regions, we can spatialize the invasive species pressures to OHIBC regions.

### Intersect MEOW with OHIBC

Read in MEOW regions shapefile and intersect with OHIBC regions; calculate area of each OHIBC region that falls within each MEOW region.

``` {r calc_areas_of_meows}

bc_rgn <- read_sf(dir_spatial, 'ohibc_rgn')

meow_rgn <- read_sf(file.path(dir_M, 'git-annex/bcprep/_raw_data/meow'),
                        'meow_ecos') %>%
  st_transform(st_crs(bc_rgn))

meow_in_bc <- st_intersection(bc_rgn, meow_rgn) %>%
  setNames(tolower(names(.))) %>%
  select(rgn_id, ecoregion)

meow_in_bc$area_km2 <- st_area(meow_in_bc) / 1e6

bc_meow_df <- meow_in_bc %>%
  as.data.frame() %>%
  group_by(rgn_id) %>%
  mutate(tot_area = sum(area_km2),
         eco_area = area_km2 / tot_area) %>%
  select(-geometry, -area_km2, -tot_area) %>%
  ungroup()
  
write_csv(bc_meow_df, file.path(dir_goal, 'int/invspp_bc_meow_rgns.csv'))

ggplot(meow_in_bc) +
  geom_sf(aes(fill = ecoregion), size = .25) +
  scale_fill_brewer(palette = 'Dark2')

```

### Which species are included as invasive in BC?

``` {r determine_BC_invasives}

spp_threat <- read_csv(file.path(dir_goal, 'int/invspp_threat_scored.csv'))
spp_ecorgn_raw <- read_csv(file.path(dir_goal, 'int/invspp_occurrences_raw.csv')) %>%
  clean_df_names() 

meow_list <- read_csv(file.path(dir_goal, 'int/invspp_meow_ids.csv')) %>%
  clean_df_names() %>%
  select(ecoregion_code = eco_code,
         # eco_code = eco_code_x,
         ecoregion)

bc_meow_df <- read_csv(file.path(dir_goal, 'int/invspp_bc_meow_rgns.csv'))

spp_names <- read_csv(file.path(dir_goal, 'int/invspp_spp_names.csv')) %>%
  setNames(tolower(names(.)) %>% str_replace('[^a-z]', '_')) %>%
  select(species_id, species_name, higher_taxa, common_name)

spp_bc <- spp_ecorgn_raw %>%
  left_join(meow_list, by = 'ecoregion_code') %>%
  filter(ecoregion %in% bc_meow_df$ecoregion) %>%
  select(species_id, ecoregion) %>%
  group_by(species_id) %>%
  summarize(ecorgns = paste(ecoregion, collapse = '; ')) %>%
  ungroup() %>%
  left_join(spp_names, by = 'species_id') %>%
  left_join(spp_threat, by = 'species_id') %>%
  select(species_name, common_name, higher_taxa, 
         ecorgns, threat_txt, pressure) %>%
  arrange(desc(pressure))

write_csv(spp_bc, file.path(dir_goal, 'int/invspp_bc_spp_list.csv'))

DT::datatable(spp_bc)
```


### Assign invasives pressures to OHIBC rgns

For each ecoregion, calculate the proportional invasive species pressure relative to the global maximum; combine this result with the ecoregion area dataframe.  For each OHIBC region, calculate the area-weighted mean of invasive species pressures.

``` {r read_basic_invasives_data}

invspp_prs <- read_csv(file.path(dir_goal, 'int/invspp_meow_pressures.csv')) %>%
  mutate(ecoregion   = tolower(ecoregion))

n_ref <- max(invspp_prs$n_spp)
prs_ref <- max(invspp_prs$pressure_sum)

bc_meow_df <- read_csv(file.path(dir_goal, 'int/invspp_bc_meow_rgns.csv')) %>%
  mutate(ecoregion = tolower(ecoregion))

invspp_bc <- left_join(bc_meow_df, invspp_prs, by = 'ecoregion') 

prs_invspp_bc <- invspp_bc %>%
  group_by(rgn_id) %>%
  summarize(n_spp = sum(n_spp * eco_area),
            pressure = sum(pressure_sum * eco_area) / prs_ref) %>%
  ungroup()
  
write_csv(prs_invspp_bc, file.path(dir_goal, 'output', 'prs_inv_spp.csv'))

DT::datatable(prs_invspp_bc %>% 
                mutate(pressure  = round(pressure, 4),
                       n_spp     = round(n_spp, 2)))
```

``` {r plot_pressures}

prs_invspp_bc <- read_csv(file.path(dir_goal, 'output', 'prs_inv_spp.csv'))

prs_df <- prs_invspp_bc %>%
  left_join(get_rgn_names(), by = 'rgn_id')

prs_plot <- ggplot(prs_df, aes(x = rgn_name, y = pressure)) +
  ggtheme_plot() +
  geom_bar(stat = 'identity', show.legend = FALSE) +
  scale_color_brewer(palette = 'Dark2') +
  ylim(0, NA) +
  theme(axis.text.x = element_text(angle = 22.5, hjust = .75)) +
  labs(title = 'Invasive species pressure',
       x = 'region',
       y = 'Rescaled pressure score')

print(prs_plot)

```


-----

``` {r results = 'asis'}

prov_wrapup(commit_outputs = FALSE)

```
