---
title: 'OHIBC: Pressures for logging and mining'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

source('~/github/ohibc/src/R/common.R')  ### an OHIBC specific version of common.R;
  ### includes library(tidyverse); library(stringr)
library(sf)

dir_git     <- '~/github/ohibc'
dir_spatial <- file.path(dir_git, 'prep/_spatial') 
dir_anx     <- file.path(dir_M, 'git-annex/bcprep')

### goal specific folders and info
goal      <- '_pressures'
scenario  <- 'v2017'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario)
dir_goal_anx <- file.path(dir_anx, goal, scenario)
dir_data  <- file.path(dir_anx, '_raw_data/databc/logging_cutblocks')

### provenance tracking
library(provRmd); prov_setup()

### set up the default BC projection to be BC Albers
p4s_bcalb <- c('bcalb' = '+init=epsg:3005')

```

# Summary

Generates pressures layers for logging activity.

Also generates a layer for mining activity which will be set to zeros for all observations, but preserves a placeholder for future mining pressures.

-----

# Data Source 

* __Reference__: https://catalogue.data.gov.bc.ca/dataset/harvested-areas-of-bc-consolidated-cutblocks-
* __Downloaded__: 11/29/2017
* __Description__:  This is spatial data depicting the cut block boundaries and year of harvest for crown lands within British Columbia. It is created from the provincial Forest Cover, from The RESULTS Reporting system, from Forest Tenures applications and from satellite imagery using change detection processes. It represents the gross opening area when derived from forest cover, RESULTS or Forest Tenure data and does not exclude non harvested 'reserved' areas. Satellite image change detection represents the area harvested each year and may not be the entire cutblock area and will not include non harvested 'reserve' areas.
* __Time range__: 1900 - 2016
* __Format__:  .shp

-----
  
# Methods: Logging

The logging cutblocks polygons dataset includes the following attributes:

 [1] "OPENINGID"  "FTAREASQM"  "FTLENGTHM"  "SHAPE"      "OBJECTID"   "S_NN_CD_DT" "HARVESTYR"  "DSTRBSTDT"  "DSTRBEDDT"  "DATASOURCE" "AREAHA"    
[12] "CUTBLOCKID" "geometry" 

In addition to the spatial information contained in `geometry`, these fields seem likely to be useful:

* `HARVESTYR` is the year of harvest, derived from the disturbance start date of the data source or from the year the disturbance was detected using Landsat change detection.
* `DSTRBSTDT`	(DISTURBANCE_START_DATE) is derived from the data source. RESULTS describes it as; derived from the first reported disturbance activity reporting. This date is used to calculate the silviculture obligation milestones.
* `DSTRBEDDT`	(DISTURBANCE_END_DATE) is derived from the data source. RESULTS describes it as; derived from the last disturbance activity reporting record based on the disturbance activity's completion date.

Assuming a 20-year period for functional regrowth (linear), we cut out any forestry blocks with no activity later than 1981.  From updated BC CHI: "As in Ban et al (2010), forest cutblocks included were those for the past 20 years but the date range advanced (1993-2013). All records retired before 1993 were removed from the dataset as they were considered functionally restored (Anderson 1996)."

``` {r read_spatial_filter_years_and_intersect_to_ohibc}

lc_ohibc_file <- file.path(dir_goal_anx, 'shp', 'log_cuts_ohibc.shp')

if(!file.exists(lc_ohibc_file)) {
  ### logging cutblocks poly
  lc_poly <- read_sf(file.path(dir_data, 'VEG_CONSOLIDATED_CUT_BLOCKS_SP', 
                               'CNS_CUT_BL_polygon.shp'))
  
  lc_recent <- lc_poly %>%
    dplyr::select(cutb_id = CUTBLOCKID,
                  hrv_yr  = HARVESTYR, 
                  dst_st  = DSTRBSTDT, 
                  dst_end = DSTRBEDDT) %>%
    mutate(hrv_end = as.integer(str_extract(dst_end, '^[0-9]{4}')),
           hrv_end = ifelse(is.na(hrv_end), hrv_yr, hrv_end),
           hrv_end = ifelse(hrv_end < hrv_yr, hrv_yr, hrv_end)) %>%
    filter(hrv_end > 1981)
  
  ohibc_rgn <- read_sf(file.path(dir_spatial, 'ohibc_rgns_unclipped.shp')) %>%
    st_transform(st_crs(lc_recent))
  
  lc_ohibc <- st_intersection(lc_recent, ohibc_rgn) %>%
    select(-rgn_name, -rgn_code)
  ### still a big file - 153 MB - store remotely
  
  st_write(lc_ohibc, lc_ohibc_file, delete_layer = TRUE)

} else {
  git_prov(lc_ohibc_file, filetype = 'output')
}

```

From this trimmed dataset, we will create a dataframe of years vs. impact for each geometry; then a dataframe of area (of each cutblock) within each region.  Comparing this to total inland area will give us a good estimate of the logging impact intensity.

``` {r create_impact_df}

lc_ohibc_file <- file.path(dir_goal_anx, 'shp', 'log_cuts_ohibc.shp')

lc_ohibc <- st_read(lc_ohibc_file)
  git_prov(lc_ohibc_file, filetype = 'input')

lc_ohibc_df <- lc_ohibc %>%
  as.data.frame() %>%
  select(cutb_id, hrv_yr, hrv_end, rgn_id)

yrs <- 1981:2016
cutb_ids <- unique(lc_ohibc_df$cutb_id)

yr_df <- data.frame(cutb_id = rep(cutb_ids, times = length(yrs)),
                    year    = rep(yrs, each = length(cutb_ids)))

lc_yrs_df <- lc_ohibc_df %>%
  left_join(yr_df, by = 'cutb_id') %>%
  mutate(hrv_length = hrv_end - hrv_yr,
         impact = 1 - (year - hrv_end)/20,
         impact = ifelse(impact < 0, 0, impact),
         impact = ifelse(year < hrv_yr, 0, impact),
         impact = ifelse(year >= hrv_yr & year <= hrv_end, 1, impact)) %>%
  filter(impact > 0) %>%
  select(cutb_id, year, impact) %>%
  distinct()

lc_areas_df <- lc_ohibc %>%
  select(rgn_id, cutb_id) %>%
  mutate(cutb_area_km2 = as.numeric(st_area(.$geometry)) / 1e6) %>%
  as.data.frame() %>%
  select(-geometry) %>%
  filter(cutb_area_km2 > .1) ### ignore cutblocks smaller than .1 km^2...

lc_impacts <- lc_areas_df %>%
  left_join(lc_yrs_df, by = 'cutb_id') %>%
  filter(!is.na(year)) %>%
  group_by(rgn_id, year) %>%
  summarize(tot_impact = sum(impact * cutb_area_km2))

write_csv(lc_impacts, file.path(dir_goal, 'int/logging_impacts.csv'))

DT::datatable(lc_impacts)
```

``` {r get_rgn_land_area}
### get area of inland regions
ohibc_a_inland <- read_sf(file.path(dir_spatial, 'ohibc_rgns_unclipped.shp')) %>%
  mutate(tot_area_km2 = st_area(geometry) / 1e6) %>%
  as.data.frame() %>%
  select(rgn_id, tot_area_km2)
ohibc_a_rgn    <- read_sf(file.path(dir_spatial, 'ohibc_rgn.shp')) %>%
  mutate(ocean_area_km2 = st_area(geometry) / 1e6) %>%
  as.data.frame() %>%
  select(rgn_id, ocean_area_km2)

ohibc_inland_area <- ohibc_a_inland %>%
  full_join(ohibc_a_rgn, by = 'rgn_id') %>%
  mutate(inland_area_km2 = as.numeric(tot_area_km2 - ocean_area_km2))

logging_intensity <- read_csv(file.path(dir_goal, 'int/logging_impacts.csv')) %>%
  left_join(ohibc_inland_area, by = 'rgn_id') %>%
  filter(rgn_id != 7) %>%
  mutate(intensity = tot_impact / inland_area_km2,
         pressure  = intensity / max(intensity)) %>%
  select(rgn_id, year, intensity, pressure) %>%
  filter(year >= 2001)

write_csv(logging_intensity, file.path(dir_goal, 'output', 'prs_logging.csv'))

DT::datatable(logging_intensity)
```

``` {r visualize_logging_intensity}
logging_intensity <- read_csv(file.path(dir_goal, 'output', 'prs_logging.csv')) %>%
  left_join(get_rgn_names(), by = 'rgn_id')

ggplot(logging_intensity, aes(x = year, y = pressure)) +
  ggtheme_plot() +
  geom_line(aes(group = rgn_name)) +
  facet_wrap( ~ rgn_name)
```

# Methods: Mining

Create a dataframe of zeros for region ID as a placeholder for future mining pressures.

``` {r}

mining_df <- data.frame(rgn_id = 1:8,
                        intensity = 0,
                        pressure  = 0)

write_csv(mining_df, file.path(dir_goal, 'output', 'prs_mining.csv'))

```

-----

``` {r prov_footer, results = 'asis'}
prov_wrapup(commit_outputs = FALSE)
```


