---
title: 'OHIBC: Commercial Fishing Pressure layers prep'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

#Summary

The commercial fishing pressure layers are created from spatialized catch by gear data provided by Watson (2017).

Two layers are created here:

1. **Landings** - representing the pressure that targeted catch exerts on the system
2. **Discards** - representing the pressure that non-targeted catch exerts on the system

The reference point used for each is 110% of the maximum catch rate (tons/km2) within each sub region found across the entire time series (1950 - 2014).

***

#Data Source

**Reference**: [Watson (2017)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5387926/)

**Downloaded**: April 21, 2017

**Description**:  Catch per half degree cell (raw values are in tons per km^2^)  

**Native data resolution**: 0.5 degree    

**Time range**: 1950 - 2014 

**Format**:  Tabular  

***
  
#Methods

##Setup

Load all relevant libraries 

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',
                      echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(ggplot2)
library(raster)
library(doParallel)
library(foreach)
library(seaaroundus)
library(sf)

dir_git <- '~/github/ohibc'
source(file.path(dir_git, 'src/R/common.R'))

dir_spatial <- path.expand(file.path(dir_git, 'prep/_spatial'))

### goal specific folders and info
goal      <- '_pressures'
scenario  <- 'v2017'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario)
dir_goal_anx <- file.path(dir_M, 'git-annex/bcprep', goal, scenario)

#setting up provenance
# devtools::install_github('oharac/provRmd')
# library(provRmd)
# prov_setup()

#library(provRmd); prov_setup()

# BC region polygon

bc_poly <- st_read(dsn = file.path(dir_git, "prep/_spatial"), layer = "ohibc_rgn_wgs84") %>%
           as(.,"Spatial")
```


First get the template raster with a resolution of 0.5 degree cells. The `getcells()` function comes from the [seaaroundus R package](https://github.com/ropensci/seaaroundus).

The values of these cells are the Cell ID numbers. In the fisheries catch dataset, these Cell ID numbers match up with the "Seq" numbers.

```{r}

  saup_cells <- getcells("POLYGON ((-180 90, 180 90, 180 -90, -180 -90, -180 90))")

   saup_rast <- raster::raster(ncol=720, nrow=360)
   saup_rast[] <- saup_cells

   #crop global cells to just those in the BC area
   bc_rast <- crop(saup_rast, bc_poly) 
   
   bc_cells <- getValues(bc_rast) #get all cells in the area to help filter out spatial catch data
   
   plot(bc_rast)
   plot(bc_poly, add=T)
   
# base raster
rast_base <- raster(file.path(dir_spatial, 'raster/ohibc_rgn_raster_1000m.tif'))

# assign the total area of each cell to the cells using data from Watson

raw <- readRDS(paste0(file.path(dir_M,'marine_threats/impact_acceleration/stressors/comm_fish/int/catch_data_2014.rds'))) %>%
          filter(Seq %in% bc_cells) %>% #select only those cells in the BC region
          dplyr::select(Seq, OceanArea) %>%
          distinct()

cell_areas <- subs(bc_rast, raw, by = 1, which = 2, subsWithNA=T)
```

# Load data

Read in the data, filter it between landings and discards. The data differentiates the catch into four categories of catch rate:

1. SSF_CR (small scale fisheries catch rate)  
2. LSF_CR (Large scale fisheries catch rate)  
3. IUU_CR (illegal, unregulated and unreported)  
4. Discards_CR (discards)  

The first three are aggregated by cell to represent total landings per cell. The last categorie, Discards_CR, is separated and used to represent the total tons of discarded catch.

An important note is that the data is provided as a "Catch Rate" meaning tons/km2. Since our final layer has a resolution of 1km^2^, we don't need to multiply these catch rates by the total ocean area.

```{r, eval = F}

#register cores for parallel processing
registerDoParallel(cores = 8)

years = c(1950:2014)

foreach(yr = years) %dopar%{
  
  #read in raw data for the year
  raw <- readRDS(paste0(file.path(dir_M,'marine_threats/impact_acceleration/stressors/comm_fish/int/catch_data_'), yr, '.rds')) %>%
          filter(Seq %in% bc_cells) #select only those cells in the BC region
  
  #landings
  landings <- raw %>%
          rowwise() %>%
          mutate(catch = sum(SSF_CR, LSF_CR, IUU_CR))%>%
          group_by(Seq)%>%
          summarise(cell_catch = sum(catch))
          
    #rasterize catch by swapping cell ids with 
raster::subs(bc_rast, landings, by = 1, which = 2, subsWithNA=TRUE, filename = paste0('fish_rasts/watson/landings/landings_',yr,'.tif'),overwrite=T) 
  
  #discards
  discards <- raw%>%
          mutate(catch = Discards_CR)%>%
          group_by(Seq)%>%
          summarise(cell_catch = sum(catch))

raster::subs(bc_rast, discards, by = 1, which = 2, subsWithNA=TRUE, filename = paste0('fish_rasts/watson/discards/discards_',yr,'.tif'),overwrite=T) 

}

```

## Five year means

Mean catch per cell is calculated over a rolling window of 5 years to account for interannual variability.

```{r mean_catch, eval = F}

land <- list.files('fish_rasts/watson/landings', full.names=T, pattern = ".tif")
disc <- list.files('fish_rasts/watson/discards', full.names=T, pattern = ".tif")

foreach (i = 1950:2010) %dopar%{
  
  yrs <- c(i:(i+4))
  
  out_land <- land[which(substr(land, nchar(land)-7, nchar(land)-4) %in% yrs)]%>%
            stack()%>%
            calc(fun = function(x){mean(x, na.rm = T)}, 
                 filename = paste0('fish_rasts/watson/landings/five_year_means/mean_catch_', yrs[1], '_', yrs[5], '.tif'), overwrite = T)
  
  out_disc <- disc[which(substr(disc, nchar(land)-7, nchar(land)-4) %in% yrs)]%>%
            stack()%>%
            calc(fun = function(x){mean(x, na.rm = T)}, 
                 filename = paste0('fish_rasts/watson/discards/five_year_means/mean_catch_', yrs[1], '_', yrs[5], '.tif'), overwrite = T)
}

```


## Reference Point

Look at all catch data standardized by NPP from 2003 - 2014 and use 110% * the maximum value as the reference point.

```{r}

land_5 <- list.files('fish_rasts/watson/landings/five_year_means', full.names=T) %>% 
              stack() %>%
              overlay(cell_areas, fun=function(x,y){x*y})
names(land_5) <- c(1954:2014)
disc_5 <- list.files('fish_rasts/watson/discards/five_year_means', full.names=T) %>% 
              stack() %>%
              overlay(cell_areas, fun=function(x,y){x*y})
names(disc_5) <- c(1954:2014)

#Landings

#run zonal stats on each raster
land_prs <- extract(land_5, bc_poly, na.rm = TRUE, df = TRUE, weights = TRUE, normalizeWeights = TRUE) %>%
     rename(rgn_id = ID) %>%
     gather(key = "year", value = "cell_catch", -rgn_id, - weight) %>%
     mutate(year = substr(year, 2, 5),
            rgn_id = as.character(rgn_id),   #forcing into character to join with bc_poly@data$rgn_id
            catch_weighted = cell_catch * weight) %>% 
     group_by(rgn_id, year) %>%
     summarize(rgn_catch = sum(catch_weighted, na.rm=T)) %>%
     left_join(bc_poly@data, by = "rgn_id") %>%
     mutate(catch_by_area = rgn_catch/area_km2) %>% #calculate regional catch sums weighted by area
     group_by(rgn_id) %>%
     mutate(max_catch = max(catch_by_area)) %>% #get the maximum catch/area for each region across all years
     mutate(prs_value = catch_by_area/(1.1* max_catch)) %>%
     ungroup() %>%
    mutate(year = as.numeric(year))

disc_prs <- extract(disc_5, bc_poly, na.rm = TRUE, df = TRUE, weights = TRUE, normalizeWeights = TRUE) %>%
     rename(rgn_id = ID) %>%
     gather(key = "year", value = "cell_catch", -rgn_id, - weight) %>%
     mutate(year = substr(year, 2, 5),
            rgn_id = as.character(rgn_id),   #forcing into character to join with bc_poly@data$rgn_id
            catch_weighted = cell_catch * weight) %>% 
     group_by(rgn_id, year) %>%
     summarize(rgn_catch = sum(catch_weighted, na.rm=T)) %>%
     left_join(bc_poly@data, by = "rgn_id") %>%
     mutate(catch_by_area = rgn_catch/area_km2) %>% #calculate regional catch sums weighted by area
     group_by(rgn_id) %>%
     mutate(max_catch = max(catch_by_area)) %>% #get the maximum catch/area for each region across all years
     mutate(prs_value = catch_by_area/(1.1* max_catch)) %>%
     ungroup() %>%
    mutate(year = as.numeric(year))

```


***

# Results


``` {r calc_rgn_means}

#landings

ohibc_rgn_land <- land_prs %>%
                  dplyr::select(rgn_id, year, pressure = prs_value)

write_csv(ohibc_rgn_land, file.path(dir_goal, 'output/prs_landings_fish_layer.csv'))

#discards

ohibc_rgn_disc <- disc_prs %>%
                  dplyr::select(rgn_id, year, pressure = prs_value)

write_csv(ohibc_rgn_disc, file.path(dir_goal, 'output/prs_discards_fish_layer.csv'))

```

# Results

```{r}

#Landings

lplot <- ggplot(land_prs, aes(x = year,y = prs_value, color = rgn_id, goup = 1))+
  geom_line()+
  labs(color = "Region",
       y = "Pressure score",
       x = "Year",
       title = "Landings fishing pressure")

#Discards

dplot <- ggplot(disc_prs, aes(x = year,y = prs_value, color = rgn_id, goup = 1))+
  geom_line()+
  labs(color = "Region",
       y = "Pressure score",
       x = "Year",
       title = "Discards fishing pressure")

cowplot::plot_grid(lplot, dplot, ncol = 2)

```


***

#Citation information  
Watson, R. A. (2017). A database of global marine commercial, small-scale, illegal and unreported fisheries catch 1950–2014. *Scientific Data*, 4.
