---
title: 'OHIBC: CW trash layers prep'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = FALSE, message = FALSE, warning = FALSE)

library(rgdal)
library(raster)
source('~/github/ohibc/src/R/common.R')  ### an OHIBC specific version of common.R

dir_git     <- '~/github/ohibc'
dir_spatial <- file.path(dir_git, 'prep/spatial')  ### github: general buffer region shapefiles
dir_anx     <- file.path(dir_M, 'git-annex/bcprep')

### goal specific folders and info
goal     <- 'cw'
scenario <- 'v2016'
dir_goal <- file.path(dir_git, 'prep', goal, scenario)
dir_goal_anx <- file.path(dir_anx, goal, scenario)

### provenance tracking
library(provRmd); prov_setup()

### support scripts
source(file.path('~/github/ohibc/src/R/map_scores.R'))
  ### score plotting scripts
source(file.path(dir_git, 'src/R/rast_tools.R')) 
  ### raster plotting and analyzing scripts

### set up the default BC projection to be BC Albers
p4s_bcalb <- c('bcalb' = '+init=epsg:3005')

```

#Summary

This data is incorporated into the OHI British Columbia Clean Waters (CW) goal.

***

#Data Source

**Reference**: [van Sebille et al, 2015](http://iopscience.iop.org/article/10.1088/1748-9326/10/12/124006/meta)

**Downloaded**: September 8, 2016 from Figshare: https://figshare.com/articles/data_of_microplastic_abundance_and_mass_from_Figure_3_of_Van_Sebille_et_al_2015_ERL_paper/1613256

**Native Data Resolution**:   1.0° in WGS84 projection

**Values**: Count (number/km2) and mass density (g/km2), based upon three different ocean circulation models (Lebreton, Maximenko, van Sebille)  

**Time Range**: N/A

**Format**: CSV

***

# Methods

## Read in data

from: van Sebille et al 2015. A global inventory of small floating plastic debris.

``` {r read_vanSebille_data}
dir_data_vs <- file.path(dir_M, 'git-annex/bcprep/_raw_data/vanSebille_plastics')

vs_files_all <- list.files(dir_data_vs, full.names = TRUE) %>%
  setNames(basename(.) %>% str_replace('.csv', ''))

# [1] "gridcell_areas"            "latitudes"                 "lebretonmodel_abundance"  
# [4] "lebretonmodel_mass"        "longitudes"                "maximenkomodel_abundance" 
# [7] "maximenkomodel_mass"       "vansebillemodel_abundance" "vansebillemodel_mass"     

vs_files <- vs_files_all[c(3, 4, 6:9)] ### select just data layers for models

vs_rasts <- vector('list', length = length(vs_files)) %>%
  setNames(names(vs_files))

for (i in 1:length(vs_files)) { ### i = 1
  vs_file <- vs_files[i]
  vs_fn <- names(vs_files)[i]
  
  vs_matrix <- read_csv(vs_file, col_names = FALSE) %>%
    data.matrix() %>%
    .[2:181, 1:360]
    ### it appears that row 361 is glitchy; ditch it, and ditch +90 N row as well
  rownames(vs_matrix) <- c(90:-89)
  colnames(vs_matrix) <- c(0:359)
  
  vs_rast <- raster(vs_matrix)
  extent(vs_rast) <- c(0, 360, -90, 90)
  vs_rast <- rotate(vs_rast)
  crs(vs_rast) <- CRS('+init=epsg:4326')
  
  vs_rasts[[vs_fn]] <- vs_rast
}

vs_stack <- stack(vs_rasts)

bc_rgn_rast <- raster(file.path(dir_spatial, 'raster', 'ohibc_rgn_wgs84_30min.tif'))

vs_stack_crop <- crop(vs_stack, bc_rgn_rast)
plot(vs_stack_crop)

### Assume go with van Sebille model by mass
writeRaster(vs_stack_crop[['vansebillemodel_mass']], 
            file.path(dir_goal_anx, 'int/cw_trash_mass_vansebille.tif'), 
            overwrite = TRUE)

```

Crop to BC and interpolate using Thin Plate Spline

``` {r interpolate_bc_sla}

trash_dens <- raster(file.path(dir_goal_anx, 'int/cw_trash_mass_vansebille.tif'))

trash_dens_int_file <- file.path(dir_goal_anx, 'int/cw_trash_mass_int.tif')

if(!file.exists(trash_dens_int_file)) {
  message('Interpolating BC trash mass data')

  xy <- data.frame(xyFromCell(trash_dens, 1:ncell(trash_dens)))

  v  <- getValues(trash_dens)
  tmpdf <- cbind(xy, v) %>%
    filter(!is.na(v)) 
  xy1 <- tmpdf[ , 1:2]
  v1  <- tmpdf[ , 3]
  tps_model <- fields::Tps(xy1, v1)

  trash_dens_int <- interpolate(trash_dens, tps_model)

  trash_dens_gapfilled <- trash_dens
  values(trash_dens_gapfilled)[is.na(values(trash_dens))] <- values(trash_dens_int)[is.na(values(trash_dens))] 

  message('Writing gapfilled BC trash mass to: ', trash_dens_int_file)
  writeRaster(trash_dens_gapfilled, trash_dens_int_file, overwrite = TRUE)
  
} else {
  message('Reading gapfilled BC trash mass from: ', trash_dens_int_file)

  trash_dens_gapfilled <- raster(trash_dens_int_file)
}

plot(trash_dens_gapfilled)

```

Reproject into BC Albers at 1000 m resolution

```{r reproject}

rast_dens_int <- raster(file.path(dir_goal_anx, 'int/cw_trash_mass_int.tif'))

### reproject to BC Albers
rast_base_bcalb   <- raster(file.path(dir_spatial, 'raster/ohibc_rgn_raster_1000m.tif'))

trash_dens_bcalb <- projectRaster(from = rast_dens_int, to = rast_base_bcalb, method = 'ngb')
values(trash_dens_bcalb)[is.na(values(rast_base_bcalb))] <- NA


writeRaster(trash_dens_bcalb, file.path(dir_goal_anx, 'int/cw_trash_mass_bcalb.tif'), overwrite = TRUE)

``` 

``` {r plot plastic pollution rasters}

plot(trash_dens_bcalb, main = 'Mass density (g/km2)')

```


## Log Transform

Transform the values to a log(x + 1) value.

```{r log_transform}

trash_dens_bcalb <- raster(file.path(dir_goal_anx, 'int/cw_trash_mass_bcalb.tif'))

trash_dens_log <- calc(trash_dens_bcalb, fun = function(x) {log(x + 1)})

writeRaster(trash_dens_log, file.path(dir_goal_anx, 'int/cw_trash_mass_log.tif'), overwrite = TRUE)

plot(trash_dens_log, main = 'Mass density\nlog(g/km2)')

```

## Rescale zero to 1

Using the 99.99%ile, rescale all values to range from zero to 1.

```{r quantile}

trash_dens_log <- raster(file.path(dir_goal_anx, 'int/cw_trash_mass_log.tif'))

trash_dens_ref <- quantile(trash_dens_log, prob = c(0.001, 0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99, 0.999, 0.9999))

trash_dens_99 <- trash_dens_ref[10]

print(hist(trash_dens_log, main = 'Mass density (log(g/km2))'))

trash_dens_rescale <- calc(trash_dens_log, fun = function(x) {ifelse(x > trash_dens_99, 1, x / trash_dens_99)}) %>%
  setNames('log_mass')

writeRaster(trash_dens_rescale, file.path(dir_goal_anx, 'rescaled_layers/cw_trash_dens_rescale.tif'), overwrite = TRUE)

plot(trash_dens_rescale, main = 'CW trash pressure layer: rescaled mass density')

```

