---
title: 'OHIBC: Ocean Acidification Pressure layers prep'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(sp)        # the classes and methods that make up spatial ops in R
library(rgdal)
library(raster)
library(DT)

source('src/R/common.R')  
  ### an OHIBC specific version of common.R; must be in ohibc working dir

dir_git <- '~/github/ohibc'
dir_anx  <- file.path(dir_neptune_data, 'git-annex/bcprep')
dir_rgn <- file.path(dir_git, 'prep/regions')  ### github: general buffer region shapefiles

### goal specific folders and info
goal      <- 'pressures'
scenario  <- 'v2016'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario)
dir_goal_anx <- file.path(dir_anx, goal, scenario)

### provenance tracking
source('~/github/ohibc/src/R/prov.R')      
  ### Provenance tracking functions: must source at start to initialize prov_track
dir_prov <- file.path(dir_goal, 'prov') 
  ### set a provenance folder for this script
this_script_file <- file.path(dir_goal, 'pressures_lyr_prep.Rmd') 
  ### can't recognize the current script name on its own :(
prov_run_tag <- 'standard run'

### goal-specific source scripts
source(file.path(dir_goal, 'pressures_lyr_fxns.R'))

### other support functions
source(file.path(dir_git, 'src/R/rast_tools.R'))

### set up proj4string options: BC Albers and WGS84
p4s_wgs84 <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0')
p4s_bcalb <- '+proj=aea +lat_1=50 +lat_2=58.5 +lat_0=45 +lon_0=-126 +x_0=1000000 +y_0=0 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0'

reload = FALSE
```

# Summary

There are two parts to creating this layer:

1. Data prep to get raw data into the correct format:
    * Aggregate monthly values for aragonite saturation state to an annual mean value per cell, for baseline decade and current study period.
    * Clip rasters to extents of OHIBC region, and reproject into BC Albers projection.
    * Determine baseline by calculating decadal mean for 1880-1889.
2. Creating the pressure layers for OHIBC:
    * Determine change in aragonite saturation state since baseline for each year of study period (2005-2014)
    * Rescale to values from 0 - 1.
    * Interpolate to coastline at 1 km resolution.
    
This process is completed entirely within this script.

# Updates

Not applicable.

# Data

Raw data was provided by Woods Hole on December 16, 2014. This data is an update to the work done by Feely et al. (2009).

Native Data Resolution: 1 degree
Values: Surface Ω aragonite saturation state
Time Range: 1880-1889 and 2005-2014 (monthly data was provided for each year)
Format: NetCDF

# Methods

## Read raw data

Raw data is read in for 1880 - 1889 and 2005 - 2014 datasets.  Latitude, longitude, and aragonite saturation state $\Omega$ are extracted for each year in each dataset.  Land cells are set to NA.

``` {r read_aragonite_lyrs, echo = FALSE}

library(ncdf4)

### read in 1880-1889 data
oa_data_dir <- file.path(dir_neptune_data, 'git-annex/globalprep/Pressures_OceanAcidification/v2015')
data_nc_18 <- nc_open(file.path(oa_data_dir, 'input/cesm_co2sys_1880-1889.nc'))
# print(data_nc_18)

### read in 2005-2014 data
data_nc_20 = nc_open(file.path(oa_data_dir, 'input/cesm_co2sys_2005-2014.nc'))
# print(data_nc_20)

### longitude values are stored in the variable 'TLONG'; lat in 'TLAT'
long <- ncvar_get(data_nc_18, varid = 'TLONG') #same lat and long for both data sets so just reading in from data_nc_18 works
lat  <- ncvar_get(data_nc_18, varid = 'TLAT')

### select the surface aragonite variable (OARG)
arag_18 <- ncvar_get(data_nc_18, varid = 'OARG')
arag_20 <- ncvar_get(data_nc_20, varid = 'OARG')

### land cells are given large, strange value, so set these to NA
### arag cell values range from zero to 5; set all out of that to NA?
land_val <- max(arag_18, na.rm = TRUE)
arag_18[arag_18 >= land_val] <- NA
range_a18 <- round(range(arag_18, na.rm = TRUE), 3)
arag_20[arag_20 >= land_val] <- NA 
range_a20 <- round(range(arag_20, na.rm = TRUE), 3)

```

Range of $\Omega_{1880-1889}$ for global dataset: `r paste(range_a18, collapse = ' - '))`.

Range of $\Omega_{2005-2014}$ for global dataset: `r paste(range_a20, collapse = ' - '))`.

## Convert raw global data to OHIBC region data

The raw data, in WGS84 coordinate ref system, is rasterized, clipped down to a bounding box around the OHIBC region, and reprojected into BC Albers projection.  This data is then resampled to 1 km resolution using nearest neighbor method.

The resulting raster layers (in rasterBrick format) include annual means for each coarse cell (approx 1 degree, from original resolution) for each year in the dataset.  The two decades (1880-1889 and 2005-2014) are assigned to two separate rasterBricks.

``` {r load_rgn, echo = FALSE}

rgn_lyr <- 'ohibc_rgn'

cat(sprintf('Reading OHIBC regions shapefile...\n  %s/%s.shp', dir_rgn, rgn_lyr))
rgn_poly <- readOGR(dsn = path.expand(dir_rgn), layer = rgn_lyr,
                       verbose = FALSE, stringsAsFactors = FALSE)

rgn_poly_wgs84 <- readOGR(dsn = path.expand(dir_rgn), 
                         layer = paste(rgn_lyr, '_wgs84', sep = ''),
                         verbose = FALSE, stringsAsFactors = FALSE)

```

``` {r create_rast_bricks, echo = FALSE}
### generate raster bricks for decades 1880-1889 and 2005-2014
### then: clip to extents of region (in WGS 84 CRS)
### then: reproject to BC Albers projection
rb_a18_file <- file.path(dir_goal, 'int/oa/rbrick_a18_1km_raw.tif')
rb_a20_file <- file.path(dir_goal, 'int/oa/rbrick_a20_1km_raw.tif')

if(!file.exists(rb_a18_file) | reload) {
  rbrick_arag_18 <- annual_mean(arag_18, yr_init = 1880) %>%
    raster::crop(rgn_poly_wgs84, snap = 'out') %>%
    raster::projectRaster(crs = p4s_bcalb)
  rbrick_arag_20 <- annual_mean(arag_20, yr_init = 2005) %>%
    crop(rgn_poly_wgs84, snap = 'out') %>%
    raster::projectRaster(crs = p4s_bcalb)

  ### resample to 1 km... use base raster from regions folder (note: 500m base raster available too)
  rast_base <- raster(file.path(dir_rgn, 'ohibc_base_raster_1000m.tif')) %>%
    raster::projectRaster(crs = p4s_bcalb) ### why is this not bc albers to begin with?
  
  ### why does resample fail on a raster brick:
  # rb_arag_18_resamp <- raster::resample(rbrick_arag_18, rast_base,
  #                              method   = 'ngb',
  #                              progress = 'text',
  #                              filename = file.path(dir_goal, 'int/oa/rbrick_arag_18_raw.tif'),
  #                              overwrite = TRUE)
  # rb_arag_20_resamp <- resample(rbrick_arag_20, rast_base,
  #                              method   = 'ngb',
  #                              progress = 'text',
  #                              filename = file.path(dir_goal, 'int/oa/rbrick_arag_20_raw.tif'),
  #                              overwrite = TRUE)
  
  rb_a18_resamp <- resample_brick(rbrick_arag_18, rast_base,
                                 filename = rb_a18_file)
  rb_a20_resamp <- resample_brick(rbrick_arag_20, rast_base,
                                 filename = rb_a20_file)
} else {
  rb_a18_resamp <- brick(rb_a18_file)
  rb_a20_resamp <- brick(rb_a20_file)
  
  names(rb_a18_resamp) <- paste('y', c(1880:1889), sep = '')
  names(rb_a20_resamp) <- paste('y', c(2005:2014), sep = '')
}

plot_rast(rb_a18_resamp[['y1880']], 
          title = 'Annual mean arag. sat state, 1880', 
          scale_label = 'omega',
          rev_scale = FALSE)

plot_rast(rb_a20_resamp[['y2014']], 
          title = 'Annual mean arag. sat state, 2014', 
          scale_label = 'omega',
          rev_scale = FALSE)

```

## Rescale values

Baseline aragonite saturation state $\Omega_{base}$ is based upon the decadal average from 1880-1889.

Deviation from aragonite saturation state is determined for each year in the study period:

$$\delta \Omega_{year} = (\Omega_{base} - \Omega_{year}) / \Omega_{year} $$

Note that current is subtracted from the baseline; this way, a reduction in $\Omega $ becomes a positive pressure value.  It is then normalized by the current mean state; so a decrease in $\Omega $ while the current state is high indicates less pressure than the same decrease when the current state is near 1. 

$\delta \Omega_{year} $ is then modified to account for increases in aragonite saturation state (pressure = 0) and arag sat state less than 1 (pressure = 1).

``` {r rescale_values, echo = FALSE}
### Get decadal mean of 1880-1889 as baseline
oa_mean_baseline <- calc(rb_a18_resamp, mean)
# plot(oa_mean_baseline)
# plot(rgn_poly, add = TRUE)
plot_rast(oa_mean_baseline, 
          title = 'Mean arag. sat state, 1880-1889', 
          scale_label = 'omega',
          rev_scale = FALSE)


### get raster brick of change in aragonite sat state.
### Note that by dividing by the current amount, we are
### normalizing - so a given drop in omega, with a high
### current sat state, will result in a lower pressure than
### the same drop at a lower saturation state.

# oa_delta <- oa_mean_baseline - rb_a20_resamp
oa_delta <- (oa_mean_baseline - rb_a20_resamp)/rb_a20_resamp
names(oa_delta) <- names(rb_a20_resamp)
# plot(oa_delta)

### If current mean sat state > 1, set pressure to 1.  If delta
### is negative (current state > baseline state) set to zero. 
oa_rescaled <- oa_delta
oa_means <- rb_a20_resamp

values(oa_rescaled)[values(oa_means) <= 1] <- 1    
  ### all values at or less than 1 are given a value of 1
values(oa_rescaled)[values(oa_delta)  < 0] <- 0
  ### all delta less than 0 (indicating a decrease in acidity) are capped at 0
plot_rast(oa_rescaled[['y2014']], 
          title = 'OA rescaled 2014', 
          scale_label = 'oa',
          rev_scale = TRUE)
```

## Interpolate $\delta \Omega_{year}$ to coastline at 1 km

Now that $\delta \Omega_{year}$ has been determined at a 1 degree cell scale, we interpolate that to the coastlines to fill NAs.  Using `raster::interpolate()`, we select a random sample of points from the $\delta \Omega_{year}$ raster (for speed and memory) and apply the thin plate spline model (`fields::Tps()`).

``` {r interpolate_mask_save, echo = FALSE}
oa_raster_file <- c(file.path(dir_goal, 'data/oa/oa_rescaled_2005-2014.tif'))

if(!file.exists(oa_raster_file) | reload) {
  ### from raster bricks, interpolate/extrapolate to extent of raster.  First,
  ### set up thin plate spline model from x-y of raster and values of raster
  oa_rescaled_int <- interpolate_brick(oa_rescaled, subset_n = 2000)
    ### sampling to avoid memory problems; the Tps model seems to choke
    ### with too many points
  
  # plot(oa_rescaled_int)
  # plot(oa_rescaled)
  
  ### mask to OHIBC regions
  oa_resc_int_rgn <- mask(oa_rescaled_int, rgn_poly)
  # plot(oa_resc_int_rgn[[1]])
  # plot(rgn_poly, add = TRUE)
  
  writeRaster(oa_resc_int_rgn,
              filename  = oa_raster_files, 
              bylayer   = FALSE, 
              bandorder = 'BIL',
              overwrite = TRUE)
}


# rbrick <- brick(oa_raster_file)
# names(rbrick) <- str_replace(names(rbrick), 'oa_rescaled_', 'y')

plot_rast(rbrick[['y2014']], 
          title = 'OA rescaled 2014, interpolated to 1 km', 
          scale_label = 'oa',
          rev_scale = TRUE)

```

-----

``` {r child = file.path(dir_git, 'src/templates/ohibc_prov_ftr.Rmd')}
```
