---
title: 'OHIBC: Fisheries modified Kobe plots'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

source('~/github/ohibc/src/R/common.R')  ### an OHIBC specific version of common.R;
### includes library(tidyverse); library(stringr)
library(ggplot2)


dir_git     <- '~/github/ohibc'

### goal specific folders and info
goal      <- 'fis'
scenario  <- 'kobe'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario)

source(file.path(dir_goal, 'kobe_fxns.R'))
### where the functions are defined

```

Kobe plot based upon B' and F', which are in turn based upon $F/F_{MSY}$ and $B/B_{MSY}$ relative to thresholds

$$x_{fis} = \displaystyle\sum_{i=1}^{n}w_i*\left(\frac{F' + B'}{2}\right)$$

* To calculate $B'$, we define categories in which a given stock is overfished, underfished, or adequately fished based on $B/B_{MSY}$.  These are based on $B/B_{MSY}$ thresholds of $\theta_{OF}$ = 0.8 (overfished) and $\theta_{UF}$ = 1.5 (underfished): 

| $B'$ calculation        | Stock condition                                |
| :---------------------: | :--------------------------------------------- |
| $B' = \frac{B/B_{MSY}}{0.8}$ | Overfished ($B/B_{MSY} < \theta_{OF}$) |
| $B' = 1$                | Appropriately fished ($\theta_{OF} <= B/B_{MSY} < \theta_{UF}$) |
| $B' = \frac{(B/B_{MSY})_{max} - (B/B_{MSY})}{(B/B_{MSY})_{max} - 1.5}$ | Underfished ($B/B_{MSY} >= \theta_{UF}$) |

* To calculate F', we examine fishing pressure within two regimes: overfished stock ($B/B_{MSY} < \theta_{OF}$) and not overfished stock ($B/B_{MSY} >= \theta_{OF}$).  Within each regime we divide fishing pressure into overfishing, underfishing, and appropriate fishing pressures based on $F/F_{MSY}$, including under- and overfishing $F/F_{MSY}$ thresholds of $\psi_{UF}$ = 0.8 and $\psi_{OF}$ = 1.2, respectively.

| $F'$ calculation        | Stock condition    | Fishing pressure                       |
| :---------------------: | :----------------- | :---------------------------           |
| $F' = 0$                | Overfished         | Gross overfishing ($F/F_{MSY} > B/B_{MSY} + 1.5$) |
| $F' = \frac{(B/B_{MSY} + 1.5) - F/F_{MSY}}{1.5}$ | Overfished | Moderate overfishing ($B/B_{MSY} + 0.2 < F/F_{MSY} <= B/B_{MSY} + 1.5$) |
| $F' = 1$                | Overfished         | Appropriate fishing ($B/B_{MSY} - 0.2 <= F/F_{MSY} < B/B_{MSY} + 0.2$) |
| $F' = \frac{F/F_{MSY}}{B/B_{MSY} - 0.2}$  | Overfished | Underfishing ($F/F_{MSY} < B/B_{MSY} - 0.2$) |
| $F' = \frac{F/F_{MSY}}{\psi_{UF}}$ | _Not_ overfished | Underfishing ($F/F_{MSY} < \psi_{UF}$)
| $F' = 1$                | _Not_ overfished   | Appropriate fishing ($\psi_{UF} <= F/F_{MSY} < \psi_{OF}$) |
| $F' = \frac{(F/F_{MSY})_{max} - F/F_{MSY}}{(F/F_{MSY})_{max} - \psi_{OF}}$  | _Not_ overfished | Overfishing ($F/F_{MSY} >= \psi_{OF}$) |


The following plots assume a local maximum (across all fisheries within region of study) $B/B_{MSY}$ of 3.5, and a local maximum $F/F_{MSY}$ of 2.5.

``` {r generate_kobe_plot, eval = TRUE}

f_fmsy_max <- 2.5
b_bmsy_max <- 3.5
reso = 0.01

fish_stat_df <- data.frame(f_fmsy = rep(seq(0, f_fmsy_max, reso), each  = b_bmsy_max/reso + 1),
                           b_bmsy = rep(seq(0, b_bmsy_max, reso), times = f_fmsy_max/reso + 1))

kobe <- fish_stat_df %>% 
  rescale_bprime(overfished_th = 0.8) %>%
  rescale_fprime(overfished_th = 0.8) %>%
  mutate(x = (fPrime + bPrime)/2)

kobe_fPrime_plot <- ggplot(data = kobe, aes(x = b_bmsy, y = f_fmsy, fill = fPrime)) +
  ggtheme_plot + 
  geom_raster(alpha = .8) +
  scale_fill_distiller(palette = 'RdYlGn', direction = 1) +
  labs(title = 'F\' = f(B/Bmsy, F/Fmsy)',
       fill = 'F\' value',
       x = 'B/Bmsy',
       y = 'F/Fmsy')

print(kobe_fPrime_plot)
ggsave(filename = file.path(dir_goal, 'kobe_fPrime.png'), width = 8, height = 6, units = 'cm')

kobe_fis_score_plot <- ggplot(data = kobe, aes(x = b_bmsy, y = f_fmsy, fill = x)) +
  ggtheme_plot + 
  geom_raster(alpha = .8) +
  scale_fill_distiller(palette = 'RdYlGn', direction = 1) +
  labs(title = 'FIS = f(B/Bmsy, F/Fmsy)',
       fill = 'FIS score',
       x = 'B/Bmsy',
       y = 'F/Fmsy')

print(kobe_fis_score_plot)
ggsave(filename = file.path(dir_goal, 'kobe_fis_score.png'), width = 8, height = 6, units = 'cm')
```

***

The following plots incorporate a "critical" zone for $B/B_{MSY}$ below which fishing is prohibited.  BC Precautionary approach sets the critical value at 40% of $B_{MSY}$.  For our calculations we will allow a -.1/+.3 tolerance around this to allow for uncertainty, though the tolerances can certainly be adjusted as necessary.  The overfishing values will also be adjusted to drop to zero for all cases where B/Bmsy falls below (B/Bmsy critical - tolerance).

The following plots assume a local maximum (across all fisheries within region of study) $B/B_{MSY}$ of 3.5, and a local maximum $F/F_{MSY}$ of 2.5.

``` {r generate_kobe_plot_crit}


kobe <- generate_kobe_df(f_fmsy_max = 2.5,
                         b_bmsy_max = 3.0,
                         bmax_val = 1/3)

hcr_df <- data.frame(b_bmsy = c(0, .4, .8, 3),
                     f_fmsy = c(0,  0,  1, 1))

kobe_fPrime_plot <- ggplot(data = kobe, aes(x = b_bmsy, y = f_fmsy)) +
  ggtheme_plot + 
  geom_raster(alpha = .8, aes(fill = fPrime)) +
  scale_fill_distiller(palette = 'RdYlGn', direction = 1) +
  geom_line(data = hcr_df, aes(x = b_bmsy, y = f_fmsy), color = 'black', size = 1.5, alpha = .6) +
  labs(title = 'F\' = f(B/Bmsy, F/Fmsy)',
       fill = 'F\' value',
       x = 'B/Bmsy',
       y = 'F/Fmsy')

print(kobe_fPrime_plot)
ggsave(filename = file.path(dir_goal, 'kobe_fPrime_crit.png'), width = 8, height = 6, units = 'cm')

kobe_bPrime_plot <- ggplot(data = kobe, aes(x = b_bmsy, y = f_fmsy)) +
  ggtheme_plot + 
  geom_raster(alpha = .8, aes(fill = bPrime)) +
  scale_fill_distiller(palette = 'RdYlGn', direction = 1) +
  geom_line(data = hcr_df, aes(x = b_bmsy, y = f_fmsy), color = 'black', size = 1.5, alpha = .6) +
  labs(title = 'B\' = f(B/Bmsy)',
       fill = 'B\' value',
       x = 'B/Bmsy',
       y = 'F/Fmsy')

print(kobe_bPrime_plot)
ggsave(filename = file.path(dir_goal, 'kobe_bPrime_crit.png'), width = 8, height = 6, units = 'cm')

kobe_fis_score_plot <- ggplot(data = kobe, aes(x = b_bmsy, y = f_fmsy)) +
  ggtheme_plot + 
  geom_raster(alpha = .8, aes(fill = x_geom)) +
  scale_fill_distiller(palette = 'RdYlGn', direction = 1) +
  geom_line(data = hcr_df, aes(x = b_bmsy, y = f_fmsy), color = 'black', size = 1.5, alpha = .6) +
  labs(title = 'FIS = f(B/Bmsy, F/Fmsy)',
       fill = 'FIS score',
       x = 'B/Bmsy',
       y = 'F/Fmsy') +
  annotate(geom = 'text', label = 'critical', x = .15, y = -.1, 
           size = 2, 
           color = 'grey20') + 
  annotate(geom = 'text', label = 'cautious', x =  .5, y = -.1, 
           size = 2, 
           color = 'grey20') + 
  annotate(geom = 'text', label = 'healthy',  x = 1.2, y = -.1, 
           size = 2, 
           color = 'grey20') + 
  annotate(geom = 'text', label = 'underexploited',  x = 2.5, y = -.1, 
           size = 2, 
           color = 'grey20')

print(kobe_fis_score_plot)
ggsave(filename = file.path(dir_goal, 'kobe_fis_score_crit.png'), width = 8, height = 6, units = 'cm')

kobe_fis_score_plot2 <- ggplot(data = kobe, aes(x = b_bmsy, y = f_fmsy)) +
  ggtheme_plot + 
  geom_raster(alpha = .8, aes(fill = x_arith)) +
  scale_fill_distiller(palette = 'RdYlGn', direction = 1) +
  geom_line(data = hcr_df, aes(x = b_bmsy, y = f_fmsy), color = 'black', size = 1.5, alpha = .6) +
  labs(title = 'FIS = f(B/Bmsy, F/Fmsy)',
       fill = 'FIS score',
       x = 'B/Bmsy',
       y = 'F/Fmsy') +
  annotate(geom = 'text', label = 'critical', x = .15, y = -.1, 
           size = 2, 
           color = 'grey20') + 
  annotate(geom = 'text', label = 'cautious', x =  .5, y = -.1, 
           size = 2, 
           color = 'grey20') + 
  annotate(geom = 'text', label = 'healthy',  x = 1.2, y = -.1, 
           size = 2, 
           color = 'grey20') + 
  annotate(geom = 'text', label = 'underexploited',  x = 2.5, y = -.1, 
           size = 2, 
           color = 'grey20')

print(kobe_fis_score_plot2)
ggsave(filename = file.path(dir_goal, 'kobe_fis_score_crit_arith.png'), width = 8, height = 6, units = 'cm')

```

*** 

