---
title: 'OHIBC: land use raster prep'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(raster)
library(rgdal)
library(rgeos)

source('~/github/ohibc/src/R/common.R')  ### an OHIBC specific version of common.R

dir_git     <- '~/github/ohibc'
dir_spatial <- file.path(dir_git, 'prep/spatial')  ### github: general buffer region shapefiles
dir_anx     <- file.path(dir_M, 'git-annex/bcprep')

### goal specific folders and info
goal      <- 'hab_cs_cp'
scenario  <- 'v2017'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario)
dir_goal_anx <- file.path(dir_anx, goal, scenario)

### provenance tracking
library(provRmd); prov_setup()

### support scripts
source(file.path(dir_git, 'src/R/rast_tools.R')) 
  ### gdal_rast2, plotting

### set up the default BC projection to be BC Albers
p4s_bcalb <- c('bcalb' = '+init=epsg:3005')

```

# Summary

Merge landuse rasters, convert to BC Albers projection, and crop to BC extents

# Updates from previous assessment

***

# Data Source [NOTE: can be copied from README.md in rawdata file]
**Reference**: [citation for source data; website, literature, contact information. Version of data (if relevant). Screenshots if a series of menus had to be navigated to obtain the data.]

**Downloaded**: [date downloaded or received]

**Description**:  [e.g., surface aragonite state]

**Native data resolution**: [e.g., 1 degree, 30 m, etc.]   

**Time range**: [e.g., 1880-1899, monthly data provided for each year] 

**Format**:  [e.g. NetCDF]

**classifications**:

- 11 Unclassified: Areas not classified due to clouds
- 21 Settlement: Built-up and urban
- 25 Roads: Primary, secondary and tertiary
- 31 Water: Natural and human-made
- 41 Forest: Treed areas >1 ha in size
- 42 Forest Wetland: Wetland with forest cover
- 45 Trees: Treed areas <1 ha in size
- 46 Treed Wetland: Wetland with tree cover
- 51 Cropland: Annual and perennial
- 61 Grassland Managed: Natural grass and shrubs used for cattle grazing
- 62 Grassland Unmanaged: Natural grass and shrubs with no apparent use (forest openings, alpine meadows, tundra, etc.)
- 71 Wetland: Undifferentiated wetland
- 73 Wetland Shrub: Wetland with shrub cover
- 74 Wetland Herb: Wetland with grass cover
- 91 Other land: Rock, beaches, ice, barren land 

***
  
# Prepare rasters

Combine three UTM region maps (8, 9, 10) for each of three years (1990, 2000, 2010).  Each raw raster is transformed to BC Albers for equal area projection, and then the three rasters for each year are merged into one complete raster.

``` {r reproject and merge land cover rasters}

### set up pathnames for final rasters
rast_landuse_files <- c('1990' = file.path(dir_goal_anx, 'habs/land_cover_1990_30m.tif'),
                        '2000' = file.path(dir_goal_anx, 'habs/land_cover_2000_30m.tif'),
                        '2010' = file.path(dir_goal_anx, 'habs/land_cover_2010_30m.tif'))
rast_data_dir <- file.path(dir_anx, '_raw_data/gov_of_canada/d2016/land_use_90_00_10')

### look for final rasters with no associated file  
rast_landuse_files <- rast_landuse_files[!file.exists(rast_landuse_files)]

for(i in 1:length(rast_landuse_files)) {
  # i <- 3
  
  year <- names(rast_landuse_files[i])
  
  ### find all raster files for this year; there should be three zones
  rast_yr <- list.files(rast_data_dir,
                        full.names = TRUE,
                        pattern = year)
  rast_yr <- rast_yr[str_detect(rast_yr, '.tif$')]
  
  # rast_list <- lapply(rast_yr, raster) ### read 'em all in

  ### have to reproject; each is in a different UTC projection, project all to BC Albers
  # y <- parallel::mclapply(rast_list, mc.cores = 12,
  #   FUN = function(x) {
  #     if(minValue(x) < 1) values(x)[values(x) < 1] <- NA
  #     projectRaster(x, res = res(x), method = 'ngb', crs = p4s_bcalb,
  #                   filename = file.path(dir_goal_anx, 'habs', paste0(names(x), '.tif')),
  #                   overwrite = TRUE,
  #                   progress = 'text')
  # })
  x1 <- file.path(dir_goal_anx, 'habs', basename(rast_yr))
  y <- lapply(x1, raster)
  # for(j in 1:length(y)) { # j <- 1
  #   cat(paste0('Checking ', names(y[[j]]), '\n'))
  #   if(minValue(y[[j]]) < 1) {
  #     cat('  Converting zeroes...\n')
  #     values(y[[j]])[values(y[[j]]) < 1] <- NA
  #   }
  # }
  
  ### merge the reprojected rasters; the origins are off by ~15 m, or about half
  ### the resolution, so tolerance = 0.5 to allow for the mismatch.
  ### Saves completed rasters in git-annex/bcprep/hab_cs_cp/v2016/habs
  ### b/c the UTM 9 raster has zeros instead of NAs, merge this one last; merge will keep earlier non-NA values over the zeros
  z <- raster::merge(y[[1]], y[[3]], y[[2]],  
                     tolerance = 0.5,
                     filename = rast_landuse_files[i],
                     progress = 'text')
}

```

-----

``` {r prov_footer, results = 'asis'}
prov_wrapup()
```

