---
title: 'OHIBC: Habitat goals - saltmarsh and coastal forest layers prep'
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(raster)
library(rgdal)

source('~/github/ohibc/src/R/common.R')  ### an OHIBC specific version of common.R

dir_git     <- path.expand('~/github/ohibc') %>% path.expand()
dir_spatial <- file.path(dir_git, 'prep/spatial')  ### github: general buffer region shapefiles
dir_anx     <- file.path(dir_M, 'git-annex/bcprep')

### goal specific folders and info
goal      <- 'hab_cs_cp'
scenario  <- 'v2017'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario)
dir_v2016_anx <- file.path(dir_anx, goal, 'v2016')
dir_goal_anx <- file.path(dir_anx, goal, scenario)
dir_spatial_anx <- file.path(dir_anx, 'spatial')

### provenance tracking
library(provRmd); prov_setup()

### support scripts
source(file.path(dir_git, 'src/R/rast_tools.R')) 
  ### gdal_rast2, plotting

### set up the default BC projection to be BC Albers
p4s_bcalb <- c('bcalb' = '+init=epsg:3005')

```

# Summary

Prepare existing habitat layers (salt marsh and forest raster land use raster) for HAB, CP, CS goals for OHIBC assessment.  For CP, calculate values based on exposure, elevation, and protective potential weighting as required for different protective habitats.  For CS, calculate values based on carbon storage potential.

Assessment areas for both habitats will be limited to the coastal watersheds, i.e. BC subwatersheds that intersect a 1 km buffer from the coastline.  For coastal protection, these will further be limited by an exposure band only 2 km wide.

***

# Data Sources

<!-- * **Seagrass and salt marsh polygons** -->
<!--     * **Reference**: CEC. 2016. North America’s Blue Carbon: Assessing Seagrass, Salt Marsh and Mangrove Distribution and Carbon Sinks. Montreal, Canada: Commission for Environmental Cooperation. 54 pp. -->
<!--     * **Downloaded**: http://www.cec.org/sites/default/files/Atlas/Files/blue_carbon/BlueCarbon_Shapefile.zip (downloaded 8/25/16) -->
<!--     * **Description**:  shapefiles of marine seagrass, mangrove, and salt marsh habitat in North America -->
<!--     * **Native data resolution**: NA    -->
<!--     * **Time range**: NA (data were collected over many years but no time series available) -->
<!--     * **Format**:  vector shapefiles -->
* **Land cover raster 1990-2000-2010**
    * **Reference**: Land Use 1990, 2000 and 2010. 2015. Agriculture and Agri-Foods Canada.
    * **Downloaded**: http://open.canada.ca/data/en/dataset/18e3ef1a-497c-40c6-8326-aac1a34a0dec (downloaded 10/7/16)
    * **Description**:  Land cover raster covering all of Canada south of 60° N
    * **Native data resolution**: 30 meters   
    * **Time range**: 1990, 2000, 2010
    * **Format**:  GeoTIFF
* **Exposure polylines and coastal class polylines**
    * **Reference**: 
        * British Columbia Marine Conservation Analysis Project Team. 2011. Shorezone exposure. Marine Atlas of Pacific Canada: A Product of the British Columbia Marine Conservation Analysis. Available from www.bcmca.ca.
        * British Columbia Marine Conservation Analysis Project Team. 2011. Coastal classes. Marine Atlas of Pacific Canada: A Product of the British Columbia Marine Conservation Analysis. Available from www.bcmca.ca.
    * **Downloaded**: http://bcmca.ca/datafiles/sets/bcmca_eco_set_physical_atlas.zip (downloaded 11/5/15)
    * **Description**:  Marxan-prepared polylines representing exposure and coastal class along BC coastline
    * **Native data resolution**: NA; shorezone reference units are 400 m
    * **Time range**: NA
    * **Format**:  Polyline shapefiles
* **Elevation raster**
    * **Reference**: National Geophysical Data Center. 2006. 2-minute Gridded Global Relief Data (ETOPO2) v2. National Geophysical Data Center, NOAA. doi:10.7289/V5J1012Q [7/26/16].
    * **Downloaded**: https://www.ngdc.noaa.gov/mgg/global/etopo2.html (downloaded 7/26/16)
    * **Description**:  Global elevation and bathymetry
    * **Native data resolution**: 2 arc-minutes   
    * **Time range**: NA
    * **Format**:  NetCDF, GeoTIFF
<!-- * **Population density raster** -->
<!--     * **NOTE**: Population weighting was considered based on InVEST coastal vulnerability model; however, after discussion with Andrew Day and Karin Bodtker of Vancouver Aquarium, population weighting will not be used.  -->
<!--     * **Reference**: Center for International Earth Science Information Network - CIESIN - Columbia University. 2016. Documentation for the Gridded Population of the World, Version 4 (GPWv4). Palisades NY: NASA Socioeconomic Data and Applications Center (SEDAC). http://dx.doi.org/10.7927/H4D50JX4 Accessed 25 Oct 2016. -->
<!--     * **Downloaded**: http://sedac.ciesin.columbia.edu/data/collection/gpw-v4/sets/browse (downloaded 10/25/16) -->
<!--     * **Description**:  Population density raster, adjusted to UN 2015 values; pop density in persons per km^2^ -->
<!--     * **Native data resolution**: ???? 2.5 arc-minutes    -->
<!--     * **Time range**: 2000, 2005, 2010, 2015, 2020 (only 2015 used for this analysis) -->
<!--     * **Format**:  GeoTIFF -->
* **British Columbia coastal watersheds**
    * **Reference**:
    * **Downloaded**:  (downloaded 11/30/16)
    * **Description**:  Watershed outlines for British Columbia
    * **Native data resolution**: NA
    * **Time range**: NA
    * **Format**:  ESRI Shapefile modified from geodatabase, rasterized to 500 m resolution.
    
***
  
# Methods

* reproject elevation raster to region raster parameters
* Extrapolate exposure polylines
    * identify appropriate exposures and rank from very protected (1) to very exposed (6)
    * buffer out to 2 km; reorder by exposure, so very exposed is bottom of list
    * rasterize using gdal_rast2 function, last value is priority
* crop land cover raster to region raster extents and mask to coastal regions (3 nm offshore, coastal watersheds inland)

## Prepare support rasters

### check for inland and offshore buffer area rasters

This assessment will use prepared rasters at 30 m resolution for inland (coastal watershed buffer) and offshore.  These rasters will be used to mask analysis areas.  The 30 m resolution matches the resolution of the land use data.

``` {r check_region_rasters}

### 
rast_inland_file   <- file.path(dir_spatial_anx, 'raster/ohibc_coastal_watersheds_30m.tif')
rast_offshore_file <- file.path(dir_spatial_anx, 'raster/ohibc_offshore_3nm_30m.tif')

if(any(!file.exists(c(rast_inland_file, rast_offshore_file))))
  stop('One or more of the region raster files is missing')

rast_rgn_file <- file.path(dir_goal_anx, 'other_spatial/rast_in_off_30m.tif')

if(!file.exists(rast_rgn_file)) {
  ### combine inland and offshore raster values
  rast_rgn_inland_30m   <- raster(rast_inland_file)
  rast_rgn_offshore_30m <- raster(rast_offshore_file)
  rast_in_off_30m <- rast_rgn_inland_30m
  values(rast_in_off_30m)[!is.na(values(rast_rgn_offshore_30m))] <- values(rast_rgn_offshore_30m)[!is.na(values(rast_rgn_offshore_30m))]
  writeRaster(rast_in_off_30m, file.path(dir_goal_anx, 'other_spatial/rast_in_off_30m.tif'), overwrite = TRUE)
}

```

### create elevation raster

Prepare elevation raster by reprojecting global raster to the extents, projection, and resolution of OHIBC base region raster.  This raster is used as a weighting factor in the Coastal Protection goal.

``` {r prep_elev_rasters}

rast_rgn_30m <- raster(rast_rgn_file)

topo_rast_file  <- file.path(dir_goal_anx, 'other_spatial/ohibc_topo_30m.tif')

if(!file.exists(topo_rast_file)) {
  message('Creating file: \n  ', topo_rast_file)
  
  
  ### ETOPO v2 in Mazu:
  ### git-annex/bcprep/_raw_data/bathymetry/d2015/ETOPO2v2_bathymetry
  etopo <- raster(file.path(dir_anx, '_raw_data/bathymetry/d2015/ETOPO2v2_bathymetry',
                            'ETOPO2v2c.tif'))
  ### add CRS and extent info:
  crs(etopo) <- '+init=epsg:4326'
  extent(etopo) <- c(-180, 180, -90, 90)
  
  rast_topo <- projectRaster(etopo, rast_rgn_30m)
  
  ### trim to just values near sea level, for file size
  values(rast_topo)[values(rast_topo) > 10 | values(rast_topo) < -10] <- NA
  
  writeRaster(rast_topo, topo_rast_file, overwrite = TRUE) 
  
} else {
  
  message('File exists: \n  ', topo_rast_file)
  git_prov(topo_rast_file, filetype = 'output')
  
}


```

### Prepare exposure raster

Exposure raster is created by extending a 2 km buffer from shoreline polylines to create polygons with values indicating levels of wave exposure, from 1 (very protected) to 6 (very exposed).  The buffer polygons are rasterized, and where two buffers overlap (at boundaries between two exposure classes), the polygons within the shapefile have been ordered to raster out to the maximum exposure class.

Because the exposure raster is based off a 2 km buffer from shoreline polylines, it also acts as a masking layer to avoid including inland habitats.

``` {r prep_exposure_raster}

rast_rgn_30m <- raster(rast_rgn_file)

expos_rast_file  <- file.path(dir_goal_anx, 'other_spatial/ohibc_expos_30m.tif')

if(!file.exists(expos_rast_file)) {
  message('Creating file: \n  ', expos_rast_file)
  
  ohibc_expos_2km_shp <- file.path(dir_goal_anx, 'other_spatial', 'ohibc_expos_2km')
  
  if(!file.exists(ohibc_expos_2km_shp %>% paste0('.shp'))) {
    dir_bcmca <- file.path(dir_anx, '_raw_data/bcmca/d2015/bcmca_eco_set_physical_complete/MarxanData_Physical')
  
    # poly_coastal_class <- readOGR(dir_bcmca, 'BCMCA_ECO_Physical_CoastalClasses_MARXAN') %>%
    #   crop(rast_rgn)
    poly_exposure <- readOGR(dir_bcmca, 'BCMCA_ECO_Physical_Exposure_MARXAN')
    
    poly_exp_2km <- poly_exposure %>%
      gBuffer(width = 2000, byid = TRUE, capStyle = 'FLAT')
    
    ### hunch: third digit is exposure: 
    # 6: Very Exposed
    # 5: Exposed
    # 4: Semi-exposed
    # 3: Semi-protected
    # 2: Protected
    # 1: Very Protected
    
    poly_exp_2km@data <- poly_exp_2km@data %>%
      mutate(exposure = ifelse(ID < 20000, 
                               floor((ID - 2400)/10), ### most instances
                               floor((ID - 24000)/100))) ### instances where last position is two digits
    
    poly_exp_2km_ordered <- poly_exp_2km[order(poly_exp_2km@data$exposure), ]
    
    writeOGR(poly_exp_2km_ordered, 
             dsn = dirname(ohibc_expos_2km_shp),
             layer = basename(ohibc_expos_2km_shp),
             driver = 'ESRI Shapefile',
             overwrite_layer = TRUE) 
  }
  
  ### ordering by exposure then rasterizing with gdal_rasterize means the
  ### cell will receive the highest exposure class present.
  rast_expos <- gdal_rast2(src = ohibc_expos_2km_shp,
                           rast_base = rast_rgn_30m,
                           dst = expos_rast_file,
                           value = 'exposure',
                           override_p4s = TRUE)

} else {
  
  message('File exists: \n  ', expos_rast_file)
  git_prov(expos_rast_file, filetype = 'output')
  
}

```

## Analysis of coastal protection

Coastal protection for salt marsh will be mitigated by exposure; habitat in areas with higher exposure will be valued as providing more protection.

Coastal protection for trees/forest will be mitigated by exposure and elevation (0 m = 1.0, 5 m or higher = 0).  Maximum inland reach is 2 km.  Carbon storage values will include all trees within the entire coastal watersheds regions.

### Seagrass status and trend

Seagrass coastal protection value will currently be left out of the calculation.  Nutrient pressure is insignificant within BC; pressures for log booms, development, and sedimentation are likely more significant, but will require serious rethinking of the seagrass model.

### Salt Marsh status and trend

For salt marshes, we will determine rate of salt marsh change within 2 km inland and offshore.  Salt marsh presence will be established using categories 71, 73, and 74 from the Government of Canada 30 m land use rasters for 1990, 2000, and 2010.  

Salt marsh health status for coastal protection is determined by summing the weighted vulnerability value of salt marshes over the whole Howe Sound area.  Vulnerability value is weighted by exposure (0.25 = very protected, 1.0 = semi exposed, categories from BCMCA).  As saltmarsh is lost, the sum drops; saltmarsh lost in exposed and/or highly populated areas counts as a higher penalty.  The reference point for a score of 1 will be the 1990 salt marsh sum(weighted vulnerability value), adjusted to reflect a 30% drop in salt marsh since historical baselines.

Simple area of salt marsh extent will be calculated as well, for calculation of Carbon Storage and Habitat goals.

``` {r crop_landuse_rasters}

rast_rgn_30m <- raster(rast_rgn_file)

landuse_files <- c('lu1990' = file.path(dir_goal_anx, 'habs/land_cover_1990_30m_crop.tif'),
                   'lu2000' = file.path(dir_goal_anx, 'habs/land_cover_2000_30m_crop.tif'),
                   'lu2010' = file.path(dir_goal_anx, 'habs/land_cover_2010_30m_crop.tif'))

if(any(!file.exists(landuse_files))) {
  landuse_files_raw <- c('lu1990' = file.path(dir_goal_anx, 'habs/land_cover_1990_30m.tif'),
                         'lu2000' = file.path(dir_goal_anx, 'habs/land_cover_2000_30m.tif'),
                         'lu2010' = file.path(dir_goal_anx, 'habs/land_cover_2010_30m.tif'))

  stack_landuse_raw <- stack(landuse_files_raw) %>%
    crop(rast_rgn_30m)
  stack_landuse <- stack_landuse_raw %>%
    mask(rast_rgn_30m)
  writeRaster(stack_landuse, 
              filename = landuse_files,
              bylayer = TRUE, overwrite = TRUE)
}
```
  
``` {r process_saltmarsh_rast}

sm_rast_files <- file.path(dir_goal_anx, sprintf('habs/ohibc_sm_%s_30m.tif', c(1990, 2000, 2010)))

if(any(!file.exists(sm_rast_files))) {

  stack_landuse <- stack(landuse_files)
  
  stack_sm <- filter_rast(stack_landuse, vals = c(71, 73, 74), revalue = 1)
  
  ### write final rasters
  writeRaster(stack_sm, filename = sm_rast_files, bylayer = TRUE, overwrite = TRUE)
} else {
  git_prov(landuse_files, filetype = 'input')
  git_prov(sm_rast_files, filetype = 'output')
}

```

``` {r interpolate_years}


interpolate_hab <- function(df, col) { 
### df <- read_csv(file.path(dir_goal, 'output', 'hab_sm_area.csv'))
  df <- df %>%
    arrange(year)
  
  ### col <- 'sm_area_km2'
  names(df)[names(df) == col] <- 'tmp_val'
  
  year_vec <- df$year %>% unique()
  
  df_expanded <- data.frame()
  for (i in 1:(length(year_vec) - 1)) { ### i <- 1
    
    year_a <- year_vec[i]; year_z <- year_vec[i + 1]
    
    df_tmp <- df %>%
      filter(year %in% c(year_a, year_z)) %>%
      group_by(rgn_id) %>%
      complete(year = year_a:year_z) %>%
      mutate(year_index = (year - first(year)) / (last(year) - first(year)),
             tmp_val    = first(tmp_val) * (1 - year_index) + last(tmp_val) * year_index) %>%
      dplyr::select(-year_index) %>%
      ungroup()
    
    names(df_tmp)[names(df_tmp) == 'tmp_val'] <- col
    
    df_expanded <- df_expanded %>%
      bind_rows(df_tmp) %>%
      arrange(rgn_id, year) %>%
      distinct()
  }
  
  return(df_expanded)
}

```


With the rasters for raw and exposure-mediated saltmarsh presence, use crosstab() to count cells per region, using the coastal watersheds regions.  Because the exposure raster limits the calculations to the 2 km buffer anyway, the calculation will count cells in the set $saltmarsh \cap 2km \cap coastal watershed$.

``` {r count_sm_cells_by_rgn}

sm_area_file <- file.path(dir_goal, 'output/hab_sm_area.csv')
sm_exp_file  <- file.path(dir_goal, 'output/cp_sm_exposure.csv')

if(any(!file.exists(c(sm_area_file, sm_exp_file)))) {
    
  rast_rgn_30m   <- raster(rast_rgn_file) %>%
    setNames('rgn_id')
  
  if(!exists('rast_elev_30m')) {
    rast_elev_30m <- raster(topo_rast_file)
    values(rast_elev_30m)[values(rast_elev_30m) > 5] <- NA
    rast_elev_30m <- rast_elev_30m / rast_elev_30m ### set all elevs below 5 m to 1.0
  }
  
  if(!exists('rast_expos_30m')) {
    rast_expos_30m <- raster(expos_rast_file) ### NOTE: exposure is 1-6
    rast_expos_30m <- (rast_expos_30m - 1) / 5 ### rescale to 0-1 values, where 
      ### 0 is "very protected", and 1 is "very exposed"
  }
  
  ### adjust the raw saltmarsh raster by elevation under 5 m;
  ### then for coastal protection, adjust by exposure
  sm_rast_files  <- file.path(dir_goal_anx, sprintf('habs/ohibc_sm_%s_30m.tif', c(1990, 2000, 2010)))
  stack_sm       <- stack(sm_rast_files) * rast_elev_30m # clip by elev for HAB, CS and CP
  stack_sm_expos <- stack_sm * rast_expos_30m # adj by exposure for CP
  
  saltmarsh_yrs <- c(1990, 2000, 2010)
  
  ### crosstab the raw stack to get raw count of saltmarsh area
  sm_area_df <- data.frame()
  for(i in 1:length(saltmarsh_yrs)) {
    message('Processing ', saltmarsh_yrs[i])
    sm_area_tmp <- crosstab(stack_sm[[i]] %>%
                              setNames('sm_present'), 
                            rast_rgn_30m, 
                            useNA  = TRUE,
                            digits = 1,
                            progress = 'text') %>%
      as.data.frame() %>%
      mutate(year       = saltmarsh_yrs[i],
             rgn_id     = as.integer(as.character(rgn_id)), ### factors! ugh
             sm_present = as.integer(as.character(sm_present))) %>%
      select(year, rgn_id, sm_present, n_cells = Freq)
    sm_area_df <- bind_rows(sm_area_df, sm_area_tmp)
  }
  
  ### * crosstab the exposure-adjusted stack to get exposure-weighted area
  sm_exp_df <- data.frame()
  for(i in 1:length(saltmarsh_yrs)) {
    message('Processing ', saltmarsh_yrs[i])
    sm_exp_tmp <- crosstab(stack_sm_expos[[i]] %>%
                             setNames('sm_expos'), 
                           rast_rgn_30m, 
                           useNA  = TRUE,
                           digits = 1,
                           progress = 'text') %>%
      as.data.frame() %>%
      mutate(year     = saltmarsh_yrs[i],
             rgn_id   = as.integer(as.character(rgn_id)), ### factors! ugh
             sm_expos = as.numeric(as.character(sm_expos))) %>%
      select(year, rgn_id, sm_expos, n_cells = Freq)
    sm_exp_df <- bind_rows(sm_exp_df, sm_exp_tmp)
  }
  
  sm_area_df2 <- sm_area_df %>%
    group_by(rgn_id, year) %>%
    mutate(sm_area_km2 = n_cells * 0.03^2,
           area_km2_tot = sum(sm_area_km2)) %>% ### 30-meter resolution
    filter(!is.na(sm_present) & !is.na(rgn_id)) %>%
    select(-sm_present, -n_cells) %>%
    ungroup()
  sm_exp_df2 <- sm_exp_df %>%
    mutate(area_km2 = n_cells * 0.03^2,
           sm_expos_area = sm_expos * area_km2) %>%
    group_by(rgn_id, year) %>%
    summarize(area_km2_tot = sum(area_km2),
              sm_expos_area_tot = sum(sm_expos_area, na.rm = TRUE)) %>%
    filter(!is.na(rgn_id)) %>%
    ungroup()
  
  # sm_area_df2 <- read_csv(file.path(dir_goal, 'output/hab_sm_area.csv')) %>% filter(year %in% c(1990, 2000, 2010)) %>% distinct()
  # sm_exp_df2  <- read_csv(file.path(dir_goal, 'output/cp_sm_exposure.csv')) %>% filter(year %in% c(1990, 2000, 2010)) %>% distinct()
  
  sm_area_df3 <- interpolate_hab(sm_area_df2, 'sm_area_km2') %>%
    fill(area_km2_tot)
  sm_exp_df3 <- interpolate_hab(sm_exp_df2, 'sm_expos_area_tot') %>%
    fill(area_km2_tot)
  
  write_csv(sm_area_df3, sm_area_file)
  write_csv(sm_exp_df3,  sm_exp_file)
  
} else {
  
  message('Salt marsh area files already exist: \n  ',
          sm_area_file, '\n  ', sm_exp_file)
  git_prov(c(sm_area_file, sm_exp_file), filetype = 'output')
  
}

```

### Coastal forest status and trend

Coastal protection values for coastal forests are calculated similar to salt marsh; vulnerability factors include exposure and elevation.  Anything below 5 m is considered, though not weighted by elevation since our elevation raster is at a rather coarse resolution.  This 5 m value is taken from InVEST, due to the inundation/storm surge from an extreme event.

Simple area of extent for coastal forests will be included as well, for use in calculating Carbon Storage and Habitat goals.

``` {r process_coastal_forest_rast}

cf_rast_files <- file.path(dir_goal_anx, sprintf('habs/ohibc_cf_%s_30m.tif', c(1990, 2000, 2010)))

if(any(!file.exists(cf_rast_files))) {

  stack_landuse <- stack(landuse_files)
  
  stack_cf <- filter_rast(stack_landuse, vals = c(41, 42, 45, 46), revalue = 1)
  
  ### write final rasters
  writeRaster(stack_cf, filename = cf_rast_files, bylayer = TRUE, overwrite = TRUE)
} else {
  git_prov(landuse_files, filetype = 'input')
  git_prov(cf_rast_files, filetype = 'output')
}
```

``` {r count_cf_cells_by_rgn}

cf_area_file <- file.path(dir_goal, 'output/hab_cf_area.csv')
cf_exp_file  <- file.path(dir_goal, 'output/cp_cf_exposure.csv')

if(any(!file.exists(c(cf_area_file, cf_exp_file)))) {

  rast_rgn_30m   <- raster(rast_rgn_file) %>%
    setNames('rgn_id')
  
  if(!exists('rast_elev_30m')) {
    rast_elev_30m <- raster(topo_rast_file)
    values(rast_elev_30m)[values(rast_elev_30m) > 5] <- NA
    rast_elev_30m <- rast_elev_30m / rast_elev_30m ### set all elevs below 5 m to 1.0
  }
  
  if(!exists('rast_expos_30m')) {
    rast_expos_30m <- raster(expos_rast_file) ### NOTE: exposure is 1-6
    rast_expos_30m <- (rast_expos_30m - 1) / 5 ### rescale to 0-1 values, where 
      ### 0 is "very protected", and 1 is "very exposed"
  }
  
  ### adjust coastal forest rasters by elevation and exposure
  cf_rast_files  <- file.path(dir_goal_anx, sprintf('habs/ohibc_cf_%s_30m.tif', c(1990, 2000, 2010)))
  stack_cf       <- stack(cf_rast_files) # not adj by elev for CS - use whole watershed
  stack_cf_expos <- stack_cf * rast_expos_30m * rast_elev_30m # adj by elev and expos for CP
  
  coastalforest_yrs <- c(1990, 2000, 2010)
  
  ### crosstab the raw stack to get raw count of saltmarsh area
  cf_area_df <- data.frame()
  for(i in 1:length(coastalforest_yrs)) {
    message('Processing ', coastalforest_yrs[i])
    cf_area_tmp <- crosstab(stack_cf[[i]] %>%
                              setNames('cf_present'), 
                            rast_rgn_30m, 
                            useNA  = TRUE,
                            digits = 1,
                            progress = 'text') %>%
      as.data.frame() %>%
      mutate(year       = coastalforest_yrs[i],
             rgn_id     = as.integer(as.character(rgn_id)), ### factors! ugh
             cf_present = as.integer(as.character(cf_present))) %>%
      select(year, rgn_id, cf_present, n_cells = Freq)
    cf_area_df <- bind_rows(cf_area_df, cf_area_tmp)
  }
  
  ### * crosstab the exposure-adjusted stack to get exposure-weighted area
  cf_exp_df <- data.frame()
  for(i in 1:length(coastalforest_yrs)) {
    message('Processing ', coastalforest_yrs[i])
    cf_exp_tmp <- crosstab(stack_cf_expos[[i]] %>%
                             setNames('cf_expos'), 
                           rast_rgn_30m, 
                           useNA  = TRUE,
                           digits = 1,
                           progress = 'text') %>%
      as.data.frame() %>%
      mutate(year     = coastalforest_yrs[i],
             rgn_id   = as.integer(as.character(rgn_id)), ### factors! ugh
             cf_expos = as.numeric(as.character(cf_expos))) %>%
      select(year, rgn_id, cf_expos, n_cells = Freq)
    cf_exp_df <- bind_rows(cf_exp_df, cf_exp_tmp)
  }
  
  cf_area_df2 <- cf_area_df %>%
    group_by(rgn_id, year) %>%
    mutate(cf_area_km2 = n_cells * 0.03^2,
           area_km2_tot = sum(cf_area_km2)) %>% ### 30-meter resolution
    filter(!is.na(cf_present) & !is.na(rgn_id)) %>%
    select(-cf_present, -n_cells) %>%
    ungroup()
  cf_exp_df2 <- cf_exp_df %>%
    mutate(area_km2 = n_cells * 0.03^2,
           cf_expos_area = cf_expos * area_km2) %>%
    group_by(rgn_id, year) %>%
    summarize(area_km2_tot = sum(area_km2),
              cf_expos_area_tot = sum(cf_expos_area, na.rm = TRUE)) %>%
    filter(!is.na(rgn_id)) %>%
    ungroup()
  
  # cf_area_df2 <- read_csv(file.path(dir_goal, 'output/hab_cf_area.csv')) %>% filter(year %in% c(1990, 2000, 2010)) %>% distinct()
  # cf_exp_df2  <- read_csv(file.path(dir_goal, 'output/cp_cf_exposure.csv')) %>% filter(year %in% c(1990, 2000, 2010)) %>% distinct()
  
  cf_area_df3 <- interpolate_hab(cf_area_df2, 'cf_area_km2') %>%
    fill(area_km2_tot)
  cf_exp_df3 <- interpolate_hab(cf_exp_df2, 'cf_expos_area_tot') %>%
    fill(area_km2_tot)
  
  write_csv(cf_area_df3, file.path(dir_goal, 'output/hab_cf_area.csv'))
  write_csv(cf_exp_df3,  file.path(dir_goal, 'output/cp_cf_exposure.csv'))

} else {
  
  message('Coastal forest area files already exist: \n  ',
          cf_area_file, '\n  ', cf_exp_file)
  git_prov(c(cf_area_file, cf_exp_file), filetype = 'output')
  
}
```

## results

### saltmarsh

`r DT::datatable(read_csv(sm_area_file, nogit = TRUE))`

`r DT::datatable(read_csv(sm_exp_file, nogit = TRUE))`

### coastal forests

`r DT::datatable(read_csv(cf_area_file, nogit = TRUE))`

`r DT::datatable(read_csv(cf_exp_file, nogit = TRUE))`

***

``` {r prov_footer, results = 'asis'}
prov_wrapup()
```

