---
title: 'OHIBC Howe Sound: Habitat goal score'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(raster)
library(rgdal)
library(rgeos)

source('~/github/ohibc/src/R/common.R')  ### an OHIBC specific version of common.R

dir_git     <- path.expand('~/github/ohibc')
dir_spatial <- file.path(dir_git, 'prep/spatial')  ### github: general buffer region shapefiles
dir_anx     <- file.path(dir_M, 'git-annex/bcprep')

### goal specific folders and info
goal      <- 'hab_cs_cp'
scenario  <- 'vHS'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario)
dir_v2016_anx <- file.path(dir_anx, goal, 'v2016')
dir_goal_anx <- file.path(dir_anx, goal, scenario)

### provenance tracking
library(provRmd); prov_setup()

### support scripts
source(file.path(dir_git, 'src/R/rast_tools.R')) 
  ### gdal_rast2, plotting

### set up the default BC projection to be BC Albers
p4s_bcalb <- c('bcalb' = '+init=epsg:3005')

```

# Summary

Pull data layers to calculate Howe Sound Habitats goal

- Seagrass
    - excluded from analysis: nutrient pressure insignificant; no model yet for log booms/development/sedimentation pressure
- Saltmarsh
    - land use change
- Coastal forests
    - excluded from analysis: does not directly support marine biodiversity
- Soft bottom
    - trawl pressure
- EBSAs? 
    - excluded from analysis: none present in Howe Sound
- Howe Sound-specific sponges
    - protection by closures, reference point 100% protected
    - alternative model based on trawl pressure

# Data Sources

* **Seagrass and salt marsh polygons**
    * **Reference**: CEC. 2016. North America’s Blue Carbon: Assessing Seagrass, Salt Marsh and Mangrove Distribution and Carbon Sinks. Montreal, Canada: Commission for Environmental Cooperation. 54 pp.
    * **Downloaded**: http://www.cec.org/sites/default/files/Atlas/Files/blue_carbon/BlueCarbon_Shapefile.zip (downloaded 8/25/16)
    * **Description**:  shapefiles of marine seagrass, mangrove, and salt marsh habitat in North America
    * **Native data resolution**: NA   
    * **Time range**: NA (data were collected over many years but no time series available)
    * **Format**:  vector shapefiles
* **Land cover raster 1990-2000-2010**
    * **Reference**: Land Use 1990, 2000 and 2010. 2015. Agriculture and Agri-Foods Canada.
    * **Downloaded**: http://open.canada.ca/data/en/dataset/18e3ef1a-497c-40c6-8326-aac1a34a0dec (downloaded 10/7/16)
    * **Description**:  Land cover raster covering all of Canada south of 60° N
    * **Native data resolution**: 30 meters   
    * **Time range**: 1990, 2000, 2010
    * **Format**:  GeoTIFF
* Sponge reef closures
* Sponge reef locations
* Soft bottom habitat
* Shrimp trawl pressures


***

# Methods

Biodiverse habitats are scored based on the relative extent and health.  They are not weighted when combined - each habitat is assumed to have equal importance to the others.  Habitats not present in a region will not be counted against the score.

## Coastal Forests

Since they do not directly support marine biodiversity, coastal forests are not included in this goal.

## Salt marsh

Condition of salt marsh habitat will be scored based on its current extent relative to the extent in 1990.

$$\frac{C_{sm}}{C_{sm_{ref}}} = \frac{A_c}{A_{ref}}$$

where $A_c$ is the set of current extent of habitat, $A_{ref}$ is the reference extent of habitat.  Using extent rasters at equal area projection, the condition can be calculated simply using the ratio of cell counts for current and reference time periods.

## Seagrasses

Since seagrass condition is based on pressure rather than area, the model is different from saltmarsh and forests.

$$\frac{C_{sg}}{C_{sg_{ref}}} = \displaystyle\sum_{j=1}^{n_c}((1 - p_j) * A_j) / A_{total}$$

where $p$ represents the nutrient pressure (compared to a reference of zero human-introduced nutrients).  Using an equal area projection raster and a fixed habitat extent, we can calculate this more simply by $1 - \bar{p}$.  Seagrasses are currently excluded from analysis as model does not currently account for large concerns including log booms.

## Soft-Bottom Sediment

Soft-bottom sediment condition is based on trawling pressure from the shrimp trawl fishery.

$$\frac{C_{sb}}{C_{sb_{ref}}} = \displaystyle\sum_{j=1}^{n_c}((1 - p_j) * A_j) / A_{total}$$

where $p$ represents the trawling pressure, where trawl effort $E_{trawl}$ is based upon the annual mean of trawl hours from 2005-2015:

$$p_{trawl} = E_{trawl}/E_{trawl_{max}}$$

## Sponge reef

Sponge reef conditions are calculated using a simple ratio of protected area to total area.  Protections are afforded by sponge reef closures but these only cover certain reefs.

$$\frac{C_{sp}}{C_{sp_{ref}}} = A_{closure} / A_{total}$$

## Goal model

The habitat goal model is the unweighted mean of all biodiversity-supporting marine habitats.  

$$x_{HAB} = \displaystyle\sum_{i=1}^{n}\left(\frac{C_i}{C_{i_{ref}}} * \frac{A_i}{A_T}\right)$$

``` {r load_hab_data}

hab_cf <- read_csv(file.path(dir_goal, 'int/howe_forest_status_trend.csv')) %>%
  select(status_yr = year, area_km2 = area_nonzero_cf_km2) %>%
  arrange(status_yr) %>%
  mutate(hab = 'cf',
         c_cref = area_km2 / first(area_km2),
         trend = (c_cref - lag(c_cref))/((status_yr - lag(status_yr)) * lag(c_cref)))

  ### trend is in absolute change per year

hab_sm <- read_csv(file.path(dir_goal, 'int/howe_saltmarsh_status_trend.csv')) %>%
  select(status_yr = year, area_km2 = area_sm_km2) %>%
  arrange(status_yr) %>%
  mutate(hab = 'sm',
         c_cref = area_km2 / first(area_km2),
         trend = (c_cref - lag(c_cref))/((status_yr - lag(status_yr)) * lag(c_cref)))
  ### trend is in proportional change per year

# hab_sg <- read_csv(file.path(dir_goal, 'int/howe_seagrass_status_trend.csv')) %>%
#   select(status_yr = year, area_km2 = sg_area_km2, pres_mean_sg) %>%
#   arrange(status_yr) %>%
#   mutate(hab = 'sg',
#          c_cref = 1 - pres_mean_sg) %>%
#   select(-pres_mean_sg)
# hab_sg$trend <- lm(formula = c_cref ~ status_yr, 
#                    data = hab_sg %>% 
#                      filter(status_yr %in% c(2003:2013)))[['coefficients']]['status_yr']
  ### trend is in absolute change per year

hab_sp <- read_csv(file.path(dir_goal, 'int/sponge_closure_summary.csv')) %>%
  mutate(hab = 'sp',
         c_cref = area_prot_km2 / area_tot_km2,
         status_yr = 2015) %>%
  select(hab, area_km2 = area_tot_km2, c_cref, status_yr)

hab_sb <- read_csv(file.path(dir_goal, 'int/soft_btm_trawl_summary.csv')) %>%
  mutate(hab = 'sb',
         c_cref = 1 - weighted_prs,
         status_yr = 2015) %>%
  select(hab, area_km2 = area_tot_km2, c_cref, status_yr)

hab_scores <- hab_cf %>%
  bind_rows(hab_sm) %>%
  # bind_rows(hab_sg) %>%
  bind_rows(hab_sb) %>%
  bind_rows(hab_sp)
  
hab_scores_summary <- hab_scores %>%
  group_by(hab) %>%
  filter(hab != 'cf') %>%
  filter(status_yr == max(status_yr)) %>%
  ungroup() %>% 
  mutate(hab_st_tot = sum(c_cref) / n(),
         hab_tr_tot = sum(trend, na.rm = TRUE) / sum(!is.na(trend)))
  
write_csv(hab_scores_summary, file.path(dir_goal, 'output', 'hab_scores_summary.csv'))

knitr::kable(hab_scores_summary, caption = 'Habitats')

hab_scores_layer <- hab_scores_summary %>%
  select(hab_status = hab_st_tot,
         hab_trend  = hab_tr_tot) %>%
  distinct()

write_csv(hab_scores_layer, file.path(dir_goal, 'output', 'hab_scores.csv'))

```     

***

``` {r prov_footer, results = 'asis'}
prov_wrapup()
```

