---
title: 'OHIBC: Regression causal chain model with fixed effects'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes:
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library('plm') ### before Tidyverse - 'lead' and 'lag' function conflict
source('https://raw.githubusercontent.com/oharac/src/master/R/common.R')

dir_ohibc  <- '~/github/ohibc'
dir_calc   <- file.path(dir_ohibc, 'calc_ohibc')
dir_master <- file.path(dir_calc, 'master')

source(file.path(dir_calc, 'calc_scores_fxns.R'))

### provenance tracking
# library(provRmd); prov_setup()

```


# Regression model

## Change in pressure vs. resilience

The first link in the chain connects a management action now (time $t$) to changes in pressures in the future (time $t + \lambda$).  Framing this as proportional change in pressure, $\frac{p_{t+\lambda} - p_t}{p_{t}}$, examine the coefficient and significance of regulatory resilience.

$$\%\Delta p_t = \frac{p_{t+\lambda} - p_t}{p_{t}} = \alpha + \beta r_{reg,t}$$
$$\%\Delta p_t = \frac{p_{t+\lambda} - p_t}{p_{t}} = \alpha + \beta_1 r_{reg,t} + \beta_2 r_{soc,t} + \beta_3 r_{ecol,t}$$

We will include fixed effects on region and year.  Time effects were shown to be significant for all goals except for Clean Waters; for simplicity we will include time effects on all goals.  Fixed effects and random effects in various combinations were also shown to be significant for most goals; again, for simplicity (though not parsimony) we will include fixed effects for all goals.  Note: Econometricians Doug Steigerwald and Kyle Meng both recommended fixed effects in general over random effects.  See `regr_causal_chain_re_fe_te.Rmd` for details of testing time effects and random/fixed effects.

To enable testing of all these combinations of goal/lag year/model variant, we will set up a data structure of lists within lists.

* List level 1: goal/subgoal (x12)
    * List level 2: lag year (x6)
        * list level 3: nested model variant (x8) (7 plus intercept only?)
        
``` {r regress_prs_coefficients_res_components}

### Set up basic data frame - make sure prs and res are 0-1; status
### range not important here since we'll take a ratio inside a loop.

res_reg <- read_csv(file.path(dir_calc, 'res_reg_only.csv')) %>%
  select(goal, region_id, year, res_reg = score)
res_soc <- read_csv(file.path(dir_calc, 'res_soc_only.csv')) %>%
  select(goal, region_id, year, res_soc = score)
res_ecol <- read_csv(file.path(dir_calc, 'res_ecol_only.csv')) %>%
  select(goal, region_id, year, res_ecol = score)
prs_no_social <- read_csv(file.path(dir_calc, 'prs_no_social.csv')) %>%
  select(goal, region_id, year, pressures = score)

scores_rgn_res <- read_csv(file.path(dir_calc, 'scores_all.csv')) %>%
# scores_rgn_res <- read_csv(file.path(dir_calc, 'scores_all.csv')) %>%
  filter(region_id != 0) %>%
  spread(dimension, score) %>%
  select(goal, region_id, year, status, trend) %>%
  left_join(res_reg, by = c('goal', 'region_id', 'year')) %>%
  left_join(res_soc, by = c('goal', 'region_id', 'year')) %>%
  left_join(res_ecol, by = c('goal', 'region_id', 'year')) %>%
  left_join(prs_no_social, by = c('goal', 'region_id', 'year')) %>%
  group_by(goal, region_id) %>%
  arrange(year) %>%
  mutate(res_reg   = res_reg / 100,
         res_soc   = res_soc / 100,
         res_ecol  = res_ecol / 100,
         pressures = pressures / 100) %>%
  filter(!all(is.na(res_reg), is.na(res_soc), is.na(res_ecol))) %>%
  filter(!is.na(pressures)) %>%
  left_join(get_rgn_names(), by = c('region_id' = 'rgn_id'))

goals <- scores_rgn_res$goal %>% 
  unique() 

```

```{r helper functions}

calc_d_prs <- function(goalname, scores_df, yr_lag) {
  d_prs_df <- scores_df %>%
    filter(goal == goalname) %>%
    mutate(obs_future_prs = lead(pressures, yr_lag),
           d_prs = (obs_future_prs - pressures) / pressures,
           intercept = 1) %>%
    filter(!is.na(d_prs))
  return(d_prs_df)
}

calc_d_status <- function(goalname, scores_df, yr_lag) {
  d_stat_df <- scores_df %>%
      filter(goal == goalname) %>%
      mutate(obs_future_status = lead(status, yr_lag),
             d_status = (obs_future_status - status) /status) %>%
      filter(!is.na(d_status)) %>%
      filter(!is.infinite(d_status))
  return(d_stat_df)
}

tidy_model <- function(mdl) {
  mdl_df <- mdl %>%
    broom::tidy() %>%
    mutate(adj_r_squared = summary(mdl)$adj.r.squared,
           aic    = AIC(mdl), 
           yrs    = sum(str_detect(term, 'year')),
           rgns   = sum(str_detect(term, 'rgn')),
           params = rgns + yrs + n() - 1)
  return(mdl_df)
}

```


``` {r resilience regr function}

calc_res_regr <- function(df) {
  ### For a given dataframe filtered to goal and lag year, loop over all
  ### combinations of resilience regression models (except intercept only).
  ### For each base model, calculate fixed and random effects on region, and
  ### for each of those, calculate with and without time effect.
  
  res_mdls <- list(reg  = c('res_reg'),
                   ecol = c('res_ecol'),
                   soc  = c('res_soc'),
                   reg_ecol = c('res_reg', 'res_ecol'),
                   reg_soc  = c('res_reg', 'res_soc'),
                   ecol_soc = c('res_ecol', 'res_soc'),
                   reg_ecol_soc = c('res_reg', 'res_ecol', 'res_soc'))

  ### initialize list for this combo of goal and lag
  res_mdl_list <- vector('list', length = length(res_mdls) + 1)

  ### This loop calculates all the permutations of the resilience component
  ### model.  
  for(i in seq_along(res_mdls)) {
    ### i <- 1
    
    ### create text versions of the regression formulas for the base regression
    ### (fixed or random effect determined in plm call) and regression with time
    ### effects.
    mdl_text <- sprintf('d_prs ~ %s + factor(year) + factor(rgn_code)', paste(res_mdls[[i]], collapse = ' + '))

    message('  ', i, ':', mdl_text)
    names(res_mdl_list)[i] <- mdl_text

    ### Check that there are values for each component, otherwise devolves to different model:
    if(any(colSums(is.na(df[ , res_mdls[[i]]])) == nrow(df))) {
      message('Model', mdl_text, 'has NA columns - setting to "NA model"!')
      res_mdl_list[[i]] <- 'NA model'
      next()
    }
  
    ### Check that model columns have some variation over time to avoid
    ### degenerate models.  Compare range of each column to sqrt of machine precision
    ### (could also just compare to a small regular number...)
    if(any(sapply(res_mdls[[i]], 
                  FUN = function(x) {
                    diff(range(df[[x]], na.rm = TRUE)) < .Machine$double.eps^.5
                    }
                  )) & !str_detect(mdl_text, 'intercept')) {
      message('Model', mdl_text, 'has degenerate columns - setting to "degenerate model"!')
      res_mdl_list[[i]] <- 'degenerate model'
      next()
    }
    
    mdl_fe <- lm(as.formula(mdl_text), data = df)
    # mdl_fe <- plm(as.formula(mdl_text), data = df, 
    #                     index = c('rgn_code', 'year'),
    #                     model = 'within')

    res_mdl_list[[i]] <- mdl_fe
  }
  
  ### add intercept only model; index is i+1:
  message('  ', i + 1, ': intercept only')
  mdl_fe <- lm(d_prs ~ 1 + factor(year) + factor(rgn_code), data = df)
  res_mdl_list[[i+1]] <- mdl_fe
  names(res_mdl_list)[i + 1] <- 'intercept only'

  return(res_mdl_list)
}

```

``` {r loop over resilience regressions}

goal_regr_list <- vector('list', length = length(goals))
names(goal_regr_list) <- goals

for(goalname in goals) {
  # goalname <- 'FIS'
  
  yr_lag_regr_list <- vector('list', length = 6)
  names(yr_lag_regr_list) <- paste0('lag', 1:6)
  
  for(yr_lag in 1:6) {
    # yr_lag <- 5
    d_prs_df <- calc_d_prs(goalname, scores_rgn_res, yr_lag)
    
    if(nrow(d_prs_df) == 0) next() ### skip degenerate case
    
    message('Calculating', goalname, 'at lag', yr_lag)
    tmp <- calc_res_regr(d_prs_df)
    
    yr_lag_regr_list[[yr_lag]] <- tmp
  }
  
  goal_regr_list[[goalname]] <- yr_lag_regr_list
}

```

# Comparing models

## Notes about AIC and BIC:

From https://methodology.psu.edu/AIC-vs-BIC:

> So what’s the bottom line? In general, it might be best to use AIC and BIC together in model selection. For example, in selecting the number of latent classes in a model, if BIC points to a three-class model and AIC points to a five-class model, it makes sense to select from models with 3, 4 and 5 latent classes. AIC is better in situations when a false negative finding would be considered more misleading than a false positive, and BIC is better in situations where a false positive is as misleading as, or more misleading than, a false negative.

from https://robjhyndman.com/hyndsight/aic/:

> The AIC does not require nested models. One of the neat things about the AIC is that you can compare very different models. However, make sure the likelihoods are computed on the same data. For example, you cannot compare an ARIMA model with differencing to an ARIMA model without differencing, because you lose one or more observations via differencing. That is why auto.arima uses a unit root test to choose the order of differencing, and only uses the AIC to select the orders of the AR and MA components.

Seems like AIC (and BIC?) should be fine with a standard linear model using OLS?  We're not assuming anything fancy about the data.  BUT: the lag relationship will mean AIC comparing lag 4 vs. lag 5 (e.g.) are looking at different data.  So: use AIC to choose the best model for each goal for a specific lag, but then look at the lags in some other manner.  If they're all saying the same model form, then use a different criterion to choose the best lag model.

## Compare all pressure ~ resilience models

From the `goal_regr_list` list of lists, sort through to answer the following:

1) For each goal/lag year (using fixed effects on region and time), determine best model by comparing AIC.
2) For each goal, compare lag year models using R^2^ as a heuristic.

For each loop, assemble a dataframe with results by model; bind together by lag year and goal.  Maybe not memory efficient but whatevs.

```{r}

### initialize dataframe over all goals, lag years, and models
mdls_df <- data.frame()

for(goalname in goals) {
  # goalname <- goals[1]
  
  ### initialize dataframe for this goal over all lag years and models
  mdls_goal_df <- data.frame()
  
  for(yr_lag in 1:6) {
    # yr_lag <- 1
    lag_txt <- paste0('lag', yr_lag)
    
    ### select models list for this goal and lag year combo
    mdls_to_test <- goal_regr_list[[goalname]][[lag_txt]]
    
    ### initialize dataframe for this goal and lag year over all models
    mdls_goal_lag_df <- data.frame()
    for(test_mdl in names(mdls_to_test)) {
      # test_mdl <- names(mdls_to_test)[1]
    
      mdl <- mdls_to_test[[test_mdl]]
      
      if(class(mdl) == 'character') {
        ### test if NA model or degenerate model
        tmp_df <- data.frame(goal  = goalname, 
                             model = test_mdl,
                             lag   = yr_lag, 
                             model_error = mdl)
      } else {

        tmp_df <- tidy_model(mdl) %>%
          mutate(goal   = goalname,
                 model  = test_mdl,
                 lag    = yr_lag)
      }
      
      mdls_goal_lag_df <- mdls_goal_lag_df %>%
        bind_rows(tmp_df)
    }
    
    mdls_goal_df <- mdls_goal_df %>%
      bind_rows(mdls_goal_lag_df)
  }
  
  mdls_df <- mdls_df %>%
    bind_rows(mdls_goal_df)
}

write_csv(mdls_df, 'int/prs_res_results_fixed_effects.csv')
```

```{r}

prs_res_mdl <- read_csv('int/prs_res_results_fixed_effects.csv') %>%
  mutate(est_text = formatC(round(estimate, 4), digits = 4, format = 'f'),
         est_text = case_when(p.value < .001 ~ paste0(est_text, '***'),
                              p.value <  .01 ~ paste0(est_text, '**'),
                              p.value <  .05 ~ paste0(est_text, '*'),
                              p.value <  .10 ~ paste0(est_text, '°'),
                              TRUE ~ est_text)) %>%
  select(goal, model, lag, yrs, rgns, adj_r_squared, aic, term, est_text, model_error) %>%
  spread(term, est_text) %>%
  clean_df_names() %>%
  select(-na) ### holdover from spreading NA rows

write_csv(prs_res_mdl, 'int/prs_res_results_fixed_effects_formatted_p.csv')

DT::datatable(prs_res_mdl %>% mutate_if(is.numeric, ~ round(., 4)))

```

## Change in status vs. pressure

The second link in the causal chain connects a change in pressure now (time $t$) to changes in status in the future (time $t + \lambda$).  Framing this as proportional change in status, $\frac{x_{t+\lambda}}{x_{t}}$, examine the coefficient and significance of pressure.

$$\%\Delta X_t = \frac{x_{t+\lambda} - x_t}{x_{t}} = \alpha + \beta p_{t}$$

``` {r setup d status vs pressures}

### Set up basic data frame - make sure prs and res are 0-1; status
### range not important here since we'll take a ratio inside a loop.

scores_rgn_d_status <- read_csv(file.path(dir_calc, 'scores_all.csv')) %>%
# scores_rgn <- read_csv(file.path(dir_calc, 'scores_all.csv')) %>%
  filter(region_id != 0) %>%
  spread(dimension, score) %>%
  select(goal, region_id, year, status, pressures) %>%
  left_join(prs_no_social %>% rename(prs_omit_soc = pressures), 
            by = c('goal', 'region_id', 'year')) %>%
  group_by(goal, region_id) %>%
  arrange(year) %>%
  filter(!is.na(pressures)) %>%
  left_join(get_rgn_names(), by = c('region_id' = 'rgn_id'))

goals <- scores_rgn_d_status$goal %>% 
  unique()

```

``` {r regress d status vs pressures all}
d_status_prs <- data.frame()

for(yr_lag in 1:6) {
  # yr_lag <- 1
  for(goalname in goals) { # goalname <- 'FIS'

    d_status_df <- calc_d_status(goalname, scores_rgn_d_status, yr_lag)
    
    if(nrow(d_status_df) == 0) next()
    if(length(unique(d_status_df$year)) <= 1) next()
    
    d_status_mdl <- lm(d_status ~ pressures + factor(rgn_code) + factor(year), data = d_status_df)
    d_status_prs_lm <- tidy_model(d_status_mdl) %>%
      mutate(goal   = goalname,
             model  = test_mdl,
             lag    = yr_lag)

    d_status_prs <- bind_rows(d_status_prs, d_status_prs_lm)
  }
}

write_csv(d_status_prs, 'int/status_prs_results_incl_soc.csv')

```

```{r}

st_prs_mdl <- read_csv('int/status_prs_results_incl_soc.csv') %>%
  mutate(est_text = formatC(round(estimate, 4), digits = 4, format = 'f'),
         est_text = case_when(p.value < .001 ~ paste0(est_text, '***'),
                              p.value <  .01 ~ paste0(est_text, '**'),
                              p.value <  .05 ~ paste0(est_text, '*'),
                              p.value <  .10 ~ paste0(est_text, '°'),
                              TRUE ~ est_text)) %>%
  select(goal, model, lag, yrs, rgns, adj_r_squared, aic, term, est_text) %>%
  spread(term, est_text) %>%
  clean_df_names()

write_csv(st_prs_mdl, 'int/status_prs_results_incl_soc_formatted_p.csv')

DT::datatable(st_prs_mdl %>% mutate_if(is.numeric, ~ round(., 4)))

```

``` {r regress d status vs pressures no social}
d_status_prs <- data.frame()

for(yr_lag in 1:6) {
  # yr_lag <- 2
  for(goalname in goals) { # goalname <- 'LSP'

    d_status_df <- calc_d_status(goalname, scores_rgn_d_status, yr_lag)
    
    if(nrow(d_status_df) == 0) next()
    if(length(unique(d_status_df$year)) <= 1) next()
    if(all(is.na(d_status_df$prs_omit_soc))) next()
    
    d_status_mdl <- lm(d_status ~ prs_omit_soc + factor(rgn_code) + factor(year), data = d_status_df)
    d_status_prs_lm <- tidy_model(d_status_mdl) %>%
      mutate(goal  = goalname,
             model = test_mdl,
             lag   = yr_lag)

    d_status_prs <- bind_rows(d_status_prs, d_status_prs_lm)
  }
}

write_csv(d_status_prs, 'int/status_prs_results_no_soc.csv')

```

```{r}

st_prs_mdl <- read_csv('int/status_prs_results_no_soc.csv') %>%
  mutate(est_text = formatC(round(estimate, 4), digits = 4, format = 'f'),
         est_text = case_when(p.value < .001 ~ paste0(est_text, '***'),
                              p.value <  .01 ~ paste0(est_text, '**'),
                              p.value <  .05 ~ paste0(est_text, '*'),
                              p.value <  .10 ~ paste0(est_text, '°'),
                              TRUE ~ est_text)) %>%
  select(goal, model, lag, yrs, rgns, adj_r_squared, aic, term, est_text) %>%
  spread(term, est_text) %>%
  clean_df_names()

write_csv(st_prs_mdl, 'int/status_prs_results_no_soc_formatted_p.csv')

DT::datatable(st_prs_mdl %>% mutate_if(is.numeric, ~ round(., 4)))

```

#### Change in status vs. pressure

`r DT::datatable(d_status_prs %>% mutate_if(is.double, ~ round(., 4)))`


