---
title: 'OHIBC: Regression causal chain model with fixed effects: status vs pressure'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes:
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library('plm') ### before Tidyverse - 'lead' and 'lag' function conflict
source('https://raw.githubusercontent.com/oharac/src/master/R/common.R')
library(caret) ### for cross validation

dir_ohibc  <- '~/github/ohibc'
dir_calc   <- file.path(dir_ohibc, 'calc_ohibc')
dir_master <- file.path(dir_calc, 'master')

source(file.path(dir_calc, 'calc_scores_fxns.R'))

### provenance tracking
# library(provRmd); prov_setup()

```


## Change in status vs. pressure

The second link in the causal chain connects a change in pressure now (time $t$) to changes in status in the future (time $t + \lambda$).  Framing this as proportional change in status, $\frac{x_{t+\lambda}}{x_{t}}$, examine the coefficient and significance of pressure.

$$\%\Delta X_t = \frac{x_{t+\lambda} - x_t}{x_{t}} = \alpha + \beta p_{t}$$

```{r helper functions}

calc_d_status <- function(goalname, scores_df, yr_lag) {
  d_stat_df <- scores_df %>%
      filter(goal == goalname) %>%
      mutate(obs_future_status = lead(status, yr_lag),
             d_status = (obs_future_status - status) /status) %>%
      filter(!is.na(d_status)) %>%
      filter(!is.infinite(d_status))
  return(d_stat_df)
}

tidy_model <- function(mdl) {
  mdl_df <- mdl %>%
    broom::tidy() %>%
    mutate(adj_r_squared = summary(mdl)$adj.r.squared,
           aic    = AIC(mdl), 
           yrs    = sum(str_detect(term, 'year')),
           rgns   = sum(str_detect(term, 'rgn')),
           params = rgns + yrs + n() - 1)
  return(mdl_df)
}

```

## Regress changes in status vs. all pressures

### set up dataframe

``` {r setup d status vs pressures}

### Set up basic data frame - make sure prs and res are 0-1; status
### range not important here since we'll take a ratio inside a loop.

prs_no_social <- read_csv(file.path(dir_calc, 'prs_no_social.csv')) %>%
  select(goal, region_id, year, pressures = score)

scores_rgn_d_status <- read_csv(file.path(dir_calc, 'scores_all.csv')) %>%
# scores_rgn <- read_csv(file.path(dir_calc, 'scores_all.csv')) %>%
  filter(region_id != 0) %>%
  spread(dimension, score) %>%
  select(goal, region_id, year, status, pressures) %>%
  left_join(prs_no_social %>% rename(prs_omit_soc = pressures), 
            by = c('goal', 'region_id', 'year')) %>%
  group_by(goal, region_id) %>%
  arrange(year) %>%
  filter(!is.na(pressures)) %>%
  left_join(get_rgn_names(), by = c('region_id' = 'rgn_id'))

goals <- scores_rgn_d_status$goal %>% 
  unique()

```

### Run regression for all goals, all lags

Include a cross validation step.

``` {r regress d status vs pressures all}

### Define training control
set.seed(123)
k <- 5
reps <- 5
train.control <- trainControl(method = "repeatedcv", number = k, repeats = reps)

### initialize dataframes
d_status_prs <- data.frame()
crossval_df  <- data.frame()

for(yr_lag in 1:6) {
  # yr_lag <- 1
  for(goalname in goals) { # goalname <- 'FIS'

    d_status_df <- calc_d_status(goalname, scores_rgn_d_status, yr_lag)
    
    if(nrow(d_status_df) == 0) next()
    if(length(unique(d_status_df$year)) <= 1) next()
    
    test_df <- d_status_df %>%
      group_by(rgn_code) %>%
      summarize(range_vals = max(status) - min(status))
    if(all(test_df$range_vals == 0)) {
      d_status_prs_lm <- data.frame(goal = goalname, lag = yr_lag, model_error = 'no change in status')
      crossval_results <- d_status_prs_lm
    } else {
    
      d_status_mdl <- lm(d_status ~ pressures + factor(rgn_code) + factor(year), data = d_status_df)
      d_status_prs_lm <- tidy_model(d_status_mdl) %>%
        mutate(goal   = goalname,
               lag    = yr_lag)
      
      cross_val_mdl <- train(d_status ~ pressures + factor(year) + factor(rgn_code), data = d_status_df, 
                             method = "lm",
                             trControl = train.control)
      crossval_results <- cross_val_mdl$results %>%
        mutate(goal   = goalname,
               lag    = yr_lag) %>%
        clean_df_names()
    }

    d_status_prs <- bind_rows(d_status_prs, d_status_prs_lm)
    crossval_df <- bind_rows(crossval_df, crossval_results)
  }
}

write_csv(d_status_prs, 'int/status_prs_results_incl_soc.csv')
write_csv(crossval_df, 'int/status_prs_crossval_incl_soc.csv')

```

### Clean up results

```{r}

crossval_df <- read_csv('int/status_prs_crossval_incl_soc.csv') %>%
  select(-intercept, -model_error)

st_prs_mdl <- read_csv('int/status_prs_results_incl_soc.csv') %>%
  mutate(est_text = formatC(round(estimate, 4), digits = 4, format = 'f'),
         est_text = case_when(p.value < .001 ~ paste0(est_text, '***'),
                              p.value <  .01 ~ paste0(est_text, '**'),
                              p.value <  .05 ~ paste0(est_text, '*'),
                              p.value <  .10 ~ paste0(est_text, '°'),
                              TRUE ~ est_text)) %>%
  select(goal, lag, yrs, rgns, adj_r_squared, aic, term, est_text) %>%
  spread(term, est_text) %>%
  left_join(crossval_df, by = c('goal', 'lag')) %>%
  clean_df_names()

write_csv(st_prs_mdl, 'tables/status_prs_results_incl_soc_formatted_p.csv')

DT::datatable(st_prs_mdl %>% mutate_if(is.numeric, ~ round(., 4)))

```

## Repeat the above for data with social pressures excluded

### run regression across all goals and lags

including a cross validation

``` {r regress d status vs pressures no social}
### initialize dataframes
d_status_prs <- data.frame()
crossval_df  <- data.frame()

for(yr_lag in 1:6) {
  # yr_lag <- 2
  for(goalname in goals) { # goalname <- 'LSP'

    d_status_df <- calc_d_status(goalname, scores_rgn_d_status, yr_lag)
    
    if(nrow(d_status_df) == 0) next()
    if(length(unique(d_status_df$year)) <= 1) next()
    if(all(is.na(d_status_df$prs_omit_soc))) next()
    test_df <- d_status_df %>%
      group_by(rgn_code) %>%
      summarize(range_vals = max(status) - min(status))
    
    if(all(test_df$range_vals == 0)) {
      
      d_status_prs_lm <- data.frame(goal = goalname, lag = yr_lag, model_error = 'no change in status')
      crossval_results <- d_status_prs_lm
      
    } else {
      
      d_status_mdl <- lm(d_status ~ prs_omit_soc + factor(rgn_code) + factor(year), data = d_status_df)
      d_status_prs_lm <- tidy_model(d_status_mdl) %>%
        mutate(goal  = goalname,
               lag   = yr_lag)
  
      cross_val_mdl <- train(d_status ~ prs_omit_soc + factor(year) + factor(rgn_code), data = d_status_df, 
                             method = "lm",
                             trControl = train.control)
      crossval_results <- cross_val_mdl$results %>%
        mutate(goal   = goalname,
               lag    = yr_lag) %>%
        clean_df_names()    
      
    }

    d_status_prs <- bind_rows(d_status_prs, d_status_prs_lm)
    crossval_df <- bind_rows(crossval_df, crossval_results)
  }
}

write_csv(d_status_prs, 'int/status_prs_results_no_soc.csv')
write_csv(crossval_df, 'int/status_prs_crossval_no_soc.csv')

```

### clean up results

```{r}

crossval_df <- read_csv('int/status_prs_crossval_no_soc.csv') %>%
  select(-intercept, -model_error)

st_prs_mdl <- read_csv('int/status_prs_results_no_soc.csv') %>%
  mutate(est_text = formatC(round(estimate, 4), digits = 4, format = 'f'),
         est_text = case_when(p.value < .001 ~ paste0(est_text, '***'),
                              p.value <  .01 ~ paste0(est_text, '**'),
                              p.value <  .05 ~ paste0(est_text, '*'),
                              p.value <  .10 ~ paste0(est_text, '°'),
                              TRUE ~ est_text)) %>%
  select(goal, lag, yrs, rgns, adj_r_squared, aic, term, est_text) %>%
  spread(term, est_text) %>%
  left_join(crossval_df, by = c('goal', 'lag')) %>%
  rename(pressures = prs_omit_soc) %>%
  clean_df_names()

write_csv(st_prs_mdl, 'tables/status_prs_results_no_soc_formatted_p.csv')

DT::datatable(st_prs_mdl %>% mutate_if(is.numeric, ~ round(., 4)))

```


