---
title: "OHIBC Howe Sound: Lasting Special Places"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(sp)        # the classes and methods that make up spatial ops in R
library(maptools)  # tools for reading and manipulating spatial objects
library(rgeos)
library(rgdal)
library(raster)


dir_git  <- '~/github/ohibc'         ### set wd to work in Github OHIBC location
source(file.path(dir_git, 'src/R/common.R'))  ### an OHIBC specific version of common.R

scenario <- 'vHS'

dir_spatial <- file.path(dir_git, 'prep/spatial') ### github: general buffer region shapefiles
dir_goal    <- file.path(dir_git, 'prep/lsp', scenario)
dir_rast    <- file.path(dir_goal, 'raster')            ### goal-specific raster files are small
source(file.path(dir_goal, 'lsp_fxns.R'))
source(file.path(dir_git, 'src/R/rast_tools.R'))

dir_anx <- file.path(dir_M, 'git-annex/bcprep') 
dir_goal_anx <- file.path(dir_anx, 'lsp', 'v2016')  
  ### git-annex: goal-specific large files; note in v2016 scenario instead of vHS

### provenance tracking
library(provRmd); prov_setup()

### set up the default BC projection to be BC Albers
p4s_bcalb <- c('bcalb' = '+init=epsg:3005')

```

# Summary: OHIBC Lasting Special Places subgoal (Sense of Place)

The Lasting Special Places goal model for Howe Sound is similar to the OHI Global model: a region's status is based upon percent of protected area within 1 km inland buffer and percent of protected area within 3 nautical mile offshore buffer, compared to a reference point of 30% protected area.

$$X_{LSP} = \frac{\frac{pA_{CMPA}}{pA_{refCMPA}} + \frac{pA_{CP}}{pA_{refCP}}}{2}$$

*pA* = percent of area within the inland or offshore buffer; *CMPA* = coastal marine protected area (3nm offshore); *CP* = coastline protected (1km inland); and *refCMPA* = *refCP* = 30% reference point for both measures.

For Howe Sound, based upon conversations with Andrew Day and Karin Bodtker, the weighting between onshore and offshore protected area will be area-weighted rather than a simple 50/50 weighting as in the OHI Global.

An alternative will also examine protected areas within the entire watershed that feeds Howe Sound, rather than simply a 1 km inland buffer.

-----

# Data Sources

WDPA database <citation info?>

BC Parks, Ecological Reserves, and Protected Areas data 

-----

# Methods

## Read in BC WDPA-MPA shapefile

We will create a BC-specific polygon subset of the WDPA global dataset, then rasterize to BC Albers at 500 m resolution.

NOTE: If BC WDPA file does not yet exist, `get_wdpa_poly()` creates it from the original WDPA-MPA file.  This takes a long time, due to reading in the full WDPA-MPA geodatabase into a SpatialPolygonsDataFrame.

``` {r set up the extents for Howe Sound, based on Howe Sound region outline and watershed}
# hs_ws <- readOGR(dsn = file.path(dir_goal, 'watershed'), layer = 'howe_sound_watersheds')
# hs_rgn <- readOGR(dsn = file.path(dir_spatial), layer = 'howe_sound_rgn')
# bbox(hs_ws)
# #         min       max
# # x 1141118.3 1238724.7
# # y  471385.5  616826.2
# bbox(hs_rgn)
# #         min       max
# # x 1169571.2 1227753.5
# # y  473548.3  534590.4
ext_howe <- extent(c(1140000, 1240000, 470000, 620000))
```

``` {r setup_BC_WDPA_polygons}
poly_hs_wdpa <- get_wdpa_poly(p4s_bcalb, reload = FALSE) %>%  ### defaults to BC Albers
  spTransform(CRS(p4s_bcalb)) %>%
  crop(ext_howe)

writeOGR(poly_hs_wdpa,
         dsn = path.expand(file.path(dir_goal, 'spatial')),
         layer = 'hs_wdpa_poly',
         driver = 'ESRI Shapefile',
         overwrite_layer = TRUE)

```

## Read in BC Parks, Ecological Reserves, and Protected Areas data

DataBC offers spatial data on BC protected areas with small differences from the WDPA dataset.  We will use this as well.

``` {r setup_BC_PEP_polygons}
poly_hs_pep <- readOGR(dsn = file.path(dir_anx, '_raw_data/databc/TA_PEP_SVW'),
                       layer = 'TA_PEP_SVW_polygon',
                       stringsAsFactors = FALSE) %>%
  spTransform(CRS(p4s_bcalb)) %>%
  crop(ext_howe)

writeOGR(poly_hs_pep,
         dsn = path.expand(file.path(dir_goal, 'spatial')),
         layer = 'hs_pep_poly',
         driver = 'ESRI Shapefile',
         overwrite_layer = TRUE)

```

-----

## Rasterize the BC WDPA-MPA shapefile to Howe Sound extents

``` {r rasterize_HS_WDPA_and_PEP}

rast_base <- raster(file.path(dir_spatial, 'raster/ohibc_rgn_raster_500m.tif')) %>%
  crop(ext_howe)

hs_wdpa_shp_file  <- file.path(dir_goal, 'spatial', 'hs_wdpa_poly.shp')
hs_wdpa_rast_file <- file.path(dir_goal, 'spatial', 'hs_wdpa_rast_500m.tif')
hs_pep_shp_file   <- file.path(dir_goal, 'spatial', 'hs_pep_poly.shp')
hs_pep_rast_file  <- file.path(dir_goal, 'spatial', 'hs_pep_rast_500m.tif')

rast_wdpa <- gdal_rast2(src = hs_wdpa_shp_file,
                        rast_base = rast_base, 
                        dst = hs_wdpa_rast_file, 
                        value = 'STATUS_YR', 
                        override_p4s = TRUE)

rast_pep  <- gdal_rast2(src = hs_pep_shp_file,
                        rast_base = rast_base, 
                        dst = hs_pep_rast_file, 
                        value = 'OBJECTID', ### no year field available
                        override_p4s = TRUE)

```


-----

## Set up coastal buffer rasters

Buffer shapefiles are located in `github/ohibc/prep/spatial`.  LSP uses 1 km inland and 3nm offshore buffers, while resilience requires analysis over the entire EEZ.  

Analysis will be done using raster::crosstab() comparing the WDPA raster to various region rasters.  Using a 500 m raster is the coarsest that should be used on a 1 km feature; a base raster is available at `~/github/ohibc/prep/spatial/ohibc_rgn_raster_500m.tif`.

* If rasters are not already available for 1 km inland, 3 nm offshore, and EEZ:
    * Read in buffer shapefiles to SpatialPolygonsDataFrames
    * rasterize to same extents/resolution as 500m base raster.

```{r get_analysis_rasters}

### check for presence of buffer rasters
rast_3nm_file <- file.path(dir_goal, 'spatial/hs_offshore_3nm_raster_500m.tif')
rast_1km_file <- file.path(dir_goal, 'spatial/hs_inland_1km_raster_500m.tif')

reload <- TRUE

if(!file.exists(rast_3nm_file) | !file.exists(rast_1km_file) | reload == TRUE) {
  message('Creating region buffer rasters from region buffer shapefiles')

  rast_base <- raster(file.path(dir_spatial, 'raster/ohibc_rgn_raster_500m.tif')) %>%
    crop(ext_howe)

  poly_3nm_file <- file.path(dir_spatial, 'howe_sound_offshore_3nm.shp')
  poly_1km_file <- file.path(dir_spatial, 'howe_sound_inland_1km.shp')

  rast_1km  <- gdal_rast2(src = poly_1km_file,
                          rast_base = rast_base, 
                          dst = rast_1km_file, 
                          value = 'rgn_id',
                          override_p4s = TRUE)
  rast_3nm  <- gdal_rast2(src = poly_3nm_file,
                          rast_base = rast_base, 
                          dst = rast_3nm_file, 
                          value = 'rgn_id',
                          override_p4s = TRUE)
} 

```

``` {r setup_watershed_raster}
rast_base <- raster(file.path(dir_spatial, 'raster/ohibc_rgn_raster_500m.tif')) %>%
  crop(ext_howe)

rast_watershed <- gdal_rast2(src = file.path(dir_spatial, 'watershed/howe_sound_watersheds'),
                                 rast_base = rast_base,
                                 dst = file.path(dir_goal, 'spatial/howe_sound_watershed_500m.tif'),
                                 # value = 'WATERSHED_',
                                 override_p4s = TRUE)
```

``` {r plot raster}

### hide git prov outputs for this chunk for now...
rast_3nm  <- raster(file.path(dir_goal, 'spatial/hs_offshore_3nm_raster_500m.tif'),
                    nogit = TRUE)
rast_1km  <- raster(file.path(dir_goal, 'spatial/hs_inland_1km_raster_500m.tif'),
                    nogit = TRUE)
rast_ws   <- raster(file.path(dir_goal, 'spatial/howe_sound_watershed_500m.tif'),
                    nogit = TRUE)
rast_wdpa <- raster(file.path(dir_goal, 'spatial/hs_wdpa_rast_500m.tif'),
                    nogit = TRUE)
rast_pep  <- raster(file.path(dir_goal, 'spatial/hs_pep_rast_500m.tif'),
                    nogit = TRUE)

library(tmap)

rast_map <- tm_shape(rast_3nm, is.master = TRUE) +
  tm_raster(alpha = 1, palette = 'blue', legend.show = FALSE) + 
  tm_shape(rast_ws) +
  tm_raster(alpha = 1, palette = 'purple', legend.show = FALSE) + 
  tm_shape(rast_1km) +
  tm_raster(alpha = .5, palette = 'green', legend.show = FALSE) + 
  tm_shape(rast_wdpa) +
  tm_raster(alpha = .5, palette = 'Reds') +
  tm_shape(rast_pep) +
  tm_raster(alpha = .5, palette = 'Oranges') +
  tm_layout(legend.outside = TRUE)


print(rast_map)

```

-----

# Calculate goal model

``` {r lsp_zonal_stats}

rast_3nm  <- raster(file.path(dir_goal, 'spatial/hs_offshore_3nm_raster_500m.tif'))
values(rast_3nm)[!is.na(values(rast_3nm))] <- 1
rast_1km  <- raster(file.path(dir_goal, 'spatial/hs_inland_1km_raster_500m.tif'))
values(rast_1km)[!is.na(values(rast_1km))] <- 1
rast_ws   <- raster(file.path(dir_goal, 'spatial/howe_sound_watershed_500m.tif'))
values(rast_ws)[!is.na(values(rast_ws))] <- 1

rast_pep  <- raster(file.path(dir_goal, 'spatial/hs_pep_rast_500m.tif'))
rast_wdpa_pep <- raster(file.path(dir_goal, 'spatial/hs_wdpa_rast_500m.tif'))
  ### will adjust values of this to include pep areas

### for any non-WDPA cells with PEP values, set to 1 (distinct from WDPA values, which are all years)
values(rast_wdpa_pep)[!is.na(values(rast_pep)) & is.na(values(rast_wdpa_pep))] <- 1

zonal_3nm_file <- file.path(dir_goal, 'int', 'zonal_stats_3nm.csv')
zonal_1km_file <- file.path(dir_goal, 'int', 'zonal_stats_1km.csv')
zonal_ws_file  <- file.path(dir_goal, 'int', 'zonal_stats_ws.csv')

stats_3nm <- raster::crosstab(rast_wdpa_pep, rast_3nm, useNA = TRUE, progress = 'text') %>%
  as.data.frame() %>%
  setNames(c('year', 'rgn_id', 'n_cells')) %>%
  mutate(year   = as.integer(as.character(year)),
         rgn_id = as.integer(as.character(rgn_id))) %>%
  filter(!is.na(rgn_id)) %>%
  arrange(rgn_id, year)

stats_1km <- raster::crosstab(rast_wdpa_pep, rast_1km, useNA = TRUE, progress = 'text') %>%
  as.data.frame() %>%
  setNames(c('year', 'rgn_id', 'n_cells')) %>%
  mutate(year   = as.integer(as.character(year)),
         rgn_id = as.integer(as.character(rgn_id))) %>%
  filter(!is.na(rgn_id)) %>%
  arrange(rgn_id, year)

stats_ws <- raster::crosstab(rast_wdpa_pep, rast_ws, useNA = TRUE, progress = 'text') %>%
  as.data.frame() %>%
  setNames(c('year', 'rgn_id', 'n_cells')) %>%
  mutate(year   = as.integer(as.character(year)),
         rgn_id = as.integer(as.character(rgn_id))) %>%
  filter(!is.na(rgn_id)) %>%
  arrange(rgn_id, year)

write_csv(stats_3nm, zonal_3nm_file)
write_csv(stats_1km, zonal_1km_file)
write_csv(stats_ws,  zonal_ws_file)

```

Once the WDPA raster is cross-tabulated against the OHI region rasters (both 3 nm offshore and 1 km inland) we have the number of protected cells, identified by year of protection, within each region.  NA values are unprotected cells.

### Summary of zonal stats dataframes (3 nm offshore):

``` {r}
print(summary(stats_3nm))
```

### Summary of zonal stats dataframes (1 km inland):

``` {r}
print(summary(stats_1km))
```

### Summary of zonal stats dataframes (inland, full watershed):

``` {r}
print(summary(stats_ws))
```

-----

## Calculate protected area and total area by region

Grouping by rgn_id, the total number of cells per region is determined by summing cell counts across ALL years, including cells with year == NA (unprotected cells).  We can then determine the protected area for each year by looking at the cumulative sum of cells up to any given year.

Since the cells are 500 m on a side, we can easily calculate area by multiplying cell count * 0.25 km^2^ per cell.

Finally we can calculate the status of a region for any given year by finding the ratio of protected:total and normalizing by the goal's target of 30% protected area.


``` {r summarize_zonal_stats, eval = TRUE}

stats_3nm <- read_csv(file.path(dir_goal, 'int', 'zonal_stats_3nm.csv'))
stats_1km <- read_csv(file.path(dir_goal, 'int', 'zonal_stats_1km.csv'))
stats_ws  <- read_csv(file.path(dir_goal, 'int', 'zonal_stats_ws.csv'))

lsp_thresh <- 0.30

### Determine total cells per region (n_cells_tot) and then a cumulative
### total of cells per region
prot_1km <- stats_1km %>%
  filter(!is.na(rgn_id)) %>% ### ditch cells outside region boundaries
  mutate(n_cells_tot = sum(n_cells),
         n_cells_cum = cumsum(n_cells),
         a_tot_km2   = n_cells_tot / 4,
         a_prot_km2  = n_cells_cum / 4) %>%
  filter(!is.na(year))  %>% ### this ditches non-protected cell counts but already counted in n_cells_tot
  mutate(pct_prot   = round(n_cells_cum / n_cells_tot, 4),
         lsp_status = round(ifelse(pct_prot > lsp_thresh, 100, (pct_prot / lsp_thresh) * 100), 2)) %>%
  distinct() %>%
  complete(year = c(1, 1960:2016)) %>%
  fill(rgn_id:lsp_status)

prot_3nm <- stats_3nm %>%
  filter(!is.na(rgn_id)) %>% ### ditch cells outside region boundaries
  mutate(n_cells_tot = sum(n_cells),
         n_cells_cum = cumsum(n_cells),
         a_tot_km2   = n_cells_tot / 4,
         a_prot_km2  = n_cells_cum / 4) %>%
  filter(!is.na(year))  %>% ### this ditches non-protected cell counts but already counted in n_cells_tot
  mutate(pct_prot   = round(n_cells_cum / n_cells_tot, 4),
         lsp_status = round(ifelse(pct_prot > lsp_thresh, 100, (pct_prot / lsp_thresh) * 100), 2)) %>%
  distinct()  %>%
  complete(year = c(1, 1960:2016)) %>%
  fill(rgn_id:lsp_status)

prot_ws <- stats_ws %>%
  filter(!is.na(rgn_id)) %>% ### ditch cells outside region boundaries
  mutate(n_cells_tot = sum(n_cells),
         n_cells_cum = cumsum(n_cells),
         a_tot_km2   = n_cells_tot / 4,
         a_prot_km2  = n_cells_cum / 4) %>%
  filter(!is.na(year))  %>% ### this ditches non-protected cell counts but already counted in n_cells_tot
  mutate(pct_prot   = round(n_cells_cum / n_cells_tot, 4),
         lsp_status = round(ifelse(pct_prot > lsp_thresh, 100, (pct_prot / lsp_thresh) * 100), 2)) %>%
  distinct()  %>%
  complete(year = c(1, 1960:2016)) %>%
  fill(rgn_id:lsp_status)

write_csv(prot_ws, file.path(dir_goal,  'int', 'area_protected_ws.csv'))
write_csv(prot_3nm, file.path(dir_goal, 'int', 'area_protected_3nm.csv'))
write_csv(prot_1km, file.path(dir_goal, 'int', 'area_protected_1km.csv'))

```


### Protected areas and status (latest year only):

`r knitr::kable(prot_3nm %>% filter(year == max(year)) %>% select(-rgn_id, -contains('cell')), caption = '3 nautical mile offshore zone - area in km^2^')`

`r knitr::kable(prot_1km %>% filter(year == max(year)) %>% select(-rgn_id, -contains('cell')), caption = '1 kilometer inland zone - area in km^2^')`

`r knitr::kable(prot_ws %>% filter(year == max(year)) %>% select(-rgn_id, -contains('cell')), caption = 'watershed inland zone - area in km^2^')`

***

## Combine scores for inland and offshore, and writing output layers

The status is based on an area-weighted average of the offshore (3nm) and inland (watershed) scores. 

``` {r combine_inland_and_offshore, eval = TRUE}

prot_3nm <- read_csv(file.path(dir_goal, 'int', 'area_protected_3nm.csv'))
prot_1km <- read_csv(file.path(dir_goal, 'int', 'area_protected_1km.csv'))
prot_ws  <- read_csv(file.path(dir_goal, 'int', 'area_protected_ws.csv'))

prot_df <- prot_1km %>%
  dplyr::select(rgn_id, year,
                lsp_st_1km = lsp_status,
                a_prot_1km = a_prot_km2,
                a_tot_1km  = a_tot_km2) %>%
  full_join(prot_3nm %>% 
              dplyr::select(rgn_id, year,
                            lsp_st_3nm = lsp_status,
                            a_prot_3nm = a_prot_km2,
                            a_tot_3nm  = a_tot_km2),
            by = c('rgn_id', 'year')) %>%
  full_join(prot_ws %>% 
              dplyr::select(rgn_id, year,
                            lsp_st_ws = lsp_status,
                            a_prot_ws = a_prot_km2,
                            a_tot_ws  = a_tot_km2),
            by = c('rgn_id', 'year')) %>%
  fill(lsp_st_1km:a_tot_ws) %>% 
    ### this catches any years where NAs exist for one or the other region;
    ### the complete() and fill() from the dataframes should catch this anyway
  mutate(lsp_st_1km = ifelse(is.na(lsp_st_1km), 0, lsp_st_1km),
         lsp_st_3nm = ifelse(is.na(lsp_st_3nm), 0, lsp_st_3nm),
         lsp_st_ws  = ifelse(is.na(lsp_st_ws),  0, lsp_st_ws),
         lsp_status_areawt  = (lsp_st_ws * a_tot_ws + lsp_st_3nm * a_tot_3nm) / (a_tot_ws + a_tot_3nm),
         lsp_status_3nm_1km = (lsp_st_3nm + lsp_st_1km) / 2,
         lsp_status         = (lsp_st_ws  + lsp_st_3nm) / 2) %>%
  distinct()

write_csv(prot_df, file.path(dir_goal, 'output', 'area_protected_total.csv'))

```

`r DT::datatable(prot_df %>% select(-rgn_id))`

``` {r estimate status and trend by year, eval = TRUE}
status_file <- file.path(dir_goal, 'output', 'lsp_status.csv')
trend_file  <- file.path(dir_goal, 'output', 'lsp_trend.csv')

year_span <- 20

status_df <- read_csv(file.path(dir_goal, 'output', 'area_protected_total.csv')) %>% 
  select(rgn_id, year, lsp_status) %>%
  filter(year > max(year) - year_span)
write_csv(status_df, status_file)

trend_df <- read_csv(file.path(dir_goal, 'output', 'area_protected_total.csv')) %>% 
  select(rgn_id, year, lsp_status) %>%
  filter(year > max(year) - year_span)

trend_df$trend_lm <- lm(lsp_status ~ year, data = trend_df)$coefficients[[2]]

### convert to percent change per annum and mult by 5 for 5-year prediction
trend_df <- trend_df %>%
  mutate(lsp_trend = trend_lm/first(lsp_status) * 5) %>%
  select(-lsp_status, -trend_lm)

write_csv(trend_df, trend_file)

```

Year-by-year status and trend estimates will be saved:

* ``r status_file``: estimate of status by region since 1980
* ``r trend_file``: estimate of trend by region since 1990

### Status and trend estimates:

``` {r calc_lsp_status_trend_summary}
lsp_status_trend_summary <- status_df %>% 
  left_join(trend_df,  by = c('rgn_id', 'year')) %>%
  arrange(desc(year), rgn_id)
```

`r DT::datatable(lsp_status_trend_summary, caption = 'LSP status and trend estimates')`

-----

``` {r, results = 'asis'}
prov_wrapup()
```

