---
title: 'OHIBC data prep: Salmon'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(raster)
library(sf)

source('~/github/ohibc/src/R/common.R')  ### an OHIBC specific version of common.R

scenario <- 'v2017'
goal     <- 'sal'
dir_git  <- path.expand('~/github/ohibc')
dir_goal <- file.path(dir_git, 'prep', goal, scenario)
dir_spatial <- file.path(dir_git, 'prep/_spatial')

dir_data_bc  <- file.path(dir_M, 'git-annex/bcprep', '_raw_data')

library(provRmd); prov_setup()

reload <- FALSE


```

# Summary: OHIBC Salmon

This script prepares layers Salmon sub-goals for food provision and First Nations Resource Access Opportunity in British Columbia's coastal regions.  

-----

# Data sources

* DFO Salmon catch and escapement data

-----

# Methods

## Extract catch and escapement info

This is from Karen's cleaned salmon data spreadsheet.

``` {r}

salmon_cleaned_file <- file.path(dir_goal, 'int/salmon_data_cleaned.csv')


salmon_xls  <- file.path(dir_data_bc, 'dfo_salmon/salmon_data_cleaned_KLH_Aug12.xlsx')

salmon_df <- readxl::read_excel(salmon_xls) %>%
  clean_df_names() %>%
  filter(!is.na(stock)) %>%
  transmute(year, 
            stock, 
            esc      = as.numeric(escapement),
            esc_tgt  = as.numeric(escapement_target),
            c_er     = as.numeric(realized_catch_or_exploitation_rate),
            c_er_tgt = as.numeric(harvest_target_or_exploitation_rate)) %>%
  group_by(stock) %>%
  mutate(esc_avail = sum(!is.na(esc) & !is.na(esc_tgt)),
         catch_avail = sum(!is.na(c_er) & !is.na(c_er_tgt))) %>%
  ungroup()
  
write_csv(salmon_df, salmon_cleaned_file)

```

### Exclude stocks with inadequate data

From the raw salmon data, we will include only stocks with at least five observations for (realized catch/expl rate AND harvest/expl rate target) and/or (escapement AND escapement target).  Stocks with neither will be dropped.

``` {r identify stocks with inadequate data}

salmon_data_df <- read_csv(salmon_cleaned_file)

stocks_dropped <- salmon_data_df %>%
  filter(esc_avail < 5 & catch_avail < 5) %>%
  select(stock) %>% distinct()

knitr::kable(stocks_dropped, caption = 'Stocks dropped due to inadequate data')

```


## Write output layers

Output layers for `functions.R` will be simply the year, stock, E/Et, C/Ct; these will each be assigned equally across all regions.

* `sal_catch.csv`
* `sal_escapes.csv`
* `ao_sal_catch.csv` is a dupe of `sal_catch` but treated as a separate layer in case of future changes that differentiate between FP and AO purposes.
* `ao_sal_escapes.csv` is similarly a dupe of `sal_escapes`


``` {r}

stocks <- read_csv(file.path(dir_goal, 'int/salmon_data_cleaned.csv'))

yrs <- stocks$year %>% unique() %>% sort()
rgn_yrs <- data.frame(rgn_id = rep(c(1:6, 8), times = length(yrs)),
                      year   = rep(yrs, each = 7))

stock_layers <- stocks %>%
  mutate(E_Et = (esc / esc_tgt) %>% round(5),
         C_Ct = (c_er / c_er_tgt) %>% round(5),
         C_Ct = ifelse(is.infinite(C_Ct), NA, C_Ct))

catch_layer <- stock_layers %>%
  filter(catch_avail >= 5) %>% 
    ### arbitrary data quality threshold for keeping a stock
  filter(!is.na(C_Ct)) %>%
  select(year, stock, C_Ct) %>%
  left_join(rgn_yrs, by = 'year')

escapes_layer <- stock_layers %>%
  filter(esc_avail >= 5) %>% 
    ### arbitrary data quality threshold for keeping a stock
  filter(!is.na(E_Et)) %>%
  select(year, stock, E_Et) %>%
  left_join(rgn_yrs, by = 'year')

write_csv(catch_layer,   file.path(dir_goal, 'output', 'sal_catch.csv'))
write_csv(escapes_layer, file.path(dir_goal, 'output', 'sal_escapes.csv'))
write_csv(catch_layer,   file.path(dir_goal, 'output', 'ao_sal_catch.csv'))
write_csv(escapes_layer, file.path(dir_goal, 'output', 'ao_sal_escapes.csv'))

```

-----

``` {r provenance, results = 'asis'}

prov_wrapup()

```
