---
title: 'OHIBC data prep: Salmon'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(raster)
library(sf)

source('~/github/ohibc/src/R/common.R')  ### an OHIBC specific version of common.R

scenario <- 'v2017'
goal     <- 'sal'
dir_git  <- path.expand('~/github/ohibc')
dir_goal <- file.path(dir_git, 'prep', goal, scenario)
dir_spatial <- file.path(dir_git, 'prep/_spatial')

dir_data_bc  <- file.path(dir_M, 'git-annex/bcprep', '_raw_data')

library(provRmd); prov_setup()

reload <- FALSE


```

# Summary: OHIBC Salmon

This script prepares layers Salmon sub-goals for food provision and First Nations Resource Access Opportunity in British Columbia's coastal regions.  

From Halpern et al. (2012) :


-----

# Data sources

* DFO Salmon catch and escapement data

-----

# Methods

## Extract catch and escapement info

``` {r}

salmon_raw_file <- file.path(dir_goal, 'int/salmon_data_raw.csv')

if(!file.exists(salmon_raw_file)) {
  
  salmon_sum_xls  <- file.path(dir_data_bc, 'dfo_salmon/Salmon Data Summary.xlsx')
  salmon_sum_cols <- readxl::read_excel(salmon_sum_xls, 1) %>%
    clean_df_names()
  salmon_sum_avail <- readxl::read_excel(salmon_sum_xls, 2) %>%
    clean_df_names()
  
  salmon_xls <- file.path(dir_data_bc, 'dfo_salmon/OHI Data Canada 29 Dec 2017.xlsx')
  
  ltr_to_num <- function(x) {
    if(is.na(x)) return(NA)
    if(nchar(x) == 1) {
      return (which(letters == tolower(x)))
    }
    if(nchar(x) == 2) {
      x1 <- substr(x, 1, 1)
      x2 <- substr(x, 2, 2)
      x1_num <- which(letters == tolower(x1))
      x2_num <- which(letters == tolower(x2))
      return( x1_num * 26 + x2_num)
    }
  }
  
  get_col_data <- function(y, type = 'c') {
    if(is.na(y)) {
      return('no column given')
    } else if (type == 'c') {
      return(as.character(data_raw[ , y]))
    } else {
      return(data_raw[ , y])
    }
  }
  
  salmon_stocks <- salmon_sum_cols$stock_name
    ### NOTE: Stikine sheet has two stocks; go by stock rather than sheet!
  salmon_list <- vector('list', length = length(salmon_stocks))
  
  for(i in seq_along(salmon_stocks)) { ### i <- 10
    
    col_info <- salmon_sum_cols %>%
      filter(stock_name == salmon_stocks[i])
    avail_info <- salmon_sum_avail %>%
      filter(stock_name == salmon_stocks[i])
  
    skip_rows     <- col_info$data_start - 1
    ### Column IDs are given as Excel column letters; convert to number
    esc_tgt_col   <- ltr_to_num(col_info$escapement_target_col_id)
    esc_col       <- ltr_to_num(col_info$escapement_col_id)
    catch_tgt_col <- ltr_to_num(col_info$catch_target_col_id)
    catch_col     <- ltr_to_num(col_info$catch_col_id)
    year_col      <- ltr_to_num(col_info$yr_col)
    
    data_raw <- readxl::read_excel(salmon_xls, 
                                   sheet = col_info$sheet_name,
                                   skip = skip_rows,
                                   col_names = FALSE) %>%
      as.matrix()
    
    stock_data <- data.frame('stock'     = col_info$stock_name,
                             'avail'     = avail_info$availability,
                             'year'      = get_col_data(year_col),
                             'esc_tgt'   = get_col_data(esc_tgt_col),
                             'escapes'   = get_col_data(esc_col),
                             'catch_tgt' = get_col_data(catch_tgt_col),
                             'catch'     = get_col_data(catch_col))
  
    salmon_list[[i]] <- stock_data
  
  }
  
  salmon_data_raw <- bind_rows(salmon_list) %>%
    filter(!is.na(year))
  
  write_csv(salmon_data_raw, salmon_raw_file)
  
} else {
  
  git_prov(salmon_raw_file, 'output')

}

```

From the raw salmon data, we will include only stocks with columns for (catch & catch target) and/or (escapement & escapemenet target).  Stocks with neither will be dropped.  
``` {r identify stocks with inadequate data}
salmon_data_raw <- read_csv(salmon_raw_file)

stocks_dropped <- salmon_data_raw %>%
  filter(avail == 3) %>%
  select(stock) %>% distinct()

knitr::kable(stocks_dropped, caption = 'Stocks dropped due to inadequate data')
```

Due to the methods for extracting data from the spreadsheets, some columns have text-based data that needs to be extracted or cleaned.  Extract embedded data from text where applicable, and use "as.numeric" to convert text w/o data to NA.

``` {r identify stocks with inadequate data}
salmon_data_raw <- read_csv(salmon_raw_file)

stocks <- salmon_data_raw %>%
  filter(avail != 3)

cat('\nProblems in "year":\n')
print(stocks$year[str_detect(stocks$year, '[^0-9]')])
cat('\nProblems in "esc_tgt":\n')
print(stocks$esc_tgt[str_detect(stocks$esc_tgt, '[^0-9\\.]')] %>% unique())
cat('\nProblems in "escapes":\n')
print(stocks$escapes[str_detect(stocks$escapes, '[^0-9\\.]')] %>% unique())
cat('\nProblems in "catch_tgt":\n')
print(stocks$catch_tgt[str_detect(stocks$catch_tgt, '[^0-9\\.]')] %>% unique())
cat('\nProblems in "catch":\n')
print(stocks$catch[str_detect(stocks$catch, '[^0-9\\.]')] %>% unique())

knitr::kable(stocks_dropped, caption = 'Stocks dropped due to inadequate data')
```

``` {r}

salmon_data_raw <- read_csv(salmon_raw_file)

stocks <- salmon_data_raw %>%
  filter(avail != 3) %>%
  mutate(year = as.integer(year),       
           ### no actual problems here
         escapes = as.numeric(escapes), 
           ### no actual problems here
         esc_tgt = str_replace_all(esc_tgt, '[^0-9-]', ''), 
           ### ditch non-digits, spaces, and commas
         esc_tgt = str_replace_all(esc_tgt, '-.*', ''),     
           ### use hyphen to drop anything after
         esc_tgt = as.numeric(esc_tgt), 
           ### no problems remaining
         catch_tgt = ifelse(str_detect(catch_tgt, '[^0-9\\.]'), 
                                           str_extract(catch_tgt, '[0-9]*\\%'), 
           ### if non-digits in string, extract number attached to %; this extracts the 5 from "3-5%"
                                           catch_tgt),
           ### otherwise assign all-digits catch target field
         catch_tgt = as.numeric(str_replace(catch_tgt, '%', '')),
         catch = ifelse(catch == 'uncertain (10.1 to 19.4 depending on methods)', 10.1, catch), 
           ### manually assign this one
         catch = as.numeric(catch))

write_csv(stocks, file.path(dir_goal, 'int/salmon_data_cleaned.csv'))

```

-----

``` {r provenance, results = 'asis'}

prov_wrapup()

```
