---
title: 'OHIBC goal prep: Iconic Species (Sense of Place subgoal)'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(sp)        # the classes and methods that make up spatial ops in R
library(rgdal)
library(raster)
library(data.table)

dir_git <- '~/github/ohibc'

source(file.path(dir_git, 'src/R/common.R'))  ### an OHIBC specific version of common.R

dir_anx <- file.path(dir_neptune_data, 'git-annex/bcprep') ### git-annex: goal-specific large files
dir_anx_global <- file.path(dir_neptune_data, 'git-annex/globalprep/SPP_ICO')
dir_spp <- file.path(dir_git, 'prep/spp_ico')
dir_rgn <- file.path(dir_git, 'prep/regions')  ### github: general buffer region shapefiles

dir_data_am    <- file.path(dir_neptune_data, 'git-annex/globalprep/_raw_data', 'aquamaps/v2014') 
dir_data_iucn  <- file.path(dir_neptune_data, 'git-annex/globalprep/_raw_data', 'iucn_spp') 

scenario <- 'v2016'

source(file.path(dir_git, 'src/R/prov.R'))   
  ### Provenance tracking functions: must source at start to initialize prov_track
dir_prov <- file.path(dir_spp, 'prov') 
  ### set a provenance folder for this script
this_script_file <- file.path(dir_spp, 'data_prep_ico.Rmd')
  ### can't recognize the current script name on its own :(
prov_run_tag <- 'standard run'

source(file.path(dir_git, 'src/R/poly_plot_scores.R'))
### for plotting final scores

source(file.path(dir_spp, 'R/spp_fxn.R'))
git_prov(file.path(dir_spp, 'R/spp_fxn.R'), type = 'sourced script')
### SPP-specific and ICO-specific functions

### set up proj4string options: BC Albers and WGS84
p4s_opts <- c('EPSG:3005 NAD83/BC Albers' = '+proj=aea +lat_1=50 +lat_2=58.5 +lat_0=45 +lon_0=-126 +x_0=1000000 +y_0=0 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0',
              'EPSG:4326 WGS 84'          = '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0')
p4s_bcalb <- p4s_opts[1]
```


# Summary: OHIBC Iconic Species Subgoal (Sense of Place)

This script prepares scores (status and trend) for Iconic Species in 
British Columbia's coastal regions.  Spatial data from IUCN and Aquamaps is
combined with extinction risk information from IUCN and conservation rank
info based on NatureServe categories.

Currently, the Iconic Species sub-goal model is identical to the OHI Global 
model: a region's status is based upon an unweighted average of species
health for all 'iconic' species found within each BC reporting region.

From Halpern et al (2012):

> Iconic species are those that are relevant to local cultural identity through a species’ relationship to one or more of the following: 1) traditional activities such as fishing, hunting or commerce; 2) local ethnic or religious practices; 3) existence value; and 4) locally-recognized aesthetic value (e.g., touristic attractions/common subjects for art such as whales). Habitat-forming species are not included in this definition of iconic species, nor are species that are harvested solely for economic or utilitarian purposes (even though they may be iconic to a sector or individual). ...

> Ultimately, almost any species can be iconic to someone, and so the intent with this goal was to focus on those species widely seen as iconic within a country, and iconic from a cultural or existence value (rather than for a livelihoods or extractive reason). ...

> The reference point is to have the risk status of all assessed species as Least Concern (i.e., a goal score = 1.0)

The Status of this sub-goal (X~ICO~) is then the % of iconic species in each threat category (as defined by the IUCN Red List, and for OHIBC, NatureServe), such that:

$$X_{ICO} = \frac{\displaystyle\sum_{category}S_{cat}*w_{cat}}{\displaystyle\sum_{category}S_{cat}}$$

where for each IUCN threat category (or analogous NatureServe conservation rank):

* *S~cat~* is the number of assessed species in the category
* *w~cat~* is the status weight assigned for that category (note, these are the inverse of the risk value used in the SPP calculations):
    * 'LC' = 1.0, 'NT' = 0.8, 'VU' = 0.6, 'EN' = 0.4, 'CR' = 0.2, 'EX' = 0.0
    * NatureServe ranks are assigned analogous scores
    * Where available, NatureServe province-scale ranks take precedence over IUCN ranks as they are a better representation of the risk at the scale of British Columbia.

ICO trend is calculated in a similar manner, but weightings are assigned according to IUCN population trend: 'Decreasing' = -0.5, 'Stable' = 0.0, 'Increasing' = +0.5.  NatureServe information on population trend was not readily available.

# Data Sources

For information on IUCN data (range maps, extinction risk categories, population trends), NatureServe (conservation rankings), and AquaMaps (range maps), please see the OHIBC SPP data prep document, since this ICO goal is calculated from the intermediate outputs of that workflow.

Iconic Species List:

* The OHI global iconic species list was created by combining three species lists from WWF Global: global priorities, regional and local priorities, and flagship species
* <citation>

-----
# Methods

## Prepare master list of Iconic Species

Get list of all species listed as "iconic" for BC.  For now, the analysis is based on the global list of Iconic Species, filtered to just the species that occur in BC.

To this global ICO species list, we attach `spp_all_bcsee` df to acquire IUCN category & trend, and NatureServe global and provincial conservation rank information.

``` {r read iconic species list, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
ico_list_file <- file.path(dir_spp, scenario, 'ico_list_ohibc.csv')

message(sprintf('Reading OHIBC iconic species list from: \n  %s\n', ico_list_file))
ico_list_raw <- read.csv(ico_list_file, stringsAsFactors = FALSE) %>%
  select(rgn_name   = Country, 
         comname    = Specie.Common.Name,  
         sciname    = Specie.Scientific.Name, 
         ico_flag   = Flagship.Species,
         ico_local  = Priority.Species_Regional.and.Local,
         ico_global = Priority.Species_Global,
         ico_rgn    = Nation.Specific.List
  )
git_prov(ico_list_file)

### clean up names
ico_list_raw <- ico_list_raw %>%
  mutate(rgn_name = str_trim(rgn_name),
         sciname  = str_trim(sciname),
         comname  = str_trim(comname)) %>%
  filter(sciname != '')  

### ditch local species except where rgn_name == 'Canada', then ditch global and regional flags.
ico_list <- ico_list_raw %>%
  filter(is.na(ico_local) | rgn_name == 'Canada') %>%
  select(comname, sciname) %>%
  unique()

### SPP data_prep reads global spp_all, appends BC NatureServe ranks, and
### saves to OHIBC goal directory.  This function uses that updated file.
### NOTE: Category score already accounts for status_pr_score by now.
spp_all <- read.csv(file.path(dir_spp, scenario, 'int/spp_all_bcsee.csv'), 
                    stringsAsFactors = FALSE)
git_prov(file.path(dir_spp, scenario, 'int/spp_all_bcsee.csv'))

# join ico_list to spp_all to incorporate category info and parent/subpop info.
ico_list <- ico_list %>%
  left_join(spp_all %>%
              select(sciname, am_sid, iucn_sid, 
                     popn_trend, popn_category, category_score, trend_score, 
                     status_gl, status_pr, status_gl_score, status_pr_score),
            by = 'sciname')
ico_list <- ico_list %>%
  filter(!is.na(popn_category) | !(is.na(status_pr)))
```

-----

## Find BC-specific cell locations for all species on ICO list

Using same raster method as for SPP, identify .5 deg cells within BC regions.
Then filter IUCN and AM lists to just species on ico_list, and just cells within BC regions.

``` {r set up BC rasters, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}

### Read in OHIBC polygons, in WGS84 CRS
rgn_lyr <- 'ohibc_rgn_wgs84'

message(sprintf('Reading OHIBC regions shapefile...\n  %s/%s.shp', dir_rgn, rgn_lyr))
poly_bc_rgn <- readOGR(dsn = path.expand(dir_rgn), layer = rgn_lyr,
                       verbose = FALSE, stringsAsFactors = FALSE)
git_prov(file.path(dir_rgn, sprintf('%s.shp', rgn_lyr)))

### From OHIBC polygons, extract against AquaMaps LOICZID raster
### to get BC-specific cells (using spp_rgn2cell() function)
rgn_p4s <- proj4string(poly_bc_rgn)
names(rgn_p4s) <- names(p4s_opts[p4s_opts == rgn_p4s])
message(sprintf('Region loaded: CRS = %s \n  (%s)\n', names(rgn_p4s), rgn_p4s))

rgn2cell_list <- spp_rgn2cell(poly_bc_rgn, reload  = FALSE)

rgn2cell_df      <- rgn2cell_list[[1]]
loiczid_rast     <- rgn2cell_list[[2]]

```

``` {r Get spatial distributions, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}

### Get spatial distributions of species from both Aquamaps and IUCN shps

am_cells_spp   <- spp_get_am_cells(rgn2cell_df, n_max = -1, prob_filter = 0, reload = FALSE)
iucn_cells_spp <- spp_get_iucn_cells(rgn2cell_df)

iucn_ico_bc <- iucn_cells_spp %>%
  filter(sciname %in% ico_list$sciname) %>%
  select(rgn_id, loiczid, rgn_name, sciname) %>%
  mutate(sp_source = 'iucn') %>%
  unique()
### 3413 species/cell combinations

am_ico_bc <- am_cells_spp %>%  
  left_join(spp_all %>% 
              select(am_sid, sciname), 
            by = 'am_sid') %>%
  filter(sciname %in% ico_list$sciname) %>%
  select(rgn_id, loiczid, rgn_name, sciname) %>%
  mutate(sp_source = 'am') %>%
  unique()
```

At this point we have `r nrow(iucn_ico_bc)` species/cell instances based on IUCN spatial data, and `r nrow(am_ico_bc)` species/cell instances based on AM spatial data.

**NOTE:** Unlike the SPP goal, in which IUCN data was preferentially used to identify cell locations for a species, for ICO I am allowing both IUCN and Aquamaps datasets to indicate presence of an iconic species in a given cell.  

For example, if in cell #65000, for Tursiops truncatus, IUCN indicates presence and Aquamaps does not, it will be registered as present.  But if in cell #65001, Aquamaps indicates presence and IUCN does not, it will also by registered as present (whereas in SPP it would register as NOT present).

Because the spatially explicit range is not critical, but rather just whether the species is present or absent within a give assessment region, it seems reasonable to consider both spatial sources.

``` {r find ico by cell, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}

ico_bc_cell_rgn <- rbind(iucn_ico_bc, am_ico_bc) %>%
  left_join(spp_all %>%
              filter(!str_detect(spatial_source, 'subpop') &
                     !str_detect(spatial_source, 'alias')) %>%
              select(sciname, category_score, trend_score),
            by = 'sciname')

```

Iconic Species currently identified within British Columbia waters: 
*`r paste(ico_bc_cell_rgn$sciname %>% unique(), collapse = ', ')`*.

### Map of species count by cell

``` {r ico plot species count raster, echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE, fig.height = 4, fig.width = 6, fig.align = 'center'}
ico_bc_cells <- ico_bc_cell_rgn %>%
  select(-rgn_id, -rgn_name, -sp_source) %>%
  unique()

ico_summary_cells <- ico_bc_cells %>%
  group_by(loiczid) %>%
  summarize(n_ico = n(), 
            mean_cat = mean(category_score, na.rm = TRUE), 
            mean_trend = mean(trend_score, na.rm = TRUE)) %>%
  as.data.frame()

spp_plot_raster(ico_summary_cells, loiczid_rast, which_id = 'n_ico', 
                poly_rgn = poly_bc_rgn,
                title = 'Iconic Species count', scale_label = 'ico count', 
                scale_limits = c(0, max(ico_summary_cells$n_ico)))


```

-----

# Calculate Goal Model

## Summarize mean category/trend per cell/region, IUCN

For each raster cell, we generate a summary of number of species included, average extinction risk score of all species in cell, and average trend score of all species in cell with population trend information.  This section calculates extinction risk score based solely on IUCN category ranks, which excludes subpopulation scores due to lack of spatially explicit distribution info for subpopulations.

From raster-level summaries, we generate a region-level summary based on area-weighted average extinction risk and population trend.

``` {r iconic species by rgn, echo = FALSE, message = FALSE, warning = FALSE}
ico_bc_rgn <- ico_bc_cell_rgn %>%
  select(-loiczid, -sp_source) %>%
  unique()

ico_summary_rgn <- ico_bc_rgn %>%
  group_by(rgn_id, rgn_name) %>%
  summarize(n_ico = n(), 
            mean_cat = mean(category_score, na.rm = TRUE), 
            mean_trend = mean(trend_score, na.rm = TRUE)) %>%
  as.data.frame()

DT::datatable(ico_bc_rgn %>% arrange(rgn_id),
              caption = 'Iconic Species by region',
              rownames = FALSE,
              class = 'stripe hover compact',
              options  = list(dom = 'tp'))

```

``` {r summing by cells and rgn, echo = FALSE, message = FALSE, warning = FALSE}

DT::datatable(ico_summary_rgn,
              caption = 'summary by region',
              rownames = FALSE,
              class = 'stripe hover compact',
              options  = list(dom = 'tp'))

# DT::datatable(ico_summary_cells,
#               caption = 'summary by cell (LOICZID)',
#               rownames = FALSE,
#               class = 'stripe hover compact',
#               options  = list(dom = 'tp'))
```


### Map of mean extinction risk category and population trend by cell

*Category value range: least concern = 0, extinct = 1*

*Trend value range: population increasing = +0.5, stable = 0, decreasing = -0.5*

``` {r ico plot cell score raster, echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE,  fig.height = 4, fig.width = 6, fig.align = 'center'}

spp_plot_raster(ico_summary_cells, loiczid_rast, which_id = 'mean_cat', 
                poly_rgn = poly_bc_rgn,
                title = 'Iconic Species mean rank', scale_label = 'ico rank', 
                scale_limits = c(min(ico_summary_cells$mean_cat), 
                                 max(ico_summary_cells$mean_cat)),
                rev_scale = TRUE)

spp_plot_raster(ico_summary_cells, loiczid_rast, which_id = 'mean_trend', 
                poly_rgn = poly_bc_rgn,
                title = 'Iconic Species mean trend', scale_label = 'ico trend', 
                scale_limits = c(min(ico_summary_cells$mean_trend), 
                                 max(ico_summary_cells$mean_trend)))

```

-----
  
## Create final output files

From the region-level summaries of extinction risk and population trend, we calculate regional status and trend scores, and save as .csv.

`r sprintf('Writing ICO status and trend to:\n  %s\n  %s\n', 
            file.path(dir_spp, scenario, 'data/ico_status.csv'),
            file.path(dir_spp, scenario, 'data/ico_trend.csv'))`

``` {r save status and trend, echo = FALSE, message = FALSE, warning = FALSE}

ico_status <- ico_summary_rgn %>%
  mutate(score = 1 - mean_cat) %>%
  select(rgn_id, score)
ico_trend <- ico_summary_rgn %>%
  select(rgn_id, score = mean_trend)

write_csv(ico_status, file.path(dir_spp, scenario, 'data/ico_status.csv'))
git_prov(file.path(dir_spp, scenario, 'data/ico_status.csv'), type = 'output')
write_csv(ico_trend,  file.path(dir_spp, scenario, 'data/ico_trend.csv'))
git_prov(file.path(dir_spp, scenario, 'data/ico_trend.csv'), type = 'output')

```

## Compare status and trend by region

``` {r Compare status and trend, echo = FALSE, message = FALSE, fig.height = 4, fig.width = 6, fig.align = 'center'}

ico_status_100 <- ico_status %>% 
  mutate(score = round(score*100, 2))

poly_plot_scores(ico_status_100, 
                 scale_label = 'ICO Status', 
                 title = 'OHIBC ICO Status')

poly_plot_scores(ico_trend, 
                 scale_label = 'ICO Trend', 
                 title = 'OHIBC ICO Trend', 
                 scale_limits = c(min(ico_trend$score, 0), 
                                  max(ico_trend$score, 0)))

DT::datatable(ico_status_100 %>% 
                mutate(status = score) %>%
                left_join(ico_trend %>%
                            mutate(trend = round(score, 3)), 
                          by = 'rgn_id') %>%
                left_join(poly_bc_rgn@data, by = 'rgn_id') %>%
                dplyr::select(rgn_id, rgn_name, status, trend),
              caption = 'ICO Status and Trend by region',
              rownames = FALSE,
              class = 'stripe hover compact',
              options  = list(dom = 't'))

```


``` {r child = file.path(dir_git, 'src/templates/ohibc_prov_ftr.Rmd')}
```
