---
title: 'OHIBC: readme.md generator'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

source('~/github/ohibc/src/R/common.R')  ### an OHIBC specific version of common.R;
  ### includes library(tidyverse); library(stringr)

dir_git     <- '~/github/ohibc'
dir_anx     <- file.path(dir_M, 'git-annex/bcprep')

```

# Summary

Walks through file structure of OHIBC and checks to see whether folders have a readme.  Eventually, will generate a readme for each folder based on a template.

# Method 

## Check folders for Readme.md files

Identify all folders that *should* have readme files.

TO DO: 

* For those that do, do a quick check on the quality: file size, difference from template, etc.
* For those that don't, flag them for next steps.

``` {r check github}

check_folders <- c(# 'calc_ohibc', 
                   # 'regionHoweSound', 
                   'prep')

git_dirs <- data.frame(repo = dir_git,
                       dir  = list.dirs(dir_git),
                       stringsAsFactors = FALSE)  %>%
  filter(!str_detect(dir, '/\\.'))

dir_check <- git_dirs %>%
  mutate(level = str_count(dir, '/') - str_count(path.expand(dir_git), '/')) %>%
  filter(level <= 3) %>%
  rowwise() %>%
  mutate(readme = list.files(dir, pattern = 'readme', ignore.case = TRUE) %>%
           paste(collapse = ', ')) %>%
  ungroup()
  
html_check <- dir_check %>%
  rowwise() %>%
  mutate(html = list(list.files(dir, pattern = 'htm[l]?$', ignore.case = TRUE))) %>%
  ungroup() %>%
  unnest(html) %>%
  select(dir, html) %>%
  mutate(file_base = str_replace(tolower(html), '.htm[l]?$', ''))

rmd_check <- dir_check %>%
  rowwise() %>%
  mutate(rmd = list(list.files(dir, pattern = 'rmd$', ignore.case = TRUE))) %>%
  ungroup() %>%
  unnest(rmd) %>%
  select(dir, rmd) %>%
  mutate(file_base = str_replace(tolower(rmd), '.rmd$', '')) 

check_join <- html_check %>%
  full_join(rmd_check, by = c('dir', 'file_base')) %>%
  select(-file_base)

dir_html <- dir_check %>%
  left_join(check_join, by = 'dir')

write_csv(dir_html, 'dir_html_github.csv')

DT::datatable(dir_html %>%
                mutate(dir = str_replace(dir, path.expand(repo), repo)) %>%
                select(-repo)) %>% 
  DT::formatStyle(1:6, fontSize = '12px', fontFamily = 'Helvetica')

```


## Scan for Rmd summaries

For each folder in the repo, scan the existing .Rmd files for "summary" information.  The following info is used for generating a new readme.md:

* Rmd path
* Rmd title (from yaml header)
* Rmd summary (from "# Summary" section up until next header or hr)

``` {r scan_for_rmd_summary}

dir_html <- read_csv('dir_html_github.csv')

rmd_sum_df <- dir_html %>% 
  filter(!is.na(rmd)) %>%
  mutate(rmd_path = file.path(dir, rmd))

get_rmd_info <- function(rmd_file) {
  ### rmd_file <- rmd_sum_df$rmd_path[11]
  rmd_text <- readr::read_file(rmd_file)

  rmd_sum <- rmd_text %>% 
    str_split('#+|```') %>%
    as.data.frame() %>%
    setNames('text') %>%
    filter(str_detect(text, 'Summary')) %>%
    mutate(text_clean = str_replace_all(text, 'Summary[:]?', ''),
             ### remove "summary" title
           text_clean = str_replace_all(text_clean, '\\*{3,}|-{3,}', ''), 
             ### remove horizontal rule markers (*** or ---); note this could
             ### also affect bold-italics e.g. ***bold italics*** but whatevs
           text_clean = str_trim(text_clean))
             ### remove line breaks
  
  rmd_title <- rmd_text %>% 
    str_split('#+') %>%
    as.data.frame() %>%
    setNames('text') %>%
    filter(str_detect(tolower(text), 'title')) %>%
    mutate(text_clean = str_extract(text, '(?<=(title:)).+'),
           text_clean = str_replace_all(text_clean, '"', ''),
           text_clean = str_replace_all(text_clean, "'", ''),
           text_clean = str_trim(text_clean))

  rmd_info_df <- data.frame(rmd_path    = rmd_file,
                            rmd_title   = rmd_title$text_clean[1],
                            rmd_summary = rmd_sum$text_clean[1],
                            stringsAsFactors = FALSE)
}

rmd_info_df <- rmd_sum_df$rmd_path %>%
  lapply(get_rmd_info) %>%
  bind_rows()

rmd_sum_df <- rmd_sum_df %>%
  left_join(rmd_info_df, by = 'rmd_path')

### include a flag for Rmd files with no summary (NA) or a short 
### summary (less than X characters long).

rmd_sum_df <- rmd_sum_df %>%
  mutate(rmd_summary_length = nchar(rmd_summary),
         rmd_summary = ifelse(is.na(rmd_summary), 'NA', rmd_summary),
         rmd_summary_issue = ((rmd_summary == 'NA') | (rmd_summary_length < 100)))

write_csv(rmd_sum_df, 'rmd_title_summary.csv')

DT::datatable(rmd_sum_df %>% 
                mutate(dir = str_replace(dir, path.expand(repo), repo)) %>%
                select(dir, html, rmd, 
                       rmd_title, rmd_summary, 
                       rmd_summary_length, rmd_summary_issue)) %>% 
  DT::formatStyle(1:7, fontSize = '12px', fontFamily = 'Helvetica')

```


## Assemble readme information 

Generate field values for template from the text of the Rmd titles, summaries, and file paths.  This also identifies Rmd and HTML files present in each directory and creates a bulleted list of these.

``` {r assemble_readme}

rmd_sum_df <- read_csv('rmd_title_summary.csv')
dir_sum_df <- read_csv('dir_html_github.csv') %>%
  left_join(rmd_sum_df)

### collapse all by directory; modify only prep folders (excluding spatial) for now
readme_df <- dir_sum_df %>%
  mutate(github_url = str_replace(dir, path.expand(dir_git), 
                                  'https://github.com/OHI-Science/ohibc/blob/master'),
         rawgit_url = str_replace(dir, path.expand(dir_git), 
                                  'https://rawgit.com/OHI-Science/ohibc/master'),
         readme_title = case_when(!is.na(rmd_title) ~ rmd_title,
                                  !is.na(rmd)       ~ rmd,
                                  !is.na(html)      ~ html,
                                  TRUE ~ str_replace(dir, path.expand(dir_git), '')),
         readme_rmd = ifelse(is.na(rmd), 'no Rmd present', file.path(github_url, rmd)),
         readme_html = ifelse(is.na(html), 'no HTML present', file.path(rawgit_url, html))) %>%
  mutate(text_block = sprintf('## %s\n\n* __Rmd file:__ %s \n* __HTML file:__ %s\n\n### Summary:\n\n%s',
                              readme_title, 
                              readme_rmd,
                              readme_html,
                              rmd_summary)) %>%
  group_by(dir) %>%
  summarize(text_block_all = paste(text_block, collapse = '\n\n-----\n\n')) %>%
  ungroup() %>%
  mutate(dir_short = str_replace(dir, path.expand(dir_git), ''))

### add an internal files field for display at the 'prep' level.  Note this
### still includes provenance files
readme_df <- readme_df %>%
  rowwise() %>%
  mutate(internal_files = paste(list.files(dir, pattern = 'rmd$|htm[l]?$', 
                                           ignore.case  = TRUE,
                                           recursive    = TRUE,
                                           full.names   = TRUE), collapse = '\n'),
         internal_files = str_replace_all(internal_files, paste0(dir, '/'), '* ')) %>%
  mutate(goal_dir = str_extract(dir, '(?<=prep/).+'),
         goal_dir = str_split(goal_dir, '/')[[1]][1]) %>%
  ungroup() %>%
  left_join(read_csv('goal_dir_lookup.csv'), by = 'goal_dir')

write_csv(readme_df, 'readme_df.csv')

```

## Create readmes in specified directories

### Identify the directories to be included

The full readme dataframe includes information for all folders in the repo (as specified by dir_git).  Here we can specify directories in which we want Readmes to be created, as directories to _include_ and to _exclude_.  Use regex patterns if desired; be aware of trailing slashes, e.g. do you want to write a Readme for `ohibc/prep` or only its subfolders? in which case specify `/prep/` to avoid the base `prep` folder.

``` {r define include-exclude directories}
### Include the following directories (str_detect looks for this in file path)
dirs_incl <- '/prep/'
### Exclude the following directories (str_detect looks for this in file path)
dirs_excl <- '/_spatial'

readme_df <- read_csv('readme_df.csv') %>%
  filter(str_detect(dir, paste(dirs_incl, collapse = '|'))) %>%
  filter(!str_detect(dir, paste(dirs_excl, collapse = '|')))

readme_probs <- read_csv('rmd_title_summary.csv') %>%
  filter(dir %in% readme_df$dir) %>%
  filter(rmd_summary_issue) %>%
  mutate(rmd_path = str_replace(rmd_path, path.expand(repo), repo)) %>%
  select(contains('rmd_'), -rmd_summary_issue)

```

Directories with the following strings will be _included_: 

* `r paste(dirs_incl, collapse = ', ')`

Directories with the following strings will be _excluded_:

* `r paste(dirs_excl, collapse = ', ')`

Note that the following .Rmd files may have issues with their summaries, either non-existent or suspiciously short.  Please take a look at the Rmds to make sure they are up to OHI standards!

<font size = 1>
`r DT::datatable(readme_probs) %>% DT::formatStyle(1:4, fontSize = '12px', fontFamily = 'Helvetica')`
</font>

### Loop over the specified directories and create readmes in each.

``` {r loop over directories}

readme_df <- read_csv('readme_df.csv') %>%
  filter(str_detect(dir, paste(dirs_incl, collapse = '|'))) %>%
  filter(!str_detect(dir, paste(dirs_excl, collapse = '|'))) %>%
  mutate(internal_files = ifelse(is.na(internal_files), 'no .Rmd or .html files present', internal_files))

for(i in seq_along(readme_df$dir)) {
  ### Check whether prep, year, or generic
  # i <- 28

  prep_dir <- readme_df$dir[i]
  
  # message('Creating README files for ', prep_dir)
  
  ### Identify the type of readme from folder name: 
  ### * generic prep folder (e.g. ohibc/prep/cw)
  ### * generic prep/scenario folder (e.g. ohibc/prep/cw/v2017)
  ### * generic other folder (e.g. ohibc/calc_ohibc)
  readme_type <- case_when(str_detect(basename(prep_dir), 'v[0-9]{4}|vHS') ~ 'generic_readme_year.md',
                           str_detect(prep_dir, '/prep/[_]?[a-zA-Z]{2}')   ~ 'generic_readme_prep.md',
                           TRUE ~ 'generic_readme.md')
  
  ### Use str_replace to replace field names in the generic template with 
  ### appropriate text from the readme_df dataframe.
  
  if(readme_type == 'generic_readme_year.md') {
    
    message('Creating goal/year README file for ', prep_dir)
    readme_text <- read_file(file.path(dir_git, 'src/templates', readme_type)) %>%
      str_replace('_DIRNAME_', readme_df$dir_short[i]) %>%
      str_replace('_GOALNAME_', readme_df$goal_name[i]) %>%
      str_replace('_GOALURL_', readme_df$goal_url[i]) %>%
      str_replace('_SUMMARYINFO_', readme_df$text_block_all[i])
    
  } else if(readme_type == 'generic_readme_prep.md') {
    
    message('Creating goal README file for ', prep_dir)
    readme_text <- read_file(file.path(dir_git, 'src/templates', readme_type)) %>%
      str_replace('_DIRNAME_', readme_df$dir_short[i]) %>%
      str_replace('_GOALNAME_', readme_df$goal_name[i]) %>%
      str_replace('_GOALURL_', readme_df$goal_url[i]) %>%
      str_replace('_FILEPATHINFO_', readme_df$internal_files[i])
    
  } else { ### go with generic
    
    message('Creating misc README file for ', prep_dir)
    readme_text <- read_file(file.path(dir_git, 'src/templates', readme_type)) %>%
      str_replace('_DIRNAME_', readme_df$dir_short[i]) %>%
      str_replace('_FILEPATHINFO_', readme_df$internal_files[i])
    
  }
  
  ### write out the readme.md to the appropriate directory
  readr::write_file(readme_text, file.path(readme_df$dir[i], 'README.md'))
  
}

```

``` {r delete old readmes if needed, eval = FALSE}

# x <- list.files(dir_git, recursive = TRUE, pattern = 'readme_old.md', full.names = TRUE, ignore.case = TRUE)
# unlink(x)

```

-----

## Create readmes for git-annex folders?

not yet...

``` {r check_git-annex, eval = FALSE}

anx_dirs <-  data.frame(repo = dir_anx,
                        dir  = list.dirs(dir_anx),
                        stringsAsFactors = FALSE) 

dir_check <- anx_dirs %>%
  mutate(dir = str_replace(dir, path.expand(dir_anx), ''),
         level = str_count(dir, '/')) %>%
  filter(level > 0 & level <= 3) %>%
  mutate(readme = list.files())

```
