---
title: 'OHIBC: data prep for wild-capture fisheries: DFO catch data'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, message = FALSE, warning = FALSE)

options(scipen = 999)

dir_git     <- path.expand('~/github/ohibc')
source(file.path(dir_git, 'src/R/common.R'))  
dir_spatial <- file.path(dir_git, 'prep/_spatial')  
dir_anx     <- file.path(dir_M, 'git-annex/bcprep')

### goal specific folders and info
goal      <- 'fis'
scenario  <- 'v2017'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario)
dir_goal_anx <- file.path(dir_anx, goal, scenario)
dir_spatial  <- file.path(dir_git, 'prep/spatial')
dir_dfo_data <- file.path(dir_anx, '_raw_data/dfo_khunter')


### provenance tracking
library(provRmd); prov_setup()
```

# Summary

This script preps the DFO catch data for use in the BC Fisheries model.

***

# Data Source 

**Reference**: Department of Fisheries & Oceans

**Downloaded**: May - June, 2018 (downloaded from Dropbox link shared by Karen Hunter at DFO)

**Description**:  Shapefiles for individual stocks representing catch per OHI BC region.

**Time range**: 2007-2015

**Format**:  Shapefiles

***
  
# Explore

Explore data: this first chunk just looks at the data to see what we are working with.

```{r}
library(tidyverse)
library(sf)
```

Let's take a look at one file to see the data format

```{r}
halibut <- st_read(file.path(dir_anx, "_raw_data/dfo_khunter/fisheries/d2018/halibut/Halibut_2015.shp"), quiet = T) %>% 
  as.data.frame() %>%
  select(-geometry)

#data held in the shapefile
DT::datatable(head(halibut))
```

Do they all look like this? Let's look at shrimp

```{r}
shrimp <- st_read(file.path(dir_anx, "_raw_data/dfo_khunter/fisheries/d2018/shrimp/Shrimp_2009.shp"), quiet = T) %>%
  as.data.frame() %>%
  select(-geometry)

#data held in the shapefile
DT::datatable(head(shrimp))
```

I emailed Karen at DFO to get the details. Here is the important information:

Column to use for all, if available is `Sum_LANDED` which has Kilograms as units. 

>For shrimp:   
> "You need to sum columns _pinks thru spny for shrimp catch"
>
>For Dungeness:  
> "Use sum calc_wg for crab catch"
>
>Column to use for **Spot prawn** is `sum_prwn_w`. The units are in kilograms.
>
>Column to use for **Herring** is `Sum_Catch`. The units are in MT.
>
>**Boccacio** – I do not understand why this is not `Sum_LANDED` (KGS) like the other groundfish. I’ll take a look at this. Could be an artifact of bycatch records in the DB.  Looks like there is a `sum_catch` in the output. Sorry about that. I was going quickly and didn’t catch this strange output.
>
>Column to use for **Geoduck** is `Tl_landed`. The units are in Pounds (LB).  
>
> Behold, the new tuna scientist confirmed that tuna catch has always been reported in pounds not kgs. I think I told you kgs. I either was given the wrong info previously or wrote it down incorrectly. 

***

# Prep work

Created a function, `read_files` to read in each annual file for each species, aggregate, and make a dataframe with the columns we are interested in (region and catch).
```{r}
spp_folders <- list.files(file.path(dir_anx, '_raw_data/dfo_khunter/fisheries/d2018'))

read_files <- function(folder, column){
  
  files <- list.files(paste0(dir_anx, '/_raw_data/dfo_khunter/fisheries/d2018/', folder), full.names = T)
  
  out <- data.frame()
  
  for(yr in 2007:2015){
    #get the shapefile for yr
    yr_files <- files[str_detect(files, as.character(yr))]
    
    yr_shp_file <- yr_files[str_detect(yr_files, as.character(".shp"))]
    
    yr_shp <- st_read(yr_shp_file, quiet = T) %>%
    as.data.frame() %>%
      select(rgn_name, rgn_id, rgn_code, column) %>%
      mutate(year = yr)
    
    out <- rbind(yr_shp, out)
  }
    return(out)
    
}
```

Now we can get a single dataset for each stock by running the `read_files` function. We are still waiting on two more years of data for Albacore tuna which is why I have it commented out. Otherwise this would throw an error since not all years are there. 

```{r}
#albacore
tuna <- read_files("albacore_tuna", "Sum_CatchW") %>%
  mutate(species = "albacore_tuna",
         units = "lbs") %>%
  rename(harvest = "Sum_CatchW")
#arrowtooth flounder
af <- read_files("arrowtooth_flounder", "Sum_LANDED") %>%
  mutate(species = "arrowtooth_flounder",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#boccacio
boc <- read_files("bocaccio", "Sum_LANDED") %>%
  mutate(species = "bocaccio",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#dogfish
dog <- read_files("dogfish", "Sum_LANDED") %>%
  mutate(species = "dogfish",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#dungness crab
crab <- read_files("dungeness_crab", "Sum_clc_wg") %>%
  mutate(species = "dungeness_crab",
         units   = "kg") %>%
  rename(harvest = "Sum_clc_wg")
#hake
hake <- read_files("hake", "Sum_LANDED") %>%
  mutate(species = "hake",
         units = "kg") %>%
  rename(harvest = "Sum_LANDED")
#halibut
hal <- read_files("halibut", "Sum_LANDED") %>%
  mutate(species = "halibut",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#herring
herr <- read_files("herring", "Sum_Catch") %>%
  mutate(species = "herring",
         units   = "metric_tons") %>%
  rename(harvest = "Sum_Catch")
#geoduck
geo <- read_files("geoduck", "Sum_Tt_lnd") %>%
  mutate(species = "geoduck",
         units   = "lbs") %>%
  rename(harvest = "Sum_Tt_lnd")
#lingcod
ling <- read_files("lingcod", "Sum_LANDED") %>%
  mutate(species = "lingcod",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#pacific cod
cod <- read_files("pacific_cod", "Sum_LANDED") %>%
  mutate(species = "pacific_cod",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#pacific ocean perch
pop <- read_files("pop", "Sum_LANDED") %>%
  mutate(species = "pacific_ocean_perch",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#rockfish
rfish <- read_files("rockfish", "Sum_LANDED") %>%
  mutate(species = "rockfish",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#sablefish
sable <- read_files("sablefish", "Sum_LANDED") %>%
  mutate(species = "sablefish",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#skates
skates <- read_files("skates", "Sum_LANDED") %>%
  mutate(species = "skates",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#spot prawn
prawn <- read_files("spot_prawn", "Sum_prwn_w") %>%
  mutate(species = "spot_prawn",
         units   = "kg") %>%
  rename(harvest = "Sum_prwn_w")
#walleye pollock
pollock <- read_files("walleye_pollock", "Sum_LANDED") %>%
  mutate(species = "walleye_pollock",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#soles (without southern rock sole)
soles <- read_files("soles", "Sum_LANDED") %>%
  mutate(species = "sole",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
#southern rock sole
srsole <- read_files("southernrocksole", "Sum_LANDED") %>%
  mutate(species = "srsole",
         units   = "kg") %>%
  rename(harvest = "Sum_LANDED")
```

Shrimp are trickier - we need to add across columns. So I'm doing this outside the `read_files` function. 

Karen says: 

> "You need to sum columns _pinks thru spny for shrimp catch"

```{r shrimp}

files <- list.files(paste0(dir_anx, '/_raw_data/dfo_khunter/fisheries/d2018/shrimp'), full.names = T)
  
  out <- data.frame()
  
  for(yr in 2007:2015){
    #get the shapefile for yr
    yr_files <- files[str_detect(files, as.character(yr))]
    
    yr_shp_file <- yr_files[str_detect(yr_files, as.character(".shp"))]
    
    yr_shp <- st_read(yr_shp_file, quiet = T) %>%
      as.data.frame() %>%
      mutate(harvest = Sum_pinks + Sum_sdstrp + Sum_prawns + Sum_hmpbck + Sum_docks + Sum_smth_p + Sum_flxd_p +Sum_othr_w + Sum_spny_p) %>%
      select(rgn_name, rgn_id, rgn_code, harvest) %>%
      mutate(year = yr)
    
    out <- rbind(yr_shp, out)
  }
  
shrimp <- out %>%
  mutate(species = "shrimp",
         units   = "kg")
```

Some of these species have multiple stocks within BC. For example, there are two Southern Rock Sole stocks (RSOLE5AB and RSOLEHSTR). But the catch data we have for rock sole does not differentiate between the stocks. So, for these stocks, we can use the proportion of each stock per OHI region. For example, Aristazabal Island has both RSOLEHSTR and RSOLE5AB located in the region. But what propotion of Aristazabal Island region contains one vs the other? We can use the `2_ram_stock_to_ohibc_areas.csv` data which used stock specific shapefiles to calculate proportion of each stock range in each OHI region. 

I only do this for the stocks I know are not BC wide. Otherwise it causes some funkiness later on.

```{r}

stock_rgn_prop <- readr::read_csv("ram/2_ram_stock_to_ohibc_areas.csv") %>%
  mutate(species = 
           case_when(
             stockid %in% c("PCOD5AB", "PCODHS","PCODWCVANI") ~ "pacific_cod",
             stockid %in% c("RSOLE5AB", "RSOLEHSTR") ~ "srsole",
             stockid == "LINGCODSOG" ~ "lingcod",
             stockid %in% c("PERCHQCI", "PERCHWCVANI") ~ "pacific_ocean_perch"
           )) %>%
  select(-a_prop, -tot_area_km2) %>%   #remove these unnecessary columns
  group_by(rgn_name, species) %>%      
  mutate(stock_area = sum(area_km2),           # we want the total area of the species within each BC region and then we calculate
         stock_rgn_prop = area_km2/stock_area) # the proportion of each BC region that contains the stock. This ensures that the proportional area the two (or three) stocks in a region will add up to one

```

Now I'm:  

1. combining all species catch dataframes into one  
2. converting tons and lbs to kilograms  
3. joining with the `stock_rgn_prop` dataframe
4. adding in stockids to link to assessment info  
5. calculating catch of each stock within each region by multiplying catch in each region by the % of that region that contains the stock. This only really affects the species with multiple stocks (cod, perch and sole)  
6. calculate total catch per region per year and then  
7. get the proportional catch for each stock
8. Save as `dfo_catch.csv`

```{r}
species_catch <- af %>%
  rbind(boc, dog, crab, cod, hal, herr, geo, ling, pop, hake, 
        rfish, sable, skates, prawn, pollock, soles, srsole, shrimp, tuna) %>%
  mutate(harvest_kgs = ifelse(units == "lbs", harvest/0.453592,  #pounds to kg
                              ifelse(units == "metric_tons", harvest*1000, harvest))) %>%
  select(-units, -harvest) %>%
  left_join(stock_rgn_prop, by = c("species", "rgn_name", "rgn_id")) %>% #this adds in the calculated prop for multiple stocks but we need to add in the stockids for the other stocks
  mutate(stockid =
           case_when(
             species == "sablefish" ~ "SABLEFPCAN",
             species == "bocaccio" ~ "BOCACCBCW",
             species == "albacore_tuna" ~ "ALBANPAC",
             species == "hake" ~ "PHAKEPCOAST",
             species == "halibut" ~ "PHALNPAC",
             species %in% c("pacific_cod", "pacific_ocean_perch", "lingcod", "srsole") ~ stockid
           ),
          catch = ifelse(!is.na(stock_rgn_prop), harvest_kgs * stock_rgn_prop, harvest_kgs)) %>% # calculate the proportion of each stock harvested in each region. This is only different for species with multiple stocks in BC
  mutate(assessed = ifelse(is.na(stockid), 0, 1)) %>%  #identify stocks that are assessed
  group_by(rgn_name, rgn_id, rgn_code, year, assessed) %>%
  mutate(rgn_catch  = sum(catch)) %>% #this calculates the total catch (in kgs) for each region of either assessed or unassessed species
  ungroup() %>%
  mutate(rgn_ass_catch_prop = ifelse(is.na(stockid), NA, catch/rgn_catch)) %>% ## catch_prop gives the proportion of catch that harvest contributes towards the overall assessed or unassessed catch. We are really only going to use the assessed part of this.
  select(rgn_name, rgn_id, year, species, stockid, catch, rgn_catch, rgn_ass_catch_prop)
  
write_csv(species_catch, "output/dfo_catch.csv")
```

We also need to calculate the total proportion of unassessed catch per region per year. I use a simple (but correct) rule here that if a species does not have a `stockid` then it is not assessed. This works because I've carefully added in only those stockids for stocks that I know are assessed.

```{r}
rgn_catch <- species_catch %>%
  group_by(year, rgn_id, rgn_name) %>%
  summarize(rgn_catch = sum(catch))

unassessed <- species_catch %>%
  filter(is.na(stockid)) %>%  #if a stockid is na it is not assessed. 
  group_by(rgn_name, rgn_id, year) %>%
  summarize(unass_catch = sum(catch))

catch_summary <- rgn_catch %>%
  left_join(unassessed) %>%
  mutate(unass_catch_prop = unass_catch/rgn_catch,
          ass_catch_prop  = 1-unass_catch_prop)

write.csv(catch_summary, "output/rgn_catch_summary.csv")
```

# Plot

I'm going to plot each species catch over time. Sent to Karen to look at. The Geoduck trends look questionable, here is what she says:

> Geoduck operates as a rotational fishery and I understand that catch does fluctuate a lot – but I agree those values are drastic!  The same operation was performed on all years.  I’ll check because maybe you received only a portion of the 2009 data?   There a paper by Khan et al. from the late 2000s that describes the geoduck fishery or you can read the Integrated Fishery Management Plan available online.

> Shrimp catch spiked recently so that sounds right.

```{r plot, fig.width = 12, fig.height = 8}

ggplot(species_catch, aes(x = year, y = catch, color = rgn_name)) +
  geom_line() +
  facet_wrap(~species, scales = 'free')
```


-----

``` {r provenance, results = 'asis'}
prov_wrapup(commit_outputs = FALSE)
```


