---
title: "Calculating FIS, MAR, and SAL weights from DFO's Year in Review reports"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

# Setup

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

dir_git <- '~/github/ohibc'
source(file.path(dir_git, 'src/R/common.R'))  ### an OHIBC specific version of common.R
dir_rgn <- file.path(dir_git, 'prep/regions')  ### github: general buffer region shapefiles
dir_anx <- file.path(dir_M, 'git-annex/bcprep')


### goal specific folders and info
goal      <- 'fis'
scenario  <- 'v2017'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario)
dir_spatial  <- file.path(dir_git, 'prep/_spatial')

### provenance tracking
library(provRmd); prov_setup()

### packages
library(readxl)
```

This script cleans up annual catch data by species for BC and calculates the weights of wild-caught salmon (SAL), other wild-caught fish (FIS), and mariculture (MAR). The data was translated, by hand, from pdf copies of the Year in Review reports as shared by DFO into excel. 

# Clean data

```{r}
catch_xl <- file.path(dir_anx, '_raw_data/DFO_YearinReview/Data_YearinReview_RAW.xlsx')
excel_sheets(catch_xl)
```


## Collect all sheets

Each group of years is located in distinct sheets. We need to read each in, clean them and then aggregate into one long-format dataframe, keeping only the landings and harvest observations, rather than economic value.

```{r}

landings_df <- data.frame()

shts <- excel_sheets(catch_xl)

for(sht in shts) { ### sht <- shts[1]
  
  sheet <- read_xlsx(catch_xl, sheet = sht) 
  
  ### take the first row, which has the years and find the minimum.
  yrs <- slice(sheet, 1) %>% 
           unlist(., use.names = FALSE) %>%
    unique() 
  
  yr_min <- min(yrs, na.rm = TRUE)
  
  ### need to rename columns
  colnames(sheet) <- paste0(colnames(sheet), '_', sheet[1, ])
  colnames(sheet)[1] <- 'species'
  
  df <- sheet %>%
    clean_df_names() %>%
    gather(key = 'variable', value = 'value', -species) %>% 
    mutate(year     = str_extract(variable, '[0-9]{4}'),
           variable = str_replace_all(variable, '_[0-9_]+', '')) %>%
    mutate(species  = tolower(species),
           value    = ifelse(value == '<.1', 0, value),
           value    = as.numeric(value),
           year     = as.numeric(year)) %>%
    filter(!is.na(species)) %>%
    filter(str_detect(variable, 'landings|harvest|value')) %>%
    mutate(xl_sheet = sht)
  ### values are in thousands of tons. since we're getting a prop we don't 
  ### need to multiply by 1000 here. Different years have landings or harvest 
  ### as column names


  ### For most of the sheets we only want to use the earliest year, as we have
  ### more trust in these being true values given the time lag to update them. 
  ### But there are some years where no report was created, including 2007, 
  ### 2008, and 2014. We need to make some exceptions in filtering out these 
  ### data years.
  
  yr_shts <- data.frame()

  if(yr_min %in% c(2004, 2013)) {
    df <- df %>%
      filter(!is.na(year))
  }
  if(yr_min == 2011) {
    df <- df %>%
      filter(year %in% c(2011, 2012))
  }
  if(yr_min %in% c(1998:2003, 2005:2010)){
    df <- df %>%
      filter(year == yr_min)
  }

  ### bind df to landings_df
  landings_df <- bind_rows(df, landings_df)
  
}

DT::datatable(landings_df)

write_csv(landings_df, file.path(dir_goal, 'int/dfo_yr_in_review_totals.csv'))

```

## Identify by FIS, SAL, MAR

We will need to identify which rows are totals vs. individual species, and where each species falls into the three subgoals.

Calculate total catch for each year for the species we're interested in (basically remove salmon and farmed species).

```{r}

landings_df <- read_csv(file.path(dir_goal, 'int/dfo_yr_in_review_totals.csv'))

landings_clean <- landings_df %>%
  mutate(species = str_replace_all(species, '[^a-z]', ''),
         species = str_replace(species, 'cultured', 'farmed'))
  
### Check that target rows are present in all years
spp_yr_df <- landings_clean %>% 
  select(species, year) %>%
  arrange(year) %>%
  group_by(species) %>%
  summarize(yrs = ifelse(sum(year %in% 1998:2015) == 18, 'all', paste0(year, collapse = ', ')))
### * 'herring', 'groundfish', 'farmedsalmon', 'wildsalmon', 
###   'wildshellfish': all years
### * 'farmedshellfish' and 'shellfishfarmed': all years (together)
### 
### 'othersum' sometimes refers to tuna, sardines, and other; sometimes 
### it excludes tuna/sardines.
### * Use 'othernonfarmed' where available (created to sum tuna, sardines, 
###   and other where the original data did not), 
### * otherwise use 'othersum' where the original data included those species.

wild_spp <- c('herring', 'groundfish', 'wildsalmon', 'wildshellfish', 'othersum', 'othernonfarmed')
farm_spp <- c('farmedsalmon', 'farmedshellfish', 'shellfishfarmed', 'otherfarmed')

landings_by_category <- landings_clean %>%
  group_by(year) %>%
  mutate(fis_mar_sal = case_when(
    species %in% farm_spp   ~ 'mar',
    species == 'wildsalmon' ~ 'sal',
    species == 'othersum' & 
      any(str_detect(species, 'othernonfarmed')) ~ NA_character_,
    species %in% wild_spp   ~ 'fis',
    TRUE                  ~ NA_character_)) %>%
  ungroup() %>%
  filter(!is.na(fis_mar_sal) & !is.na(year)) %>%
  mutate(variable = str_replace(variable, 'harvest|landings', 'weight')) %>%
  select(-xl_sheet)

write_csv(landings_by_category, file.path(dir_goal, 'int/dfo_yr_in_review_fis_mar_sal.csv'))

```

## Examine plots of harvest, landed value, wholesale value
``` {r}

landings_by_category <- read_csv(file.path(dir_goal, 'int/dfo_yr_in_review_fis_mar_sal.csv'))
  
landings_prop_df <- landings_by_category %>%
  group_by(year, variable) %>%
  mutate(total = sum(value)) %>%
  group_by(year, variable, fis_mar_sal) %>%
  summarize(total_category = sum(value),
            total = first(total),
            prop_category = total_category / total) %>%
  ungroup()

DT::datatable(landings_prop_df)

write_csv(landings_prop_df, 'output/catch_props_fis_mar_sal.csv')

ggplot(landings_prop_df, aes(x = year, y = prop_category, color = fis_mar_sal)) + 
  geom_line() + 
  theme_bw() + 
  ylim(c(0, NA)) +
  labs(y = 'Proportion of total',
       x = 'Year', 
       title = 'Proportion of weight or value by goal',
       color = 'Goal') +
  facet_grid( . ~ variable)
```




