---
title: 'OHIBC: plot FIS scores by stock and region'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 8, fig.height = 6, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

dir_git <- '~/github/ohibc'
source(file.path(dir_git, 'src/R/common.R'))  ### an OHIBC specific version of common.R
dir_rgn <- file.path(dir_git, 'prep/regions')  ### github: general buffer region shapefiles
dir_anx <- file.path(dir_M, 'git-annex/bcprep')


### goal specific folders and info
goal      <- 'fis'
scenario  <- 'v2017'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario)
dir_spatial  <- file.path(dir_git, 'prep/_spatial')

### provenance tracking
library(provRmd); prov_setup()

### support scripts
source(file.path(dir_git, 'src/R/rast_tools.R')) 
source("~/github/ohibc/calc_ohibc/calc_scores_fxns.R")
  ### raster plotting and analyzing scripts
```

# Summary

This script takes the output from `functions.R` FIS model and plots individual stock scores by year and relative contribution to score.

## Stock scores and overall scores by region

``` {r assemble_dataframe}

## plotting fishery catch weighting by region
stock_plot_df <- read_csv(file.path(dir_goal, 'summary/fis_from_functions.csv')) %>%
  filter(!is.na(score)) %>%
  group_by(region_id, year) %>%
  mutate(total_catch = sum(rgn_catch),
         rgn_catch_pct = rgn_catch / total_catch,
         total_score = sum(score * rgn_catch) / total_catch) %>%
  ungroup() %>%
  left_join(get_rgn_names(), by = c('region_id' = 'rgn_id')) %>%
  filter(rgn_catch > 0) %>%
  filter(year > 1990)

```

``` {r plot_fis_scores}

ggplot(stock_plot_df, aes(x = year, y = score)) +
    ggtheme_plot() +
    geom_line(aes(group = stock_id, color = stock_id,
                  size = rgn_catch_pct),
              lineend = 'round', alpha = .5) +
    labs(color = 'Stock ID',
         y     = 'Stock Score') +
    guides(colour = guide_legend(override.aes = list(size = 3)),
           size = 'none') +
    facet_wrap(~rgn_name)

```

## Proportion of hake vs non hake catch over time

```{r}
hake <- stock_plot_df %>%
  mutate(hake = ifelse(stock_id == "PHAKEPCOAST", "Hake", "Non-hake")) %>%
  group_by(hake, year) %>%
  summarize(catch = sum(rgn_catch)) %>%
  filter(year>1997)

ggplot(hake, aes(x=year, y=catch, fill = hake)) +
  geom_bar(stat="identity") +
  ylab("Catch (tons)") +
  xlab("Year") +
  labs(fill='')
```


## Catch by stock and region

``` {r plot_catch}

ggplot(stock_plot_df, aes(x = year, y = rgn_catch)) +
    ggtheme_plot() +
    geom_line(aes(group = stock_id, color = stock_id),
              alpha = .5, size = 2) +
    labs(color = 'Stock ID',
         y     = 'Stock reported catch (tons)') +
    guides(colour = guide_legend(override.aes = list(size = 3)),
           size = 'none') +
  facet_wrap(~rgn_name )

```

### Without Hake

``` {r plot_catch_no_hake}

ggplot(filter(stock_plot_df, stock_id!="PHAKEPCOAST"), aes(x = year, y = rgn_catch)) +
    ggtheme_plot() +
    geom_line(aes(group = stock_id, color = stock_id),
              alpha = .5, size = 2) +
    labs(color = 'Stock ID',
         y     = 'Stock reported catch (tons)') +
    guides(colour = guide_legend(override.aes = list(size = 3)),
           size = 'none') +
  facet_wrap(~rgn_name)

```

## Fisheries status scores

I copied in the FIS function from functions.R and made small changes to read in the layers directly from the output folder in the fis/v2017 folder rather than layers.csv. 

```{r}
# 
# FIS <- function(layers) {
# 

  ##### Gather parameters and layers #####
  ### * ram_b_bmsy, ram_f_fmsy, ram_catch
  ### * rgn_stock_wt_uniform, _saup, _dfo
  ### * ram_dfo_saup_lookup.csv

  # status_year    <- layers$data$scenario_year
  # data_year      <- status_year
  # status_yr_span <- layers$data$status_year_span

  ram_b_bmsy        <- read_csv("output/ram_b_bmsy.csv") %>%
    rename(b_bmsy = value) %>%
    select(-param)
  ram_f_fmsy        <- read_csv("output/ram_f_fmsy.csv") %>%
    rename(f_fmsy = value) %>%
    select(stock_id, year, f_fmsy)
  hake_catch        <- read_csv("output/hake_catch.csv")   #hake catch from DFO YOR
  ram_catch         <- read_csv("output/ram_catch.csv") %>%
    rename(catch = value) %>%
    select(stock_id, year, catch) %>%
    filter(stock_id != "PHAKEPCOAST") %>%     #we need to replace RAM Hake catch with catch from DFO
    bind_rows(hake_catch)
  rgn_stock_area  <- read_csv("output/ram_stock_area.csv") %>%
    rename(region_id = rgn_id, stock_id = stockid)
  dfo_assessed_catch <- read_csv("output/dfo_assessed_catch_props.csv") %>%
    select(year, penalty = prop_land_ass)
  
  status_yr_span <- c(2001:2016)
  
ggplot(dfo_assessed_catch, aes(x = year, y = penalty, group = 1)) +
  geom_line() +
  theme_bw() +
  ylab("Penalty (equal to % of catch assessed)") +
  ggtitle("Proportion of total wild, non-salmon, catch assessed")
```

We see a large dip in later years because many species only have stock assessments up until 2010-2014. Sablefish assessments end at 2010, Tuna and Bocaccio end in 2012, Ocean Perch ends in 2013, Soles, Halibut and Pacific and Ling cod end in 2014. Pacific Hake is the only stock with assessments through 2016. 

```{r}

ts_ram <- ram_b_bmsy %>%
  filter(year > 1997)

ggplot(ts_ram, aes(x = year, y = b_bmsy, color = stock_name)) +
  geom_line() +
  theme_bw()+
  annotate("rect", xmin = 2010, xmax = 2016, ymin = 0, ymax = 4.2,
        alpha = .2) +
  ylab("B/Bmsy")

```

```{r}
  ### These parameters are based on conversation with Ian Perry, Karen Hunter,
  ### and Karin Bodtker on May 24 2017.
  b_bmsy_underexploit_penalty <- 0.25
  b_bmsy_underexploit_thresh  <- 3.00
  f_fmsy_underfishing_penalty <- 0.25
  f_fmsy_overfishing_thresh   <- 2.00

  ### Apply rolling mean to F/Fmsy
  ram_f_fmsy <- ram_f_fmsy %>%
    mutate(f_fmsy_raw = f_fmsy) %>%
    arrange(stock_id, year) %>%
    group_by(stock_id) %>%
    mutate(f_fmsy = zoo::rollmean(f_fmsy_raw, k = 4, align = 'right', fill = NA)) %>%
    ungroup()

  stock_status_layers <- ram_b_bmsy %>%
    full_join(ram_f_fmsy, by = c('year', 'stock_id'))

  ########################################################.
  ##### run each fishery through the Kobe plot calcs #####
  ########################################################.
  ### * ram_b_bmsy, ram_f_fmsy


  ### Function for converting B/Bmsy values into a 0 - 1 score
  rescale_bprime_crit <- function(fish_stat_df,
                                  bmax, bmax_val) {

    ### parameter from DFO harvest control rule:
    overfished_th  <- 0.8
    ### parameter from OHI California Current:
    underfished_th <- 1.5

    bmax_adj <- (bmax - underfished_th) / (1 - bmax_val) + underfished_th
    ### this is used to create a "virtual" B/Bmsy max where score drops
    ### to zero.  If bmax_val == 0, this is bmax; if bmax_val > 0, bmax_adj
    ### extends beyond bmax, to create a gradient where bmax_val occurs at bmax.

    fish_stat_df <- fish_stat_df %>%
      # group_by(stock) %>% ### grouping by stock will set b_max by max per stock, instead of max overall
      mutate(b_max     = max(b_bmsy, na.rm = TRUE)) %>%
      ungroup() %>%
      mutate(bPrime = NA,
             bPrime = ifelse(b_bmsy < overfished_th,  ### overfished stock
                             b_bmsy / overfished_th,
                             bPrime),
             bPrime = ifelse(b_bmsy >= overfished_th & b_bmsy < underfished_th,
                             1,                       ### appropriately fished stock
                             bPrime),
             bPrime = ifelse(b_bmsy >= underfished_th,
                             (bmax_adj - b_bmsy) / (bmax_adj - underfished_th), ### underfished stock
                             bPrime),
             bPrime = ifelse(bPrime < 0, 0, bPrime))
    return(fish_stat_df)
  }


  ### Function to create vertical gradient based on distance from
  ### ideal F/Fmsy value to actual F/Fmsy value
  f_gradient <- function(f, over_f, under_f, fmax, fmin_val) {
    x <- ifelse(f < over_f & f > under_f, 1, NA)
    x <- ifelse(f <= under_f, (f * (1 - fmin_val) / under_f + fmin_val), x)
    x <- ifelse(f >= over_f,  (fmax - f) / (fmax - over_f), x)
    x <- ifelse(f > fmax, NA, x)
    return(x)
  }

  ### Function to convert F/Fmsy values into 0 - 1 score
  rescale_fprime_crit <- function(fish_stat_df,
                                  fmax, fmin_val) {

    ### params from DFO harvest control rule:
    Bcrit <- 0.4; overfished_th <- 0.8
    ### params from OHI California Current:
    underfishing_th <- 0.8; overfishing_th  <- 1.2

    bcritslope = 1 / (overfished_th - Bcrit)
    ### connecting from (Bcrit, 0) to (overfished_th, 1)

    fish_stat_df <- fish_stat_df %>%
      mutate(fPrime = ifelse(b_bmsy < overfished_th & f_fmsy < fmax,
                             f_gradient(f_fmsy + (overfished_th - b_bmsy) * bcritslope,
                                        over_f = overfishing_th,
                                        under_f = underfishing_th,
                                        fmax = fmax,
                                        fmin_val = fmin_val),
                             NA),
             fPrime = ifelse(b_bmsy >= overfished_th & f_fmsy < fmax,
                             f_gradient(f_fmsy,
                                        over_f = overfishing_th,
                                        under_f = underfishing_th,
                                        fmax = fmax,
                                        fmin_val = fmin_val),
                             fPrime),
             fPrime = ifelse(is.na(fPrime), 0, fPrime), ### fill zeros everywhere unscored
             fPrime = ifelse(is.na(f_fmsy), NA, fPrime) ### but if no f_fmsy, reset to NA
      )
    return(fish_stat_df)
  }

  stock_status_df <- stock_status_layers %>%
    rescale_bprime_crit(bmax     = b_bmsy_underexploit_thresh,
                        bmax_val = b_bmsy_underexploit_penalty) %>%
    rescale_fprime_crit(fmax     = f_fmsy_overfishing_thresh,
                        fmin_val = f_fmsy_underfishing_penalty) %>%
    mutate(x_prod = ifelse(!is.na(fPrime), (fPrime * bPrime), bPrime),
           basis  = ifelse(!is.na(fPrime), 'F_Fmsy, B_Bmsy', 'B_Bmsy only')) %>%
    dplyr::select(year, stock_id,
                  score = x_prod,
                  basis,
                  bPrime, fPrime,
                  b_bmsy, f_fmsy)  %>%
    group_by(stock_id) %>%
    complete_years(status_yr_span, method = 'carry', dir = 'forward') %>%
    ungroup()
  
ggplot(filter(stock_status_df, year > 1994), aes(x = year, y = score, color = stock_id)) +
  geom_line() +
  theme_bw() +
  ylab("Stock Score") +
  ggtitle("Stock scores based on F' and/or B'")
```

This plot shows all stocks are given a stock score for every year, even beyond their last assessment. Each stock goes to 2016. This might be an argument for extending our assessment penalty to 2016, basically assuming all stocks assessed in 2010 are assessed from 2010 - 2016...

### Evaluating penalties

Here are the scores with and without the penalty factor. The trends are very similar across time and regions since the penalty is simply a proportional decrease to the score, but it does change the story in, for example, 2000 and 2010.

```{r}
  ##############################################################.
  ##### calculate distribution of fishery catch to regions #####
  ##############################################################.

  ### calculate weights within each region by regional catch
  catch_df <- ram_catch %>%
    left_join(rgn_stock_area, by = 'stock_id') %>%
    group_by(region_id, stock_id) %>%
    complete_rgn_years(status_yr_span, method = 'carry', dir = 'forward') %>%
    ungroup() %>%
    mutate(rgn_catch = catch * a_prop,
           rgn_catch = ifelse(is.na(rgn_catch), 0, rgn_catch))

  stock_score_df <- stock_status_df %>%
    filter(!is.na(score)) %>%
    group_by(stock_id) %>%
    arrange(stock_id, year) %>%
    fill(score, .direction = c('down')) %>%
    fill(score, .direction = c('up')) %>%
    ungroup() %>%
    full_join(catch_df, by = c('stock_id', 'year')) %>%
    select(region_id, year, stock_id, score, rgn_catch, basis) %>%
    filter(!is.na(region_id))

  # if(data_year == max(status_yr_span)) {
  #   ### note, this contains all years, but only write it once
  #   write_csv(stock_score_df, '~/github/ohibc/prep/fis/v2017/summary/fis_from_functions.csv')
  # }

  score_df <- stock_score_df %>%
    filter(!is.na(score)) %>%
    group_by(region_id, year) %>%
    summarize(total_catch = sum(rgn_catch),
              n_stocks    = sum(rgn_catch > 0),
              ass_score = sum(score * rgn_catch) / total_catch) %>% #score based on assessed stocks
    left_join(dfo_assessed_catch) %>%
    filter(!is.na(penalty)) %>%
    rowwise() %>%
    mutate(unassess_score =  ass_score * penalty, #score based on unassessed stocks (penalty factor * assessed)
           total_score = (ass_score + unassess_score)/2) %>%
    ungroup()

  rgns <- get_rgn_names()
  
  fis_status <- score_df %>%
    select(region_id = region_id, year, ass_score, total_score) %>%
    mutate(no_penalty = ass_score * 100,
           score_w_penalty = total_score * 100,
           goal      = 'FIS',
           dimension = 'status') %>%
    left_join(rgns, by = c("region_id" = "rgn_id"))

  
nopen <- ggplot(fis_status, aes(x = year, y = no_penalty, color = rgn_name)) +
  geom_line() +
  theme_bw() +
  ggtitle("Region fisheries goal status \nwithout a penalty") +
  ylab("Status") +
  theme(legend.position="none")

pen <- ggplot(fis_status, aes(x = year, y = score_w_penalty, color = rgn_name)) +
  geom_line() +
  theme_bw() +
  ggtitle("Region fisheries goal status \nwith a penalty") +
  ylab("Status") +
  theme(legend.position="none")

cowplot::plot_grid(nopen, pen)

```

-----

``` {r provenance, results = 'asis'}
prov_wrapup(commit_outputs = FALSE)
```
