---
title: 'OHIBC: Goal prep for wild-capture fisheries'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(raster)
library(rgdal)

dir_git <- '~/github/ohibc'
source(file.path(dir_git, 'src/R/common.R'))  ### an OHIBC specific version of common.R
dir_spatial <- file.path(dir_git, 'prep/spatial')  ### github: general buffer region shapefiles
dir_anx     <- file.path(dir_M, 'git-annex/bcprep')


### goal specific folders and info
goal      <- 'fis'
scenario  <- 'v2017'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario) 
dir_goal_anx <- file.path(dir_anx, goal, scenario)

### provenance tracking
library(provRmd); prov_setup()

### Kobe plot functions
source(file.path(dir_goal, 'kobe_fxns.R'))

### set up proj4string options: BC Albers and WGS84
p4s_wgs84 <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
p4s_bcalb <- '+proj=aea +lat_1=50 +lat_2=58.5 +lat_0=45 +lon_0=-126 +x_0=1000000 +y_0=0 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0'

```

# Summary

Calculate annual status for wild-capture fisheries (food provision) goal for OHIBC.  

* Each fishery will be scored based upon B/Bmsy and F/Fmsy similar to the modified Kobe plot from OHI California Current.  
    * B/Bmsy and F/Fmsy values will be taken from the updated RAM database with modeled scores through 2013.
    * F/Fmsy values are smoothed using a four-year rolling average: status year and prior three years.
    * For stocks without B/Bmsy and/or F/Fmsy values in RAM, e.g. salmon, other methods of determining stock status?
* For each region, fishery scores will be combined based upon proportion of the contribution of each fishery's catch to the overall catch for the region.  
    * In OHICC, weighting was based on average catch over all data years.  For OHIBC, weighting will be based on smoothed catch from SAUP reconstructions, smoothed using a four-year rolling average.

***

# Data Source [NOTE: can be copied from README.md in rawdata file]
**Reference**: [citation for source data; website, literature, contact information. Version of data (if relevant). Screenshots if a series of menus had to be navigated to obtain the data.]

**Downloaded**: [date downloaded or received]

**Description**:  [e.g., surface aragonite state]

**Native data resolution**: [e.g., 1 degree, 30 m, etc.]   

**Time range**: [e.g., 1880-1899, monthly data provided for each year] 

**Format**:  [e.g. NetCDF]

***
  
# Methods

## Create time series of stock status scores

Apply modified Kobe plot calculations to RAM time series. Parameters used to calculate scores (B' and F') based on B/Bmsy and F/Fmsy:

* `b_bmsy_max` is the B/Bmsy for a generalized unfished stock.
* `bmax_val` is the rescaled B' score at a B/Bmsy value of `b_bmsy_max`.  This allows for softening the penalty to underfished stocks by assigning a non-zero value (between 0-1) to an unfished stock.
* `f_fmsy_max` is the highest allowable fishing pressure, beyond which a stock's F' score is zero.
* `fmin_val` is rescaled F' score at an F/Fmsy of zero (no fishing pressure).  This allows for softening the penalty to underfishing by assigning a non-zero value (between 0-1).
* underfished/overfished and underfishing/overfishing thresholds provide acceptable windows around F/Fmsy = 1 and B/Bmsy = 1.  
* `Bcrit` provides a fisheries management threshold between "cautious" and "critical" stocks, while `Bcautious` provides threshold between "cautious" and "healthy" stocks.

In this formulation, the stock score is the product of F' and B' for that stock in that year.

$$x_{fis} = \displaystyle\sum_{i=1}^{n}w_i*(F' * B')$$

* To calculate $B'$, we use DFO-defined categories in which a given stock is considered underfished, overfished, or healthy.  These are based on $B/B_{MSY}$ thresholds of $B_{overfished}$ = 0.8 and $B_{underfished}$ = 1.5.  For these calculations, we have assumed a $(B/B_{MSY})_{max}$ for an unfished stock to be 3.0 (i.e. $B/B_{MSY}$ is 33% of unfished biomass).

| $B'$ calculation        | Stock condition                                         |
| :---------------------: | :------------------------------------------------------ |
| $B' = \frac{B/B_{MSY}}{0.8}$ | Overfished ($B/B_{MSY} < B_{overfished}$)          |
| $B' = 1$                | Healthy ($B_{cautious} <= B/B_{MSY} < B_{underfished}$) |
| $B' = \frac{(B/B_{MSY})_{max} - (B/B_{MSY})}{(B/B_{MSY})_{max} - 1.5}$ | Underfished ($B/B_{MSY} >= B_{underfished}$) |

* To calculate F', we examine fishing pressure within three regimes defined by DFO fisheries management: critical stock ($B/B_{MSY} < B_{critical}$), cautious stock ($B_{cautious} <= B/B_{MSY} < B_{cautious}$) and healthy stock ($B/B_{MSY} >= B_{cautious}$).  Within each regime we divide fishing pressure into overfishing, underfishing, and appropriate fishing pressures based on $F/F_{MSY}$, including under- and overfishing $F/F_{MSY}$ thresholds of $F_{underfishing}$ = 0.8 and $F_{overfishing}$ = 1.2, respectively.  

| $F'$ calculation        | Stock condition    | Fishing pressure                       |
| :---------------------: | :----------------- | :---------------------------           |
| $F' = 0$                | critical           | any                                    |
| $F' = \frac{(B/B_{MSY} + 1.5) - F/F_{MSY}}{1.5}$ | Cautious | overfishing ($B/B_{MSY} + 0.2 < F/F_{MSY} <= B/B_{MSY} + 1.5$) |
| $F' = 1$                | cautious         | Appropriate fishing ($B/B_{MSY} - 0.2 <= F/F_{MSY} < B/B_{MSY} + 0.2$) |
| $F' = \frac{F/F_{MSY}}{B/B_{MSY} - 0.2}$  | Overfished | Underfishing ($F/F_{MSY} < B/B_{MSY} - 0.2$) |
| $F' = \frac{F/F_{MSY}}{\psi_{UF}}$ | healthy | Underfishing ($F/F_{MSY} < \psi_{UF}$)
| $F' = 1$                | healthy   | Appropriate fishing ($\psi_{UF} <= F/F_{MSY} < \psi_{OF}$) |
| $F' = \frac{(F/F_{MSY})_{max} - F/F_{MSY}}{(F/F_{MSY})_{max} - \psi_{OF}}$  | healthy | Overfishing ($F/F_{MSY} >= \psi_{OF}$) |

``` {r rescore_stock_status}

stocks_ts <- read_csv(file.path(dir_goal, 'ram/stocks_ram_timeseries.csv'))

stocks_ts <- stocks_ts %>%
  rename(b_bmsy = ts_bbmsy,
         f_fmsy = mean_ffmsy,
         year   = tsyear) %>%
  rescale_bprime_crit(overfished_th = 0.8,
                      underfished_th = 1.5,
                      bmax = 3.0,
                      bmax_val = .25) %>%
  rescale_fprime_crit(Bcrit = 0.4,
                      Bcautious = 0.8,
                      underfishing_th = 0.8,
                      overfishing_th = 1.2,
                      fmax  = 2.0,
                      fmin_val = .25) %>%
  mutate(x_prod = (fPrime * bPrime)) %>%
  dplyr::select(year, stockid, stocklong, areaid,
         score = x_prod,
         b_bmsy, f_fmsy)

write_csv(stocks_ts, file.path(dir_goal, 'int/stock_scores_ts.csv'))

```

## Determine spatial allocation of RAM stocks to SAUP cells

Stock assessments from RAM database are spatially differentiated by either the entire BC coast or subregions within the BC EEZ: Hecate Strait, Strait of Georgia, Queen Charlotte Islands, West Coast Vancouver Island, etc.  These RAM DFO regions are outlined in a shapefile (ram/ram_dfo_areaid.shp) which is used to extract SAUP cell IDs.  The `dfo_rgn_weight` value is the proportion of DFO region area to ocean area for the cell.

``` {r ram_rgns_to_saup_cells}

saup_bc_rast <- raster::raster(file.path(dir_goal, 'saup/saup_bc_rast.tif'))

eez_poly <- readOGR(dsn = dir_spatial %>% path.expand,
                    layer = 'ohibc_eez') %>%
  spTransform(crs(saup_bc_rast))
  
cell_area_df <- raster::extract(saup_bc_rast, eez_poly,
                                  weights = TRUE,
                                  normalizeWeights = FALSE) %>%
  lapply(FUN = function(x) as.data.frame(x, stringsAsFactors = FALSE)) %>%
  bind_rows() %>%
  rename(cell_id = value, ocean_pct = weight)

ram_dfo_rgn <- readOGR(dsn   = path.expand(file.path(dir_goal, 'ram')),
                       layer = 'ram_dfo_areaid') %>%
  spTransform(crs(saup_bc_rast))

### the dfo area ID polygons need to be clipped to just ocean area.
# x <- raster::intersect(ram_dfo_rgn, eez_poly)
# writeOGR(x, 
#          dsn   = path.expand(file.path(dir_goal, 'ram')),
#          layer = 'ram_dfo_areaid',
#          driver = 'ESRI Shapefile',
#          overwrite_layer = TRUE) 

ram_rgn_to_saup_list <- raster::extract(saup_bc_rast, ram_dfo_rgn, 
                                     weights = TRUE,
                                     normalizeWeights = FALSE) %>%
  lapply(FUN = function(x) as.data.frame(x, stringsAsFactors = FALSE)) %>%
  setNames(tolower(ram_dfo_rgn@data$areaid))

### weight is how much of cell is in RAM DFO region; but this also includes area
### lost to land, not just to other regions.  Normalize cell weights
### to total *ocean* area
ram_rgn_to_saup <- ram_rgn_to_saup_list %>%
  bind_rows(.id = 'areaid') %>%
  rename(cell_id = value) %>%
  left_join(cell_area_df, by = 'cell_id') %>%
  mutate(dfo_rgn_weight = weight / ocean_pct) %>%
  select(-weight)

write_csv(ram_rgn_to_saup, file.path(dir_goal, 'ram/ram_rgn_to_saup.csv'))

```

Using a manually-matched lookup table of Sea Around Us taxa to RAM Database stocks, match RAM stock assessments to SAUP catch values.

The method of dividing catch across OHIBC regions and DFO regions at the same time: 

* multiply the catch in a given cell by the proportion of the cell falling within a DFO region.  This assumes equal distribution of catch within a cell. 
    * For catch divided across multiple spatial stocks, e.g. Hecate Strait and Queen Charlotte Islands, this assumes that the stock boundaries do not overlap.  If the stock boundaries overlap, there could be double-counting.
* Multiple the DFO-region-weighted catch by the proportion of cell falling within an OHIBC region.
    * This results in the `stock_scores_by_cell` dataframe
    * This is saved in `int/stock_scores_by_cell.csv`
* For each fishery and year, sum the DFO rgn and OHIBC rgn weighted catch across all cells within the OHIBC region.
    * This results in the `stock_scores_by_rgn` dataframe
    * This is saved in `int/stock_scores_by_rgn.csv`

``` {r connect_ram_to_saup}

saup_to_ram_ids <- read_csv(file.path(dir_goal, 'raw/saup_to_ram_ids.csv')) %>%
  filter(!is.na(saup_taxonkey) & !is.na(ram_stockid))

ram_stock_scores <- read_csv(file.path(dir_goal, 'int/stock_scores_ts.csv')) %>%
  filter(!is.na(f_fmsy))

saup_catch_ts <- read_csv(file.path(dir_anx, 'fis/v2017/saup/saup_bc_clean.csv')) %>%
  group_by(year, cell_id, taxonkey, taxonsciname, taxoncomname, rgn_id, weight) %>%
  summarize(cellcatch = sum(allocatedcatch)) %>%
  ungroup()

ram_rgn_to_saup <- read_csv(file.path(dir_goal, 'ram/ram_rgn_to_saup.csv'))

stock_scores_by_cell <- saup_catch_ts %>%
  inner_join(saup_to_ram_ids %>% 
              dplyr::select(taxonkey = saup_taxonkey, stockid = ram_stockid, areacode = ram_areacode),
            by = 'taxonkey') %>%
  inner_join(ram_stock_scores, by = c('year', 'stockid')) %>%
  inner_join(ram_rgn_to_saup, by = c('cell_id', 'areacode' = 'areaid'))

stock_scores_by_rgn <- stock_scores_by_cell %>% 
  mutate(catch = cellcatch * weight * dfo_rgn_weight) %>%
    ### this is not strictly spatially accurate.  
  dplyr::select(year, cell_id, rgn_id, 
           stockid, taxonkey, stocklong, 
           score, catch) %>%
  group_by(year, rgn_id, stockid, taxonkey, score) %>%
  summarize(rgn_catch = sum(catch))

write_csv(stock_scores_by_cell, file.path(dir_goal, 'int/stock_scores_by_cell.csv'))
write_csv(stock_scores_by_rgn,  file.path(dir_goal, 'int/stock_scores_by_rgn.csv'))

```

From `stock_scores_by_rgn`, find the catch-weighted mean across all stocks by year.  This will result in the mean FIS status per region.

``` {r calculate_FIS}

rgn_names <- read_csv(file.path(dir_spatial, 'rgn_names.csv'))

stock_scores_by_rgn <- read_csv(file.path(dir_goal, 'int/stock_scores_by_rgn.csv')) %>%
  filter(year >= 1990) %>%
  group_by(year, rgn_id) %>%
  mutate(rgn_weight = rgn_catch/sum(rgn_catch)) %>%
  ungroup()

scores_by_rgn <- stock_scores_by_rgn %>%
  group_by(year, rgn_id) %>%
  summarize(status = sum(score * rgn_catch) / sum(rgn_catch)) %>%
  left_join(rgn_names, by = 'rgn_id') %>%
  ungroup()

status_plot <- ggplot(scores_by_rgn, aes(x = year, y = status)) + 
  geom_line(data = stock_scores_by_rgn, aes(x = year, y = score, 
                                     group = stockid, color = stockid,
                                     size = rgn_weight),
            lineend = 'round', alpha = .5) +
  scale_size_continuous(guide = FALSE) +
  geom_line(size = 1.5)  + 
  trelliscope::facet_wrap(~ rgn_name)

stock_scores <- stock_scores_by_rgn %>%
  filter(rgn_id == 1)

scores <- scores_by_rgn %>%
  filter(rgn_id == 1)

status_plot <- ggplot(scores, aes(x = year, y = status)) + 
  geom_line(data = stock_scores, aes(x = year, y = score, 
                                     group = stockid, color = stockid,
                                     size = rgn_weight),
            lineend = 'round', alpha = .5) +
  scale_size_continuous(guide = FALSE) +
  geom_line(size = 1.5)  +
  labs(title = scores$rgn_name[1])

print(status_plot)

```
***

``` {r prov_footer, results = 'asis'}
prov_wrapup(commit_outputs = FALSE)
```

