---
title: 'OHIBC: CW chemical layers prep'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(rgdal)
library(raster)
source('~/github/ohibc/src/R/common.R')  ### an OHIBC specific version of common.R

dir_git     <- '~/github/ohibc'
dir_spatial <- file.path(dir_git, 'prep/spatial')  ### github: general buffer region shapefiles
dir_anx     <- file.path(dir_M, 'git-annex/bcprep')

### goal specific folders and info
goal     <- 'cw'
scenario <- 'v2016'
dir_goal <- file.path(dir_git, 'prep', goal, scenario)
dir_goal_anx <- file.path(dir_anx, goal, scenario)

### provenance tracking
library(provRmd); prov_setup()

### support scripts
source(file.path('~/github/ohibc/src/R/map_scores.R'))
  ### score plotting scripts
source(file.path(dir_git, 'src/R/rast_tools.R')) 
  ### raster plotting and analyzing scripts

### set up the default BC projection to be BC Albers
p4s_bcalb <- c('bcalb' = '+init=epsg:3005')

```

#Summary

These data were prepared for the global 2016 assessment.  For each layer, we are simply reprojecting the data to the BC EEZ and rescaling to the local 99.99th percentile.

The Status of chemical pollution was measured via four global datasets: land-based organic from agricultural pesticide use (reported to FAO), inorganic pollution from runoff from impervious surfaces, and ocean-based pollution from commercial shipping and ports.

The shipping, ports and harbors, and inorganic pollution have not changed over time.  However, the organic (pesticide) pollution is updated yearly, with the most recent data from 2013.  The raw shipping, ports and harbors, and inorganic pollution data are located [here](https://knb.ecoinformatics.org/#view/doi:10.5063/F1S180FS).

The most recent land-based organic data is on a NCEAS server: Mazu:marine_threats/impact_layers_2013_redo/impact_layers/work/land_based/before_2007/raw_global_results/

***

# Data Sources

* Land based organics from pesticides
    * __Reference__: 
    * __Downloaded__: accessed from Mazu
    * __Native Data Resolution__: Mollweide 934 m
    * __Values__: 
    * __Time Range__: 2002 - 2013
    * __Format__: GeoTIFF
* Ocean-based from shipping, ports, harbors
    * __Reference__: 
    * __Downloaded__:  accessed from Mazu
    * __Native Data Resolution__: Mollweide 934 m
    * __Values__:  Pollution pressures 0 - 1, log-transformed and rescaled to 99.99%ile
    * __Time Range__: NA
    * __Format__: GeoTIFF
* Land-based inorganics from impervious surfaces
    * __Reference__: 
    * __Downloaded__:  accessed from Mazu
    * __Native Data Resolution__: Mollweide 934 m
    * __Values__: Pollution pressures 0 - 1, log-transformed and rescaled to 99.99%ile
    * __Time Range__: NA
    * __Format__: GeoTIFF

***

# Methods

## Process individual layers

### Pesticides

These data cover a time period of 2002-2013.  Each layer is read in, reprojected to BC Albers at 1 km resolution and clipped to OHIBC EEZ.  Then each layer is log-transformed (log(x+1)) and then rescaled to the 99.99%ile (rescaled) value found across all years to create layer values from zero to 1.

#### Read in data

``` {r read_and_clip_pesticide_raster}

poly_eez_base <- readOGR(path.expand(dir_spatial), 'ohibc_rgn')
rast_base <- raster(file.path(dir_spatial, 'raster/ohibc_rgn_raster_1000m.tif')) %>%
  crop(poly_eez_base)

### Pesticide pressures layers are on Mazu here:
### Mazu:marine_threats/impact_layers_2013_redo/impact_layers/work/land_based/before_2007/raw_global_results

dir_pest <- file.path(dir_M, 'marine_threats/impact_layers_2013_redo/impact_layers',
                      'work/land_based/before_2007',
                      'raw_global_results')

pest_files <- list.files(dir_pest, pattern = '*pest*', full.names = TRUE)
### basename: raw_global_results/global_plumes_pest_2002_raw.tif

new_rast_files <- file.path(dir_goal_anx, 'tmp', basename(pest_files)) %>%
    str_replace('global_plumes', 'cw')

reload <- FALSE
if(any(!file.exists(new_rast_files)) | reload) { # i = 1
  for (i in 1:length(pest_files)) {
    message('Processing ', basename(pest_files[i]))
    raw_rast <- raster(pest_files[i])
    new_rast_file <- new_rast_files[i]
    new_rast <- raster::projectRaster(raw_rast, rast_base,
                          overwrite = TRUE)
    values(new_rast)[is.na(values(rast_base))] <- NA
    writeRaster(new_rast, new_rast_file, overwrite = TRUE)
  }
  
} else {
  ### force git provenance registration
  git_prov(pest_files, filetype = 'input')
  git_prov(new_rast_files, filetype = 'output')
}

```

#### Log Transform pesticides

```{r log}

pest_layers_files <- list.files(file.path(dir_goal_anx, 'tmp'), pattern = '*pest*', full.names = TRUE)
pest_log_files <- file.path(dir_goal_anx, 'int', str_replace(basename(pest_layers_files), '.tif', '_log.tif'))

pest_stack <- stack(pest_layers_files)

pest_log <- calc(pest_stack, fun = function(x) {log(x + 1)}) %>%
  setNames(basename(pest_log_files) %>% str_replace('.tif', ''))

writeRaster(pest_log, filename = pest_log_files, bylayer = TRUE, overwrite = TRUE)

plot(pest_log)

```

#### Rescale pesticides to 99.99th quantile

```{r quantile}

pest_log_files <- list.files(file.path(dir_goal_anx, 'int'), 
                             pattern = '*pest*',
                             full.names = TRUE)
pest_log_files <- pest_log_files[str_detect(pest_log_files, 'log')]

pest_rescale_files <- file.path(dir_goal_anx, 'rescaled_layers',
                                basename(pest_log_files) %>%
                                  str_replace('raw_log', 'rescaled'))

pest_log <- stack(as.list(pest_log_files))

pest_ref <- quantile(pest_log, prob = 0.9999) %>% 
  max()
  ### quantile returns vector of 99.99%iles by layer; find max over all layers

pest_rescale <- calc(pest_log, fun = function(x) {
    ifelse(x > pest_ref, 1, x / pest_ref)
  }) %>%
  setNames(basename(pest_rescale_files) %>% str_replace('.tif', ''))

writeRaster(pest_rescale, filename = pest_rescale_files, 
            bylayer = TRUE,
            overwrite = TRUE)

plot(pest_rescale)

```


### Ocean pollution

Data are already log-transformed and normalized.  Here we reproject and clip to just OHIBC EEZ, and rescale to a new max value based on highest pressure within OHIBC EEZ.

``` {r ocean pollution and land-based inorganics}
## ocean pollution (shipping and ports) and land-based inorganic pollution layers
# only one ocean pollution raster for both time periods (so only normalized by one time period)

dir_threats <- file.path(dir_M, 
                         'marine_threats/impact_layers_2013_redo/impact_layers',
                         'final_impact_layers/threats_2013_final')

ocean_pollution <- raster(file.path(dir_threats, 
                          'normalized_by_one_time_period/ocean_pollution.tif'))

op_bcalb <- projectRaster(ocean_pollution, rast_base)
values(op_bcalb)[is.na(values(rast_base))] <- NA ### only OHIBC EEZ

### max is 0.8973; re-rescale to new 99.99%ile?
op_newref <- quantile(op_bcalb, prob = 0.9999)

op_bcalb_rescale <- calc(op_bcalb, fun = function(x) {
    ifelse(x > op_newref, 1, x / op_newref)
  }) %>%
  setNames('ocean_pollution')

writeRaster(op_bcalb_rescale, 
            file.path(dir_goal_anx, 'rescaled_layers', 'cw_ocean_poll_rescaled.tif'),
            overwrite = TRUE)

```

###  Land-based inorganics

Data are already log-transformed and normalized.  Here we reproject and clip to just OHIBC EEZ, and rescale to a new max value based on highest pressure within OHIBC EEZ.

``` {r inorganic land-based pollutants}

# two rasters for inorganic pollution (2003-2006 and 2007-2010)
# I used the 2007-2010 raster (normalized by both time periods):

dir_threats <- file.path(dir_M, 
                      'marine_threats/impact_layers_2013_redo/impact_layers',
                      'final_impact_layers/threats_2013_final')

inorganic_poll_07_10 <- raster(file.path(dir_threats, 
                             'normalized_by_two_time_periods/inorganic.tif'))
                   
ip_bcalb <- projectRaster(inorganic_poll_07_10, rast_base)
values(ip_bcalb)[is.na(values(rast_base))] <- NA ### only OHIBC EEZ

### max is 0.8973; re-rescale
ip_newref <- quantile(ip_bcalb, prob = 0.9999)

ip_bcalb_rescale <- calc(ip_bcalb, fun = function(x) {
    ifelse(x > ip_newref, 1, x / ip_newref)
  }) %>%
  setNames('inorganic')

writeRaster(ip_bcalb_rescale, 
            file.path(dir_goal_anx, 'rescaled_layers', 'cw_inorg_rescaled.tif'),
            overwrite = TRUE)

```

## Combine layers (sum method)

Because one layer (pesticides) is calculated for each year from 2002-2013, we will perform this as a loop:  stack each year of pesticides with the static layers for inorganic and ocean pollution, and sum for each year, saving each as a separate raster by year.

NOTE: sum seems to penalize regions harder than geometric mean in this case.

``` {r combine_chem_pollution_by_sum}

pest_rescale_files <- list.files(file.path(dir_goal_anx, 'rescaled_layers'), pattern = '*pest*', full.names = TRUE)
pest_rescale <- stack(as.list(pest_rescale_files))
ocean_rescale <- raster(file.path(dir_goal_anx, 'rescaled_layers', 'cw_ocean_poll_rescaled.tif'))
inorg_rescale <- raster(file.path(dir_goal_anx, 'rescaled_layers', 'cw_inorg_rescaled.tif'))


chem_total <- vector('list', length = nlayers(pest_rescale))

for (i in 1:nlayers(pest_rescale)) { # i = 1
  year <- names(pest_rescale[[i]]) %>%
               str_replace('cw_pest', 'cw_chem_sum')
  message('Processing ', year)
  chem_year <- stack(list(ocean_rescale, inorg_rescale, pest_rescale[[i]])) %>%
    calc(fun = sum, na.rm = TRUE) %>%
    setNames(year)
  values(chem_year)[is.na(values(rast_base))] <- NA
  chem_total[[i]] <- chem_year
}
chem_stack_sum <- stack(chem_total)

writeRaster(chem_stack_sum, 
            filename  = file.path(dir_goal_anx, 'rescaled_layers', paste0(names(chem_stack_sum), '.tif')),
            bylayer   = TRUE,
            overwrite = TRUE)

```

## Combine layers (geometric mean method)

Same as sum method, but calc geometric mean for each year, saving each as a separate raster by year.

``` {r combine_chem_pollution_by_geom_mean}

pest_rescale_files <- list.files(file.path(dir_goal_anx, 'rescaled_layers'), pattern = '*pest*', full.names = TRUE)
pest_rescale <- stack(as.list(pest_rescale_files))
ocean_rescale <- raster(file.path(dir_goal_anx, 'rescaled_layers', 'cw_ocean_poll_rescaled.tif'))
inorg_rescale <- raster(file.path(dir_goal_anx, 'rescaled_layers', 'cw_inorg_rescaled.tif'))

chem_total <- vector('list', length = nlayers(pest_rescale))

for (i in 1:nlayers(pest_rescale)) { # i = 1
  year <- names(pest_rescale[[i]]) %>%
               str_replace('cw_pest', 'cw_chem_gmean')
  message('Processing ', year)
  chem_year <- stack(list(ocean_rescale, inorg_rescale, pest_rescale[[i]]))
  
  ### count # of non-NA layers per cell
  chem_year_layers <- calc(chem_year, fun = function(x) {sum(!is.na(x))})
  
  ### since chem_year layers are by pressure, convert to score for geometric-mean,
  ### then calculate product, take appropriate root value, and subtract from one
  chem_year_inv <- 1 - chem_year
  
  chem_year_prod <- chem_year_inv %>%
    calc(fun = prod, na.rm = TRUE)
  
  chem_year_gm <- 1 - (chem_year_prod ^ (1/chem_year_layers)) 
    ### ^ (1/chem_year_layers) term finishes converting product to geometric mean,
    ### based on how many layers were used to calculate product
  
  ### set to NA any non-EEZ cells
  values(chem_year_gm)[is.na(values(rast_base))] <- NA
  
  chem_total[[i]] <- chem_year_gm %>%
    setNames(year)
}

chem_stack_gm <- stack(chem_total)

writeRaster(chem_stack_gm, 
            filename  = file.path(dir_goal_anx, 'rescaled_layers', paste0(names(chem_stack_gm), '.tif')),
            bylayer   = TRUE,
            overwrite = TRUE)

```


***

``` {r prov_footer, results = 'asis'}
prov_wrapup()
```

