---
title: 'OHIBC: Clean Waters goal prep'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(rgdal)
library(raster)
source('~/github/ohibc/src/R/common.R')  ### an OHIBC specific version of common.R

dir_git     <- '~/github/ohibc'
dir_spatial <- file.path(dir_git, 'prep/spatial')  ### github: general buffer region shapefiles
dir_anx     <- file.path(dir_M, 'git-annex/bcprep')

### goal specific folders and info
goal     <- 'cw'
scenario <- 'v2017'
dir_goal <- file.path(dir_git, 'prep', goal, scenario)
dir_goal_anx <- file.path(dir_anx, goal, scenario)

### provenance tracking
library(provRmd); prov_setup()

### support scripts
source(file.path('~/github/ohibc/src/R/map_scores.R'))
  ### score plotting scripts
source(file.path(dir_git, 'src/R/rast_tools.R')) 
  ### raster plotting and analyzing scripts

### set up the default BC projection to be BC Albers
p4s_bcalb <- c('bcalb' = '+init=epsg:3005')

```

# Summary

This script pulls together data from the four components of Clean Waters and outputs layers for toolbox:

* chemical pollution, which includes land-based organic chemicals (pesticides), land-based inorganic chemicals (impervious surface runoff), and ocean-based pollution from shipping and harbors.
    * processed by `cw_chem_prep.Rmd`; log-transformed and normalized by 99.99%tile within EEZ.
    * as in the global analysis, this is assessed using the 3 nm coastal buffer zone.
    * details: 
        * years       : 2002 - 2013
        * resolution  : 1000, 1000  (x, y) in m
        * coord. ref. : `+proj=aea +lat_1=50 +lat_2=58.5 +lat_0=45 +lon_0=-126 +x_0=1000000 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs` 
        * values      : 0, 2.889625  (min, max) (for first layer)
* nutrient pollution, which includes nutrient runoff from agriculture
    * processed by `cw_nutrient_prep.Rmd`; log-transformed and normalized by 99.99%tile within EEZ.
    * as in the global analysis, this is assessed at the 3 nm coastal buffer zone.
    * details: 
        * years       : 2002 - 2013
        * resolution  : 1000, 1000  (x, y) in m
        * coord. ref. : `+proj=aea +lat_1=50 +lat_2=58.5 +lat_0=45 +lon_0=-126 +x_0=1000000 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs` 
        * values      : 0, 1  (min, max)
* marine debris, which includes marine plastics debris surface density by mass
    * processed by `cw_trash_prep.Rmd`; log-transformed and normalized by 99.99%tile within __NE Pacific__ rather than BC EEZ.
    * as in the global analysis, this is assessed across the entire EEZ.
    * details: 
        * years       : no time series
        * resolution  : 1000, 1000  (x, y) in m
        * coord. ref. : `+proj=aea +lat_1=50 +lat_2=58.5 +lat_0=45 +lon_0=-126 +x_0=1000000 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs` 
        * values      : 0.2782703, 1  (min, max)

* pathogens, processed TBD.

This script pulls in finalized raster layers for each component, determines a score for each component by OHIBC region, then determines the overall score using a geometric mean calculation.

***

# Data Source -- REMOVE from here? if adjusting the other preps to be children of this, each has its own data source section...

* **Reference**: 
* **Downloaded**: 
* **Native Data Resolution**:   
* **Values**: 
* **Time Range**: 
* **Format**: 

***

# Methods

## Calculate nutrient scores

Nutrient scores are determined using the 3 nm offshore region.  Scores are calculated as (1 - mean nutrient pressure); since nutrient pressure scales from 0 to 1, a region with mean pressures near zero will result in a region score near 1.

``` {r define_rgn_extract}
rgn_extract <- function(rast, rgn_poly) {
  message('Extracting raster ', names(rast))
  ptm <- proc.time()
  rgn_vals <- raster::extract(rast, rgn_poly, weights = FALSE)
  message('... elapsed time: ', round((proc.time() - ptm)[3], 2), ' sec')
  
  names(rgn_vals) <- rgn_poly@data$rgn_id
  
  ### For the dataframe without cell weights, each list is just a
  ### vector of values, so we can simply assign that to a column in
  ### the data frame.
  rgn_vals_df <- data.frame()
  for (rgn_id in names(rgn_vals)) {
    temp_df <- data.frame(rgn_id   = as.numeric(rgn_id), 
                          cell_val = unlist(rgn_vals[[rgn_id]]))
    rgn_vals_df <- rbind(rgn_vals_df, temp_df)
  }
  
  return(rgn_vals_df)
}

```

``` {r calc_nutrient_scores}

### nutrient: rescaled fertilizers from 2002-2013

### 1. load regions
if(!exists('bc_rgn_poly')) 
  bc_rgn_poly <- readOGR(dsn = path.expand(dir_spatial),
                         layer = 'ohibc_rgn')
if(!exists('bc_3nm_poly')) 
  bc_3nm_poly <- readOGR(dsn = path.expand(dir_spatial),
                         layer = 'ohibc_offshore_3nm')
### 2. load layers
nutr_layer_files <- list.files(file.path(dir_goal_anx, 'rescaled_layers'),
                               pattern = 'fert',
                               full.names = TRUE)
nutr_stack <- stack(nutr_layer_files) %>%
  setNames(nutr_layer_files %>% basename %>% str_replace('.tif', '')) %>%
  raster::crop(bc_rgn_poly)

years <- names(nutr_stack) %>%
  str_extract('[0-9]{4}') %>%
  as.integer()

rgns_list <- list('po_nutrient'     = bc_rgn_poly,
                  'po_nutrient_3nm' = bc_3nm_poly)

reload <- FALSE

for (i in seq_along(rgns_list)) { # i <- 1
  po_filename <- file.path(dir_goal, 'output', paste0(names(rgns_list)[i], '.csv'))
  if(file.exists(po_filename) & reload == FALSE) {
    message('Pressure file exists: \n  ', po_filename)
    next()
  }
  
  message('Calculating nutrient pressures from ', names(rgns_list)[i])
  nutr_list <- vector('list', length = length(years))
  for (j in 1:length(years)) {
    # j <- 1
    message('In loop; processing layer ', j, ' out of ', length(years))
    tmp_df <- rgn_extract(nutr_stack[[j]], rgns_list[[i]])
    nutr_list[[j]] <- tmp_df %>%
      rename(nutr_val = cell_val) %>%
      mutate(year = years[j])
  }
  
  nutr_df <- bind_rows(nutr_list)
  
  po_nutr <- nutr_df %>%
    group_by(year, rgn_id) %>%
    summarize(nutr_pressure = mean(nutr_val, na.rm = TRUE))
  
  write_csv(po_nutr, po_filename)
  
}

nutr_map <- plot_rast_map(nutr_stack[[nlayers(nutr_stack)]],
                          rgn_poly = bc_rgn_poly,
                          rev_scale = TRUE,
                          title = 'Nutrient pressures')


```

## Calculate chemical scores

Chemical scores are determined using the 3 nm offshore region.  Since these rasters are sums of multiple rescaled chemical pollution source layers, values can exceed 1.0.  So the cell values must be clipped to 1.0 (simple clip, not a new normalization)

``` {r calc_chem_scores}

### chemical: sum of rescaled inorganic (no TS) + ocean (no TS) + pesticides (2002-2013)

### 1. load regions
bc_rgn_poly <- readOGR(dsn = path.expand(dir_spatial),
                       layer = 'ohibc_rgn')
bc_3nm_poly <- readOGR(dsn = path.expand(dir_spatial),
                       layer = 'ohibc_offshore_3nm')

### 2. load layer(s)
chem_layer_files <- list.files(file.path(dir_goal_anx, 'rescaled_layers'),
                               pattern = 'chem_gmean',
                               full.names = TRUE)
chem_stack <- stack(chem_layer_files) %>%
  setNames(chem_layer_files %>% 
             basename %>% 
             str_replace('\\.tif', '')) %>%
  raster::crop(bc_rgn_poly)

### NOTE:  The chem layers are sums of multiple sources of chemical pollution.
### These sums exceed 1.0; anything over 1 must be rescaled.
chem_stack_clipped <- chem_stack
values(chem_stack_clipped)[values(chem_stack_clipped) > 1] <- 1

years <- names(chem_stack_clipped) %>%
  str_extract('[0-9]{4}') %>%
  as.integer()

rgns_list <- list('po_chemical'     = bc_rgn_poly,
                  'po_chemical_3nm' = bc_3nm_poly)

reload <- FALSE

for (i in seq_along(rgns_list)) { # i <- 1
  po_filename <- file.path(dir_goal, 'output', paste0(names(rgns_list)[i], '.csv'))
  if(file.exists(po_filename) & reload == FALSE) {
    message('Pressure file exists: \n  ', po_filename)
    next()
  }
  
  message('Calculating chemical pressures from ', names(rgns_list)[i])
  chem_list <- vector('list', length = length(years))
  for (j in 1:length(years)) {
    # j <- 1
    message('In loop; processing layer ', j, ' out of ', length(years))
    tmp_df <- rgn_extract(chem_stack_clipped[[j]], rgns_list[[i]])
    chem_list[[j]] <- tmp_df %>%
      rename(chem_val = cell_val) %>%
      mutate(year = years[j])
  }
  
  chem_df <- bind_rows(chem_list)
  
  po_chem <- chem_df %>%
    group_by(year, rgn_id) %>%
    summarize(chem_pressure = mean(chem_val, na.rm = TRUE))
  
  write_csv(po_chem, po_filename)
  
}

chem_map <- plot_rast_map(chem_stack[[nlayers(chem_stack)]],
                          rgn_poly = bc_rgn_poly,
                          rev_scale = TRUE,
                          title = 'Chem pollution pressures')

```

## Calculate trash scores

Trash scores are calculated across the entire EEZ.  There is no time series for this layer.

``` {r calc_trash_scores}

### trash: rescaled marine plastics (no TS)

### 1. load regions
bc_rgn_poly <- readOGR(dsn = path.expand(dir_spatial),
                       layer = 'ohibc_rgn')
bc_3nm_poly <- readOGR(dsn = path.expand(dir_spatial),
                       layer = 'ohibc_offshore_3nm')

### 2. load layer(s)
trash_layer_files <- list.files(file.path(dir_goal, 'rescaled_layers'),
                               pattern = 'trash',
                               full.names = TRUE)

if(length(trash_layer_files) > 1) {
  message('More than one trash rescaled layer file! \n  ', paste(trash_layer_files, collapse = '\n  '))
  
  ### select max, linear scale, EEZ ref point
  quantile_tag <- str_detect(trash_layer_files, '_max_')
  region_tag   <- str_detect(trash_layer_files, '_eez')

  trash_layer_file <- trash_layer_files[quantile_tag & region_tag]
  message('Using: \n  ', trash_layer_file)
} else {
  trash_layer_file <- trash_layer_files[1]
}
  

trash_rast <- raster(trash_layer_file) %>%
  setNames(trash_layer_file %>% 
             basename %>% 
             str_replace('.tif', '')) %>%
  raster::crop(bc_rgn_poly)


rgns_list <- list('po_trash'     = bc_rgn_poly,
                  'po_trash_3nm' = bc_3nm_poly)

for (i in seq_along(rgns_list)) { # i <- 1
  
  message('Calculating marine debris pressures from ', names(rgns_list)[i])
  trash_df <- rgn_extract(trash_rast, rgns_list[[i]]) %>%
    rename(trash_val = cell_val)

  po_trash_df <- trash_df %>%
    group_by(rgn_id) %>%
    summarize(trash_pressure = mean(trash_val, na.rm = TRUE)) 
  
  write_csv(po_trash_df, file.path(dir_goal, 'output', paste0(names(rgns_list)[i], '.csv')))
}

trash_map <- plot_rast_map(trash_rast,
                          rgn_poly = bc_rgn_poly,
                          rev_scale = TRUE,
                          title = 'Marine plastic debris pressures')

```

### pathogens: still TBD


## Calculate geometric mean of all scores.

``` {r calc_geom_means}
nutr_score_df  <- read_csv(file.path(dir_goal, 'output', 'po_nutrient_3nm.csv'))
chem_score_df  <- read_csv(file.path(dir_goal, 'output', 'po_chemical_3nm.csv'))
trash_score_df <- read_csv(file.path(dir_goal, 'output', 'po_trash.csv')) %>%
  left_join(nutr_score_df %>% select(rgn_id, year),
            by = 'rgn_id')
path_score_df <- read_csv(file.path(dir_git, 'prep/ao/v2017/output', 'ao_closures.csv')) %>%
    mutate(days_in_year = ifelse(lubridate::leap_year(year), 366, 365)) %>%
    group_by(rgn_id, year) %>%
    summarize(path_pressure = (sum(days_avg) / first(days_in_year))) %>%
    ungroup()

cw_score_df <- nutr_score_df %>%
  full_join(chem_score_df,  by = c('year', 'rgn_id')) %>%
  full_join(path_score_df,  by = c('year', 'rgn_id')) %>%
  full_join(trash_score_df, by = c('year', 'rgn_id')) %>%
  gather(type, value, nutr_pressure:trash_pressure) %>%
  filter(!is.na(value)) %>%
  mutate(component_score = 1 - value)

cw_score_summary <- cw_score_df %>%
  group_by(rgn_id, year) %>%
  summarize(n_sources = n(), 
            prod = prod(component_score), 
            score = prod^(1/n_sources), ### this finishes our geometric mean
            sources = paste(type, collapse = ', '),
            values  = paste(round(value, 4), collapse = ', ')) %>%
  ungroup()
  
write_csv(cw_score_summary, file.path(dir_goal, 'output', 'cw_score_summary.csv'))

DT::datatable(cw_score_summary)

```

``` {r plot them}

cw_status_components <- cw_score_summary %>%
  select(rgn_id, year, score) %>%
  mutate(type = 'total') %>%
  bind_rows(cw_score_df %>% rename(score = component_score) %>% select(-value)) %>%
  left_join(get_rgn_names(), by = c('rgn_id'))


cw_status_plot <- ggplot(cw_status_components %>% filter(type != 'total'), 
                           aes(x = year, y = score, color = type, group = type)) +
  ggtheme_plot() +
  geom_line(size = 1, alpha = .8) +
  scale_colour_brewer(palette = 'BrBG') +
  labs(x = 'year',
       y = 'Clean Waters status',
       title = 'Clean Waters status',
       color = 'Region') +
  facet_wrap( ~ rgn_name)

print(cw_status_plot)


```


***

``` {r prov_footer, results = 'asis'}
prov_wrapup()
```

