---
title: 'OHIBC: CW trash layers prep'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(rgdal)
library(raster)
source('~/github/ohibc/src/R/common.R')  ### an OHIBC specific version of common.R

dir_git     <- '~/github/ohibc'
dir_spatial <- file.path(dir_git, 'prep/spatial')  ### github: general buffer region shapefiles
dir_anx     <- file.path(dir_M, 'git-annex/bcprep')

### goal specific folders and info
goal     <- 'cw'
scenario <- 'v2017'
dir_goal <- file.path(dir_git, 'prep', goal, scenario)
dir_goal_anx <- file.path(dir_anx, goal, scenario)

### provenance tracking
library(provRmd); prov_setup()

### support scripts
source(file.path('~/github/ohibc/src/R/map_scores.R'))
  ### score plotting scripts
source(file.path(dir_git, 'src/R/rast_tools.R')) 
  ### raster plotting and analyzing scripts

### set up the default BC projection to be BC Albers
p4s_bcalb <- c('bcalb' = '+init=epsg:3005')

```

#Summary

This data is incorporated into the OHI British Columbia Clean Waters (CW) goal.

***

#Data Source

**Reference**: [van Sebille et al, 2015](http://iopscience.iop.org/article/10.1088/1748-9326/10/12/124006/meta)

**Downloaded**: September 8, 2016 from Figshare: https://figshare.com/articles/data_of_microplastic_abundance_and_mass_from_Figure_3_of_Van_Sebille_et_al_2015_ERL_paper/1613256

**Native Data Resolution**:   1.0° in WGS84 projection

**Values**: Count (number/km2) and mass density (g/km2), based upon three different ocean circulation models (Lebreton, Maximenko, van Sebille)  

**Time Range**: N/A

**Format**: CSV

***

# Methods

## Read in data

from: van Sebille et al 2015. A global inventory of small floating plastic debris.  For this analysis we will use only the Van Sebille model for mass density rather than abundance.

``` {r read_vanSebille_data}
dir_data_vs <- file.path(dir_M, 'git-annex/bcprep/_raw_data/vanSebille_plastics')

vs_files_all <- list.files(dir_data_vs, full.names = TRUE) %>%
  setNames(basename(.) %>% str_replace('.csv', ''))

# [1] "gridcell_areas"            "latitudes"                 "lebretonmodel_abundance"  
# [4] "lebretonmodel_mass"        "longitudes"                "maximenkomodel_abundance" 
# [7] "maximenkomodel_mass"       "vansebillemodel_abundance" "vansebillemodel_mass"     

vs_files <- vs_files_all[str_detect(vs_files_all, 'vansebillemodel_mass')] 
### select just Van Sebille mass layer for models

vs_rasts <- vector('list', length = length(vs_files)) %>%
  setNames(names(vs_files))

for (i in 1:length(vs_files)) { ### i = 1
  vs_file <- vs_files[i]
  vs_fn <- names(vs_files)[i]
  
  vs_matrix <- read_csv(vs_file, col_names = FALSE) %>%
    data.matrix() %>%
    .[2:181, 1:360]
    ### it appears that row 361 is glitchy; ditch it, and ditch +90 N row as well
  rownames(vs_matrix) <- c(90:-89)
  colnames(vs_matrix) <- c(0:359)
  
  vs_rast <- raster(vs_matrix)
  extent(vs_rast) <- c(0, 360, -90, 90)
  vs_rast <- rotate(vs_rast)
  crs(vs_rast) <- CRS('+init=epsg:4326')
  
  vs_rasts[[vs_fn]] <- vs_rast
}

vs_stack <- stack(vs_rasts)

### get BC raster at WGS84 projection to provide extents for cropping
bc_rgn_rast <- raster(file.path(dir_spatial, 'raster', 'ohibc_rgn_wgs84_30min.tif'))

bc_ext <- extent(bc_rgn_rast) %>% round()

vs_stack_crop <- crop(vs_stack, bc_ext)
plot(vs_stack_crop)

### Save raster of van Sebille model by mass
writeRaster(vs_stack_crop[['vansebillemodel_mass']], 
            file.path(dir_goal, 'int/cw_trash_mass_vansebille.tif'), 
            overwrite = TRUE)

### create North East Pacific raster: north of Equator; west of 100 W and east of 100 E
n_pac_ext <- extent(c(-180, -100, 0, 65))
vs_mass_npac <- crop(vs_stack[['vansebillemodel_mass']], n_pac_ext)
plot(vs_mass_npac)

### Save the global Van Sebille mass map cropped to the North Pacific for additional reference point options
writeRaster(vs_mass_npac, 
            file.path(dir_goal, 'int/cw_trash_mass_vansebille_n_pac.tif'), 
            overwrite = TRUE)

```

Crop to BC and interpolate using Thin Plate Spline to gapfill Strait of Georgia and coastal fjords.

``` {r interpolate_bc_sla}

trash_dens <- raster(file.path(dir_goal, 'int/cw_trash_mass_vansebille.tif'))

trash_dens_interp_file <- file.path(dir_goal, 'int/cw_trash_mass_interp.tif')

if(!file.exists(trash_dens_interp_file)) {
  message('Interpolating BC trash mass data')

  xy <- data.frame(xyFromCell(trash_dens, 1:ncell(trash_dens)))

  v  <- getValues(trash_dens)
  tmpdf <- cbind(xy, v) %>%
    filter(!is.na(v)) 
  xy1 <- tmpdf[ , 1:2]
  v1  <- tmpdf[ , 3]
  tps_model <- fields::Tps(xy1, v1)

  trash_dens_interp <- interpolate(trash_dens, tps_model)

  trash_dens_gapfilled <- trash_dens
  values(trash_dens_gapfilled)[is.na(values(trash_dens))] <- values(trash_dens_interp)[is.na(values(trash_dens))] 

  message('Writing gapfilled BC trash mass to: ', trash_dens_interp_file)
  writeRaster(trash_dens_gapfilled, trash_dens_interp_file, overwrite = TRUE)
  
} else {
  message('Reading gapfilled BC trash mass from: ', trash_dens_interp_file)

  trash_dens_gapfilled <- raster(trash_dens_interp_file)
}

plot(trash_dens_gapfilled)

```

Reproject into BC Albers at 1000 m resolution

```{r reproject}

rast_dens_interp <- raster(file.path(dir_goal, 'int/cw_trash_mass_interp.tif'))

### reproject to BC Albers
poly_eez_base <- readOGR(path.expand(dir_spatial), 'ohibc_rgn')
rast_base_bcalb <- raster(file.path(dir_spatial, 'raster/ohibc_rgn_raster_1000m.tif')) %>%
  crop(poly_eez_base)

trash_dens_bcalb <- projectRaster(from = rast_dens_interp, to = rast_base_bcalb, method = 'ngb')
values(trash_dens_bcalb)[is.na(values(rast_base_bcalb))] <- NA

writeRaster(trash_dens_bcalb, file.path(dir_goal, 'int/cw_trash_mass_bcalb.tif'), overwrite = TRUE)

``` 

``` {r plot plastic pollution rasters}

plot(trash_dens_bcalb, main = 'Mass density (g/km2)')
plot(poly_eez_base, alpha = .5, add = TRUE)

```


## Log Transform

Transform the values to a log(x + 1) value.  Based on reference point discussions we will not do log transforms for this goal.

```{r log_transform, eval = FALSE, include = FALSE}

trash_dens_bcalb <- raster(file.path(dir_goal, 'int/cw_trash_mass_bcalb.tif'))

trash_dens_log <- calc(trash_dens_bcalb, fun = function(x) {log(x + 1)})

writeRaster(trash_dens_log, file.path(dir_goal, 'int/cw_trash_mass_log.tif'), overwrite = TRUE)

plot(trash_dens_log, main = 'Mass density\nlog(g/km2)')

```

## Rescale marine plastics pressure from zero to 1

Rescale all values to range from zero to 1, using several reference points for comparison.  These rescalings will be done on both log-transformed and raw (linear) scales.

* Max across BC EEZ
    * Using this reference point for only the BC EEZ results in high pressures of trash; the distribution is relatively uniform across the EEZ, so the reference point is relatively low compared to the mean trash density.  While this will still allow relative ranking, it may create a harsh penalty to the Clean Waters goal if the reference point is actually quite low.  Note: could do 99.99%ile but at this resolution, this is just the simple maximum.
* 110% of maximum value
    * Since the marine plastics raster from Van Sebille is relatively coarse (1 degree cells) it may be unnecessary to use a 99.99% (or other) quantile to filter out outliers, as they would presumably be averaged across a large area.  While we could simply use the max value, we apply a 110% increase on the assumption that BC trash is relatively low and so could potentially get worse over time.
* 99.99% and 100% quantile of entire North Pacific basin
    * To further examine whether the reference point based on the BC EEZ is extremely low, we examine the score based on a higher reference point - using the values estimated for the entire North Pacific ocean basin.  This should capture a larger range of debris values, providing a better idea of what is the "worst" case.  While the cells are large and should spatially average extreme values, some filtering of outliers may prove useful, so we will examine both the 99.99%ile and 100%ile.

```{r linear_rescales}

### bring in linear-scaled BC map at 1000 m BC Albers projection
trash_dens <- raster(file.path(dir_goal, 'int/cw_trash_mass_bcalb.tif'))

# print(hist(trash_dens, main = 'Mass density (log(g/km2))'))

### bring in N Pacific raster; we are not mapping so scale and projection are unimportant
trash_dens_npac <- raster(file.path(dir_goal, 'int/cw_trash_mass_vansebille_n_pac.tif'))

# print(hist(trash_dens_npac, main = 'Mass density (log(g/km2)), NE Pacific'))

trash_ref_max            <- maxValue(trash_dens) 
trash_ref_110max         <- trash_ref_max * 1.1
trash_ref_9999qtile_npac <- quantile(trash_dens_npac, prob = 0.9999) # maxValue(trash_dens_npac)

trash_max <- calc(trash_dens, fun = function(x) {
    ifelse(x > trash_ref_max, 1, x / trash_ref_max)
  }) %>%
  setNames('linear_mass')
trash_110max <- calc(trash_dens, fun = function(x) {
    ifelse(x > trash_ref_110max, 1, x / trash_ref_110max)
  }) %>%
  setNames('linear_mass')
trash_9999qtile_npac <- calc(trash_dens, fun = function(x) {
    ifelse(x > trash_ref_9999qtile_npac, 1, x / trash_ref_9999qtile_npac)
  }) %>%
  setNames('linear_mass')


writeRaster(trash_max, file.path(dir_goal, 'rescaled_layers/cw_trash_max_eez.tif'), overwrite = TRUE)
writeRaster(trash_110max, file.path(dir_goal, 'rescaled_layers/cw_trash_110max_eez.tif'), overwrite = TRUE)
writeRaster(trash_9999qtile_npac, file.path(dir_goal, 'rescaled_layers/cw_trash_9999qtile_npac.tif'), overwrite = TRUE)

plot(trash_max, main = 'CW trash pressure layer: \n mass density/max in BC EEZ')
plot(trash_9999qtile_npac, main = 'CW trash pressure layer: \n mass density/99.99%ile in NE Pac')

```

```{r log_rescales, eval = FALSE, include = FALSE}

# ### bring in log-transformed BC map at 1000 m BC Albers projection
# trash_dens_log <- raster(file.path(dir_goal, 'int/cw_trash_mass_log.tif'))
# 
# # print(hist(trash_dens_log, main = 'Mass density (log(g/km2))'))
# 
# ### bring in N Pacific raster; we are not mapping so scale and projection are unimportant
# trash_dens_npac <- raster(file.path(dir_goal, 'int/cw_trash_mass_vansebille_n_pac.tif'))
# trash_npac_log <- log(values(trash_dens_npac) + 1)
# 
# # print(hist(trash_npac_log, main = 'Mass density (log(g/km2)), NE Pacific'))
# 
# 
# trash_logref_9999qtile      <- quantile(trash_dens_log, prob = 0.9999)
# trash_logref_110max         <- max(values(trash_dens_log), na.rm = TRUE) * 1.1
# trash_logref_9999qtile_npac <- quantile(trash_npac_log, prob = 0.9999, na.rm = TRUE)
# 
# trash_log_9999qtile <- calc(trash_dens_log, fun = function(x) {ifelse(x > trash_logref_9999qtile, 1, x / trash_logref_9999qtile)}) %>%
#   setNames('log_mass')
# trash_log_110max <- calc(trash_dens_log, fun = function(x) {ifelse(x > trash_logref_110max, 1, x / trash_logref_110max)}) %>%
#   setNames('log_mass')
# trash_log_9999qtile_npac <- calc(trash_dens_log, fun = function(x) {ifelse(x > trash_logref_9999qtile_npac, 1, x / trash_logref_9999qtile_npac)}) %>%
#   setNames('log_mass')
# 
# 
# writeRaster(trash_log_9999qtile, file.path(dir_goal, 'rescaled_layers/cw_trash_9999qtile_eez_log.tif'), overwrite = TRUE)
# writeRaster(trash_log_110max, file.path(dir_goal, 'rescaled_layers/cw_trash_110max_eez_log.tif'), overwrite = TRUE)
# writeRaster(trash_log_9999qtile_npac, file.path(dir_goal, 'rescaled_layers/cw_trash_9999qtile_npac_log.tif'), overwrite = TRUE)
# 
# plot(trash_log_9999qtile, main = 'CW trash pressure layer: \nlog(mass density)/99.99%ile in BC EEZ')
# plot(trash_log_9999qtile_npac, main = 'CW trash pressure layer: \nlog(mass density)/99.99%ile in NE Pac')

```

Reference points:

* max within BC EEZ:  `r trash_ref_max` g/km^2^
    * Linear-scaled file at: ``r file.path(dir_goal, 'rescaled_layers/cw_trash_max_eez.tif')``

* 110% of max within BC EEZ:  `r trash_ref_110max` g/km^2^
    * Linear-scaled file at: ``r file.path(dir_goal, 'rescaled_layers/cw_trash_110max_eez.tif')``

* 99.99%ile within NE Pacific:  `r trash_ref_9999qtile_npac` g/km^2^
    * Linear-scaled file at: ``r file.path(dir_goal, 'rescaled_layers/cw_trash_9999qtile_npac.tif')``

***

``` {r prov_footer, results = 'asis'}
prov_wrapup()
```

