---
title: 'OHIBC: data prep for wild-capture fisheries'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(sp)
library(rgdal)
library(raster)
library(DT)

dir_git <- '~/github/ohibc'
source(file.path(dir_git, 'src/R/common.R'))  ### an OHIBC specific version of common.R
dir_rgn <- file.path(dir_git, 'prep/regions')  ### github: general buffer region shapefiles
dir_anx <- file.path(dir_M, 'git-annex/bcprep')


### goal specific folders and info
goal      <- 'fis'
scenario  <- 'v2016'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario)
dir_goal_anx <- file.path(dir_anx, goal, scenario)

### provenance tracking
source(file.path(dir_git, 'src/R/prov.R'))   
  ### Provenance tracking functions: must source at start to initialize prov_track
prov_run_tag <- 'standard run'

### support scripts
source(file.path(dir_git, 'src/R/poly_plot_scores.R')) 
  ### score plotting scripts
source(file.path(dir_git, 'src/R/rast_tools.R')) 
  ### raster plotting and analyzing scripts
source(file.path(dir_goal, 'fis_fxns.R'))
  ### goal-specific support functions

### set up proj4string options: BC Albers and WGS84
p4s_wgs84 <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
p4s_bcalb <- '+proj=aea +lat_1=50 +lat_2=58.5 +lat_0=45 +lon_0=-126 +x_0=1000000 +y_0=0 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0'

```

fisheries reporting:
* fisheries reporting is in zones (denoted by polygons)
* some fisheries are reported by cells (4 km cells?)
  * dungeness crabs
  
  
We will work at 1 km resolution (?)
* create a raster of rgn_id at 1 km resolution
  * does this already exist?
    * yes: prep/spatial/ohibc_rgn_raster_1000m.tif
      * NOTE: the raster has some problems; e.g. Haida Gwaii is filled in
    * also 500 m raster if desired
* create a raster of fishery zones (for csv reporting)
  * at 1 km resolution
* rasterize layers for each year:
  * use nearest neighbor if reclassifying
  * use cell center (default) otherwise; this allows for gdal_rasterize
  * stack together all years of a single fishery for processing

Explore layers:
``` {r, eval = FALSE}
dir_dfo <- file.path(dir_anx, '_raw_data/dfo_khunter')
dir_list <- list.files(dir_dfo)
dir_list <- dir_list[file.info(file.path(dir_dfo, dir_list))$isdir] %>%
  file.path(dir_dfo, .)
#  [1] "aquaculture"               "dungeness_crab"            "ferry_terminals"          
#  [4] "geoduck"                   "groundfish_longline"       "groundfish_mgmt_areas"    
#  [7] "groundfish_trawl"          "marina_pnt"                "pac_fishery_mgmt_areas"   
# [10] "pac_fishery_mgmt_subareas" "pacific_sardine"           "salmon_enhancement"       
# [13] "spot_prawn"  

basename(dir_list)
# shp_list <- list.files(dir_list[i], full.names = TRUE)
# shp_list <- list.files(file.path(dir_list[i], 'TENURES_PNT'), full.names = TRUE)

# shp_list <- shp_list[str_detect(shp_list, '.shp$')] %>%
#   str_replace('.shp', '')
# shp <- list()
# for (j in 1:length(shp_list)) { ### j <- 1
#   shp[j] <- rgdal::readOGR(dsn   = dirname(shp_list[j]), 
#                            layer = basename(shp_list[j]), 
#                            stringsAsFactors = FALSE)
#   print(names(shp[[j]]@data))
# }
# 
# names(shp) <- shp_list
# 
# plot(shp[[1]])
# 
# d <- shp[[1]]@data
# 
# d[1, ]
# shp_list

### ferry_terminals: CHRFRRTRMN_point
#                  ROUTE TELE_NUM TRMNL_LOC OPERATOR TRCKTTLFRR FERRY_ID   DATA_SRC COMMENTS VCLOHTTYR VCLTTYR
# 1 Victoria To Port Ang     <NA>      <NA>     <NA>          0        1 CHS CHARTS     <NA>         0       0
#   VCLUHTTYR VSSLSZLMT BERTHS PSNGRTTAUG PSNGRTTDEC PSNGRTTFEB OBJECTID PSNGRTTAPR PSNGRTTJAN       FYTRMTYP
# 1         0         0     NA          0          0          0        1          0          0 Ferry Terminal
#   AGENT PSNGRTTYR RI PSNGRTTJUL BUSTTLYR SRC_PROJ PSNGRTTJUN PSNGRTTMAR PSNGRTTMAY PSNGRTTNOV PSNGRTTOCT
# 1  <NA>         0  0          0        0     SSOG          0          0          0          0          0
#   PSNGRTTSEP
# 1          0

### marina_pnt: MARINA_PNT_point
#   FAX_NUMBER TREAT_AVL UNIT_TOTAL WSHRM_AVL EMAIL_ADDR WATER_AVL RV_SITE_TO MARINA_ID PEAK_PERID LOCATION CAMPSITE_T
# 1       <NA>      <NA>          0         U       <NA>         U          0       265       <NA>     <NA>          0
#   OBJECTID NAME MOORAGE_AV MAX_BOATS MARINA_SUP LUB_AVL LIVE_ABOAR LNDRY_AVL HOLD_TANKS GAS_AVL FUEL_AVL     F_CODE
# 1      265 <NA>          U         0          U       U          0         U          0       U        U BL17000000
#   PRI_PHNE_N DISPSL_AVL DIESEL_AV                  DESCRIPT DATA_SRC CONTACT COMMENTS CITY AVG_OCCUP AVG_DEPTH
# 1       <NA>          U         U Information Not Available     <NA>    <NA>     <NA> <NA>         0         0
#   ACCOM_TYPE SHOW_AVL SEC_PHNE_N SEASON RSTRT_AVL REL_IMPORT PROVINCE PWR_AVL POSTAL_COD SOURCE_PRJ STR_ADDRES TRSVLBL
# 1       <NA>        U       <NA>   <NA>         U          3     <NA>    <NA>       <NA>       NSOG       <NA>       U

### salmon_enhancement_facilities: Accommodation
# [1] "Echo Bay Guardian Trailer"



```

``` {r setup_fish_mgmt_zone_rasters, eval = TRUE}

### NOTE: gdal_rasterize doesn't seem to like zero values - assigns to NA.
### Check each polygon to make sure no zero value areas/regions.

rast_rgn <- raster(file.path(dir_git, 'prep/spatial/ohibc_rgn_raster_1000m.tif'))

### Rasterize Pacific Fishery Management Areas
###   STATAREA HECTARES    SQ_KM SHAPE_Leng SHAPE_Area
###          1 109760.7 1097.607   891758.4 1097606860

pfma_shp <- file.path(dir_anx, '_raw_data/dfo_khunter',
                      'pac_fishery_mgmt_areas',
                      'DFO_BC_PFMA_AREAS_50K_V3_1')

pfma_tif <- file.path(dir_anx, 'fis', scenario, 'spatial', 
                      'DFO_BC_PFMA_AREAS_50K_V3_1.tif')

rast_pfma <- good_rasterize(pfma_shp, pfma_tif, 'STATAREA')

### Rasterize Pacific Fishery Management Sub-Areas
### NOTE: DFO_BC_PFMA_SUBAREAS_CHS_V3.shp is corrupted or invalid
### DFO_BC_PFMA_SUBAREAS_50K_V3_1:
###   MGNT_AREA SUBAREA_         NAME LABEL HECTARES    SQ_KM SHAPE_Leng SHAPE_Area
###           7       22 Subarea 7-22  7-22 227.6391 2.276391   16178.94    2276387

### Need to capture both area and subarea values?
### * create new field for area/subarea ('a_suba') by taking
###   area and adding subarea as three-decimal-place fraction (subarea/1000)
### * save new poly with a '_mod' after it to not mess up raw data
### * use '_mod' file for rasterization
pfmasub_shp <- file.path(dir_anx, '_raw_data/dfo_khunter',
                      'pac_fishery_mgmt_subareas',
                      'DFO_BC_PFMA_SUBAREAS_50K_V3_1')
pfmasub_shp_mod <- file.path(dir_anx, '_raw_data/dfo_khunter',
                      'pac_fishery_mgmt_subareas',
                      'DFO_BC_PFMA_SUBAREAS_50K_V3_1_mod')
pfmasub_poly <- readOGR(dirname(pfmasub_shp), basename(pfmasub_shp))
pfmasub_poly@data <- pfmasub_poly@data %>%
  mutate(a_suba = MGNT_AREA + SUBAREA_/1000)
writeOGR(pfmasub_poly, 
         dsn = dirname(pfmasub_shp_mod), 
         layer = basename(pfmasub_shp_mod), 
         driver = 'ESRI Shapefile',
         overwrite_layer = TRUE)

pfmasub_tif <- file.path(dir_anx, 'fis', scenario, 'spatial', 
                         'DFO_BC_PFMA_SUBAREAS_50K_V3_1.tif')

rast_pfmasub <- good_rasterize(pfmasub_shp_mod, pfmasub_tif, 'a_suba')

### Rasterize Groundfish Management Areas
###   MINOR_CODE         MINOR_NAME MAJOR_CODE MAJOR_NAME GFB_MIN_CD GFB_MAJ_CD DESCR Shape_Leng Shape_Area
###            1 2A-EAST- SKIDEGATE          8         5D         01         08 5D-01   433586.1 4000732438
gfma_shp <- file.path(dir_anx, '_raw_data/dfo_khunter',
                      'groundfish_mgmt_areas',
                      'MajorMinor_GroundfishManagementAreas')

gfma_tif <- file.path(dir_anx, 'fis', scenario, 'spatial', 
                      'MajorMinor_GroundfishManagementAreas.tif')

rast_gfma <- good_rasterize(gfma_shp, gfma_tif, 'MINOR_CODE')

plot(rast_pfma)
plot(rast_pfmasub)
plot(rast_gfma)
```

``` {r create dataframe of regions to zones}

pfmsa_r_df <- data.frame(zone = values(rast_pfmasub)) %>%
  mutate(mgt_area = floor(zone),
         mgt_subarea = round((zone - floor(zone))*1000)) %>%
  group_by(mgt_area, mgt_subarea) %>%
  summarize(n_cells = n())
pfmsa_p_df <- foreign::read.dbf(paste0(str_replace(pfmasub_shp, '.shp', ''), '.dbf')) %>%
  rename(mgt_area = MGNT_AREA, mgt_subarea = SUBAREA_)

pfmsa_both <- pfmsa_p_df %>% full_join(pfmsa_r_df %>% 
                                         filter(!is.na(n_cells)),
                                       by = c("mgt_area", "mgt_subarea")) %>%
  mutate(area_check = SQ_KM/n_cells)

names(pfmsa_both) <- tolower(names(pfmsa_both))

write_csv(pfmsa_both, file.path(dir_goal, 'int', 'pfmsubareas_df.csv'))

```


``` {r fish_mgmt_zones_by_rgn}
if(!exists('rast_rgn'))
  rast_rgn <- raster(file.path(dir_git, 'prep/spatial/ohibc_rgn_raster_1000m.tif'))

### count cells of each fishery management zone that fall within each OHIBC region.
### e.g. zone 100 (300 cells) falls between region 1 (100 cells) and region 2 (200 cells)
### set up df with columns: rgn_id, pfma_id, pfmsa_id, gfma_id using getValues.
### If all rasters are same extents, then the return from getValues() should be a vector
### of cell values that correspond to the same order of cells for each vector.

rast_pfma  <- raster(file.path(dir_anx, 'fis', scenario, 'spatial', 
                              'DFO_BC_PFMA_AREAS_50K_V3_1.tif'))
rast_pfmsa <- raster(file.path(dir_anx, 'fis', scenario, 'spatial', 
                              'DFO_BC_PFMA_SUBAREAS_50K_V3_1.tif'))
rast_gfma  <- raster(file.path(dir_anx, 'fis', scenario, 'spatial', 
                              'MajorMinor_GroundfishManagementAreas.tif'))

### set up dataframe of 
if(!(extent(rast_pfma)  == extent(rast_rgn)) |
   !(extent(rast_pfmsa)  == extent(rast_rgn))) {
  message('rgn and pfma extents not equal... fix them!')
} else {
  rgn_pfma_df <- data.frame(
      rgn_id   = round(getValues(rast_rgn)),
      pfma_id  = round(getValues(rast_pfma)),
      pfmsa_id = round(getValues(rast_pfmsa), 4)) 
  rgn_pfma_df <- rgn_pfma_df %>%
    mutate(pfma_id2  = floor(pfmsa_id),
           pfmsa_id2 = round((pfmsa_id - floor(pfmsa_id))*1000)) 
  # rgn_pfma_df %>% filter(pfma_id != pfma_id2) 
  ### zero-length: so all PFMA cells match with PFMA identifiers in the PFMSA raster
  ### so: ditch the superfluous columns to avoid confusion!
  rgn_pfma_df <- rgn_pfma_df %>%
    mutate(pfmsa_id = pfmsa_id2) %>%
    select(-pfmsa_id2, -pfma_id2)
  
  rgn_pfma_df1 <- rgn_pfma_df %>%
    group_by(rgn_id) %>%
      ### count total cells in each OHIBC region
    mutate(n_cells_rgn = n()) %>%
    group_by(pfma_id) %>%
      ### count total cells in each PacFishMgmtArea
    mutate(n_pfma_tot = n()) %>%
    group_by(rgn_id, pfma_id) %>%
      ### count PFMA cells in each OHIBC region
    mutate(n_pfma_rgn = n()) %>%
    group_by(pfma_id, pfmsa_id) %>%
      ### count total cells in each PFMsubarea (within each PFMA)
    mutate(n_pfmsa_tot = n()) %>%
    group_by(rgn_id, pfma_id, pfmsa_id) %>%
      ### count PFMSA cells in each OHIBC region
    mutate(n_pfmsa_rgn = n()) %>%
    ungroup() %>%
    filter(!is.na(rgn_id)) %>%
      ### ditch the NAs (not in an OHIBC region)
    distinct()
      ### ditch the duplicates (which basically summarizes, since
      ### no cell IDs)
  
  ### check lost cells - falling outside of an OHIBC region
  # rgn_pfma_df2 <- rgn_pfma_df1 %>%
  #   group_by(pfma_id, pfmsa_id) %>%
  #   mutate(n_pfmsa_lost = n_pfmsa_tot - sum(n_pfmsa_rgn, na.rm = TRUE)) 
  # rgn_pfma_df3 <- rgn_pfma_df1 %>%
  #   select(-pfmsa_id, -n_pfmsa_rgn, -n_pfmsa_tot) %>%
  #   distinct() %>%
  #   group_by(pfma_id) %>%
  #   mutate(n_pfma_lost = n_pfma_tot - sum(n_pfma_rgn, na.rm = TRUE)) 

  write_csv(rgn_pfma_df1, file.path(dir_goal, 'rgns', 'pfmsubareas_df.csv'))

}



if(!(extent(rast_gfma) == extent(rast_rgn))) {
  message('rgn and gfma extents not equal... fix them!')
} else {
  rgn_gfma_df <- data.frame(
      rgn_id   = round(getValues(rast_rgn)),
      gfma_id  = round(getValues(rast_gfma))) %>%
    group_by(rgn_id) %>%
    mutate(n_cells_rgn = n()) %>%
    group_by(gfma_id) %>%
    mutate(n_gfma_tot = n()) %>%
    group_by(rgn_id, gfma_id) %>%
    mutate(n_gfma_rgn = n()) %>%
    filter(!is.na(rgn_id)) %>%
    distinct()
  
  ### check lost cells - falling outside of an OHIBC region
  # rgn_gfma_df1 <- rgn_gfma_df %>%
  #   group_by(gfma_id) %>%
  #   mutate(n_gfma_lost = n_gfma_tot - sum(n_gfma_rgn, na.rm = TRUE)) 

  write_csv(rgn_gfma_df, file.path(dir_goal, 'rgns', 'gfmareas_df.csv'))

}

```

``` {r child = file.path(dir_git, 'src/templates/ohibc_prov_ftr.Rmd')}
```

