---
title: 'OHIBC Howe Sound goal prep: Species'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

source('~/github/ohibc/src/R/common.R')  ### an OHIBC specific version of common.R

scenario <- 'vHS'
goal     <- 'spp_ico'
dir_git <- '~/github/ohibc'
dir_goal <- file.path(dir_git, 'prep', goal, scenario)
dir_rgn  <- file.path(dir_git, 'prep/spatial')
dir_anx        <- file.path(dir_M, 'git-annex/bcprep') ### git-annex: goal-specific large files
dir_anx_global <- file.path(dir_M, 'git-annex/globalprep/spp_ico')

library(provRmd); prov_setup()

source(file.path(dir_goal, 'spp_fxn.R'))

### set up the default BC projection to be BC Albers
p4s_bcalb <- c('bcalb' = '+init=epsg:3005')
p4s_wgs84 <- c('wgs84' = '+init=epsg:4326')

```

# Summary: OHIBC Howe Sound Species

This script prepares scores (status and trend) for species richness in 
British Columbia's Howe Sound region.  Spatial data from IUCN and Aquamaps is
combined with extinction risk information from IUCN and conservation rank
info based on province-level NatureServe categories.

Because of the small scale of Howe Sound, neither Species nor Iconic Species goals use any area weighting.  Instead, ICO and SPP status is based upon a simple average of species
health for species found within the Howe Sound region.

From Halpern et al (2012):

> The target for the Species sub-goal is to have all species at a risk status of Least Concern. We scaled the lower end of the biodiversity goal to be 0 when 75% species are extinct, a level comparable to the five documented mass extinctions and would constitute a catastrophic loss of biodiversity. The Status of assessed species was calculated as the area- and threat status-weighted average of the number of threatened species within each 0.5 degree grid cell.

**Mean risk status for Howe Sound:**

$$\bar{R} = \frac{\displaystyle\sum_{species}(Risk)}{n_{spp}}$$

**Species goal model**

$$X_{SPP} = \frac{((1 - \bar{R}_{SPP}) - 0.25)}{(1 - 0.25)} * 100%$$

where:

* $X_{SPP}$ is Species goal status
* $\bar{R}$ is mean extinction risk for identified species within Howe Sound (different subsets for ICO and SPP)
* *Risk* is scaled value for species extinction risk category, based on: 
    * 'LC' = 0.0, 'NT' = 0.2, 'VU' = 0.4, 'EN' = 0.6, 'CR' = 0.8, 'EX' = 1.0
* SPP trend is calculated as the linear trend of the average extinction risk categories over time.

-----

# Data sources

AquaMaps

* From http://www.aquamaps.org/main/home.php: "AquaMaps are computer-generated predictions of natural occurrence of marine species, based on the environmental tolerance of a given species with respect to depth, salinity, temperature, primary productivity, and its association with sea ice or coastal areas."
* Citation: Kaschner, K., J. Rius-Barile, K. Kesner-Reyes, C.Garilao, S.O. Kullander, T. Rees and R. Froese (2015). AquaMaps: Predicted range maps for aquatic species. World wide web electronic publication, www.aquamaps.org, Version 08/2015.

IUCN Red List spatial data: species range map shapefiles

* Species range map shapefiles downloaded from: http://www.iucnredlist.org/technical-documents/spatial-data
* Citation: IUCN 2015. The IUCN Red List of Threatened Species. Version 2015.3. <http://www.iucnredlist.org>. Downloaded on 21 December 2015.

IUCN Red List species index: list of all IUCN red list species, incl IUCN species ID and extinction risk category

* Extinction risk categories, past assessments, and habitat information, by species, downloaded from: http://apiv3.iucnredlist.org/
* Citation: IUCN 2015. The IUCN Red List of Threatened Species. Version 2015-3. <http://www.iucnredlist.org>. Downloaded <date of last script process>.

NatureServe/BC CDC conservation rank info from BC Species and Ecosystems Explorer: 

* BC Conservation Data Centre: http://www.env.gov.bc.ca/atrisk/toolintro.html
* NatureServe conservation status: http://explorer.natureserve.org/ranking.htm#global
* Citation: B.C. Conservation Data Centre. 2016. BC Species and Ecosystems Explorer. B.C. Ministry of Environment, Victoria B.C. Available: <http://a100.gov.bc.ca/pub/eswp/>.  Downloaded on 30 November 2016.

-----

# Methods

## Define assessment region

Using OHIBC Howe Sound region polygon, determine 0.5° raster cells corresponding to the region.

``` {r setup_region-to-cell_lookup}

### Read in OHIBC Howe Sound polygons, and transform to WGS84 CRS
poly_howesound_rgn <- readOGR(dsn = path.expand(dir_rgn), layer = 'howe_sound_rgn',
                       verbose = FALSE, stringsAsFactors = FALSE) %>%
  spTransform(CRS(p4s_wgs84))

poly_howesound_rgn@data <- poly_howesound_rgn@data %>%
  rename(rgn_id = id) %>%
  mutate(rgn_name = 'Howe Sound')

rgn2cell_list <- spp_rgn2cell(poly_howesound_rgn, 
                              rgn_tag = '_howesound',
                              reload = TRUE)

rgn2cell_df      <- rgn2cell_list[[1]]
loiczid_rast     <- rgn2cell_list[[2]]
```


## Identify Howe Sound-specific species

These are identified by comparing the full AM and IUCN cell-species files to the limited cells that appear in the Howe Sound region.

``` {r identify_HS_species}

am_cells_spp <- spp_get_am_cells(rgn2cell_df, prob_filter = 0, reload = FALSE) %>%
  filter(loiczid %in% rgn2cell_df$loiczid)

write_csv(am_cells_spp, file.path(dir_goal, 'int', 'am_cells_hs_spp.csv'))

iucn_cells_spp <- spp_get_iucn_cells(rgn2cell_df, reload = FALSE) %>%
  filter(loiczid %in% rgn2cell_df$loiczid)

write_csv(iucn_cells_spp, file.path(dir_goal, 'int', 'iucn_cells_hs_spp.csv'))

```

Number of Howe Sound species according to dataset:

* AquaMaps: `r length(am_cells_spp$am_sid %>% unique())`
* IUCN: `r length(iucn_cells_spp$iucn_sid %>% unique())`

-----

## Load species lookup table from global SPP goal

Currently this uses the global species lookup table. How this list is generated (file locations are in `~/github/ohiprep/globalprep/SPP_ICO`):

* in `ingest_iucn.R`:
    * Pull full IUCN list from http://api.iucnredlist.org/index/all.csv
    * Filter the list to marine species as indicated on each species page
    * Standardize old extinction risk categories
    * Append population trend and parent/subpop info from species web pages
    * Write to `spp_iucn_mar.csv` in git-annex.
* in `spp_fxn.R::create_spp_master_lookup()`:
    * Pull in Aquamaps species information table, create sciname field and standardize categories
    * Pull in IUCN marine species info from `ingest_iucn.R`
    * `full_join()` the two species lists
    * Determine overall category, trend, and spatial info source (preferring IUCN)
    * Remove duplicate records and data-deficient records
    * Convert text extinction risk category and popn trends to number values
    * Identify parent/subpopulations and aliases within `spatial_source` field.

``` {r read_global_species_list, eval = TRUE}
spp_global_file <- file.path(dir_anx_global, 'v2016', 'int/spp_all_cleaned.csv')
spp_info_gl <- read_csv(spp_global_file)

```

Variables in `spp_info_gl`: `r paste(names(spp_info_gl), collapse = ' | ')`

``` {r print_global_species_list, eval = TRUE}
knitr::kable(head(spp_info_gl %>% 
                dplyr::select(sciname, am_sid, iucn_sid, pop_trend, pop_cat, spatial_source)),
             align = c('l', 'c', 'c', 'c', 'c', 'c'),
             caption = 'Quick glance at global spp_info_gl dataframe:')
```

-----

## Append BC-specific species risk assessment codes

Data downloaded from BC Species and Ecosystems Explorer includes information on global status and provincial status for species, as assessed by NatureServe.

See [this table]('iucn_to_ns.html') for info on NatureServe codes

``` {r append_bcsee_scores, echo = TRUE}

spp_info <- spp_info_gl %>%
  mutate(sciname = str_replace(sciname, 'Clupea pallasii pallasii', 'Clupea pallasii')) %>%
  spp_append_bcsee()

spp_info <- spp_info %>%
  group_by(iucn_sid) %>%
  arrange(sciname) %>%
  filter(sciname == first(sciname) | is.na(iucn_sid) | !is.na(status_pr_score))
    ### this last bit gets rid of double-listed IUCN SIDs by selecting
    ### the first instance, while keeping any NAs; also keeps any species with a local status score

# z <- show_dupes(spp_info, 'iucn_sid', na.rm = TRUE)

### some species still duplicated due to subpops and/or synonyms.
### Mutate an average provincial status score; arrange by status score (to get non-NAs on top);
### then group by IUCN SID and filter out any duplicated ones (after the first).
spp_info <- spp_info %>%
  group_by(sciname) %>%
  mutate(mean_pr_score = mean(status_pr_score, na.rm = TRUE)) %>%
  arrange(status_pr_score) %>%
  group_by(iucn_sid) %>%
  filter(!duplicated(iucn_sid) | is.na(iucn_sid)) %>%
  mutate(status_pr_score = mean_pr_score) %>%
  select(-mean_pr_score)
  
# z <- show_dupes(spp_info, 'iucn_sid', na.rm = TRUE)

### These species are problematic in BC Howe Sound ICO:
# ico_spp_probs <- c('Ammodytes hexapterus', 'Ardea herodias', 'Clupea pallasii', 'Gavia immer', 'Haliaeetus leucocephalus', 'Hypomesus pretiosus')
# x <- spp_info %>% filter(sciname %in% ico_spp_probs)
### all there!

spp_info_local <- spp_info %>%
  mutate(iucn_score = cat_score,
         cat_score = ifelse(!is.na(status_pr_score), status_pr_score, cat_score)) %>%
  # filter(!is.na(cat_score)) %>%
  filter(am_sid %in% am_cells_spp$am_sid & str_detect(spatial_source, 'am') | 
           iucn_sid %in% iucn_cells_spp$iucn_sid & str_detect(spatial_source, 'iucn'))

# x <- spp_info_local %>% filter(sciname %in% ico_spp_probs)

write_csv(spp_info_local, file.path(dir_goal, 'int/spp_info_raw.csv'))

### NOTE: if we want to exclude BLI data, this is where we would have to do it
# spp_info <- spp_info %>%
#   mutate(spatial_source = ifelse(spatial_source == 'iucn-bli' & !is.na(am_sid), 'am', spatial_source)) %>%
#   filter(!spatial_source == 'iucn-bli')

### let's clean up this file and ditch the legacy columns.  Lose the reference
### columns, and to ditch multi-listings, group by am_sid and iucn_sid; then
### take the mean category and trend across all multi-listed species.
spp_clean <- spp_info_local %>%
  select(am_sid, iucn_sid, sciname, spp_group,
         #id_no, iucn_subpop, pop_trend, pop_cat, status_pr, 
         pr_score = status_pr_score,
         spatial_source, 
         cat_score, trend_score) %>%
  group_by(am_sid, iucn_sid, spatial_source) %>%
  summarize(sciname     = first(sciname), ### just to have a handy text handle?
            cat_score   = mean(cat_score,   na.rm = TRUE),
            pr_score    = mean(pr_score,    na.rm = TRUE),
            trend_score = mean(trend_score, na.rm = TRUE),
            pr_score    = ifelse(is.nan(pr_score),    NA, pr_score),
            trend_score = ifelse(is.nan(trend_score), NA, trend_score)) %>%
  ungroup()

# x <- spp_clean %>% filter(sciname %in% ico_spp_probs)

write_csv(spp_clean, file.path(dir_goal, 'int/spp_info_clean.csv'))

```

-----


``` {r get_species_past_assessments, eval = TRUE}

spp_cat_ts_file <- file.path(dir_goal, 'int', 'spp_cat_timeseries.csv')

if(!file.exists(spp_cat_ts_file)) {
  library(jsonlite)
  api_key <- 'fb71ae836f415f04f41176f6d30c4a9e4cea620d46b9e5021bf2fb142ea51bf5'
  
  spp_hist_url <- 'http://apiv3.iucnredlist.org/api/v3/species/history/id/%s?token=%s'
  
  spp_clean <- read_csv(file.path(dir_goal, 'int/spp_info_clean.csv'))

  sid_list <- spp_clean$iucn_sid[!is.na(spp_clean$iucn_sid)] %>% 
    unique() %>% 
    sort()
  
  get_past_assess <- function(sid) {
    spp_url <- sprintf(spp_hist_url, sid, api_key)
    spp_tmp <- fromJSON(spp_url)
    spp_df   <- spp_tmp$result %>%
      mutate(iucn_sid = as.integer(spp_tmp$name)) %>%
      mutate(code = toupper(code))
  }
  
  spp_past_list <- vector('list', length = length(sid_list))
  for (i in 1:length(sid_list)) { # i = 1
    sid <- sid_list[i]
    spp_past_list[[i]] <- get_past_assess(sid)
    message(i, '. Found past assessments for ', sid, ': ', paste(spp_past_list[[i]]$year, collapse = ', '))
  }
  spp_past_df <- bind_rows(spp_past_list)
  
  ### Clean up the time series
  
  spp_past_df <- spp_past_df %>%
    rename(cat_ts = code, cat_txt = category) %>%
    mutate(cat_ts = str_replace(cat_ts, 'LR/', ''),
           cat_ts = ifelse(cat_ts %in% c('K', 'I'), 'DD', cat_ts),
           cat_ts = ifelse(cat_ts == 'NR', 'NE', cat_ts),
           cat_ts = ifelse(str_detect(toupper(cat_txt), 'VERY RARE'), 'CR', cat_ts),
           cat_ts = ifelse(str_detect(toupper(cat_txt), 'LESS RARE'), 'T', cat_ts),
           cat_ts = ifelse(str_detect(toupper(cat_txt), 'STATUS INADEQUATELY KNOWN'), 'DD', cat_ts),
           cat_ts = ifelse(cat_ts == 'V', 'VU', cat_ts), 
           cat_ts = ifelse(cat_ts == 'E', 'EN', cat_ts))
  
  ### reclassifications:
  #  LC <- "LOWER RISK/LEAST CONCERN (LR/LC)"                         
  #  NT <- "LOWER RISK/NEAR THREATENED (LR/NT)"                       
  #  T  <- "THREATENED (T)" treat as "EN"
  #  VU <- "VULNERABLE (V)"                                           
  #  EN <- "ENDANGERED (E)"                                           
  #  LR/CD <- "LOWER RISK/CONSERVATION DEPENDENT (LR/CD)" treat as between VU and NT
  #  CR <- "VERY RARE AND BELIEVED TO BE DECREASING IN NUMBERS"       
  #  T  <- "LESS RARE BUT BELIEVED TO BE THREATENED-REQUIRES WATCHING"
  #  DD <- "INSUFFICIENTLY KNOWN (K)"                                 
  #  DD <- "INDETERMINATE (I)"                                        
  #  DD <- "STATUS INADEQUATELY KNOWN-SURVEY REQUIRED OR DATA SOUGHT" 
  #  NE <- "NOT RECOGNIZED (NR)"         
  pop_cat <- data.frame(cat_ts       = c("LC", "NT", "VU", "EN", "CR", "EX", "T", "CD", "NE", "DD"), 
                        cat_ts_score = c(   0,  0.2,  0.4,  0.6,  0.8,  1.0, 0.6,  0.3,   NA,  NA))
    
  spp_past_df <- spp_past_df %>% 
    left_join(pop_cat, by = 'cat_ts') %>%
    filter(!is.na(cat_ts_score) & !is.na(year)) %>%
    arrange(iucn_sid, year)
  
  # x <- spp_past_df %>% filter(sciname %in% ico_spp_probs)

  write_csv(spp_past_df, file.path(dir_goal, 'int', 'spp_cat_timeseries.csv'))
}

```

``` {r join_time_series_data_to_spp_list}
### Fill out the missing years, then join to cleaned species list
spp_past_df <- read_csv(file.path(dir_goal, 'int', 'spp_cat_timeseries.csv'))

spp_past_df <- spp_past_df %>%
  complete(year = full_seq(year, 1), nesting(iucn_sid)) %>%
  arrange(iucn_sid, year) %>%
  group_by(iucn_sid) %>%
  fill(cat_ts, cat_txt, cat_ts_score) %>% ### fills all the way to latest year (2015)
  ungroup()

spp_clean1 <- spp_clean %>% 
  left_join(spp_past_df, by = 'iucn_sid')

  # x <- spp_clean1 %>% filter(sciname %in% ico_spp_probs)

### if no time series available, time series years will be NA.  Assign a list to
### those NAs, then unnest it to create observations for those years.
spp_clean1 <- spp_clean1 %>%
  mutate(year = ifelse(is.na(year), list(c(1965:2015)), year)) %>% 
  unnest(year)

### NAs will be filled backward in time by starting from the most recent non-NA.
### To do this, we'll swap any current-year NAs with the cat_score (meaning no
### time series fill), and fill upwards instead of downwards.
spp_clean1 <- spp_clean1 %>%
  mutate(cat_ts_score = ifelse(year == 2015 & is.na(cat_ts), cat_score, cat_ts_score)) %>%
  arrange(iucn_sid, year) %>%
  group_by(iucn_sid) %>%
  fill(cat_ts, cat_ts_score, cat_txt, .direction = 'up') %>%
  ungroup()
  
### Now for species with a NatureServe provincial score (non-NA value 
### for pr_score) replace all cat_ts with pr_score; overwrite cat_score with
### overall values
spp_clean1 <- spp_clean1 %>%
  mutate(cat_score = ifelse(!is.na(pr_score), pr_score, cat_ts_score))

write_csv(spp_clean1, file.path(dir_goal, 'int/spp_info_timeseries.csv'))

```

-----

## Load Species information table from OHIBC SPP goal

Currently this uses the OHIBC species lookup table, which itself is based on the global species lookup table.  Find more info at `~/github/ohibc/prep/spp_ico/v2016/goal_prep_spp.Rmd`

-----

# Calculate goal model

## Summarize mean category & trend for all Howe Sound SPP

In this section we perform the model calculations:

* We calculate mean extinction risk score across all species identified in Howe Sound, based on NatureServe province-level categories where available, and IUCN categories elsewhere.
* From the region-level summary of risk, we determine the SPP goal status based on the goal model equation.  SPP trend is the linear trend in SPP status from 1995-2016.

``` {r summarize_spp}

am_spp <- read_csv(file.path(dir_goal, 'int', 'am_cells_hs_spp.csv')) %>%
  .$am_sid %>%
  unique()

iucn_spp <- read_csv(file.path(dir_goal, 'int', 'iucn_cells_hs_spp.csv')) %>%
  .$iucn_sid %>%
  unique()

spp_info_timeseries <- read_csv(file.path(dir_goal, '../v2016/int/spp_info_timeseries.csv')) %>%
  filter(iucn_sid %in% iucn_spp | am_sid %in% am_spp)

write_csv(spp_info_timeseries, file.path(dir_goal, 'int/spp_info_timeseries_hs.csv'))

spp_risk_ts <- spp_info_timeseries %>%
  select(pr_score, cat_ts_score, year) %>%
  mutate(cat_ts_score = ifelse(is.na(pr_score), cat_ts_score, pr_score)) %>%
  group_by(year) %>%
  summarize(mean_cat_ts = mean(cat_ts_score, na.rm = TRUE))

spp_status_ts <- spp_risk_ts %>%
  mutate(rgn_id = 1,
         status = (.75 - mean_cat_ts)/.75) %>%
  filter(year >= max(year) - 10)
  
spp_status_ts$trend <- spp_status_ts %>%
  do(trend = lm(status ~ year, data = .)$coefficients[2] * 5) %>%
  unlist()

write_csv(spp_status_ts, file.path(dir_goal, 'summary/spp_status_ts.csv'))

knitr::kable(spp_status_ts %>% arrange(desc(year)), caption = 'Howe Sound species status and trend')

```

``` {r save_howe_sound_status_trend}

spp_df <- read_csv(file.path(dir_goal, 'summary/spp_status_ts.csv')) %>%
  filter(year == max(year)) %>%
  select(-year)
  
write_csv(spp_df %>%
            select(rgn_id, score = status),
          file.path(dir_goal, 'output/spp_status.csv'))
write_csv(spp_df %>%
            select(rgn_id, score = trend),
          file.path(dir_goal, 'output/spp_trend.csv'))

```

### Category by taxonomic groups

``` {r plot_spp_by_cat}

am_spp <- read_csv(file.path(dir_goal, 'int', 'am_cells_hs_spp.csv')) %>%
  .$am_sid %>%
  unique()

iucn_spp <- read_csv(file.path(dir_goal, 'int', 'iucn_cells_hs_spp.csv')) %>%
  .$iucn_sid %>%
  unique()

spp_info_timeseries <- read_csv(file.path(dir_goal, '../v2016/int/spp_info_timeseries.csv')) %>%
  filter(iucn_sid %in% iucn_spp | am_sid %in% am_spp) %>%
  filter(!is.na(cat_score))

spp_table_info <- spp_info_timeseries %>%
  group_by(sciname) %>%
  arrange(year) %>%
  filter(year == last(year)) %>%
  ungroup() %>%
  select(am_sid, iucn_sid, 
         year,
         sciname, 
         cat_score, pr_score) %>%
  distinct() %>%
  mutate(spp_health = round(1 - cat_score, 1)) %>%
  arrange(desc(cat_score), desc(sciname))


### rearrange to get the NAs on top
spp_table_info <- spp_table_info %>%
  filter(is.na(spp_health)) %>%
  bind_rows(spp_table_info %>%
              filter(!is.na(spp_health))) %>%
  mutate(n_spp = 1:n())

### list status cats on the plot itself?
cat_list <- data.frame(spp_health = c(   0,      .1,    .2,      .3,    .4,      .5,    .6,      .7,    .8,      .9,   1.0),
                       cat_lab    = c('EX', 'CR/EX',  'CR', 'CR/EN',  'EN', 'EN/VU',  'VU', 'VU/NT',  'NT', 'NT/LC',  'LC'),
                       stringsAsFactors = FALSE)

spp_table_labeled <- spp_table_info %>%
  left_join(cat_list, by = 'spp_health') %>%
  mutate(sciname = str_replace(sciname, 'Pneumatophorus colias', 'Scomber japonicus'),
         sciname = str_replace(sciname, 'Pelamys chiliensis', 'Sarda chiliensis'),
         sciname = str_replace(sciname, 'Heteroscelus incanus', 'Tringa incana')) %>%
  mutate(cat_lab = ifelse(is.na(spp_health), 'NE/DD', cat_lab),
         cat_lab = ifelse(n_spp == last(n_spp), cat_lab, NA))

### get sciname lookup table to get class vs genus... 
scinames <- read_csv(file.path(dir_goal, 'raw/scinames.csv'))

spp_table_labeled <- spp_table_labeled %>%
  rowwise() %>%
  mutate(genus = tolower(str_split(sciname, ' ')[[1]][1])) %>%
  ungroup() %>%
  left_join(scinames, by = 'genus') %>%
  # group_by(com_class) %>%
  # mutate(proportion = 1/n()) %>%
  ungroup()

spp_plot <- ggplot(spp_table_labeled, aes(x = com_class, y = .1, fill = spp_health)) +
  theme(axis.ticks = element_blank(),
        text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
        plot.title = element_text(size = rel(1.25), hjust = 0, face = 'bold'),
        legend.key = element_rect(colour = NA, fill = NA),
        panel.border     = element_blank(),
        panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text.y  = element_blank(),
        axis.text.x = element_text(angle = 75, hjust = 1, 
                                   size = 10,
                                   margin = margin(-20)),
        legend.position = 'none') +
  geom_bar(stat = 'identity', width = .9, color = 'grey80', size = .2) +
  # geom_text(aes(label = cat_lab, y = 0, nudge_y = .02), 
  #           size = 1.5,
  #           color = 'black',
  #           # angle = 90, 
  #           hjust = 0) +
  scale_fill_gradient2(low = '#F27259', mid = '#FFCF5C', high = '#1F9B90', 
                       midpoint = 0.5,
                       breaks = c(0, .2, .4, .6, .8, 1.0),
                       labels = c('EX', 'CR', 'EN', 'VU', 'NT', 'LC'))+
  # scale_y_continuous(limits = c(0, 3)) +
  labs(fill = 'Species health')

print(spp_plot)

ggsave(file.path(dir_goal, 'spp_plot_lines2.png'),
       width = 6, height = 9, units = 'in', dpi = 300)
```

***

``` {r provenance, results = 'asis'}

prov_wrapup()

```
