---
title: 'OHIBC Howe Sound goal prep: Species'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr2.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = FALSE, message = FALSE, warning = FALSE)

source('~/github/ohibc/src/R/common.R')  ### an OHIBC specific version of common.R

scenario <- 'vHS'
goal     <- 'spp_ico'
dir_git <- '~/github/ohibc'
dir_goal <- file.path(dir_git, 'prep', goal, scenario)
dir_rgn  <- file.path(dir_git, 'prep/spatial')
dir_anx        <- file.path(dir_M, 'git-annex/bcprep') ### git-annex: goal-specific large files
dir_anx_global <- file.path(dir_M, 'git-annex/globalprep/spp_ico')



source(file.path(dir_git, 'src/R/map_scores.R'))

library(provRmd)
prov_setup()

source(file.path(dir_goal, 'spp_fxn.R'))


### set up proj4string options: BC Albers and WGS84
p4s_wgs84 <- c('EPSG:4326 WGS 84' = '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0')
p4s_bcalb <- c('EPSG:3005 NAD83/BC Albers' = '+proj=aea +lat_1=50 +lat_2=58.5 +lat_0=45 +lon_0=-126 +x_0=1000000 +y_0=0 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0')
```

# Summary: OHIBC Howe Sound Species

This script prepares scores (status and trend) for species richness in 
British Columbia's Howe Sound region.  Spatial data from IUCN and Aquamaps is
combined with extinction risk information from IUCN and conservation rank
info based on province-level NatureServe categories.

Because of the small scale of Howe Sound, neither Species nor Iconic Species goals use any area weighting.  Instead, ICO and SPP status is based upon a simple average of species
health for species found within the Howe Sound region.

From Halpern et al (2012):

> The target for the Species sub-goal is to have all species at a risk status of Least Concern. We scaled the lower end of the biodiversity goal to be 0 when 75% species are extinct, a level comparable to the five documented mass extinctions and would constitute a catastrophic loss of biodiversity. The Status of assessed species was calculated as the area- and threat status-weighted average of the number of threatened species within each 0.5 degree grid cell.

**Mean risk status for Howe Sound:**

$$\bar{R} = \frac{\displaystyle\sum_{species}(Risk)}{n_{spp}}$$

**Species goal model**

$$X_{SPP} = \frac{((1 - \bar{R}_{SPP}) - 0.25)}{(1 - 0.25)} * 100%$$

where:

* $X_{SPP}$ is Species goal status
* $\bar{R}$ is mean extinction risk for identified species within Howe Sound (different subsets for ICO and SPP)
* *Risk* is scaled value for species extinction risk category, based on: 
    * 'LC' = 0.0, 'NT' = 0.2, 'VU' = 0.4, 'EN' = 0.6, 'CR' = 0.8, 'EX' = 1.0
* SPP trend is calculated as the linear trend of the average extinction risk categories over time.

-----

# Data sources

AquaMaps

* From http://www.aquamaps.org/main/home.php: "AquaMaps are computer-generated predictions of natural occurrence of marine species, based on the environmental tolerance of a given species with respect to depth, salinity, temperature, primary productivity, and its association with sea ice or coastal areas."
* Citation: Kaschner, K., J. Rius-Barile, K. Kesner-Reyes, C.Garilao, S.O. Kullander, T. Rees and R. Froese (2013). AquaMaps: Predicted range maps for aquatic species. World wide web electronic publication, www.aquamaps.org, Version 08/2013.

IUCN Red List spatial data: species range map shapefiles

* Species range map shapefiles downloaded from: http://www.iucnredlist.org/technical-documents/spatial-data
* Citation: IUCN 2014. The IUCN Red List of Threatened Species. Version 2014.1. <http://www.iucnredlist.org>. Downloaded on 21 September 2015.

IUCN Red List species index: list of all IUCN red list species, incl IUCN species ID and extinction risk category

* Extinction risk categories downloaded from: http://api.iucnredlist.org/index/all.csv
* Habitat and subpopulation information, by species, scraped from: http://www.iucnredlist.org/details/<species ID>/0
* Citation: IUCN 2015. The IUCN Red List of Threatened Species. Version 2015-3. <http://www.iucnredlist.org>. Downloaded on 9 September 2015.

NatureServe/BC CDC conservation rank info from BC Species and Ecosystems Explorer: 

* BC Conservation Data Centre: http://www.env.gov.bc.ca/atrisk/toolintro.html
* NatureServe conservation status: http://explorer.natureserve.org/ranking.htm#global
* Citation: B.C. Conservation Data Centre. 2013. BC Species and Ecosystems Explorer. B.C. Ministry of Environment, Victoria B.C. Available: <http://a100.gov.bc.ca/pub/eswp/>.  Downloaded on 3 August 2015.

-----

# Methods

## Define assessment region

Using OHIBC Howe Sound region polygon, determine 0.5° raster cells corresponding to the region.

``` {r setup_region-to-cell_lookup}

### Read in OHIBC Howe Sound polygons, and transform to WGS84 CRS
poly_howesound_rgn <- readOGR(dsn = path.expand(dir_rgn), layer = 'howe_sound_rgn',
                       verbose = FALSE, stringsAsFactors = FALSE) %>%
  spTransform(CRS(p4s_wgs84))

poly_howesound_rgn@data <- poly_howesound_rgn@data %>%
  rename(rgn_id = id) %>%
  mutate(rgn_name = 'Howe Sound')

rgn2cell_list <- spp_rgn2cell(poly_howesound_rgn, 
                              rgn_tag = '_howesound',
                              reload = TRUE)

rgn2cell_df      <- rgn2cell_list[[1]]
loiczid_rast     <- rgn2cell_list[[2]]
```


## Identify Howe Sound-specific species

These are identified by comparing the full AM and IUCN cell-species files to the limited cells that appear in the Howe Sound region.

``` {r identify_HS_species}

am_cells_spp <- spp_get_am_cells(rgn2cell_df, prob_filter = 0, reload = FALSE) %>%
  filter(loiczid %in% rgn2cell_df$loiczid)

write_csv(am_cells_spp, file.path(dir_goal, 'int', 'am_cells_hs_spp.csv'))

iucn_cells_spp <- spp_get_iucn_cells(rgn2cell_df, reload = FALSE) %>%
  filter(loiczid %in% rgn2cell_df$loiczid)

write_csv(iucn_cells_spp, file.path(dir_goal, 'int', 'iucn_cells_hs_spp.csv'))

```

Number of Howe Sound species according to dataset:

* AquaMaps: `r length(am_cells_spp$am_sid %>% unique())`
* IUCN: `r length(iucn_cells_spp$iucn_sid %>% unique())`

-----

## Load Species information table from OHIBC SPP goal

Currently this uses the OHIBC species lookup table, which itself is based on the global species lookup table.  Find more info at `~/github/ohibc/prep/spp_ico/v2016/goal_prep_spp.Rmd`

-----

# Calculate goal model

## Summarize mean category & trend for all Howe Sound SPP

In this section we perform the model calculations:

* We calculate mean extinction risk score across all species identified in Howe Sound, based on NatureServe province-level categories where available, and IUCN categories elsewhere.
* From the region-level summary of risk, we determine the SPP goal status based on the goal model equation.  SPP trend is the linear trend in SPP status from 1995-2016.

``` {r summarize_spp}

am_spp <- read_csv(file.path(dir_goal, 'int', 'am_cells_hs_spp.csv')) %>%
  .$am_sid %>%
  unique()

iucn_spp <- read_csv(file.path(dir_goal, 'int', 'iucn_cells_hs_spp.csv')) %>%
  .$iucn_sid %>%
  unique()

spp_info_timeseries <- read_csv(file.path(dir_goal, '../v2016/int/spp_info_timeseries.csv')) %>%
  filter(iucn_sid %in% iucn_spp | am_sid %in% am_spp) %>%

spp_risk_ts <- spp_info_timeseries %>%
  select(pr_score, cat_ts_score, year) %>%
  mutate(cat_ts_score = ifelse(is.na(pr_score), cat_ts_score, pr_score)) %>%
  group_by(year) %>%
  summarize(mean_cat_ts = mean(cat_ts_score, na.rm = TRUE))

spp_status_ts <- spp_risk_ts %>%
  mutate(rgn_id = 1,
         status = (.75 - mean_cat_ts)/.75)
  
spp_status_ts$trend <- spp_status_ts %>%
  do(trend = lm(status ~ year, data = .)$coefficients[2] * 5) %>%
  unlist()

write_csv(spp_status_ts, file.path(dir_goal, 'summary/spp_status_ts.csv'))

DT::datatable(spp_status_ts %>% arrange(desc(year)), caption = 'Howe Sound species status and trend')

```

``` {r save_howe_sound_status_trend}

spp_df <- read_csv(file.path(dir_goal, 'summary/spp_status_ts.csv')) %>%
  filter(year == max(year)) %>%
  select(-year)
  
write_csv(spp_df %>%
            select(rgn_id, score = status),
          file.path(dir_goal, 'output/spp_status.csv'))
write_csv(spp_df %>%
            select(rgn_id, score = trend),
          file.path(dir_goal, 'output/spp_trend.csv'))
```

``` {r plot_spp_by_cat}

spp_info_timeseries <- read_csv(file.path(dir_goal, '../v2016/int/spp_info_timeseries.csv')) %>%
  filter(iucn_sid %in% iucn_spp | am_sid %in% am_spp) %>%
  filter(!is.na(cat_score))

spp_table_info <- spp_info_timeseries %>%
  group_by(sciname) %>%
  arrange(year) %>%
  filter(year == last(year)) %>%
  ungroup() %>%
  select(am_sid, iucn_sid, 
         year,
         sciname, 
         cat_score, pr_score) %>%
  distinct() %>%
  mutate(spp_health = round(1 - cat_score, 1)) %>%
  arrange(desc(cat_score), desc(sciname))


### rearrange to get the NAs on top
spp_table_info <- spp_table_info %>%
  filter(is.na(spp_health)) %>%
  bind_rows(spp_table_info %>%
              filter(!is.na(spp_health))) %>%
  mutate(n_spp = 1:n())

### list status cats on the plot itself?
cat_list <- data.frame(spp_health = c(   0,      .1,    .2,      .3,    .4,      .5,    .6,      .7,    .8,      .9,   1.0),
                       cat_lab    = c('EX', 'CR/EX',  'CR', 'CR/EN',  'EN', 'EN/VU',  'VU', 'VU/NT',  'NT', 'NT/LC',  'LC'),
                       stringsAsFactors = FALSE)

spp_table_labeled <- spp_table_info %>%
  left_join(cat_list, by = 'spp_health') %>%
  mutate(sciname = str_replace(sciname, 'Pneumatophorus colias', 'Scomber japonicus'),
         sciname = str_replace(sciname, 'Pelamys chiliensis', 'Sarda chiliensis'),
         sciname = str_replace(sciname, 'Heteroscelus incanus', 'Tringa incana')) %>%
  mutate(cat_lab = ifelse(is.na(spp_health), 'NE/DD', cat_lab),
         cat_lab = ifelse(n_spp == last(n_spp), cat_lab, NA))

### get sciname lookup table to get class vs genus... 
scinames <- read_csv(file.path(dir_goal, 'raw/scinames.csv')) %>%
  filter(class != 'chondrichthyes') %>%
  select(-order, -family) %>%
  filter(!(genus %in% c('eremophila', 'rondeletia', 'arenaria') & class == 'magnoliopsida')) %>%
  distinct()
### some dupes still in here: all duplicated in class magnoliopsida
### Eremophila is a bird
### Rondeletia is a fish
### Arenaria is a bird

spp_table_labeled <- spp_table_labeled %>%
  rowwise() %>%
  mutate(genus = tolower(str_split(sciname, ' ')[[1]][1])) %>%
  ungroup() %>%
  left_join(scinames, by = 'genus') %>%
  group_by(class) %>%
  mutate(proportion = 1/n())

spp_plot <- ggplot(spp_table_labeled, aes(x = class, y = proportion, fill = spp_health)) +
  ggtheme_plot + 
  theme(panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text.y  = element_blank(),
        axis.text.x = element_text(angle = 75, hjust = 1),
        legend.position = 'none') +
  geom_bar(stat = 'identity', width = .9, color = 'grey20', size = .25) +
  # geom_text(aes(label = cat_lab, y = 0, nudge_y = .02), 
  #           size = 1.5,
  #           color = 'black',
  #           # angle = 90, 
  #           hjust = 0) +
  scale_fill_gradient2(low = 'red4', mid = 'orange', high = 'green4', 
                       midpoint = 0.5,
                       breaks = c(0, .2, .4, .6, .8, 1.0),
                       labels = c('EX', 'CR', 'EN', 'VU', 'NT', 'LC'))+
  # scale_y_continuous(limits = c(0, 3)) +
  labs(fill = 'Species health')

ggsave(file.path(dir_goal, 'spp_plot.png'),
       width = 6, height = 9, units = 'cm')
```


***

``` {r provenance, results = 'asis'}

prov_wrapup()

```
