---
title: 'OHIBC: Ocean Acidification Pressure layers prep'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = FALSE, message = FALSE, warning = FALSE)

library(sp)        # the classes and methods that make up spatial ops in R
library(rgdal)
library(raster)

dir_git <- '~/github/ohibc'
source(file.path(dir_git, 'src/R/common.R'))
  ### an OHIBC specific version of common.R
dir_anx <- file.path(dir_M, 'git-annex/bcprep')
dir_rgn <- file.path(dir_git, 'prep/regions')  ### github: general buffer region shapefiles

### goal specific folders and info
goal      <- 'pressures'
scenario  <- 'v2016'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario)
dir_goal_anx <- file.path(dir_anx, goal, scenario)

### set up provenance tracking for this script:
source(file.path(dir_git, 'src/R/prov.R'))          
  ### initializes prov_track and loads functions

### goal-specific source scripts
source(file.path(dir_goal, 'pressures_lyr_fxns.R'))

### other support functions
source(file.path(dir_git, 'src/R/rast_tools.R'))

### set up proj4string options: BC Albers and WGS84
p4s_wgs84 <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
p4s_bcalb <- '+proj=aea +lat_1=50 +lat_2=58.5 +lat_0=45 +lon_0=-126 +x_0=1000000 +y_0=0 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0'

reload = FALSE
```

# Summary

There are two parts to creating this layer:

1. Data prep to get raw data into the correct format:
    * Aggregate monthly values for aragonite saturation state to an annual mean value per cell, for baseline decade and current study period.
    * Clip rasters to extents of OHIBC region, and reproject into BC Albers projection.
    * Determine baseline by calculating decadal mean for 1880-1889.
2. Creating the pressure layers for OHIBC:
    * Determine change in aragonite saturation state since baseline for each year of study period (2005-2014)
    * Rescale to values from 0 - 1.
    * Interpolate to coastline at 1 km resolution.
    
This process is completed entirely within this script.

# Updates

Not applicable.

# Data

Raw data was provided by Woods Hole on December 16, 2014. This data is an update to the work done by Feely et al. (2009).

* Native Data Resolution: 1 degree
* Values: Surface Ω aragonite saturation state
* Time Range: 1880-1889 and 2005-2014 (monthly data was provided for each year)
* Format: NetCDF

# Methods

## Data Prep

### Read raw data

Raw data is read in for 1880 - 1889 and 2005 - 2014 datasets.  Latitude, longitude, and aragonite saturation state $\Omega$ are extracted for each year in each dataset.  Land cells are set to NA.

``` {r read_aragonite_lyrs, echo = FALSE}

library(ncdf4)

### read in 1880-1889 data
oa_data_dir <- file.path(dir_M, 'git-annex/globalprep/prs_oa/v2015')
data_nc_18 <- nc_open(file.path(oa_data_dir, 'input/cesm_co2sys_1880-1889.nc'))
git_prov(file.path(oa_data_dir, 'input/cesm_co2sys_1880-1889.nc'), filetype = 'input')
# print(data_nc_18)

### read in 2005-2014 data
data_nc_20 = nc_open(file.path(oa_data_dir, 'input/cesm_co2sys_2005-2014.nc'))
git_prov(file.path(oa_data_dir, 'input/cesm_co2sys_2005-2014.nc'), filetype = 'input')
# print(data_nc_20)

### longitude values are stored in the variable 'TLONG'; lat in 'TLAT'
long <- ncvar_get(data_nc_18, varid = 'TLONG') #same lat and long for both data sets so just reading in from data_nc_18 works
lat  <- ncvar_get(data_nc_18, varid = 'TLAT')

### select the surface aragonite variable (OARG)
arag_18 <- ncvar_get(data_nc_18, varid = 'OARG')
arag_20 <- ncvar_get(data_nc_20, varid = 'OARG')

### land cells are given large, strange value, so set these to NA
### arag cell values range from zero to 5; set all out of that to NA?
land_val <- max(arag_18, na.rm = TRUE)
arag_18[arag_18 >= land_val] <- NA
range_a18 <- round(range(arag_18, na.rm = TRUE), 3)
arag_20[arag_20 >= land_val] <- NA 
range_a20 <- round(range(arag_20, na.rm = TRUE), 3)

```

Range of $\Omega_{1880-1889}$ for global dataset: `r paste(range_a18, collapse = ' - ')`.

Range of $\Omega_{2005-2014}$ for global dataset: `r paste(range_a20, collapse = ' - ')`.

### Convert raw global data to OHIBC region data

The raw data, in WGS84 coordinate ref system, is rasterized, clipped down to a bounding box around the OHIBC region, and reprojected into BC Albers projection.  This data is then resampled to 1 km resolution using nearest neighbor method.

The resulting raster layers (in rasterBrick format) include annual means for each coarse cell (approx 1 degree, from original resolution) for each year in the dataset.  The two decades (1880-1889 and 2005-2014) are assigned to two separate rasterBricks.

``` {r load_rgn, echo = FALSE}

rgn_lyr <- 'ohibc_rgn'

message(sprintf('Reading OHIBC regions shapefile...\n  %s/%s.shp', dir_rgn, rgn_lyr))
rgn_poly <- readOGR(dsn = path.expand(dir_rgn), layer = rgn_lyr,
                       verbose = FALSE, stringsAsFactors = FALSE)

rgn_poly_wgs84 <- readOGR(dsn = path.expand(dir_rgn), 
                         layer = paste(rgn_lyr, '_wgs84', sep = ''),
                         verbose = FALSE, stringsAsFactors = FALSE)

### In case we need to resample to 1 km... use base raster from 
### regions folder (note: 500m base raster available too).  Used later in
### finding zonal statistics anyway.
rast_rgn <- raster(file.path(dir_rgn, 'ohibc_rgn_raster_1000m.tif'))

```

``` {r create_rast_bricks, echo = FALSE}
### generate raster bricks for decades 1880-1889 and 2005-2014
### then: clip to extents of region (in WGS 84 CRS)
### then: reproject to BC Albers projection
rb_a18_file <- file.path(dir_goal, 'int/oa/rbrick_a18_1km_raw.tif')
rb_a20_file <- file.path(dir_goal, 'int/oa/rbrick_a20_1km_raw.tif')


if(!file.exists(rb_a18_file) | reload) {
  message('Creating raster bricks from raw data')
  rbrick_arag_18 <- annual_mean(arag_18, yr_init = 1880) %>%
    raster::crop(rgn_poly_wgs84, snap = 'out') %>%
    raster::projectRaster(crs = p4s_bcalb)
  rbrick_arag_20 <- annual_mean(arag_20, yr_init = 2005) %>%
    crop(rgn_poly_wgs84, snap = 'out') %>%
    raster::projectRaster(crs = p4s_bcalb)

  
  ### why does resample fail on a raster brick:
  # rb_arag_18_resamp <- raster::resample(rbrick_arag_18, rast_base,
  #                              method   = 'ngb',
  #                              progress = 'text',
  #                              filename = file.path(dir_goal, 'int/oa/rbrick_arag_18_raw.tif'),
  #                              overwrite = TRUE)
  # rb_arag_20_resamp <- resample(rbrick_arag_20, rast_base,
  #                              method   = 'ngb',
  #                              progress = 'text',
  #                              filename = file.path(dir_goal, 'int/oa/rbrick_arag_20_raw.tif'),
  #                              overwrite = TRUE)
  
  message('Resampling raster bricks to new resolution')
  rb_a18_resamp <- resample_brick(rbrick_arag_18, rast_rgn,
                                 filename = rb_a18_file)
  rb_a20_resamp <- resample_brick(rbrick_arag_20, rast_rgn,
                                 filename = rb_a20_file)
} else {
  message('Reading raster bricks from files')
  rb_a18_resamp <- brick(rb_a18_file)
  rb_a20_resamp <- brick(rb_a20_file)
  
  names(rb_a18_resamp) <- paste('y', c(1880:1889), sep = '')
  names(rb_a20_resamp) <- paste('y', c(2005:2014), sep = '')
}

plot_rast(rb_a18_resamp[['y1880']], 
          title = 'Annual mean arag. sat state, 1880', 
          scale_label = 'omega',
          rev_scale = FALSE)

plot_rast(rb_a20_resamp[['y2014']], 
          title = 'Annual mean arag. sat state, 2014', 
          scale_label = 'omega',
          rev_scale = FALSE)

```

### Calculate baseline

Determine the mean cell value across the entire decade from 1880-1889, as a baseline for ocean aragonite saturation levels.

```{r calculate_decadal_mean}
### Get decadal mean of 1880-1889 as baseline
oa_mean_baseline <- calc(rb_a18_resamp, mean)
# plot(oa_mean_baseline)
# plot(rgn_poly, add = TRUE)
plot_rast(oa_mean_baseline, 
          title = 'Decadal mean arag. sat state, 1880-1889', 
          scale_label = 'omega',
          rev_scale = FALSE)
```

## Rescale values as pressures

Baseline aragonite saturation state $\Omega_{base}$ is based upon the decadal average from 1880-1889.

Deviation from aragonite saturation state is determined for each year in the study period:

$$\Delta \Omega_{year} = \frac{(\Omega_{base} - \Omega_{year})}{(\Omega_{base} - 1)}$$

Note that current is subtracted from the baseline; this way, a reduction in $\Omega$ becomes a positive pressure value.  It is then normalized by the current mean state; so a decrease in $\Omega$ while the current state is high indicates less pressure than the same decrease when the current state is near 1. 

$\Delta \Omega_{year}$ is then modified to account for increases in aragonite saturation state (pressure = 0) and arag sat state less than 1 (pressure = 1).

``` {r rescale_values, echo = FALSE}

### get raster brick of change in aragonite sat state.
### Note that by dividing by the distance from baseline to 1, we are
### normalizing - so a given drop in omega, with a high current sat state, 
### will result in a lower pressure than the same drop at a lower 
### saturation state.  Dropping very close to 1 creates a pressure 
### approaching 1.0.

oa_delta <- (oa_mean_baseline - rb_a20_resamp)
oa_delta <- oa_delta/(oa_mean_baseline - 1)
names(oa_delta) <- names(rb_a20_resamp)

### If current mean sat state > 1, set pressure to 1.  If delta
### is negative (current state > baseline state) set to zero. 
oa_rescaled <- oa_delta
oa_means <- rb_a20_resamp

values(oa_rescaled)[values(oa_means) <= 1] <- 1    
  ### all values at or less than 1 are given a value of 1
values(oa_rescaled)[values(oa_delta)  < 0] <- 0
  ### all delta less than 0 (indicating a decrease in acidity) are capped at 0
plot_rast(oa_rescaled[['y2014']], 
          title = 'OA rescaled 2014', 
          scale_label = 'oa',
          rev_scale = TRUE)

```

### Interpolate layer to coastline at 1 km

Now that $\Delta \Omega_{year}$ has been determined at a 1 degree cell scale, we interpolate that to the coastlines to fill NAs.  Using `raster::interpolate()`, we select a random sample of points from the $\Delta \Omega_{year}$ raster (for speed and memory) and apply the thin plate spline model (`fields::Tps()`).

``` {r interpolate_mask_save, echo = FALSE, message = FALSE}
n_samples = 2000
oa_raster_file <- c(file.path(dir_goal, sprintf('output/oa/oa_rescaled_2005-2014_%spts.tif', n_samples)))


if(!file.exists(oa_raster_file) | reload) {
  ### from raster bricks, interpolate/extrapolate to extent of raster.  First,
  ### set up thin plate spline model from x-y of raster and values of raster
  oa_rescaled_int <- interpolate_brick(oa_rescaled, subset_n = n_samples)
    ### sampling to avoid memory problems; the Tps model seems to choke
    ### with too many points
  
  # plot(oa_rescaled_int)
  # plot(oa_rescaled)
  
  ### mask to OHIBC regions
  rbrick <- mask(oa_rescaled_int, rgn_poly)
  # plot(oa_resc_int_rgn[[1]])
  # plot(rgn_poly, add = TRUE)
  
  writeRaster(rbrick,
              filename  = oa_raster_file, 
              bylayer   = FALSE, 
              bandorder = 'BIL',
              overwrite = TRUE)
} else {
  rbrick <- brick(oa_raster_file)
  names(rbrick) <- paste('y', c(2005:2014), sep = '')
}

plot_rast(rbrick[['y2014']], 
          title = 'OA rescaled 2014, interpolated to 1 km', 
          scale_label = 'oa',
          rev_scale = TRUE)

```

## Calculate mean pressure by region by year

Using 1000 m region ID raster to run zonal statistics on the pressures layers (within brick or separately?) we can calculate the regional average pressures for each year of study, and save as a .csv.

``` {r calc_mean_pressures_by_year}

oa_rgn_mean <- zonal(rbrick, rast_rgn, fun = 'mean')

oa_rgn_mean_df <- oa_rgn_mean %>%
  as.data.frame() %>%
  gather(year, pressure, y2005:y2014) %>%
  rename(rgn_id = zone) %>%
  mutate(pressure = round(pressure, 3),
         year     = as.integer(str_replace(year, 'y', ''))) %>%
  left_join(rgn_poly@data %>% select(rgn_id, rgn_name), by = 'rgn_id') %>%
  arrange(rgn_id)

oa_plot <- ggplot(oa_rgn_mean_df, aes(x = year, y = pressure, group = rgn_name)) +
  ggtheme_plot + 
  geom_point(aes(color = rgn_name)) +
  geom_line(aes(color = rgn_name)) +
  scale_x_continuous(breaks = oa_rgn_mean_df$year) + 
  labs(y = 'oa pressure',
         title = 'OA pressure by region, 2005:2014')

print(oa_plot)

```

## Smooth annual pressures with running average

For each study year, calculate a running average of pressure score for each region, using the study year and the prior X number of years to determine the mean pressure.

``` {r calc_mean_pressures_running_avg}

num_yrs <- 3  ### calc 3 yr running average by region

study_yrs <- c(2005:2014)

oa_running_avg <- data.frame()
for (yr in study_yrs) { ### yr <- 2005 ### yr <- 2014
  yr_first <- max(yr - num_yrs + 1, min(oa_rgn_mean_df$year)) ###
  oa_block <- oa_rgn_mean_df %>%
    filter(year <= yr & year >= yr_first) %>%
    group_by(rgn_id) %>%
    summarize(year = yr, pressure_mean = mean(pressure))
  oa_running_avg <- rbind(oa_running_avg, oa_block)
}

oa_rgn_mean_df <- oa_rgn_mean_df %>% 
  left_join(oa_running_avg, by = c('rgn_id', 'year'))
  

oa_plot <- ggplot(oa_rgn_mean_df, aes(x = year, y = pressure_mean, group = rgn_name)) +
  ggtheme_plot + 
  geom_point(aes(color = rgn_name)) +
  geom_line(aes(color = rgn_name)) +
  scale_x_continuous(breaks = oa_rgn_mean_df$year) + 
  labs(y = 'oa pressure',
         title = 'OA pressure by region, 3 year avg, 2005:2014')

print(oa_plot)

write_csv(oa_rgn_mean_df, file.path(dir_goal, 'output/oa/oa_rgn_pressures.csv'))

```

-----

``` {r child = file.path(dir_git, 'src/templates/ohibc_prov_ftr.Rmd')}
```
