---
title: 'OHIBC: Calculate All Scores'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes:
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(ohicore) ### devtools::install_github('ohi-science/ohicore')

source('~/github/ohibc/src/R/common.R')

dir_ohibc  <- '~/github/ohibc'
dir_calc   <- file.path(dir_ohibc, 'calc_ohibc')
dir_master <- file.path(dir_calc, 'master')

source(file.path(dir_calc, 'calc_scores_fxns.R'))

### provenance tracking
# library(provRmd); prov_setup()

```


# Set up scenarios

Set up scenarios by assigning years.  Each scenario (eg. 'region2015') will be tied to a specific data year, not floating as in global OHI.

* Prepare scenario folders
  * create scenario dirs and subdirs
  * copy config files from master to scenarios
* Copy layers from prep to layers folder #####
  * Read in the layers.csv file with paths to the data files
  * Register layers based on include == TRUE and identify file locations
  * check that all data layers are available based on file paths
  * copy layers into current layers folder
* Check that all layers conform with details in config (`layers.csv`)

``` {r set up scenario directories and files}

scenario_years <- c(2005:2015) %>%
  setNames(paste0('region', .))

prep_scenario_dirs(dir_calc)

### copy layers from data sources to layers folder
message('Copying all layers to ', file.path(dir_calc, 'layers'))

layers_log <- read_csv(file.path(dir_master, 'layers_ohibc.csv')) %>%
  mutate(dir_prep = file.path(dir_ohibc, str_replace(dir_prep, 'ohibc:', '')))

lyrs <- register_layers(layers_log, dir_calc)

verify_layers(lyrs)

copy_layers_to_scenario(lyrs)

write_csv(lyrs %>% dplyr::select(-path_in, -path_in_exists, -path_out),
          file.path(dir_calc, 'layers.csv'))

### confirm all layers conform to details of layers.csv
conf   <- Conf(file.path(dir_calc, 'conf'))

CheckLayers(layers.csv = file.path(dir_calc, 'layers.csv'),
            layers.dir = file.path(dir_calc, 'layers'),
            flds_id    = conf$config$layers_id_fields)

```


# Iterate across scenario years

* Initialize conf, layers; store working directory in `layers$data$dir_calc` so it can be accessed by `functions.R`
* Initialize scores_all dataframe to hold scores for all years
* Calculate scores for each scenario and append to scores_all

``` {r iterate across years}

conf   <- Conf(file.path(dir_calc, 'conf'))
layers <- Layers(layers.csv = file.path(dir_calc, 'layers.csv'),
                 layers.dir = file.path(dir_calc, 'layers'))
layers$data$dir_calc    <- dir_calc

scores_all <- data.frame() ### initialize scores

for (yr_i in seq_along(scenario_years)) { ### yr_i <- 1

  scenario <- names(scenario_years)[yr_i]
  status_year <- scenario_years[yr_i]
  message('Calculating scores for ', status_year)

  ### For each run through loop, assign status_year inside
  ###   the layers object/env't so it is accessible to functions.R
  layers$data$status_year <- status_year

  scores <- CalculateAll(conf, layers) %>%
    mutate(year = status_year)

  scores_all <- scores_all %>%
    bind_rows(scores)

}

write_csv(scores_all, file.path(dir_calc, 'scores_all.csv'))

x <- scores_all %>%
  filter(dimension %in% c('status', 'trend')) %>%
  filter(!(score == 0 & dimension == 'status'))

```

``` {r plot_all_scores}

scores <- read_csv(file.path(dir_calc, 'scores_all.csv')) %>%
  filter(dimension %in% c('status', 'trend'))





```

``` {r prov_footer, results = 'asis'}
prov_wrapup(commit_outputs = FALSE)
```

