---
title: 'OHIBC: Resilience - MaPP social resilience'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = FALSE, message = FALSE, warning = FALSE)

library(sf)

dir_git <- '~/github/ohibc'
source(file.path(dir_git, 'src/R/common.R'))

dir_anx <- file.path(dir_M, 'git-annex/bcprep')
dir_spatial <- path.expand(file.path(dir_git, 'prep/_spatial'))

### goal specific folders and info
goal      <- '_resilience'
scenario  <- 'v2017'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario)
dir_goal_anx <- file.path(dir_M, 'git-annex/bcprep', goal, scenario)

library(provRmd); prov_setup()

```

# Summary

This pressure layer calculates social resilience based on the MaPP process.  

$$resilience_{rgn} = \mathbb{1}_{rgn \in MaPP} \times \frac{process + enforcement + compliance}{3}$$

From the start of the MaPP process, each region within MaPP will receive a resilience score of 1/3.  Enforcement and compliance would begin to contribute to the score once the MaPP plans are released in 2015, but since only two of the three parties (First Nations, province, but not DFO) each region will receive only partial credit.

``` {r}

df <- expand.grid(rgn_id = c(1:4, 8), year = 2001:2016) %>%
  mutate(process = ifelse(year >= 2011, 1, 0),
         enforce = ifelse(year >= 2015, 2/3, 0),
         comply  = ifelse(year >= 2015, 2/3, 0)) %>%
  group_by(rgn_id, year) %>%
  summarize(mapp_resilience = (process + enforce + comply) / 3) %>%
  ungroup()

write_csv(df, file.path(dir_goal, 'output', 'mapp_score.csv'))

```

-----

``` {r results = 'asis'}

prov_wrapup(commit_outputs = FALSE)

```
