---
title: 'OHIBC: Resilience - Food provision biomass removal resilience prep'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = FALSE, message = FALSE, warning = FALSE)

library(sf)

dir_git <- '~/github/ohibc'
source(file.path(dir_git, 'src/R/common.R'))

dir_anx <- file.path(dir_M, 'git-annex/bcprep')
dir_spatial <- path.expand(file.path(dir_git, 'prep/_spatial'))

### goal specific folders and info
goal      <- '_resilience'
scenario  <- 'v2017'
dir_goal  <- file.path(dir_git, 'prep', goal, scenario)
dir_goal_anx <- file.path(dir_M, 'git-annex/bcprep', goal, scenario)

library(provRmd); prov_setup()

```

# Summary

This script prepares a resilience layer based upon the presence, effectiveness, and compliance with fishing management regulations.  As in several other resilience layers, it is a geometric mean of three components: 

$$resilience = \sqrt[3]{reg.presence \times reg.effectiveness \times reg.compliance}$$

# Data

* __Reference__: 
* __Downloaded__: 
* __Description__:  
* __Native data resolution__: 
* __Format__:  

# Methods

Read in layers and compute geometric mean of values for each year.  Reference points:

* Regulation: 100% indicates Fisheries Act in place; this is true for all years.
* Effectiveness: Metric = Fisheries officers/vessels each year (target is max number of officers per vessels across time).
    * Number of vessels will be determined from license data.
* Compliance: Metric = Observer coverage in groundfish fishery (target 100%)

## Determine regulation score

Since regulation is long-standing, all years earn a regulation component score of 100.

## Determine effectiveness

Using officers and vessels data (from fishing license data) determine ratio of officers per vessel, and determine peak as reference point.

To determine vessel count, take the license data, filter to vessel-based licenses, clean vessel names by lower-casing and stripping all non-letter/number info, and then using distinct() to eliminate duplicates (e.g. a vessel with multiple licenses in one year).  Then summarize the count for each year.  This is not spatialized - accounts for all licensed vessels in BC waters.

``` {r}

license_file <- file.path(dir_anx, '_raw_data/fishing_license/Pacific Region Commercial Fishing Licenses En.csv')
metadata_file <- file.path(dir_anx, '_raw_data/fishing_license/Commercial License Meta Data EN (1).csv')

licensed_vessels <- read_csv(license_file) %>%
  setNames(tolower(names(.)) %>% str_replace_all('[\\s()]+', '_')) %>%
  filter(licence_type == 'VESSEL-BASED') %>%
  select(year = licence_suffix, vessel = licence_holder_vessel) %>%
  mutate(vessel = tolower(vessel) %>% str_replace_all('[^a-z0-9]', '')) %>%
  distinct() %>%
  group_by(year) %>%
  summarize(vessels = n()) %>%
  filter(year != 2017) ### drop 2017 b/c incomplete data
  

res_data_file <- file.path(dir_anx, '_raw_data/fisheries_resilience/fisheries_resilience.xlsx')
### readxl::excel_sheets(res_data_file)
### [1] "GroundfishEnforcement" "GroundfishMonitoring"  "ProtectAreaPlans"      "ReferenceNotes"       

reg_eff <- readxl::read_excel(res_data_file, sheet = 'GroundfishEnforcement') %>%
  setNames(c('year', 'officers', 'vessels')) %>%
  select(year, officers)

reg_eff <- reg_eff %>%
  left_join(licensed_vessels, by = 'year') %>%
  mutate(eff_ratio = officers / vessels,
         eff_score = eff_ratio / max(eff_ratio, na.rm = TRUE))

DT::datatable(reg_eff %>% mutate(eff_ratio = round(eff_ratio, 4), eff_score = round(eff_score, 4)))

write_csv(reg_eff, file.path(dir_goal, 'int', 'fp_reg_effectiveness.csv'))

```

## Determining compliance

``` {r} 
res_data_file <- file.path(dir_anx, '_raw_data/fisheries_resilience/fisheries_resilience.xlsx')
### readxl::excel_sheets(res_data_file)
### [1] "GroundfishEnforcement" "GroundfishMonitoring"  "ProtectAreaPlans"      "ReferenceNotes"       
reg_comp <- readxl::read_excel(res_data_file, sheet = 'GroundfishMonitoring') %>%
  setNames(c('year', 'observers')) %>%
  mutate(comp_score = observers / max(observers, na.rm = TRUE))

DT::datatable(reg_comp)
write_csv(reg_comp, file.path(dir_goal, 'int', 'fp_reg_compliance.csv'))

```

## Compute overall resilience

``` {r}
reg_comp <- read_csv(file.path(dir_goal, 'int', 'fp_reg_compliance.csv'))
reg_eff  <- read_csv(file.path(dir_goal, 'int', 'fp_reg_effectiveness.csv'))

reg_df <- reg_comp %>%
  full_join(reg_eff, by = 'year') %>%
  mutate(reg_score = 1) %>%
  gather(component, score, contains('score')) %>%
  select(year, component, score) %>%
  filter(!is.na(score)) %>%
  group_by(year) %>%
  summarize(components = paste0(component, ' = ', score, collapse = ', '),
            score = prod(score) ^ (1/n()))

write_csv(reg_df, file.path(dir_goal, 'output/fp_fishing_regs.csv'))

DT::datatable(reg_df)
```

-----

``` {r results = 'asis'}

prov_wrapup(commit_outputs = FALSE)

```
